### 2018-11-02 README.Rmd
added a README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# workingtitle

This repository contains the data and code for our paper:

> Authors, (YYYY). _Title of paper_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _Title of paper_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for 'Title of paper'_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: <https://github.com/nevrome/workingtitle/archive/master.zip>

Or you can install this compendium as an R package, workingtitle, from GitHub with:
```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("nevrome/workingtitle")
```
### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2018-11-02 analysis/paper/paper.Rmd
added analysis directory

---
title: "Title Goes Here"
author:
  - Author One
  - Author Two
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

library(workingtitle) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
```

# Introduction

Here is a citation [@Marwick2017]

# Background

```{r demo, eval = FALSE}
1 + 1
```

# Methods

# Results

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- readr::read_csv("../data/raw_data/my_csv_file.csv")
```

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```

### 2018-11-15 02_Materials_and_Methods.Rmd
changed the repo structure to a bookdown article

# Materials and Methods {#materials-and-methods}

Test

### 2018-11-15 index.Rmd
changed the repo structure to a bookdown article

---
output:
  pdf_document:
    keep_tex: true
    template: sagej_template.tex
bibliography: ["references.bib"]
biblio-style: apa
link-citations: yes
citation_package: biblatex
---

\runninghead{Smith and Wittkopf}

\title{workingtitle}

\author{Clemens Schmid\affilnum{1}}

\affiliation{\affilnum{1}Institute of Pre- and Protohistoric Archaeology, Kiel University, Kiel, DE}

\corrauth{
Clemens Schmid,
Institut für Ur- und Frühgeschichte,
Christian-Albrechts-Universität,
D - 24098 Kiel.
}

\email{clemens@nevrome.de}

\begin{abstract}
This paper is nice.
\end{abstract}

\keywords{Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission}

\maketitle

### 2018-11-15 _build.sh

@@ -1,0 +1,8 @@
+#!/usr/bin/env Rscript
+
+bookdown::render_book(
+  "index.Rmd",
+  output_file = "rendered_article/article.pdf",
+  output_dir = "rendered_article",
+  output_options = list(delete_merged_file = TRUE)
+)

### 2018-11-15 _clone_required_repos.sh

@@ -1,0 +1,9 @@
+git config --global user.email "clemens@nevrome.de"
+git config --global user.name "Clemens Schmid"
+
+cd ..
+
+git clone https://${GITHUB_PAT}@github.com/nevrome/neomod_analysis.git
+git clone https://${GITHUB_PAT}@github.com/nevrome/neiman1995.git
+
+cd neomod_textdev

### 2018-11-15 _deploy.sh

@@ -1,0 +1,10 @@
+# clone the repository to the book-output directory
+git clone -b document \
+  https://${GITHUB_PAT}@github.com/${TRAVIS_REPO_SLUG}.git \
+  document_out
+cd document_out
+git rm -rf *
+cp -r ../rendered_article/* ./
+git add --all *
+git commit -m "Update the rendered article"
+git push -q origin document

### 2018-11-16 01_Introduction.Rmd
test of reference rendering

# Introduction {#introduction}

The spread of ideas in human populations over space and time is a major topic of Cultural Evolution research [@creanzaCulturalEvolutionaryTheory2017]. Archaeology has a special perspective to offer here: It's concerned with questions and data that can embrace continent size areas and millennia of cultural development and interaction [@garveyCurrentPotentialRoles2018].

### 2018-11-19 01_Introduction.Rmd
some more text for the introduction

@@
 
-The spread of ideas in human populations over space and time is a major topic of Cultural Evolution research [@creanzaCulturalEvolutionaryTheory2017].
+The spread of ideas in human populations over space and time is a major topic of *Cultural Evolution* research [@creanzaCulturalEvolutionaryTheory2017].
 Archaeology has a special perspective to offer here: It's concerned with questions and data that can embrace continent size areas and millennia of cultural development and interaction [@garveyCurrentPotentialRoles2018].
+Archaeological data now is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and beyond that all kinds of waste and trash.
+Similarity between these artefacts within and beyond cultural-historical contexts are traditionally explained by *Cultural Transmission*: We rightfully expect humans to learn stories, traditions, ethical values and production techniques -- both functionally meaningful and seemingly pointless -- from others close to them.
+Material remains are a result of this long-term social exchange [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].

### 2018-11-19 01_Introduction.Rmd
started to collect some text for the introduction -- only preliminary

--- 2018-11-16 01_Introduction.Rmd
@@
 
-The spread of ideas in human populations over space and time is a major topic of Cultural Evolution research [@creanzaCulturalEvolutionaryTheory2017].
+The spread of ideas in human populations over space and time is a major topic of *Cultural Evolution* research [@creanzaCulturalEvolutionaryTheory2017].
 Archaeology has a special perspective to offer here: It's concerned with questions and data that can embrace continent size areas and millennia of cultural development and interaction [@garveyCurrentPotentialRoles2018].
+Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all kinds of waste and trash.
+Similarity between these artefacts within and beyond cultural-historical contexts are traditionally explained by human interaction: We rightfully expect humans to learn production as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them.
+The relatedness of historically and spatially close material remains generally is a result of long-term social exchange [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
+*Cultural Transmission Theory* is concerned with this exchange from the point of view of cultural evolution.
+
+In 1978 Robert Dunnell formulated in a contribution a previously discussed and in the following intensive to a distinction that has been taken up: In archaeological contexts and over long periods of time, ideas would have to be and behaviour patterns into two categories function (functions) and form (style).
+Functional Ideas on how the knowledge of the use of a cereal form, the competence of certain cereal Tools or weapons to manufacture or social recognition of child abundance are directly relevant to selection.
+This is where the biological evolution, the survival rate in crisis situations, or of demographic development, to favor certain ideas about their carriers.
+This Ideas are opposed by many others: Newtral traits that are not directly relevant for selection.
+such as ceramic decorations, traditional costume fashion or Jewelry shape.
+Survival or reproductive ability of a population are usually not affected by the nature of the geometric patterns on their storage vessels dependent.
+However, these characteristics pass through a partly breathtakingly fast development and allow archaeologists insights into relative chronology and social structure.
+According to Dunnell, in the context of the of Evolutionary Archaeology despite its genetically subordinate clout deserve special relevance - and a dedicated set of methods: Dunnell postulates, that functional ideas, taking into account the quality of their adaptation could be modelled, while stylistic Ideas essentially random, stochastic processes (Neutral Transmission) would follow.
+
+Dunnell's distinction fits in well with the other Context of the Dual Inheritance Theory (see Chapter 2.1.4) and has been taken up in a variety of ways, and extended.
+In reality, however, most of the Neutral variants hard selection criteria not completely sublime: Their creation causes in different Scope of costs in terms of working time and material.
+Since the distinction between stylistic and functional characteristics are not always obvious.
+it may be necessary to evaluate the functional usefulness of the of different variants with experiments and other methods of engineering science: The more similar the usefulness of all available variants in a context, the larger the probability that they will become selective relative to each other, and neutral.
+
+Fraser Neimans Article Stylistic Variation in Evolutionary Perspective: Inferences from Decorative Diversity and Interassemblage Distance in Illinois Woodland Ceramic Assemblages explores the new technologies proposed by Dunnell.
+stochastic processes and develops a simple framework for the simulation of propagation of non-functional ideas.
+In the process, the essential phenomena drift and flow (at Neiman: intergroup transmission) from population genetics in archaeology and to the interpretation of the Cultural distances between Indian ceramic inventories is used.
+Neiman's contribution was well received by the experts.
+very intensively received.
+Carl can be called Lipos works on ceramics in the Mississippi underflow, the Cultural Transmission Simulation with classical archaeological seriation, Shennans Investigation of the technological development of linear-band ceramics Ceramics in Merzbachtal or Timothy Kohlers analysis of the diversity of forms and its Role as proxy to the prehistoric social fabric in the southwest of the USA, again with ceramics .
+Also for the preparation of the simulation of this work was his contribution is an essential inspiration (see chapter 4.3.1).
+
+Evolutionary Archaeology has - parallel to the further development of the the science of archaeology - in the past 15 years of dynamism.
+The latest Developments show that they are more successful than almost any other archaeological faculty the scientific and Ideal for gradual, theoretical further development from one contribution to the next.
+During Neiman the concept of his simulation from population genetics was borrowed from other methodology for the Preparation and investigation of phylogenetic trees from cladistics to archaeological questions applied .
+Stylistic Variability is a very promising basis for modelling with regard to cultural (in comparison to spatial) distance.
+The Bayes' theorem has proven to be a potent statistical tool.
+to the model analysis and evaluation revealed .
+Archaeological data sets and their special features for Cultural evolution research is becoming more rigorous today.
+questioned and analyzed more consciously - beyond the anecdotal evidence.

### 2018-11-19 index.Rmd
new working title
small changes in build procedure

@@
 output:
-  pdf_document:
+  bookdown::pdf_book:
     keep_tex: true
@@
 
-\runninghead{Smith and Wittkopf}
+\runninghead{Cultural Transmission in Bronze Age burial rites}
 
-\title{workingtitle}
+\title{A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe}
 

### 2018-11-19 _build.sh

@@ -2,7 +2,5 @@
 
 bookdown::render_book(
   "index.Rmd",
-  output_file = "rendered_article/article.pdf",
-  output_dir = "rendered_article",
-  output_options = list(delete_merged_file = TRUE)
+  output_dir = "rendered_article"
 )

### 2018-11-20 index.Rmd
changed to rticles -- still no luck with sageapa

--- 2018-11-15 index.Rmd
@@
 ---
+title: A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe
+runninghead: Schmid.
+author:
+- name: Clemens Schmid
+  num: 1
+address:
+- num: 1
+  org: Institute of Pre- and Protohistoric Archaeology, Kiel University, Kiel, Germany
+corrauth: "Clemens Schmid, Institut für Ur- und Frühgeschichte, Christian-Albrechts-Universität, D - 24098 Kiel."
+email: clemens@nevrome.de
+abstract: "Petersilie."
+keywords: Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission
+classoption:
+  - Afour
+  - times
+bibliography: references
+bibliographystyle: apalike
 output:
-  pdf_document:
-    keep_tex: true
-    template: sagej_template.tex
-bibliography: ["references.bib"]
-biblio-style: apa
-link-citations: yes
-citation_package: biblatex
+  rticles::sage_article:
+    keep_tex: yes
 ---
 
-\runninghead{Smith and Wittkopf}
+```{r child = '01_Introduction.Rmd'}
+```
 
-\title{workingtitle}
+```{r child = '02_Materials_and_Methods.Rmd'}
+```
 
-\author{Clemens Schmid\affilnum{1}}
+```{r child = '03_Results.Rmd'}
+```
 
-\affiliation{\affilnum{1}Institute of Pre- and Protohistoric Archaeology, Kiel University, Kiel, DE}
-
-\corrauth{
-Clemens Schmid,
-Institut für Ur- und Frühgeschichte,
-Christian-Albrechts-Universität,
-D - 24098 Kiel.
-}
-
-\email{clemens@nevrome.de}
-
-\begin{abstract}
-This paper is nice.
-\end{abstract}
-
-\keywords{Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission}
-
-\maketitle
+```{r child = '04_Discussion.Rmd'}
+```

### 2018-11-20 _build.sh

@@ -1,6 +1,7 @@
 #!/usr/bin/env Rscript
 
-bookdown::render_book(
+rmarkdown::render(
   "index.Rmd",
-  output_dir = "rendered_article"
+  output_dir = "rendered_article",
+  output_file = "article.pdf"
 )

### 2018-11-24 index.Rmd
added preamble to get numbered section titles

@@
     keep_tex: yes
+    includes:
+        in_header: preamble.tex
 ---

### 2018-11-24 01_Introduction.Rmd
work on introduction text

@@
 The spread of ideas in human populations over space and time is a major topic of *Cultural Evolution* research [@creanzaCulturalEvolutionaryTheory2017].
-Archaeology has a special perspective to offer here: It's concerned with questions and data that can embrace continent size areas and millennia of cultural development and interaction [@garveyCurrentPotentialRoles2018].
-Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all kinds of waste and trash.
-Similarity between these artefacts within and beyond cultural-historical contexts are traditionally explained by human interaction: We rightfully expect humans to learn production as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them.
+Archaeology has a special perspective to offer here: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018].
+Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous remains of human live.
+Similarity between artefacts within and beyond cultural-historical contexts are traditionally explained by interaction: We rightfully expect humans to learn production as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them.
 The relatedness of historically and spatially close material remains generally is a result of long-term social exchange [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
-*Cultural Transmission Theory* is concerned with this exchange from the point of view of cultural evolution.
+*Cultural Transmission Theory* is concerned with this exchange from the point of view of Cultural Evolution.
 

### 2018-11-24 01_Introduction.Rmd
decided to start with a really introductory paragraph that gets explained/fleshed oout in the following sections

@@
 # Introduction {#introduction}
+
+This articles employs some of the methods that arose within archaeology's contemplation of *Cultural Transmission Theory* to trace the development of four major bronze age burial traditions in Europe.
+The cultural-historical evolution of archaeological cultures and ethnic groups during the european bronze age is complex and it's reconstruction an ongoing endeavour of archaeology.
+Within this history of constant emergence, displacement and demise many different forms of burial traditions were practiced.
+Most of them can be distinguished by the criteria *burial construction* and *treatment of the body*.
+Bronze age burials are either flat graves or part of a burial mound, and they contain either untreated or cremated human remains.
+Openly available bulk radiocarbon data now allows to reconstruct the trends within the distribution of these complex behaviours to test classical archaeological narratives about their respective prevalence.
+The comparison of macroregional trends also opens the perspective on long term patterns of interaction intensity on a continental scale.
 
@@
 The relatedness of historically and spatially close material remains generally is a result of long-term social exchange [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
-*Cultural Transmission Theory* is concerned with this exchange from the point of view of Cultural Evolution.
 

### 2018-11-25 01_Introduction.Rmd
started on some words about research history

@@
 The spread of ideas in human populations over space and time is a major topic of *Cultural Evolution* research [@creanzaCulturalEvolutionaryTheory2017].
+This theoretical framework is as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* to *Neoevolutionism* up until the ninteen sixties and seventies [@petermannGeschichteEthnologie2004].
+
 Archaeology has a special perspective to offer here: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018].

### 2018-11-25 01_Introduction.Rmd
more about research history

@@
 This theoretical framework is as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* to *Neoevolutionism* up until the ninteen sixties and seventies [@petermannGeschichteEthnologie2004].
+In the seventies the the discussion took another turn away from social structure to the more general observation of human behaviour.
+Three major theoretical movements emerged in this context: *Evolutionary Psychology*, *Human Behavioural Ecology* and *Dual Inheritance Theory* [@smithThreeStylesEvolutionary2000].
+The latter -- heavily popularized by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes ...
 

### 2018-11-25 01_Introduction.Rmd
more love for introduction

--- 2018-11-24 01_Introduction.Rmd
@@
 
-This articles employs some of the methods that arose within archaeology's contemplation of *Cultural Transmission Theory* to trace the development of four major bronze age burial traditions in Europe.
+This articles employs some of the methods that arose within archaeology's contemplation of *Cultural transmission theory* to trace the development of four major bronze age burial traditions in Europe.
 The cultural-historical evolution of archaeological cultures and ethnic groups during the european bronze age is complex and it's reconstruction an ongoing endeavour of archaeology.
@@
 
-The spread of ideas in human populations over space and time is a major topic of *Cultural Evolution* research [@creanzaCulturalEvolutionaryTheory2017].
-Archaeology has a special perspective to offer here: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018].
-Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous remains of human live.
-Similarity between artefacts within and beyond cultural-historical contexts are traditionally explained by interaction: We rightfully expect humans to learn production as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them.
-The relatedness of historically and spatially close material remains generally is a result of long-term social exchange [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
+The spread of ideas in human populations over space and time is a major topic of *Cultural evolution theory* [@creanzaCulturalEvolutionaryTheory2017].
+This framework is as least as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* and back to *Neoevolutionism* up until the ninteen sixties and seventies [@petermannGeschichteEthnologie2004].
+In the late seventies the the discussion took another turn away from social structure to the more general observation of human behaviour.
+Three major theoretical movements emerged in this context: *Evolutionary psychology*, *Human behavioural ecology* and *Dual inheritance theory* [@smithThreeStylesEvolutionary2000].
+The latter -- heavily popularized by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007].
+Just like in biological evolution there might be selection and mutation within *Social learning* and therefore methods of *Population genetics* could be employed within cultural anthropology -- of course with the much faster pace of cultural exchange in mind.
+A lot of theoretical and quantitative research arose from this assumptions: e.g.
+@cavalli-sforzaCulturalTransmissionEvolution1981, @boydCultureEvolutionaryProcess1985,  @teltserEvolutionaryArchaeologyMethodological1995 or @obrienEvolutionaryArchaeologyTheory1996.
+
+Archaeology has a special perspective to offer for Cultural evolution research: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018].
+Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous by-products of live.
+We rightfully expect humans to learn handcraft as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them.
+The similarity of historically and spatially close material remains can therefore be explained as a result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
 

### 2018-11-25 01_Introduction.Rmd
style vs. function

@@
 The similarity of historically and spatially close material remains can therefore be explained as a result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
+In this context a distinction of culture in function and style became an important question for *Evolutionary archaeology*, explicitly elaborated already by [@dunnell1978style].
+Some types of human behaviour obviously increase wellbeing and simplify survival, like certain hunting techniques, superior tools or sowing more fruitful crops.
+Without resorting to the inappropriate simplifications of Evolutionism their diffusion can be expected to increase over time.
+On the other hand there are behaviours that are seemingly not related to survival and reproductive capacity, like e.g.
+ceramic decorations, traditional costume fashion or jewelry design.
 
-In 1978 Robert Dunnell formulated in a contribution a previously discussed and in the following intensive to a distinction that has been taken up: In archaeological contexts and over long periods of time, ideas would have to be and behaviour patterns into two categories function (functions) and form (style).
-Functional Ideas on how the knowledge of the use of a cereal form, the competence of certain cereal Tools or weapons to manufacture or social recognition of child abundance are directly relevant to selection.
-This is where the biological evolution, the survival rate in crisis situations, or of demographic development, to favor certain ideas about their carriers.
-This Ideas are opposed by many others: Newtral traits that are not directly relevant for selection.
-such as ceramic decorations, traditional costume fashion or Jewelry shape.
-Survival or reproductive ability of a population are usually not affected by the nature of the geometric patterns on their storage vessels dependent.
 However, these characteristics pass through a partly breathtakingly fast development and allow archaeologists insights into relative chronology and social structure.

### 2018-11-25 01_Introduction.Rmd
stylistic variability

@@
 The similarity of historically and spatially close material remains can therefore be explained as a result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
-In this context a distinction of culture in function and style became an important question for *Evolutionary archaeology*, explicitly elaborated already by [@dunnell1978style].
+In this context a distinction of culture in function and style became an important question for *Evolutionary archaeology*, explicitly elaborated already by @dunnell1978style.
 Some types of human behaviour obviously increase wellbeing and simplify survival, like certain hunting techniques, superior tools or sowing more fruitful crops.
 Without resorting to the inappropriate simplifications of Evolutionism their diffusion can be expected to increase over time.
-On the other hand there are behaviours that are seemingly not related to survival and reproductive capacity, like e.g.
+On the other hand there are behaviours that are not obviously related to survival and reproductive capacity, like e.g.
 ceramic decorations, traditional costume fashion or jewelry design.
-
-However, these characteristics pass through a partly breathtakingly fast development and allow archaeologists insights into relative chronology and social structure.
-According to Dunnell, in the context of the of Evolutionary Archaeology despite its genetically subordinate clout deserve special relevance - and a dedicated set of methods: Dunnell postulates, that functional ideas, taking into account the quality of their adaptation could be modelled, while stylistic Ideas essentially random, stochastic processes (Neutral Transmission) would follow.
+Despite their seemingly secondary importance these characteristics often pass through a breathtakingly fast development and allow archaeologists sometimes precise insights into relative chronology.
+Dunnell postulates that the distribution and emergence of these stylistic cultural traits behave randomly on a population scale level.
+The stochastic processes that are at work here are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability*.
 

### 2018-11-25 01_Introduction.Rmd
some more references

@@
 Dunnell postulates that the distribution and emergence of these stylistic cultural traits behave randomly on a population scale level.
-The stochastic processes that are at work here are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability*.
+The stochastic processes that are at work here are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001].
 

### 2018-11-25 01_Introduction.Rmd
one more paragraph about burial traditions as stylistic or functional traits

@@
 Archaeology has a special perspective to offer for Cultural evolution research: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018].
-Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous by-products of live.
+Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous by-products of living.
 We rightfully expect humans to learn handcraft as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them.
@@
 ceramic decorations, traditional costume fashion or jewelry design.
-Despite their seemingly secondary importance these characteristics often pass through a breathtakingly fast development and allow archaeologists sometimes precise insights into relative chronology.
+Despite their seemingly secondary importance these characteristics often pass through a breathtakingly fast development and serve archaeologists as proxy for relative chronology.
 Dunnell postulates that the distribution and emergence of these stylistic cultural traits behave randomly on a population scale level.
@@
 
-Dunnell's distinction fits in well with the other Context of the Dual Inheritance Theory (see Chapter 2.1.4) and has been taken up in a variety of ways, and extended.
-In reality, however, most of the Neutral variants hard selection criteria not completely sublime: Their creation causes in different Scope of costs in terms of working time and material.
-Since the distinction between stylistic and functional characteristics are not always obvious.
-it may be necessary to evaluate the functional usefulness of the of different variants with experiments and other methods of engineering science: The more similar the usefulness of all available variants in a context, the larger the probability that they will become selective relative to each other, and neutral.
+In reality, beyond Dunnels distinction, most behaviour is not unambiguously stylistic or functional.
+For the sake of this case study burial rites are assumed to be stylistic, but in fact their creation causes costs in terms of working time and material to a varying extent.
+This cost might have been influenced or had their own effect on the human-nature relation: Pyres for example require a significant amount of wood and the erection of barrows several days of manpower.
+Still it can be expected that these functional disadvantages weren't critical most of the time for the general decision which kind of burial tradition is applied, but maybe for the magnitude of invested energy.
+Other factors to consider here are the level of grief [@macdonaldGriefBurialAmerican2001] or the representational intentions of the burying relatives [@henrichEvolutionCostlyDisplays2009].
 

### 2018-11-27 01_Introduction.Rmd
spelling

@@
 The spread of ideas in human populations over space and time is a major topic of *Cultural evolution theory* [@creanzaCulturalEvolutionaryTheory2017].
-This framework is as least as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* and back to *Neoevolutionism* up until the ninteen sixties and seventies [@petermannGeschichteEthnologie2004].
+This framework is as least as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* and back to *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004].
 In the late seventies the the discussion took another turn away from social structure to the more general observation of human behaviour.
@@
 Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous by-products of living.
-We rightfully expect humans to learn handcraft as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them.
+We rightfully expect humans to learn handcraft as well as building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them.
 The similarity of historically and spatially close material remains can therefore be explained as a result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].

### 2018-11-27 spellcheck.R
set default language to english

@@ -1,4 +1,4 @@
-dynspell <- function() {
+dynspell <- function(dict = hunspell::dictionary("en_GB")) {
 
   context <- rstudioapi::getSourceEditorContext()
 
@@ -16,7 +16,7 @@ dynspell <- function() {
   for (p1 in 1:length(row_texts)) {
     all_words <- unlist(stringr::str_split(row_texts[[p1]], " "))
     good_words <- stringr::str_subset(all_words, "^[^0-9]*$")
-    potentially_wrong_words <- unlist(hunspell::hunspell(good_words))
+    potentially_wrong_words <- unlist(hunspell::hunspell(good_words, dict = dict))
     if (length(potentially_wrong_words) == 0) { next }
     positions_raw <- stringr::str_locate_all(
       row_texts[p1],

### 2018-11-27 01_Introduction.Rmd
spelling

@@
 This articles employs some of the methods that arose within archaeology's contemplation of *Cultural transmission theory* to trace the development of four major bronze age burial traditions in Europe.
-The cultural-historical evolution of archaeological cultures and ethnic groups during the european bronze age is complex and it's reconstruction an ongoing endeavour of archaeology.
-Within this history of constant emergence, displacement and demise many different forms of burial traditions were practiced.
+The cultural-historical evolution of archaeological cultures and ethnic groups during the European bronze age is complex and it's reconstruction an ongoing endeavour of archaeology.
+Within this history of constant emergence, displacement and demise many different forms of burial traditions were practised.
 Most of them can be distinguished by the criteria *burial construction* and *treatment of the body*.
@@
 In this context a distinction of culture in function and style became an important question for *Evolutionary archaeology*, explicitly elaborated already by @dunnell1978style.
-Some types of human behaviour obviously increase wellbeing and simplify survival, like certain hunting techniques, superior tools or sowing more fruitful crops.
+Some types of human behaviour obviously increase well-being and simplify survival, like certain hunting techniques, superior tools or sowing more fruitful crops.
 Without resorting to the inappropriate simplifications of Evolutionism their diffusion can be expected to increase over time.
 On the other hand there are behaviours that are not obviously related to survival and reproductive capacity, like e.g.
-ceramic decorations, traditional costume fashion or jewelry design.
+ceramic decorations, traditional costume fashion or jewellery design.
 Despite their seemingly secondary importance these characteristics often pass through a breathtakingly fast development and serve archaeologists as proxy for relative chronology.
@@
 Archaeological data sets and their special features for Cultural evolution research is becoming more rigorous today.
-questioned and analyzed more consciously - beyond the anecdotal evidence.
+questioned and analysed more consciously - beyond the anecdotal evidence.

### 2018-11-27 01_Introduction.Rmd
moved remarks about stylistic nature of burial traditions to discussion

@@
 
-In reality, beyond Dunnels distinction, most behaviour is not unambiguously stylistic or functional.
-For the sake of this case study burial rites are assumed to be stylistic, but in fact their creation causes costs in terms of working time and material to a varying extent.
-This cost might have been influenced or had their own effect on the human-nature relation: Pyres for example require a significant amount of wood and the erection of barrows several days of manpower.
-Still it can be expected that these functional disadvantages weren't critical most of the time for the general decision which kind of burial tradition is applied, but maybe for the magnitude of invested energy.
-Other factors to consider here are the level of grief [@macdonaldGriefBurialAmerican2001] or the representational intentions of the burying relatives [@henrichEvolutionCostlyDisplays2009].
-
 Fraser Neimans Article Stylistic Variation in Evolutionary Perspective: Inferences from Decorative Diversity and Interassemblage Distance in Illinois Woodland Ceramic Assemblages explores the new technologies proposed by Dunnell.

### 2018-11-27 04_Discussion.Rmd
moved remarks about stylistic nature of burial traditions to discussion

# Discussion {#discussion}

In reality, beyond Dunnels distinction, most behaviour is not unambiguously stylistic or functional. For the sake of this case study burial rites are assumed to be stylistic, but in fact their creation causes costs in terms of working time and material to a varying extent. This cost might have been influenced or had their own effect on the human-nature relation: Pyres for example require a significant amount of wood and the erection of barrows several days of manpower. Still it can be expected that these functional disadvantages weren't critical most of the time for the general decision which kind of burial tradition is applied, but maybe for the magnitude of invested energy. Other factors to consider here are the level of grief [@macdonaldGriefBurialAmerican2001] or the representational intentions of the burying relatives [@henrichEvolutionCostlyDisplays2009].

### 2018-11-27 spellcheck.R
prevented spellcheck to choke on half selected words

@@ -23,6 +23,7 @@ dynspell <- function(dict = hunspell::dictionary("en_GB")) {
       paste0("([^\\p{L}])(", potentially_wrong_words, ")([^\\p{L}])")
     )
     positions <- do.call(rbind, positions_raw)
+    if (nrow(positions) == 0) { next }
     for (p2 in 1:nrow(positions)) {
       start <- rstudioapi::document_position(
         row = rows[p1],

### 2018-11-27 01_Introduction.Rmd
first sentences about the classic narrative

@@
 
-This articles employs some of the methods that arose within archaeology's contemplation of *Cultural transmission theory* to trace the development of four major bronze age burial traditions in Europe.
-The cultural-historical evolution of archaeological cultures and ethnic groups during the European bronze age is complex and it's reconstruction an ongoing endeavour of archaeology.
+This articles employs some of the methods that arose within archaeology's contemplation of *Cultural transmission theory* to trace the development of four major Bronze Age burial traditions in Europe.
+The cultural-historical evolution of archaeological cultures and ethnic groups during the European Bronze Age is complex and it's reconstruction an ongoing endeavour of archaeology.
 Within this history of constant emergence, displacement and demise many different forms of burial traditions were practised.
 Most of them can be distinguished by the criteria *burial construction* and *treatment of the body*.
-Bronze age burials are either flat graves or part of a burial mound, and they contain either untreated or cremated human remains.
+Bronze Age burials are either flat graves or part of a burial mound, and they contain either untreated or cremated human remains.
 Openly available bulk radiocarbon data now allows to reconstruct the trends within the distribution of these complex behaviours to test classical archaeological narratives about their respective prevalence.
 The comparison of macroregional trends also opens the perspective on long term patterns of interaction intensity on a continental scale.
+
+In a nutshell, the classic narrative of the development of Bronze Age burial rites in Europe presented e.g.
+by @hardingEuropeanSocietiesBronze2000a (following @hauslerBestattungssittenFruhenBronzezeit1977; @hauslerGrabUndBestattungssitten1994; @hauslerTotenorientierungUndGeographischer1996)  can be summarised as follows: In the early and middle Bronze Age inhumations dominated in different variations.
+Cremation burials in this time only occur in the Hungarian lowlands.
+Burial mounds are concentrated in parts of the Balkans as well as in eastern, western and northern Europe, while in central and southern Europe flat graves - mostly burials - predominate.
+In the Middle Bronze Age, barrows gained importance in Western and Central Europe -- a phenomenon highly linked to the emergence of the Tumulus culture in Central Europe.
+In the late Bronze Age, the Urnfield period, cremation became the most common burial custom in the context of Urnfield culture.
 
@@
 
+Burial traditions are ritualised behaviour most of the times heavily linked to religious and spiritual beliefs. For this paper they are assumed to be stylistic traits as defined by @dunnell1978style.
+
 Fraser Neimans Article Stylistic Variation in Evolutionary Perspective: Inferences from Decorative Diversity and Interassemblage Distance in Illinois Woodland Ceramic Assemblages explores the new technologies proposed by Dunnell.

### 2018-11-27 01_Introduction.Rmd
necessary additions to the classical narrative

@@
 In the late Bronze Age, the Urnfield period, cremation became the most common burial custom in the context of Urnfield culture.
+This presentation is, even when taking in account it's enormous simplification, partially misleading.
+Cremation was a lot more common already in the early and middle Bronze Age, though not in the context of the later Urnfield culture (e.g.
+in the Benelux area [@fokkensBronzeAgeLow2013] or Great-Britain [@robertsBritainIrelandBronze2013]).
 

### 2018-11-27 01_Introduction.Rmd
more work on the introduction

--- 2018-11-27 01_Introduction.Rmd
@@
 Bronze Age burials are either flat graves or part of a burial mound, and they contain either untreated or cremated human remains.
-Openly available bulk radiocarbon data now allows to reconstruct the trends within the distribution of these complex behaviours to test classical archaeological narratives about their respective prevalence.
-The comparison of macroregional trends also opens the perspective on long term patterns of interaction intensity on a continental scale.
-
-In a nutshell, the classic narrative of the development of Bronze Age burial rites in Europe presented e.g.
+In a nutshell, the classic narrative of the development of these rites in Europe presented e.g.
 by @hardingEuropeanSocietiesBronze2000a (following @hauslerBestattungssittenFruhenBronzezeit1977; @hauslerGrabUndBestattungssitten1994; @hauslerTotenorientierungUndGeographischer1996) can be summarised as follows: In the early and middle Bronze Age inhumations dominated in different variations.
 Cremation burials in this time only occur in the Hungarian lowlands.
-Burial mounds are concentrated in parts of the Balkans as well as in eastern, western and northern Europe, while in central and southern Europe flat graves - mostly burials - predominate.
-In the Middle Bronze Age, barrows gained importance in Western and Central Europe -- a phenomenon highly linked to the emergence of the Tumulus culture in Central Europe.
-In the late Bronze Age, the Urnfield period, cremation became the most common burial custom in the context of Urnfield culture.
+Burial mounds are concentrated in parts of the Balkans as well as in eastern, western and northern Europe, while in central and southern Europe flat graves -- mostly burials -- predominate.
+In the Middle Bronze Age, barrows gained importance in Western and Central Europe.
+This is highly linked to the emergence of the *Tumulus culture* in Central Europe.
+In the late Bronze Age, the Urnfield period, cremation became the most common burial custom in the context of *Urnfield culture*.
+
+Hardings narrative is, even when taking into account it's enormous simplification, partially misleading.
+Cremation, for example, was a lot more common already in the early and middle Bronze Age, though not in the context of the later Urnfield culture (e.g.
+in the Benelux area [@fokkensBronzeAgeLow2013] or Great-Britain [@robertsBritainIrelandBronze2013]).
+Beyond simple research progress, one additional reason why major misconceptions like this can appear and thrive might be a lack of unified data repositories and universally acknowledged proxies for the distribution of cultural phenomena in European archaeology: Instead of directly querrying quantitative data the typical process for the creation of large scale overviews is to excerpt the literature of all relevant local research traditions.
+The present article in contrast explores whether openly available bulk radiocarbon data could allow to reconstruct the trends within the distribution of burial rites.
+This could set the precedent for for a more reproducible and statistically sound reconstruction of the main shifts.
+The quantitative comparison of macroregional trends also opens the perspective on long term patterns of interaction intensity on a continental scale.
 

### 2018-11-27 01_Introduction.Rmd
and more

@@
 in the Benelux area [@fokkensBronzeAgeLow2013] or Great-Britain [@robertsBritainIrelandBronze2013]).
-Beyond simple research progress, one additional reason why major misconceptions like this can appear and thrive might be a lack of unified data repositories and universally acknowledged proxies for the distribution of cultural phenomena in European archaeology: Instead of directly querrying quantitative data the typical process for the creation of large scale overviews is to excerpt the literature of all relevant local research traditions.
+Beyond simple research progress, one additional reason why major misconceptions like this can appear and thrive might be a lack of unified data repositories and universally acknowledged proxies for the distribution of cultural phenomena in European archaeology: Instead of directly querying quantitative data the typical process for the creation of large scale overviews is to excerpt the literature of all relevant local research traditions.
 The present article in contrast explores whether openly available bulk radiocarbon data could allow to reconstruct the trends within the distribution of burial rites.
-This could set the precedent for for a more reproducible and statistically sound reconstruction of the main shifts.
-The quantitative comparison of macroregional trends also opens the perspective on long term patterns of interaction intensity on a continental scale.
+This could set the precedent for for a more reproducible and statistically sound observation of the main shifts.
+The quantitative comparison of macroregional trends also paves the way for an analysis of patterns of interaction intensity on a continental scale.
 
+Already the assumption that the cremation aspect of Early Bronze Age burial rituals in Britain can be directly compared to the same aspect in Hungary requires a theoretical foundation.
+*Cultural transmission theory* provides a framework where this claim can be explored as a hypothesis for further research.
 The spread of ideas in human populations over space and time is a major topic of *Cultural evolution theory* [@creanzaCulturalEvolutionaryTheory2017].
-This framework is as least as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* and back to *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004].
+This line of thought is as least as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* and back to *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004].
 In the late seventies the the discussion took another turn away from social structure to the more general observation of human behaviour.
@@
 ceramic decorations, traditional costume fashion or jewellery design.
-Despite their seemingly secondary importance these characteristics often pass through a breathtakingly fast development and serve archaeologists as proxy for relative chronology.
+Despite (or because?) their seemingly secondary importance these characteristics often pass through a breathtakingly fast development and serve archaeologists as proxy for relative chronology.
 Dunnell postulates that the distribution and emergence of these stylistic cultural traits behave randomly on a population scale level.
-The stochastic processes that are at work here are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001].
+The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001].
 

### 2018-11-27 01_Introduction.Rmd
removed displaced paragraph

@@
 Also for the preparation of the simulation of this work was his contribution is an essential inspiration (see chapter 4.3.1).
-
-Evolutionary Archaeology has - parallel to the further development of the the science of archaeology - in the past 15 years of dynamism.
-The latest Developments show that they are more successful than almost any other archaeological faculty the scientific and Ideal for gradual, theoretical further development from one contribution to the next.
-During Neiman the concept of his simulation from population genetics was borrowed from other methodology for the Preparation and investigation of phylogenetic trees from cladistics to archaeological questions applied .
-Stylistic Variability is a very promising basis for modelling with regard to cultural (in comparison to spatial) distance.
-The Bayes' theorem has proven to be a potent statistical tool.
-to the model analysis and evaluation revealed .
-Archaeological data sets and their special features for Cultural evolution research is becoming more rigorous today.
-questioned and analysed more consciously - beyond the anecdotal evidence.

### 2018-11-30 02_Materials_and_Methods.Rmd
some outline ideas for materials and methods
started to work on materials and methods

@@
 
-Test
+The analysis code of this paper is written in the statistics programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible: **doi**.
+
+The dataset used for the analysis in this paper is a subset of the openly available database Radon-B [@kneiselRadonB2013].
+It contains ^14^C dates from the European Bronze and Early Iron Age (2300 BC to 500 BC).
+Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B is a collection of more than 11,000 published dates  from around 3,500 sites from all over europe.
+
+The first step of data preparation was to download the Radon-B dataset with the R package c14bazAAR [@schmidC14bazAARDownloadPrepare2018], which allows to access several open radiocarbon archives.
+
+- Main steps of data preparation and grouping
+
+- Spatial Distance
+
+- Cultural distance
+
+The SED is defined as:
+
+$$d_{ij}^2 = \sum_{k = 1}^{n} (p_{ik} - p_{jk})^2$$
+
+- $d_{ij}^2$: SED between tqo groups $i$ and $j$
+- $k$: trait counter 
+- $n$: number of traits in the population
+- $p_{ik}$: relative frequency of the $k$th trait in group $i$
+- $p_{jk}$: relative frequency of the $k$th trait in group $j$
+
+- Simulation

### 2018-11-30 02_Materials_and_Methods.Rmd
more about radon-b

@@
 Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B is a collection of more than 11,000 published dates  from around 3,500 sites from all over europe.
+Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the the type and cultural attribution of the sampled context.
+The variable *Feature type* systematically distinguishes for example between settlements, rockshelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, etc..
+This information allows to accurately and precisely determine a moment in space and time, when a grave of a certain type was created.
+In theory the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe.
+Unfortunately the Radon-B dataset has some data quality issues: **TODO**.
+The data representativity and semantic consequences will be discussed below.
 

### 2018-12-01 02_Materials_and_Methods.Rmd
started to list the steps of data preparation

@@
 It contains ^14^C dates from the European Bronze and Early Iron Age (2300 BC to 500 BC).
-Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B is a collection of more than 11,000 published dates  from around 3,500 sites from all over europe.
+Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B (as of December 2018) is a collection of more than 11000 published dates from around 2800 sites from all over europe.
 Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the the type and cultural attribution of the sampled context.
@@
 
-The first step of data preparation was to download the Radon-B dataset with the R package c14bazAAR [@schmidC14bazAARDownloadPrepare2018], which allows to access several open radiocarbon archives.
+The following major steps of Data preparation were applied.
+Please see the code for the details of implementation.
+The value in brackets indicates the number of dates still available after the respective filtering operation.
+
+1. Downloading the Radon-B dataset with the R package c14bazAAR [@schmidC14bazAARDownloadPrepare2018], which allows to access several open radiocarbon archives. (11048)
+2. Removing all dates without age information or age information outside of the limits of the Intcal13 calibration curve: 71 - 46401 BP [@reimer_intcal13_2013]. (10956)
+3. Calibrating the ^14^C dates with the R package Bchron [@haslettSimpleMonotoneProcess2008], which implements the calibration algorithm with numerical integration.
+4.
 

### 2018-12-01 02_Materials_and_Methods.Rmd
more data preparation steps

@@
 3. Calibrating the ^14^C dates with the R package Bchron [@haslettSimpleMonotoneProcess2008], which implements the calibration algorithm with numerical integration.
-4.
+4. Reducing date selection to all dates whose $2\sigma$ calibrated age range overlap by at least one year with the predefined window of interest: 2200 -- 800 calBC. (7543)
+5. Reducing the date selection to all dates taken from graves as indicated by the *Feature type* variable. (2361)
+6. Transforming the variable *Feature type* via pattern matching to two new variables *burial type* (inhumation/cremation/unknown) and *burial construction* (mound/flat/unknown).
+7. Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwesetern Europe based on perceived spatial date distribution (Figure XXX).
+8. Reducing date selection to dates within the research area. (1894)
+9.
 

### 2018-12-01 02_Materials_and_Methods.Rmd
and even more

@@
 
-1. Downloading the Radon-B dataset with the R package c14bazAAR [@schmidC14bazAARDownloadPrepare2018], which allows to access several open radiocarbon archives. (11048)
-2. Removing all dates without age information or age information outside of the limits of the Intcal13 calibration curve: 71 - 46401 BP [@reimer_intcal13_2013]. (10956)
+1. Downloading the Radon-B dataset with the R package c14bazAAR [@schmidC14bazAARDownloadPrepare2018], which allows to access several open radiocarbon archives. (11048 dates)
+2. Removing all dates without age information or age information outside of the limits of the Intcal13 calibration curve: 71 - 46401 BP [@reimer_intcal13_2013]. (10956 dates)
 3. Calibrating the ^14^C dates with the R package Bchron [@haslettSimpleMonotoneProcess2008], which implements the calibration algorithm with numerical integration.
-4. Reducing date selection to all dates whose $2\sigma$ calibrated age range overlap by at least one year with the predefined window of interest: 2200 -- 800 calBC. (7543)
+4. Reducing date selection to all dates whose $2\sigma$ calibrated age range overlap by at least one year with the predefined window of interest: 2200 -- 800 calBC. (7543 dates)
 5. Reducing the date selection to all dates taken from graves as indicated by the *Feature type* variable. (2361)
 6. Transforming the variable *Feature type* via pattern matching to two new variables *burial type* (inhumation/cremation/unknown) and *burial construction* (mound/flat/unknown).
-7. Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwesetern Europe based on perceived spatial date distribution (Figure XXX).
-8. Reducing date selection to dates within the research area. (1894)
-9.
+7. Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial date distribution (Figure **TODO**).
+8. Reducing date selection to dates within the research area. (1894 dates)
+9. Removing ^14^C date doubles with tools of c14bazAAR. (1848 dates)
+10. Transforming data from the meaning *^14^C dates on graves* to *burial type at a certain time*. This is a more complex process, because **TODO** (1704 graves)
+11. Defining macroregions. **TODO**
+12. Reducing the date selection to the space covered by the macroregions and attributing the dates to the respective units. (1562 graves)
 
-- Main steps of data preparation and grouping
+The resulting dataset **TODO**
 

### 2018-12-01 02_Materials_and_Methods.Rmd
tightlist issue: https://tex.stackexchange.com/questions/257418/error-tightlist-converting-md-file-into-pdf-using-pandoc

@@
 1. Downloading the Radon-B dataset with the R package c14bazAAR [@schmidC14bazAARDownloadPrepare2018], which allows to access several open radiocarbon archives. (11048 dates)
+
 2. Removing all dates without age information or age information outside of the limits of the Intcal13 calibration curve: 71 - 46401 BP [@reimer_intcal13_2013]. (10956 dates)
+
 3. Calibrating the ^14^C dates with the R package Bchron [@haslettSimpleMonotoneProcess2008], which implements the calibration algorithm with numerical integration.
+
 4. Reducing date selection to all dates whose $2\sigma$ calibrated age range overlap by at least one year with the predefined window of interest: 2200 -- 800 calBC. (7543 dates)
+
 5. Reducing the date selection to all dates taken from graves as indicated by the *Feature type* variable. (2361)
+
 6. Transforming the variable *Feature type* via pattern matching to two new variables *burial type* (inhumation/cremation/unknown) and *burial construction* (mound/flat/unknown).
+
 7. Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial date distribution (Figure **TODO**).
+
 8. Reducing date selection to dates within the research area. (1894 dates)
+
 9. Removing ^14^C date doubles with tools of c14bazAAR. (1848 dates)
+
 10. Transforming data from the meaning *^14^C dates on graves* to *burial type at a certain time*. This is a more complex process, because **TODO** (1704 graves)
+
 11. Defining macroregions. **TODO**
+
 12. Reducing the date selection to the space covered by the macroregions and attributing the dates to the respective units. (1562 graves)
@@
 - $d_{ij}^2$: SED between tqo groups $i$ and $j$
+
 - $k$: trait counter
+
 - $n$: number of traits in the population
+
 - $p_{ik}$: relative frequency of the $k$th trait in group $i$
+
 - $p_{jk}$: relative frequency of the $k$th trait in group $j$

### 2018-12-01 02_Materials_and_Methods.Rmd
producing some skeleton text for the materials and methods section

@@
 
-The resulting dataset **TODO**
+The resulting dataset contains **TODO**
 
-- Spatial Distance
+This data for individual graves attributed to a macroregion can be transformed to a timeseries of trait prevalence within the respective regions during the the timeframe of interest.
+The first step is to create a absolute occurence time series **TODO**
 
-- Cultural distance
+The continous timeseries of yearwise burial rite proportions for each region allow to calculate timeseries of a measure of cultural distance for each region relation (Figure **TODO**).
+Several different measures of cultural distance could be used here following the discussion of these matters in the relevant literature (**TODO**).
+In this context the Squared Euclidian Distance (SED) seems to be a sensible statistic, because it is fairly simple, works with only few assumptions and emphasizes differences through squaring.
 
-The SED is defined as:
+The SED is defined here as:
 
@@
 
-- Simulation
+To test for correlation between the distances based on the variables *burial type* and *burial construction*, the timeframe of interest was devided into sections of 200 years.
+For each section of each the two variables, a matrix of the mean SED over time was calculated.
+The two matrizes of each section can be tested for correlation with the Mantel test, which **TODO**.
+The mean correlation for each of the $\frac{8*8}{2}-\frac{8}{2}=28$ region-region relations can be plotted in the *burial type*SED-*burial construction*SED space (Figure **TODO**).
+
+Defining a measure of spatial distance for the above defined main units of observation, the macroregions, is only possible in a highly simplified way.
+A contionus distance measure can't be employed, but it's possible to distinguish ordinal distance classes.
+For the eight regions four different categories of spatial distance were defined.
+
+Simulation **TODO**

### 2018-12-01 02_Materials_and_Methods.Rmd
more about methodology

@@
 For each section of each the two variables, a matrix of the mean SED over time was calculated.
-The two matrizes of each section can be tested for correlation with the Mantel test, which **TODO**.
+The two matrizes of each section can be tested for correlation with the Mantel test, which **TODO** [@mantelDetectionDiseaseClustering1967].
+Since both distances compared here, *burial type* and *burial construction* distance, are interval (**?**) scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test.
 The mean correlation for each of the $\frac{8*8}{2}-\frac{8}{2}=28$ region-region relations can be plotted in the *burial type*SED-*burial construction*SED space (Figure **TODO**).
@@
 
+The Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] because of the difference in the level of measurement of cultural and spatial distance.
+
 Simulation **TODO**

### 2018-12-02 02_Materials_and_Methods.Rmd
started to write on a section about the simulation

@@
 
-Simulation **TODO**
+To get a better understanding how the correlation coefficient should develop if there were strong correlation between spatial and cultural distance as measured by *burial type* and *burial construction*, a simple simulation was implemented based on the model proposed by @neimanStylisticVariationEvolutionary1995.
+This simulation assumes a population separated into $g$ groups with a group population of $N_g$ each.

### 2018-12-02 02_Materials_and_Methods.Rmd
more about the simulation

@@
 This simulation assumes a population separated into $g$ groups with a group population of $N_g$ each.
+At the beginning of the simulation there are $k$ ideas randomly, but equally distributed in this population.
+With every timestep -- $t$ defines the total number of timesteps -- all individuals learn from an other randomly chosen individual.
+The probability that this other individual is from an other group equals $m_i$, for the opposite therefore $1-m_i$.
+The partner group for each individual learning instance is selected randomly following a group interaction probability defined in the matrix $I$.

### 2018-12-02 02_Materials_and_Methods.Rmd
more work on materials&methods

@@
 The two matrizes of each section can be tested for correlation with the Mantel test, which **TODO** [@mantelDetectionDiseaseClustering1967].
-Since both distances compared here, *burial type* and *burial construction* distance, are interval (**?**) scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test.
+Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test.
 The mean correlation for each of the $\frac{8*8}{2}-\frac{8}{2}=28$ region-region relations can be plotted in the *burial type*SED-*burial construction*SED space (Figure **TODO**).
 
-Defining a measure of spatial distance for the above defined main units of observation, the macroregions, is only possible in a highly simplified way.
+Defining a measure of spatial distance for the above established main units of observation, the macroregions, is only possible in a highly simplified way.
 A contionus distance measure can't be employed, but it's possible to distinguish ordinal distance classes.
 For the eight regions four different categories of spatial distance were defined.
-
-The Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] because of the difference in the level of measurement of cultural and spatial distance.
+The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the difference in the data level of measurements (Figure **TODO**).
 
@@
 The partner group for each individual learning instance is selected randomly following a group interaction probability defined in the matrix $I$.
+The distribution of the $k$ variants can be observed for every timestep and documented in a timeseries like the one for real world data (Figure **TODO**).
+
+For the actual simulation runs, the parameters were chosen as follows: **TODO**

### 2018-12-02 02_Materials_and_Methods.Rmd
problems of Radon-B

@@
 The analysis code of this paper is written in the statistics programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible: **doi**.
+A more in-depth description and discussion of the data preparation can be found in the master thesis this article is based on **TODO**.
 
@@
 It contains ^14^C dates from the European Bronze and Early Iron Age (2300 BC to 500 BC).
-Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B (as of December 2018) is a collection of more than 11000 published dates from around 2800 sites from all over europe.
+Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B is a collection of more than 11000 published dates from around 2800 sites from all over europe (as of December 2018).
 Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the the type and cultural attribution of the sampled context.
@@
 In theory the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe.
-Unfortunately the Radon-B dataset has some data quality issues: **TODO**.
-The data representativity and semantic consequences will be discussed below.
+Unfortunately the Radon-B dataset has some quality issues mostly due to inconsistent manual data input: Incompleteness and faultiness of meta information, no supervised or well thought-out variable classification for *sample material* and *cultural attribution* and inconsistent multilingualism even within individual entries.
+These issues though do not inhibit the general data analysis strategy of this paper.
 
-The following major steps of Data preparation were applied.
-Please see the code for the details of implementation.
+To create a tidy and useful subset of Radon-B, the following major steps of data preparation were applied.
+Please see the code for details of implementation.
 The value in brackets indicates the number of dates still available after the respective filtering operation.

### 2018-12-02 02_Materials_and_Methods.Rmd
about date to grave transformation

@@
 
-10. Transforming data from the meaning *^14^C dates on graves* to *burial type at a certain time*. This is a more complex process, because **TODO** (1704 graves)
+10.
+Transforming data from the meaning *^14^C dates on graves* to *burial type at a certain time*.
+For 498 entries in the thus far assembled subset the attribution to *site* and *feature* is not unique.
+That means that for many graves multiple (up to 8) ^14^C dates have been created and entered in Radon-B.
+Dates can't be treated as semantically equal to graves.
+There's a variety of phenomena that cause this effect: In some cases safety measurements were taken -- multiple dates for one single grave that all fall within the margin of errors of each other.
+In other cases there is a set of strongly diverging dates for a collective grave that was used for hundreds of years.
+There are also cases where the *feature* information in the database is not sufficient to determine whether two dates are really from the same grave despite the identical entry.
+The following subalgorithm was applied to solve the problem automatically.
+This approach is compatible with the intention to create a cultural trait timeseries as explained below.
+(1704 graves)
+
+    1. If for an individual ^14^C date no numbers appear in it's *feature* variable value, than it is treaded as an individual grave.
+
+    2.
+For the remaining 252 problematic cases, groups were created by *site* and *feature*.
+The cummulated timespan covered by the $2\sigma$ age range of all these dates where taken as the relevant timespan for the respective grave.
 

### 2018-12-02 02_Materials_and_Methods.Rmd
about the definition of macroregions

@@
 
-11. Defining macroregions. **TODO**
+11.
+Defining macroregions to observe the development of *burial type* and *burial construction* within sensible geographical and cultural units that are appropriate to the amount of available data.
+The regions are circular and were delimited semiautomatically by creating a regular grid in accordance to main cultural-historical regions of the european Bronze Age (**citation**).
+The regions have a radius of 240km and overlap slightly (Figure **TODO**).
+That means some graves are counted towards not just one but two regions.
 

### 2018-12-02 02_Materials_and_Methods.Rmd
one more half sentence about the Mantel test

@@
 The continous timeseries of yearwise burial rite proportions for each region allow to calculate timeseries of a measure of cultural distance for each region relation (Figure **TODO**).
-Several different measures of cultural distance could be used here following the discussion of these matters in the relevant literature (**TODO**).
+Several different measures of cultural distance could be used here following the discussion of these matters in the relevant literature (**citation**).
 In this context the Squared Euclidian Distance (SED) seems to be a sensible statistic, because it is fairly simple, works with only few assumptions and emphasizes differences through squaring.
@@
 For each section of each the two variables, a matrix of the mean SED over time was calculated.
-The two matrizes of each section can be tested for correlation with the Mantel test, which **TODO** [@mantelDetectionDiseaseClustering1967].
+The two matrizes of each section can be tested for correlation with the Mantel test, which uses permutations of allegedly correlating matrizes to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967].
 Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test.

### 2018-12-02 02_Materials_and_Methods.Rmd
about the final analysis dataset

@@
 
-The resulting dataset contains **TODO**
+The resulting dataset is structured as a table with 1562 rows each representing one grave as compiled by 1701 ^14^C dates from 454 sites.
+The graves are linked to 41 different *period* terms and 25 different archaeological *cultures* as they are distinguished in Radon-B.
+1160 data were measured on sample material of bones and teeth (at least 169 of them burnt), 367 on wood and charcoal, a small part (20) of other materials such as nuts, resin or pitch.
+For the remaining 154 dates, no material information is stored.
+Regarding the variables burial type and burial construction, the conditions shown in Table (**TODO**) apply.
+
+```{r dprcrosstab, echo=FALSE, tidy=FALSE}
+data.frame(
+  flat = c(66,291,12), 
+  mound = c(96,95,117), 
+  unknown = c(241,224,559),
+  row.names = c("cremation", "inhumation", "unknown")
+) %>% 
+  knitr::kable(
+    caption = "Kreuztabelle",
+    format = "latex",
+    booktabs = T,
+    escape = F,
+    linesep = "\\addlinespace"
+  ) %>%
+  kableExtra::kable_styling(
+    full_width = T, 
+    font_size = 8,
+    latex_options = "hold_position"
+  )
+```
 

### 2018-12-02 02_Materials_and_Methods.Rmd
about timeseries

@@
 
-This data for individual graves attributed to a macroregion can be transformed to a timeseries of trait prevalence within the respective regions during the the timeframe of interest.
-The first step is to create a absolute occurence time series **TODO**
+This data for individual graves attributed to a macroregion can be transformed to a time series of trait prevalence within the respective regions.
+If a certain kind of grave can be attributed to a fuzzy point in time with radiocarbon dating, than we also know that most likely a ceremony was held around this time were the respective burial tradition was practised.
+That means that the cultural trait behind this behaviour must also have been present in one form or the other.
+An example: The idea *inhumation* must have been present in 1447 calBC in Great Britain since this year lies in the $2\sigma$ age range of the ^14^C date *NZA-32497* which was made for the Amesbury Down inhumation burial *I2639/25217*.
+For each year in the timeframe 2200 -- 800 calBC we can count the amount of graves with dates that cover this year with their $2\sigma$ age range.
+That allows the construction of regionwise time series of absolute occurrence numbers for the variables *burial type* and *burial construction* (Figure **TODO**).
+Although the number of dates is very small for some regions and timeperiods we still can attempt to calculate relative proportions of the mutually exclusiv traits (inhumation vs.
+cremation and mound grave vs.
+flat grave).
+That leads to relative timeseries of trait prevalence, which could be a powerful proxy for the development of the distribution of cultural behaviour.
 

### 2018-12-02 02_Materials_and_Methods.Rmd
spelling

@@
 It contains ^14^C dates from the European Bronze and Early Iron Age (2300 BC to 500 BC).
-Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B is a collection of more than 11000 published dates from around 2800 sites from all over europe (as of December 2018).
+Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B is a collection of more than 11000 published dates from around 2800 sites from all over Europe (as of December 2018).
 Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the the type and cultural attribution of the sampled context.
@@
 There are also cases where the *feature* information in the database is not sufficient to determine whether two dates are really from the same grave despite the identical entry.
-The following subalgorithm was applied to solve the problem automatically.
-This approach is compatible with the intention to create a cultural trait timeseries as explained below.
+The following steps were applied to solve the problem automatically.
+This approach is compatible with the intention to create a cultural trait time series as explained below.
 (1704 graves)
 
-    1. If for an individual ^14^C date no numbers appear in it's *feature* variable value, than it is treaded as an individual grave.
+    1. If for an individual ^14^C date no numbers appear in it's *feature* variable value, than it is treated as an individual grave.
 
@@
 For the remaining 252 problematic cases, groups were created by *site* and *feature*.
-The cummulated timespan covered by the $2\sigma$ age range of all these dates where taken as the relevant timespan for the respective grave.
+The cumulated timespan covered by the $2\sigma$ age range of all these dates where taken as the relevant timespan for the respective grave.
 
 11.
-Defining macroregions to observe the development of *burial type* and *burial construction* within sensible geographical and cultural units that are appropriate to the amount of available data.
-The regions are circular and were delimited semiautomatically by creating a regular grid in accordance to main cultural-historical regions of the european Bronze Age (**citation**).
+Defining macro regions to observe the development of *burial type* and *burial construction* within sensible geographical and cultural units that are appropriate to the amount of available data.
+The regions are circular and were delimited semi-automatically by creating a regular grid in accordance to main cultural-historical regions of the European Bronze Age (**citation**).
 The regions have a radius of 240km and overlap slightly (Figure **TODO**).
@@
 
-12. Reducing the date selection to the space covered by the macroregions and attributing the dates to the respective units. (1562 graves)
+12. Reducing the date selection to the space covered by the macro regions and attributing the dates to the respective units. (1562 graves)
 
@@
 
-This data for individual graves attributed to a macroregion can be transformed to a time series of trait prevalence within the respective regions.
+This data for individual graves attributed to a macro region can be transformed to a time series of trait prevalence within the respective regions.
 If a certain kind of grave can be attributed to a fuzzy point in time with radiocarbon dating, than we also know that most likely a ceremony was held around this time were the respective burial tradition was practised.
@@
 An example: The idea *inhumation* must have been present in 1447 calBC in Great Britain since this year lies in the $2\sigma$ age range of the ^14^C date *NZA-32497* which was made for the Amesbury Down inhumation burial *I2639/25217*.
-For each year in the timeframe 2200 -- 800 calBC we can count the amount of graves with dates that cover this year with their $2\sigma$ age range.
-That allows the construction of regionwise time series of absolute occurrence numbers for the variables *burial type* and *burial construction* (Figure **TODO**).
-Although the number of dates is very small for some regions and timeperiods we still can attempt to calculate relative proportions of the mutually exclusiv traits (inhumation vs.
+For each year in the time frame 2200 -- 800 calBC we can count the amount of graves with dates that cover this year with their $2\sigma$ age range.
+That allows the construction of region wise time series of absolute occurrence numbers for the variables *burial type* and *burial construction* (Figure **TODO**).
+Although the number of dates is very small for some regions and time periods we still can attempt to calculate relative proportions of the mutually exclusive traits (inhumation vs.
 cremation and mound grave vs.
 flat grave).
-That leads to relative timeseries of trait prevalence, which could be a powerful proxy for the development of the distribution of cultural behaviour.
+That leads to relative time series of trait prevalence, which could be a powerful proxy for the development of the distribution of cultural behaviour.
 
-The continous timeseries of yearwise burial rite proportions for each region allow to calculate timeseries of a measure of cultural distance for each region relation (Figure **TODO**).
+The continuous time series of year wise burial rite proportions for each region allow to calculate time series of a measure of cultural distance for each region-region relation (Figure **TODO**).
 Several different measures of cultural distance could be used here following the discussion of these matters in the relevant literature (**citation**).
@@
 
-To test for correlation between the distances based on the variables *burial type* and *burial construction*, the timeframe of interest was devided into sections of 200 years.
+To test for correlation between the distances based on the variables *burial type* and *burial construction*, the time frame of interest was divided into sections of 200 years.
 For each section of each the two variables, a matrix of the mean SED over time was calculated.
-The two matrizes of each section can be tested for correlation with the Mantel test, which uses permutations of allegedly correlating matrizes to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967].
+The two matrices of each section can be tested for correlation with the Mantel test, which uses permutations of allegedly correlating matrices to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967].
 Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test.
@@
 
-Defining a measure of spatial distance for the above established main units of observation, the macroregions, is only possible in a highly simplified way.
-A contionus distance measure can't be employed, but it's possible to distinguish ordinal distance classes.
+Defining a measure of spatial distance for the above established main units of observation, the macro regions, is only possible in a highly simplified way.
+A continuous distance measure can't be employed, but it's possible to distinguish ordinal distance classes.
 For the eight regions four different categories of spatial distance were defined.
@@
 At the beginning of the simulation there are $k$ ideas randomly, but equally distributed in this population.
-With every timestep -- $t$ defines the total number of timesteps -- all individuals learn from an other randomly chosen individual.
+With every timestep -- $t$ defines the total number of time steps -- all individuals learn from an other randomly chosen individual.
 The probability that this other individual is from an other group equals $m_i$, for the opposite therefore $1-m_i$.
 The partner group for each individual learning instance is selected randomly following a group interaction probability defined in the matrix $I$.
-The distribution of the $k$ variants can be observed for every timestep and documented in a timeseries like the one for real world data (Figure **TODO**).
+The distribution of the $k$ variants can be observed for every timestep and documented in a time series like the one for real world data (Figure **TODO**).
 

### 2018-12-08 article/01_Introduction.Rmd
started to prepare repo for code and data

--- 2018-11-27 01_Introduction.Rmd
@@
 The quantitative comparison of macroregional trends also paves the way for an analysis of patterns of interaction intensity on a continental scale.
+A core question is how cultural and spatial distance are related over the course of the European Bronze Age.
 
@@
 Dunnell postulates that the distribution and emergence of these stylistic cultural traits behave randomly on a population scale level.
-The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001].
+The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @porcicExploringEffectsAssemblage2015; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001].
 
-Burial traditions are ritualised behaviour most of the times heavily linked to religious and spiritual beliefs. For this paper they are assumed to be stylistic traits as defined by @dunnell1978style.
-
-Fraser Neimans Article Stylistic Variation in Evolutionary Perspective: Inferences from Decorative Diversity and Interassemblage Distance in Illinois Woodland Ceramic Assemblages explores the new technologies proposed by Dunnell.
-stochastic processes and develops a simple framework for the simulation of propagation of non-functional ideas.
-In the process, the essential phenomena drift and flow (at Neiman: intergroup transmission) from population genetics in archaeology and to the interpretation of the Cultural distances between Indian ceramic inventories is used.
-Neiman's contribution was well received by the experts.
-very intensively received.
-Carl can be called Lipos works on ceramics in the Mississippi underflow, the Cultural Transmission Simulation with classical archaeological seriation, Shennans Investigation of the technological development of linear-band ceramics Ceramics in Merzbachtal or Timothy Kohlers analysis of the diversity of forms and its Role as proxy to the prehistoric social fabric in the southwest of the USA, again with ceramics .
-Also for the preparation of the simulation of this work was his contribution is an essential inspiration (see chapter 4.3.1).
+@neimanStylisticVariationEvolutionary1995 takes a closer look at this processes and develops a simple framework for the simulation of non-functional idea propagation.
+Neiman identifies the essential phenomena *drift* (random dominance of equivalent traits) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.
+Furthermore he implements a measure of cultural distance to apply his theoretical insights to real world data.
+The workflow suggested by Neiman can be partially applied to the burial rites data examined in this article and was an important methodological inspiration: As burial traditions are ritualised behaviour and most of the time heavily linked to religious and spiritual beliefs they can be assumed to be stylistic traits as defined by @dunnell1978style.
+Besides @neimanStylisticVariationEvolutionary1995, another important template for the comparison of cultural and spatial was provided by @porcicSimulatingCulturalTransmission2014.

### 2018-12-08 article/03_Results.Rmd
started to prepare repo for code and data

# Results {#results}

- Characteristics and map of final dataset

- Absolute and relative temporal development of traits by region

- Cultural distance space

- Correlation of cultural distances and spatial distance

- Simulation: What is to be expected in case of influence of spatial distance?

### 2018-12-08 _build.sh

@@ -1,7 +1,7 @@
 #!/usr/bin/env Rscript
 
 rmarkdown::render(
-  "index.Rmd",
+  "article/index.Rmd",
   output_dir = "rendered_article",
   output_file = "article.pdf"
 )

### 2018-12-08 _clone_required_repos.sh

@@ -1,9 +1,0 @@
-git config --global user.email "clemens@nevrome.de"
-git config --global user.name "Clemens Schmid"
-
-cd ..
-
-git clone https://${GITHUB_PAT}@github.com/nevrome/neomod_analysis.git
-git clone https://${GITHUB_PAT}@github.com/nevrome/neiman1995.git
-
-cd neomod_textdev

### 2018-12-09 analysis/code/05_plot_scripts/map.R
even more love for the map

@@ -58,9 +58,9 @@ map_A <- ggplot() +
   ) +
   scale_size_manual(
     values = c(
-      "flat" = 10,
-      "mound" = 10,
-      "unknown" = 5
+      "flat" = 12,
+      "mound" = 12,
+      "unknown" = 6
     )
   ) +
   scale_color_manual(
@@ -151,8 +151,8 @@ map_B <- ggplot() +
 
 combined_map <- ggdraw() +
   draw_plot(map_A, 0, 0, 1, 1) +
-  draw_plot(map_B, 0.55, 0.06, 0.4, 0.4) +
-  draw_plot_label(c("A", "B"), c(0.08, 0.56), c(0.99, 0.4), size = 30)
+  draw_plot(map_B, 0.55, 0.05, 0.42, 0.42) +
+  draw_plot_label(c("A", "B"), c(0.06, 0.56), c(0.99, 0.4), size = 30)
 
 combined_map %>%
   ggsave(
@@ -161,6 +161,6 @@ combined_map %>%
     device = "jpeg",
     scale = 1,
     dpi = 300,
-    width = 330, height = 380, units = "mm",
+    width = 330, height = 400, units = "mm",
     limitsize = F
   )

### 2018-12-09 article/02_Materials_and_Methods.Rmd
added map figure to document

--- 2018-12-02 02_Materials_and_Methods.Rmd
@@
 
-7. Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial date distribution (Figure **TODO**).
+7.
+Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial date distribution (Figure \ref{map}).
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../analysis/figures/map.jpeg}
+\caption{\label{map}This is a figure.}
+\end{figure}
 

### 2018-12-09 analysis/code/run_all_scripts.R
walk should be better for run script than lapply

@@ -1,6 +1,6 @@
 library(magrittr)
 
-lapply(
+purrr::walk(
   c(
     "analysis/code/01_data_preparation_scripts",
     "analysis/code/02_dataset_description_scripts",
@@ -9,7 +9,7 @@ lapply(
     "analysis/code/05_plot_scripts"
   ),
   function(x) {
-    pbapply::pblapply(
+    purrr::walk(
       list.files(x, full.names = TRUE),
       function(y) {
         message("\n###### ", y, " ######\n")

### 2018-12-10 analysis/code/05_plot_scripts/development_burial_type.R
started to work on first development plot

@@ -1,0 +1,127 @@
+#### dependencies ####
+library(ggplot2)
+library(magrittr)
+source("analysis/code/helper_functions/geom_grob.R")
+
+##### load data ####
+load("analysis/data/tmp_data/development_amount_burial_type.RData")
+load("analysis/data/tmp_data/region_order.RData")
+load("analysis/data/tmp_data/development_proportions_burial_type.RData")
+
+#### prepare data ####
+amount_devel <- amount_development_burial_type
+regions_factor <- as.factor(amount_devel$region_name)
+amount_devel$region_name <- factor(regions_factor, levels = region_order)
+idea_factor <- as.factor(amount_devel$idea)
+amount_devel$idea <- factor(idea_factor, levels = rev(levels(idea_factor)))
+
+prop <- proportion_development_burial_type
+prop$idea <- as.factor(prop$idea)
+prop$idea <- factor(prop$idea , levels = rev(levels(prop$idea )))
+
+region_file_list <- unique(amount_devel$region_name) %>% gsub(" ", "_", ., fixed = TRUE)
+gl <- lapply(region_file_list, function(x) {
+  img <- png::readPNG(paste0("analysis/data/tmp_data/", x, ".png"))
+  g <- grid::rasterGrob(
+    img, interpolate = TRUE,
+    width = 0.14, height = 1.2
+  )
+})
+dummy <- tibble::tibble(region_name = unique(amount_devel$region_name), grob = gl )
+
+#### development_burial_type_A ####
+development_burial_type_A <- ggplot() +
+  geom_area(
+    data = amount_devel,
+    aes(x = timestep, y = n, fill = idea),
+    position = 'stack',
+    linetype = "blank"
+  ) +
+  facet_wrap(~region_name, nrow = 8) +
+  xlab("Time in years calBC") +
+  ylab("Amount of 14C dates from burials") +
+  labs(fill = "Ideas (mutually exclusive)") +
+  theme_bw() +
+  theme(
+    legend.position = "bottom",
+    panel.grid.major.x = element_line(colour = "black", size = 0.3),
+    axis.text = element_text(size = 15),
+    axis.title = element_text(size = 15),
+    strip.text.x = element_text(size = 13),
+    legend.title = element_text(size = 15, face = "bold"),
+    legend.text = element_text(size = 15)
+  ) +
+  scale_fill_manual(
[…]

### 2018-12-10 analysis/code/helper_functions/geom_grob.R

@@ -1,0 +1,51 @@
+library(ggplot2)
+library(tibble)
+library(gridExtra)
+library(grid)
+
+GeomCustom <- ggproto(
+  "GeomCustom",
+  Geom,
+  setup_data = function(self, data, params) {
+    data <- ggproto_parent(Geom, self)$setup_data(data, params)
+    data
+  },
+  
+  draw_group = function(data, panel_scales, coord) {
+    vp <- grid::viewport(x=data$x, y=data$y)
+    g <- grid::editGrob(data$grob[[1]], vp=vp)
+    ggplot2:::ggname("geom_custom", g)
+  },
+  
+  required_aes = c("grob","x","y")
+  
+)
+
+geom_custom <-  function(mapping = NULL,
+                         data = NULL,
+                         stat = "identity",
+                         position = "identity",
+                         na.rm = FALSE,
+                         show.legend = NA,
+                         inherit.aes = FALSE,
+                         ...) {
+  layer(
+    geom = GeomCustom,
+    mapping = mapping,
+    data = data,
+    stat = stat,
+    position = position,
+    show.legend = show.legend,
+    inherit.aes = inherit.aes,
+    params = list(na.rm = na.rm, ...)
+  )
+}
+
+# d <- tibble(x=1, f=rep(letters[1:11]))
+# gl <- lapply(1:4, function(c) rectGrob(gp=gpar(fill=c), width=0.2, height=0.2))
+# dummy <- tibble(f=letters[1:4], grob = gl )
+# 
+# ggplot(d, aes(x,x)) +
+#   facet_wrap(~f) +
+#   geom_custom(data=dummy, aes(grob=grob), inherit.aes = FALSE,
+#               x = 0.5, y=0.5)

### 2018-12-10 analysis/code/05_plot_scripts/development_burial_type.R
first draft of burial_type development plot

@@ -2,6 +2,7 @@
 library(ggplot2)
 library(magrittr)
 source("analysis/code/helper_functions/geom_grob.R")
+library(cowplot)
 
 ##### load data ####
 load("analysis/data/tmp_data/development_amount_burial_type.RData")
@@ -49,7 +50,8 @@ development_burial_type_A <- ggplot() +
     axis.title = element_text(size = 15),
     strip.text.x = element_text(size = 13),
     legend.title = element_text(size = 15, face = "bold"),
-    legend.text = element_text(size = 15)
+    legend.text = element_text(size = 15),
+    panel.border = element_rect(colour = "black", size = 2)
   ) +
   scale_fill_manual(
     values = c(
@@ -98,7 +100,8 @@ development_burial_type_B <- ggplot() +
     axis.title = element_text(size = 15),
     strip.text.x = element_text(size = 13),
     legend.title = element_text(size = 15, face = "bold"),
-    legend.text = element_text(size = 15)
+    legend.text = element_text(size = 15),
+    panel.border = element_rect(colour = "black", size = 2)
   ) +
   scale_fill_manual(
     values = c(
@@ -123,5 +126,27 @@ development_burial_type_B <- development_burial_type_B +
     aes(grob = grob),
     inherit.aes = FALSE,
     x = 0.1, y = 0.5
+  )
+
+#### combine plots ####
+
+combined_plot <- plot_grid(
+  development_burial_type_A,
+  development_burial_type_B,
+  labels = c("A", "B"),
+  nrow = 2,
+  align = "v",
+  label_size = 30
+)
+
+combined_plot %>%
+  ggsave(
+    "analysis/figures/development_burial_type",
+    plot = .,
+    device = "jpeg",
+    scale = 1,
+    dpi = 300,
+    width = 330, height = 850, units = "mm",
+    limitsize = F
   )
 

### 2018-12-10 article/02_Materials_and_Methods.Rmd
embedded draft

@@
 For each year in the time frame 2200 -- 800 calBC we can count the amount of graves with dates that cover this year with their $2\sigma$ age range.
-That allows the construction of region wise time series of absolute occurrence numbers for the variables *burial type* and *burial construction* (Figure **TODO**).
+That allows the construction of region wise time series of absolute occurrence numbers for the variables *burial type* and *burial construction* (Figure \ref{development_burial_type}).
 Although the number of dates is very small for some regions and time periods we still can attempt to calculate relative proportions of the mutually exclusive traits (inhumation vs.
@@
 That leads to relative time series of trait prevalence, which could be a powerful proxy for the development of the distribution of cultural behaviour.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../analysis/figures/development_burial_type.jpeg}
+\caption{\label{development_burial_type}This is a figure.}
+\end{figure}
 

### 2018-12-10 analysis/code/05_plot_scripts/development_burial_type.R

@@ -141,7 +141,7 @@ combined_plot <- plot_grid(
 
 combined_plot %>%
   ggsave(
-    "analysis/figures/development_burial_type",
+    "analysis/figures/development_burial_type.jpeg",
     plot = .,
     device = "jpeg",
     scale = 1,

### 2018-12-10 article/02_Materials_and_Methods.Rmd
development burial construction plot

@@
 For each year in the time frame 2200 -- 800 calBC we can count the amount of graves with dates that cover this year with their $2\sigma$ age range.
-That allows the construction of region wise time series of absolute occurrence numbers for the variables *burial type* and *burial construction* (Figure \ref{development_burial_type}).
+That allows the construction of region wise time series of absolute occurrence numbers for the variables *burial type* and *burial construction* (Figure \ref{development_burial_type} and \ref{development_burial_construction}).
 Although the number of dates is very small for some regions and time periods we still can attempt to calculate relative proportions of the mutually exclusive traits (inhumation vs.
@@
 \caption{\label{development_burial_type}This is a figure.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../analysis/figures/development_burial_construction.jpeg}
+\caption{\label{development_burial_construction}This is a figure.}
 \end{figure}

### 2018-12-10 analysis/code/05_plot_scripts/development_burial_construction.R

@@ -1,0 +1,151 @@
+#### dependencies ####
+library(ggplot2)
+library(magrittr)
+source("analysis/code/helper_functions/geom_grob.R")
+library(cowplot)
+
+##### load data ####
+load("analysis/data/tmp_data/development_amount_burial_construction.RData")
+load("analysis/data/tmp_data/region_order.RData")
+load("analysis/data/tmp_data/development_proportions_burial_construction.RData")
+
+#### prepare data ####
+amount_devel <- amount_development_burial_construction
+regions_factor <- as.factor(amount_devel$region_name)
+amount_devel$region_name <- factor(regions_factor, levels = region_order)
+idea_factor <- as.factor(amount_devel$idea)
+amount_devel$idea <- factor(idea_factor, levels = rev(levels(idea_factor)))
+
+prop <- proportion_development_burial_construction
+prop$idea <- as.factor(prop$idea)
+prop$idea <- factor(prop$idea , levels = rev(levels(prop$idea )))
+
+region_file_list <- unique(amount_devel$region_name) %>% gsub(" ", "_", ., fixed = TRUE)
+gl <- lapply(region_file_list, function(x) {
+  img <- png::readPNG(paste0("analysis/data/tmp_data/", x, ".png"))
+  g <- grid::rasterGrob(
+    img, interpolate = TRUE,
+    width = 0.14, height = 1.2
+  )
+})
+dummy <- tibble::tibble(region_name = unique(amount_devel$region_name), grob = gl )
+
+#### development_burial_construction_A ####
+development_burial_construction_A <- ggplot() +
+  geom_area(
+    data = amount_devel,
+    aes(x = timestep, y = n, fill = idea),
+    position = 'stack',
+    linetype = "blank"
+  ) +
+  facet_wrap(~region_name, nrow = 8) +
+  xlab("Time in years calBC") +
+  ylab("Amount of 14C dates from burials") +
+  labs(fill = "Ideas (mutually exclusive)") +
+  theme_bw() +
+  theme(
+    legend.position = "bottom",
+    panel.grid.major.x = element_line(colour = "black", size = 0.3),
+    axis.text = element_text(size = 15),
+    axis.title = element_text(size = 15),
+    strip.text.x = element_text(size = 13),
+    legend.title = element_text(size = 15, face = "bold"),
+    legend.text = element_text(size = 15),
[…]

### 2018-12-10 analysis/code/05_plot_scripts/development_burial_type.R
increased text size for all plots

@@ -46,11 +46,11 @@ development_burial_type_A <- ggplot() +
   theme(
     legend.position = "bottom",
     panel.grid.major.x = element_line(colour = "black", size = 0.3),
-    axis.text = element_text(size = 15),
-    axis.title = element_text(size = 15),
-    strip.text.x = element_text(size = 13),
-    legend.title = element_text(size = 15, face = "bold"),
-    legend.text = element_text(size = 15),
+    axis.text = element_text(size = 25),
+    axis.title = element_text(size = 25),
+    strip.text.x = element_text(size = 25),
+    legend.title = element_text(size = 25, face = "bold"),
+    legend.text = element_text(size = 25),
     panel.border = element_rect(colour = "black", size = 2)
   ) +
   scale_fill_manual(
@@ -96,11 +96,11 @@ development_burial_type_B <- ggplot() +
   theme(
     legend.position = "bottom",
     panel.grid.major.x = element_line(colour = "black", size = 0.3),
-    axis.text = element_text(size = 15),
-    axis.title = element_text(size = 15),
-    strip.text.x = element_text(size = 13),
-    legend.title = element_text(size = 15, face = "bold"),
-    legend.text = element_text(size = 15),
+    axis.text = element_text(size = 25),
+    axis.title = element_text(size = 25),
+    strip.text.x = element_text(size = 25),
+    legend.title = element_text(size = 25, face = "bold"),
+    legend.text = element_text(size = 25),
     panel.border = element_rect(colour = "black", size = 2)
   ) +
   scale_fill_manual(
@@ -118,6 +118,9 @@ development_burial_type_B <- ggplot() +
   scale_x_continuous(
     breaks = c(-2200, -2000, -1500, -1000, -800),
     limits = c(-2500, -800)
+  ) +
+  guides(
+    fill = FALSE
   )
 
 development_burial_type_B <- development_burial_type_B +
@@ -136,7 +139,7 @@ combined_plot <- plot_grid(
   labels = c("A", "B"),
   nrow = 2,
   align = "v",
-  label_size = 30
+  label_size = 35
 )
 
 combined_plot %>%

### 2018-12-10 analysis/code/05_plot_scripts/sed_region_matrix.R
more love for the matrix plot

@@ -39,7 +39,20 @@ for (row_region in region_order) {
             regionB == row_region
           )
         plot_starter <- regions_grid_subset %>%
-          ggplot()
+          ggplot() +
+          geom_line(
+            aes(time, sed, group = variable),
+            alpha = 0.3,
+            size = 0.5
+          ) +
+          geom_smooth(
+            aes(time, sed, group = variable, linetype = variable),
+            method = "loess",
+            span = 0.3,
+            size = 1,
+            colour = "black",
+            linetype = "dashed"
+          )
       } else {
         regions_grid_subset <- regions_grid %>%
           dplyr::filter(
@@ -48,21 +61,22 @@ for (row_region in region_order) {
             regionB == row_region
           )
         plot_starter <- regions_grid_subset %>%
-          ggplot()
+          ggplot()  +
+          geom_line(
+            aes(time, sed, group = variable),
+            alpha = 0.3,
+            size = 0.5
+          ) +
+          geom_smooth(
+            aes(time, sed, group = variable, linetype = variable),
+            method = "loess",
+            span = 0.3,
+            size = 1,
+            color = "black"
+          )
       }
 
       plot_list[[iter]] <- plot_starter +
-        geom_line(
-          aes(time, sed, group = variable),
-          alpha = 0.3,
-          size = 0.5
-        ) +
-        geom_smooth(
-          aes(time, sed, group = variable, linetype = variable),
-          method = "loess",
-          span = 0.3,
-          size = 1.5
-        ) +
         scale_x_continuous(
           breaks = c(-2000, -1500, -1000),
           limits = c(-2200, -800)
@@ -71,13 +85,6 @@ for (row_region in region_order) {
           limits = c(0, 2)
         ) +
         theme_bw() +
-        theme(
-          axis.text = element_text(size = 10),
-          axis.text.x = element_text(angle = 45, hjust = 1),
-          axis.title = element_text(size = 15),
-          strip.text = element_text(size = 9),
-          legend.position = "bottom"
-        ) +
         ylab("Squared Euclidian Distance") +
         xlab("Time")
 
@@ -95,7 +102,15 @@ sed_matrix <- ggmatrix(
   byrow = FALSE,
   switch = "y",
   labeller = label_wrap_gen()
-) + theme_bw()
+) + theme_bw() +
+  theme(
+    axis.text = element_text(size = 15),
[…]

### 2018-12-10 article/02_Materials_and_Methods.Rmd
added figure preliminary
better place and way to embed the plot matrix

@@
 
-The continuous time series of year wise burial rite proportions for each region allow to calculate time series of a measure of cultural distance for each region-region relation (Figure **TODO**).
+The continuous time series of year wise burial rite proportions for each region allow to calculate time series of a measure of cultural distance for each region-region relation (Figure \ref{sed_region_matrix}).
 Several different measures of cultural distance could be used here following the discussion of these matters in the relevant literature (**citation**).
 In this context the Squared Euclidian Distance (SED) seems to be a sensible statistic, because it is fairly simple, works with only few assumptions and emphasizes differences through squaring.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../analysis/figures/sed_region_matrix.jpeg}
+\caption{\label{sed_region_matrix}This is a figure.}
+\end{figure*}
 

### 2018-12-10 analysis/code/05_plot_scripts/sed_region_matrix.R

@@ -109,7 +109,7 @@ sed_matrix <- ggmatrix(
     axis.title = element_text(size = 15),
     strip.text = element_text(size = 10),
     legend.position = "bottom",
-    panel.border = element_rect(colour = "black", size = 2)
+    panel.border = element_rect(colour = "black", size = 1.5)
   )
 
 

### 2018-12-11 article/02_Materials_and_Methods.Rmd
improvements of table

@@
 Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test.
-The mean correlation for each of the $\frac{8*8}{2}-\frac{8}{2}=28$ region-region relations can be plotted in the *burial type*SED-*burial construction*SED space (Figure **TODO**).
+The test result are listed in Table \ref{distance_and_correlation_table}.
+
+```{r, echo = FALSE}
+load("../analysis/data/output_data/distance_and_correlation_table.RData")
+knitr::kable(
+  distance_and_correlation_table, 
+  "latex", 
+  caption = "\\label{distance_and_correlation_table}Demo table", 
+  booktabs = T,
+  linesep = "\\addlinespace \\hline \\addlinespace"
+) %>%
+  kableExtra::add_header_above(
+    c(" ", "burial type" = 3, "burial construction" = 3, "variable correlation" = 2)
+  ) %>%
+  kableExtra::kable_styling(
+    latex_options = c("hold_position")
+  ) %>%
+  kableExtra::column_spec(c(2,4,5,7,9), width = "1.2cm") %>%
+  kableExtra::column_spec(c(3,6,8), width = "1.8cm") %>%
+  gsub("table", "table*", .)
+```
 

### 2018-12-11 analysis/code/03_cultural_distance_scripts/07_sed_mantel_tables.R

@@ -1,24 +1,27 @@
+library(magrittr)
+
 #### burial_construction: sed - spatial distance ####
 load("analysis/data/tmp_data/mantel_sed_spatial_burial_construction.RData")
 load("analysis/data/tmp_data/squared_euclidian_distance_over_timeblocks_burial_construction.RData")
 burial_construction_distance <- sed_spatial_distance %>% dplyr::mutate(context = "burial_construction")
 burial_construction_mantel <- mantel_test_results
-names(burial_construction_mantel)[2:3] <- c("bc-spatial: Correlation coefficient", "bc-spatial: p-Value")
+names(burial_construction_mantel)[2:3] <- c("bc-sp: Correlation coefficient", "bc-sp: p-Value")
 
 #### burial_type: sed - spatial distance ####
 load("analysis/data/tmp_data/mantel_sed_spatial_burial_type.RData")
 load("analysis/data/tmp_data/squared_euclidian_distance_over_timeblocks_burial_type.RData")
 burial_type_distance <- sed_spatial_distance %>% dplyr::mutate(context = "burial_type")
 burial_type_mantel <- mantel_test_results
-names(burial_type_mantel)[2:3] <- c("bt-spatial: Correlation coefficient", "bt-spatial: p-Value")
+names(burial_type_mantel)[2:3] <- c("bt-sp: Correlation coefficient", "bt-sp: p-Value")
 
 #### burial_construction & burial_type: sed - sed ####
 load("analysis/data/tmp_data/mantel_sed_spatial_burial_type_burial_construction.RData")
 load("analysis/data/tmp_data/mantel_sed_spatial_burial_type_burial_construction.RData")
 burial_type_burial_construction_mantel <- mantel_test_results
+names(burial_type_burial_construction_mantel)[2:3] <- c("bt-bc: Correlation coefficient", "bt-bc: p-Value")
 
-#### master table ####
-master_table <- rbind(
+#### distance_and_correlation_table ####
+distance_and_correlation_table <- rbind(
   burial_type_distance,
   burial_construction_distance
 ) %>%
@@ -38,8 +41,17 @@ master_table <- rbind(
     by = "time"
   ) %>%
   dplyr::select(
-    1,2,4,5,3,6,7
+    1,3,6,7,2,4,5
+  ) %>%
+  dplyr::left_join(
+    burial_type_burial_construction_mantel,
+    by = "time"
+  ) %>%
+  dplyr::mutate_if(
+    is.numeric,
+    round,
+    3
   )
 
[…]

### 2018-12-14 analysis/code/02_dataset_description_scripts/observations_preparation.R
started to work with dynamic value integration

@@ -1,5 +1,14 @@
 storage_file <- "analysis/data/output_data/sf_prep.txt"
 
+
+
+#### bronze ####
+load("analysis/data/tmp_data/radonb.RData")
+txtstorage::store("size radonb", nrow(radonb), storage_file)
+rm(radonb)
+
+
+
 #### bronze ####
 load("analysis/data/tmp_data/bronze.RData")
 txtstorage::store("size bronze", nrow(bronze), storage_file)

### 2018-12-14 article/02_Materials_and_Methods.Rmd
started to work with dynamic value integration
replaced more numbers with code

@@
 
-1. Downloading the Radon-B dataset with the R package c14bazAAR [@schmidC14bazAARDownloadPrepare2018], which allows to access several open radiocarbon archives. (11048 dates)
+1. Downloading the Radon-B dataset with the R package c14bazAAR [@schmidC14bazAARDownloadPrepare2018], which allows to access several open radiocarbon archives. (`r r("size radonb", sf_prep)` dates)
 
-2. Removing all dates without age information or age information outside of the limits of the Intcal13 calibration curve: 71 - 46401 BP [@reimer_intcal13_2013]. (10956 dates)
+2. Removing all dates without age information or age information outside of the limits of the Intcal13 calibration curve: 71 - 46401 BP [@reimer_intcal13_2013]. (`r r("size bronze", sf_prep)` dates)
 
@@
 
-4. Reducing date selection to all dates whose $2\sigma$ calibrated age range overlap by at least one year with the predefined window of interest: 2200 -- 800 calBC. (7543 dates)
+4.
+Reducing date selection to all dates whose $2\sigma$ calibrated age range overlap by at least one year with the predefined window of interest: 2200 -- 800 calBC.
+(`r r("size bronze0", sf_prep)` dates)
 
-5. Reducing the date selection to all dates taken from graves as indicated by the *Feature type* variable. (2361)
+5. Reducing the date selection to all dates taken from graves as indicated by the *Feature type* variable. (`r r("size bronze05", sf_prep)` dates)
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{../analysis/figures/map.jpeg}
-\caption{\label{map}This is a figure.}
-\end{figure}
+8. Reducing date selection to dates within the research area. (`r r("size bronze15", sf_prep)` dates)
 
-8. Reducing date selection to dates within the research area. (1894 dates)
-
-9. Removing ^14^C date doubles with tools of c14bazAAR. (1848 dates)
+9. Removing ^14^C date doubles with tools of c14bazAAR. (`r r("size bronze16", sf_prep)` dates)
 
@@
 Transforming data from the meaning *^14^C dates on graves* to *burial type at a certain time*.
-For 498 entries in the thus far assembled subset the attribution to *site* and *feature* is not unique.
+For `r r("bronze15 multi dates one grave", sf_prep)` entries in the thus far assembled subset the attribution to *site* and *feature* is not unique.
 That means that for many graves multiple (up to 8) ^14^C dates have been created and entered in Radon-B.
@@
 This approach is compatible with the intention to create a cultural trait time series as explained below.
-(1704 graves)
+(`r r("size bronze17", sf_prep)` graves)
 
@@
     2.
-For the remaining 252 problematic cases, groups were created by *site* and *feature*.
+For the remaining `r r("bronze16 multi dates one grave with numbers", sf_prep)` problematic cases, groups were created by *site* and *feature*.
 The cumulated timespan covered by the $2\sigma$ age range of all these dates where taken as the relevant timespan for the respective grave.
@@
 
-12. Reducing the date selection to the space covered by the macro regions and attributing the dates to the respective units. (1562 graves)
+12. Reducing the date selection to the space covered by the macro regions and attributing the dates to the respective units. (`r r("gpr size", sf_desc)` graves)
 
-The resulting dataset is structured as a table with 1562 rows each representing one grave as compiled by 1701 ^14^C dates from 454 sites.
-The graves are linked to 41 different *period* terms and 25 different archaeological *cultures* as they are distinguished in Radon-B.
-1160 data were measured on sample material of bones and teeth (at least 169 of them burnt), 367 on wood and charcoal, a small part (20) of other materials such as nuts, resin or pitch.
-For the remaining 154 dates, no material information is stored.
-Regarding the variables burial type and burial construction, the conditions shown in Table (**TODO**) apply.
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../analysis/figures/map.jpeg}
+\caption{\label{map}This is a figure.}
+\end{figure}
+
+The resulting dataset is structured as a table with `r r("gpr size", sf_desc)` rows each representing one grave as compiled by `r r("dpr size", sf_desc)` ^14^C dates from `r r("dpr sites amount", sf_desc)` sites.
+The graves are linked to `r r("dpr period amount", sf_desc)` different *period* terms and `r r("dpr culture amount", sf_desc)` different archaeological *cultures* as they are distinguished in Radon-B.
+`r r("dpr material bone amount", sf_desc)` dates were measured on sample material of bones and teeth (at least `r r("dpr material cremated bones amount", sf_desc)` of them burnt), `r r("dpr material charcoal wood amount", sf_desc)` on wood and charcoal, a small part (`r r("dpr material other amount", sf_desc)`) of other materials such as nuts, resin or pitch.
+For the remaining `r r("dpr material unknown amount", sf_desc)` dates, no material information is stored.
+Regarding the variables burial type and burial construction, the conditions shown in Table \ref{tab:dprcrosstab} apply.
 
 ```{r dprcrosstab, echo=FALSE, tidy=FALSE}
-data.frame(
-  flat = c(66,291,12), 
-  mound = c(96,95,117), 
-  unknown = c(241,224,559),
-  row.names = c("cremation", "inhumation", "unknown")
-) %>% 
+load("../analysis/data/output_data/dprcrosstab.RData")
+dprcrosstab %>% 
+  unclass() %>% data.frame() %>% 
   knitr::kable(
-    caption = "Kreuztabelle",
+    caption = "Table of burial rite distribution within the active dataset",
     format = "latex",

### 2018-12-14 article/02_Materials_and_Methods.Rmd
added reference to my master thesis

@@
 The analysis code of this paper is written in the statistics programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible: **doi**.
-A more in-depth description and discussion of the data preparation can be found in the master thesis this article is based on **TODO**.
+A more in-depth description and discussion of the data preparation can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
 

### 2018-12-14 article/02_Materials_and_Methods.Rmd
better figure and table referencing

@@
 
-The analysis code of this paper is written in the statistics programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible: **doi**.
+The analysis code of this paper is written in the statistics programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible.
 A more in-depth description and discussion of the data preparation can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
@@
 7.
-Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial date distribution (Figure \ref{map}).
+Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial date distribution (Figure \ref{map} A).
 
@@
 The regions are circular and were delimited semi-automatically by creating a regular grid in accordance to main cultural-historical regions of the European Bronze Age (**citation**).
-The regions have a radius of 240km and overlap slightly (Figure **TODO**).
+The regions have a radius of 240km and overlap slightly (Figure \ref{map} B).
 That means some graves are counted towards not just one but two regions.
@@
 Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test.
-The test result are listed in Table \ref{distance_and_correlation_table}.
 
-```{r, echo = FALSE}
+Defining a measure of spatial distance for the above established main units of observation, the macro regions, is only possible in a highly simplified way.
+A continuous distance measure can't be employed, but it's possible to distinguish ordinal distance classes.
+For the eight regions four different categories of spatial distance were defined (Figure \ref{map} C).
+The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the difference in the data level of measurements.
+The test result are listed in Table \ref{tab:distanceandcorrelationtable*}.
+
+```{r distanceandcorrelationtable, echo=FALSE, tidy=FALSE}
 load("../analysis/data/output_data/distance_and_correlation_table.RData")
@@
   "latex", 
-  caption = "\\label{distance_and_correlation_table}Demo table", 
+  caption = "Mean cultural distance, correlation and mantel test results", 
   booktabs = T,
@@
 
-Defining a measure of spatial distance for the above established main units of observation, the macro regions, is only possible in a highly simplified way.
-A continuous distance measure can't be employed, but it's possible to distinguish ordinal distance classes.
-For the eight regions four different categories of spatial distance were defined.
-The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the difference in the data level of measurements (Figure **TODO**).
-
 To get a better understanding how the correlation coefficient should develop if there were strong correlation between spatial and cultural distance as measured by *burial type* and *burial construction*, a simple simulation was implemented based on the model proposed by @neimanStylisticVariationEvolutionary1995.
@@
 The partner group for each individual learning instance is selected randomly following a group interaction probability defined in the matrix $I$.
-The distribution of the $k$ variants can be observed for every timestep and documented in a time series like the one for real world data (Figure **TODO**).
+The distribution of the $k$ variants can be observed for every timestep and documented in a time series like the one for real world data.
 

### 2018-12-17 article/03_Results.Rmd
started to work on results chapter -- some discussion of data representativity. Might be better in the discussion

@@
 
-- Characteristics and map of final dataset
+Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region wise timeseries of trait occurrence.
+The figure part A with absolute numbers indicates areas and timeperiods with a lot, and others with very few points of data.
+For some there is not enough information to make any decision about the relative dominance and for many the general relevance and accuracy of the resulting timeseries can be questioned.
+There's a variety of reasons for this:
 
-- Absolute and relative temporal development of traits by region
+1. The data in Radon-B depends on the projects and interests of the involved researchers: The data was collected as needed rather then systematically for all of europe.
+
+2.
+The availability of ^14^C dates for a region and period in the literature depends on whether a modern research project took place within this context.
+Besides the research interets of researchers and institutions, there are also political reasons why an area might be better or worse explored.
+
+3.
+In the field of investigation of this work different research traditions meet, which in the past considered ^14^C data more or less essential for the construction of an absolute chronological assessment.
+In various Metal Age contexts, 14C data are inferior to the accuracy of relative chronological classifications - e.g.
+on the basis of primer typology - and are therefore used only sporadically to create absolute chronological fix points.
+Dendrochronological dating is preferred in moist soil contexts and when the necessary woods are available.
+The available ^14^C calibration curves form plateaus in some periods, which significantly limit the achievable dating accuracy of the data.
+If such an effect is known, then no data is ordered at all.
+
+4.
+^14^C dating requires a small but sufficiently contextualizable amount of organic finds.
+Due to the old wood and old water effect, short-lived sample materials such as bones of terrestrial organisms or charred cereal grains are preferred.
+The majority of ^14^C data for this work comes from bones and teeth.
+However, their preservation is subject to great differences depending on the soil [@brockPrescreeningTechniquesIdentification2010; @gordonSoilPHBone1981; @nielsen-marshBoneDiagenesisEuropean2007].
+In many mineral soil contexts no sufficiently good, dateable material is available.
+This can affect entire regions, if, for example, poor bone preservation prevails in general as a result of low-calcification soils.
 

### 2018-12-17 article/04_Discussion.Rmd
moved representativity analysis to discussion

# Discussion {#discussion}

Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region wise timeseries of trait occurrence. The figure part A with absolute numbers indicates areas and timeperiods with a lot, and others with very few points of data. For some there is not enough information to make any decision about the relative dominance and for many the general relevance and accuracy of the resulting timeseries can be questioned. There's a variety of reasons for this:

1. The data in Radon-B depends on the projects and interests of the involved researchers: The data was collected as needed for individual regions rather then systematically for all of europe.

2. The availability of ^14^C dates for a region and period in the literature depends on whether a modern research project took place within this context. Besides the research interets of researchers and institutions, there are also political reasons why an area might be better or worse explored.

3. In the field of investigation of this work different research traditions meet, which in the past considered ^14^C data more or less essential for the construction of an absolute chronological assessment. In various Metal Age contexts, 14C data are inferior to the accuracy of relative chronological classifications - e.g. on the basis of primer typology - and are therefore used only sporadically to create absolute chronological fix points. Dendrochronological dating is preferred in moist soil contexts and when the necessary woods are available. The available ^14^C calibration curves form plateaus in some periods, which significantly limit the achievable dating accuracy of the data. If such an effect is known, then no data is ordered at all.

4. ^14^C dating requires a small but sufficiently contextualizable amount of organic finds. Due to the old wood and old water effect, short-lived sample materials such as bones of terrestrial organisms or charred cereal grains are preferred. The majority of ^14^C data for this work comes from bones and teeth. However, their preservation is subject to great differences depending on the soil [@brockPrescreeningTechniquesIdentification2010; @gordonSoilPHBone1981; @nielsen-marshBoneDiagenesisEuropean2007]. In many mineral soil contexts no sufficiently good, dateable material is available. This can affect entire regions, if, for example, poor bone preservation prevails in general as a result of low-calcification soils.

5. From very few regions and periods of the European Bronze Age so many graves are documented that it can be assumed that a significant sample of the burial culture of the entire population has been researched. Instead, it is evident in several contexts that only the burials of individual social groups, gender or ethnic group were recorded: The quantity and nature of the known graves simply cannot be representative of all the deceased. In this case it can be assumed that further, deviant burial rituals were practiced, which left no remains or remains that were difficult to grasp archaeologically. These rituals are not shown in the ^14^C data from Radon-B.

- Comparison archaeological observations and reconstruction based on data

- Surprising result of no correlation: Meaning? Construction of an own story?

- General and concluding observations

The data representativity and semantic consequences will be discussed below.

In reality, beyond Dunnels distinction, most behaviour is not unambiguously stylistic or functional. For the sake of this case study burial rites are assumed to be stylistic, but in fact their creation causes costs in terms of working time and material to a varying extent. This cost might have been influenced or had their own effect on the human-nature relation: Pyres for example require a significant amount of wood and the erection of barrows several days of manpower. Still it can be expected that these functional disadvantages weren't critical most of the time for the general decision which kind of burial tradition is applied, but maybe for the magnitude of invested energy. Other factors to consider here are the level of grief [@macdonaldGriefBurialAmerican2001] or the representational intentions of the burying relatives [@henrichEvolutionCostlyDisplays2009].

### 2018-12-17 article/04_Discussion.Rmd
some more discussion of representativity

@@
 
+With regard to the primary variables *burial type* and *burial construction*, further special effects must be taken into account: The disturbance and robbery of graves was a widespread phenomenon in history and prehistory.
+Burial mounds were usually attacked more severely than flat graves due to their good visibility and could therefore be underrepresented in Radon-B.
+In many places they were also the subject of archaeological expeditions in the early days of archaeology at the end of the 19th and beginning of the 20th century.
+Of course, these excavations did not yield ^14^C-dates.
+Irrespective of this, it can be assumed that a large part of the burial mounds erected in the Bronze Age were destroyed by agricultural activity.
+This could also result in an under-representation of the archaeological findings [@hardingEuropeanSocietiesBronze2000a].
+On the other hand, if a burial mound is still present, it is more easily visible and recognized by archaeologists today: Mounds are often clearly identifiable in aerial photographs or digital terrain models, while flat graves are much more difficult to identify for today's remote sensing methods.
+It is difficult to estimate the impact of these effects in detail, but an overrepresentation of mounds seems more likely than an overrepresentation of flat graves.
+The dating of cremation graves on bone remains methodically differs from the dating of body graves because it is based on the extraction of the mineral, not the organic, fraction in the bone [@lantingDatingCrematedBones2001].
+The underlying chemical method was developed later, has only been used more frequently in the past 15 years and is associated with a number of particular difficulties [@olsenCharacterisationBlindTesting2008; @strydonck14CDatingCremated2009].
+Burnt bone is generally more susceptible to weathering than unburned bone.
+An under-representation of cremation graves in current data collections is therefore more plausible - although this effect can only be observed to a very limited extent in Radon-B.
+
 - Comparison archaeological observations and reconstruction based on data

### 2018-12-17 article/03_Results.Rmd
moved representativity analysis to discussion
decided to move region wise comparison to results section

--- 2018-12-08 article/03_Results.Rmd
@@
 
-- Characteristics and map of final dataset
-
-- Absolute and relative temporal development of traits by region
+Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region wise timeseries of trait occurrence.
+The relative proportion of each trait calculated based on ^14^C dates can be compared to the archaeological state of research:
 

### 2018-12-17 article/04_Discussion.Rmd
added conclusion segment
decided to move region wise comparison to results section

@@
 
-Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region wise timeseries of trait occurrence.
-The figure part A with absolute numbers indicates areas and timeperiods with a lot, and others with very few points of data.
+The A sections of the figures \ref{development_burial_type} and \ref{development_burial_construction} with absolute grave numbers per region and year indicate areas and timeperiods with a lot, and others with very few points of data.
 For some there is not enough information to make any decision about the relative dominance and for many the general relevance and accuracy of the resulting timeseries can be questioned.
@@
 
-- General and concluding observations
-
-The data representativity and semantic consequences will be discussed below.
-
 In reality, beyond Dunnels distinction, most behaviour is not unambiguously stylistic or functional.

### 2018-12-17 article/03_Results.Rmd
copied text from master thesis for comparison -- has to be shortend/adjusted now

@@
 The relative proportion of each trait calculated based on ^14^C dates can be compared to the archaeological state of research:
+
+`r a("Southeastern Central Europe")`^[Anzahl der in Radon-B durch ^14^C-Daten repräsentierten Gräber in dieser Region.] \newline 
+The artificial study region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, western Slovakia, Moravia and large parts of Bohemia.
+The density of ^14^C data on graves from radon B is low -- individual data are available from the various regions, but large areas must manage without evidence.
+Diachronic, the data volume is consistently low and reaches a low point in the Middle Bronze Age.
+While for the *burial type* variable a constant flow of information is given, the number of graves without information on *burial construction* increases in the Late Bronze Age to such an extent that no statement at all can be derived for them.
+To assess the relative development within the variables *burial type* and *burial construction*, as predicted with the Radon-B data set on the basis of ^14^C data, a résumé must first be drawn from the data in chapter \@ref(regions-archaeological-overview).
+This is complicated by the fact that the region intersects diverse geographical and cultural areas.
+In the southeast of Central Europe, cremation and bodily burial already existed side by side in the Early Bronze Age, and the local tradition remained stable in this respect also due to the expansion of the burial mound tradition in the Middle Bronze Age -- with a slight overall increase in cremation graves.
+The Urnfield period brought the transition to the absolute dominance of cremation burial.
+The development of the *burial type* variable in the Radon-B proxy shows this development correctly.
+The information from ^14^C data on *burial construction* correctly traces the relevance of flat burials in the Early Bronze Age, but the latter are then overemphasized in the Middle Bronze Age compared to the construction of burial mounds.
+In the Late Bronze Age, burials on flat burial grounds were frequent, although hills of the burial mound culture were also further documented.
+This circumstance -- certainly due to the small amount of data -- cannot be correctly deduced from radon-B.
+
+`r a("Poland")` \newline 
+The region *Poland* completely covers the centre and west of today's Poland.
+However, the ^14^C data from this region come almost exclusively from the voivodships of Kajuwien-Pommern, Lodsch and Lower Silesia.
+The amount of data is small overall, with significantly more data from the Early Bronze Age up to about 1700calBC.
+Almost no information is available from the Middle and Late Bronze Ages.
+This applies to both the *burial type* and the *burial construction* variables, the former being slightly better documented.
+The relative development of *burial type* approximates the *real* conditions -- i.e.
+according to the classical, archaeological narrative --: body burials were actually significantly more frequent in western Poland until the Late Bronze Age, while burials did not occur until the Urnfield Period.
+Also before, cremation graves had been laid out again and again, but in principle the tendency recorded by Radon-B is correct.
+In contrast, *burial construction* is misrepresented insofar as the burial mounds also played an important role in the Middle Bronze Age, which the dataset does not appreciate.
+
+`r a("Southern Germany")` \newline 
+The artificial district region *Southern Germany* covers Bavaria completely, but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and Western Austria.
+No ^14^C data are available from the latter regions.
+The majority of the available data comes from the Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative districts of Tübingen.
+For the Early Bronze Age a remarkable amount of data from Southern Germany are preserved in Radon-B, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves.
+The *burial type* variable is quite well understood -- although disproportionately much more of the early body graves were recorded than urnfield burials.
+This wealth of information contrasts sharply with the situation regarding the *burial construction* variable, for which almost no information from this area is available throughout the Bronze Age.
+If one considers the relative development calculated from Radon-B and compares it with the archaeological literature, then a good agreement can be found for *burial type*.
+Body burial dominated for a long time, but was then replaced by cremation from the Middle Bronze Age onwards and massively in the Late Bronze Age.
+The calculated proportions for *burial construction* could easily be dismissed as completely worthless due to the negligibly small amount of data, but in fact the proxy indicates the right tendency for early and middle Bronze Age: The dominance of flat graves was replaced by hill burials.
+The data set does not allow any statements to be made about the Late Bronze Age.
+
+`r a("Northeastern France")` \newline 
+*Northeastern France* is a district region comprising the French administrative regions Grand-Est and Bourgogne-Franche-Comté, the east of Baden-Württemberg, Rhineland-Palatinate and Saarland, as well as Luxembourg and Wallonia in Belgium.
+Accumulations of ^14^C data are only available from Württemberg and towards the Paris Basin.
+The amount of data is low overall, but very low in the early and early Middle Bronze Age.
+For *burial type*, after all, limited statements are possible after this time window, *burial construction* on the other hand must remain obscure.
+The prediction of the relative development of these variables is correspondingly error-prone.
+After all, the parallel existence of cremation and body burial in early and middle Bronze Age is correctly reproduced.
+However, this already existed in the Early Bronze Age and certainly developed more nuanced than the proxy might suggest.
+Contrary to the derivation from the ^14^C ratios, flat graves were quite common in *Northeastern France*.
+
+`r a("Northern Germany")` \newline 
+The artificial region *Northern Germany* includes all new federal states in East Germany, but also large parts of Lower Saxony and Schleswig-Holstein.
+The ^14^C data for this region come mainly from Saxony-Anhalt and neighbouring regions in Thuringia, Saxony and Lower Saxony.
+Also from the area north of Bremen some data are documented in Radon-B.
+The database does not contain as many dated graves from any other region as from northern Germany.
+Most of them date back to the Early Bronze Age, while the Middle Bronze Age from about 1800calBC to 1400calBC is a striking period for some centuries.
+Only for the Late Bronze Age is a significant number of data documented again.
+The meta information on *burial type* and *burial construction* trace this development to about the same extent.
+Since in Northern Germany influences from all four cardinal points come together and contribute significantly to the formation of local cultural phenomena, it is difficult to estimate the overall development of the primary variables here as well.
+The ^14^C proxy correctly represents *burial site* insofar as bodily burials dominated in the Early Bronze Age and are replaced by burials especially in the North from the Middle Bronze Age onwards.
+In the Late Bronze Age cremation was clearly more dominant than the data set suggests.
+With regard to *burial construction*, the current Radon-B stock records the basic dominance of flat graves well, but attributes more relevance to them in the Early and Middle Bronze Age than is usual in the archaeological assessment of the context.
+
+`r a("Southern Scandinavia")` \newline 
+The region *Southern Scandinavia* includes all of Denmark with Jutland, Funen, Zeeland and all smaller islands as well as large parts of the east of southern Sweden, and coastal regions in Schleswig-Holstein and Mecklenburg-Western Pomerania.
+The ^14^C data are very well distributed across Jutland, Funen and the Swedish Skåne region.
+Since the Bronze Age in Scandinavia begins much later than in Central Europe and is measured by another chronological system, the diachronic developments cannot simply be compared by means of the tripartite structure of the Early, Middle and Late Bronze Ages.
+The majority of data from Scandinavia are from the period IB onwards.
+In this respect, the data density for almost the entire Bronze Age is good -- also with regard to the two variables *burial type* and *burial construction*.
+The Late Neolithic from 2200calBC is a desideratum -- although at a higher level than, for example, the Early Bronze Age in *Northeastern France*.
+The adjustment in Figure \@ref(fig:development-proportions-regions-pseudoquant) is difficult, since the period which is labeled with Early Bronze Age due to the situation in Central Europe actually corresponds to the local Late Neolithic.
+If one carries out the resulting mental shift -- the Central European Middle Bronze Age corresponds to the Nordic Early Bronze Age -- and then compares the proxy of the relative grave form development with the archaeological observations, then one finds a solid agreement: body graves were the rule in periods I and II (about 1800-1300calBC), but cremations also occurred sporadically.
+From period III onwards cremation burials gained the upper hand.
+In the Scandinavian Early Bronze Age (i.e.
+Middle Bronze Age in Figure \@ref(fig:development-proportions-regions-pseudoquant)) burials in burial mounds were the absolute rule.
+In the Late Bronze Age, shallow burial fields were added.
+
+`r a("Benelux")` \newline 
+The Benelux region includes the Netherlands, northern Belgium and large parts of Lower Saxony and North Rhine-Westphalia.
+All these regions have ^14^C data on graves in Radon-B, but the data density in Belgium and the Netherlands is much higher.
+As in Southern Scandinavia, the Early Bronze Age is under-represented in the Benelux region as opposed to the Middle and Late Bronze Age, although in a slightly weaker form.
+While the information density of the *burial type* variable follows this course, *burial construction* is almost undocumented in the Late Bronze Age.
+As in the *Southern Germany* region, there is a serious difference between the information density of both variables.
+If one considers the relative development of the variables as they can be derived from the grave data in comparison to the archaeological doctrine, both agreement and deviations result here in detail.
+Here, too, the study region includes different cultural phenomena with partially diametrically different burial customs, but a general assessment of the conditions is nevertheless possible.
+In the north and east of the Benelux region body burials were the rule until the Late Bronze Age, while burials on fire dominated earlier in the south of the Meuse.
+The Radon-B proxy seems to overemphasise the dominance of cremation in the whole area.
+The relevance of hill burials in the Early and Middle Bronze Age, on the other hand, is correctly reflected, and the mixture between hill and flat graves in the context of urnfields, often resulting from the continued use of existing monumental sites, is also visible.
+
+`r a("England")` \newline 
+The *England* region stretches across England to the heights of York and Liverpool, incorporating parts of Wales and leaving out Cornwall and Devon in the southwest.
+In fact, it also touches a narrow coastal strip on the mainland, from which no ^14^C data are available.
+The available data are concentrated in the southeast of Britain, but are well distributed there.
+The diachronic data availability is relatively good in the early and middle Bronze Age, but then decreases in the late Bronze Age.
+*burial type* and *burial construction* behave accordingly.
+The relative development derived from the ^14^C data indicates the correct trend for *burial type*: Cremation graves became more and more important during the Bronze Age, while body burials were practiced for a long time in parallel.
+However, the data set emphasizes the dominance of the cremation graves too much.
+With regard to the variables *burial construction*, the basic dominance of burial mounds corresponds perfectly to the archaeological narrative, but nevertheless -- especially in the Late Bronze Age -- flat grave fields also occurred, which the data set wrongly excludes.
 

### 2018-12-17 article/03_Results.Rmd
formatting improvements

@@
 Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region wise timeseries of trait occurrence.
-The relative proportion of each trait calculated based on ^14^C dates can be compared to the archaeological state of research:
+The relative proportion of each trait calculated based on ^14^C dates can be compared to the archaeological state of research: \newline
 
-`r a("Southeastern Central Europe")`^[Anzahl der in Radon-B durch ^14^C-Daten repräsentierten Gräber in dieser Region.] \newline 
+`r a("Southeastern Central Europe")`^[Amount of graves in Radon-B for this region.]:
 The artificial study region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, western Slovakia, Moravia and large parts of Bohemia.
@@
 
-`r a("Poland")` \newline 
+`r a("Poland")`:
 The region *Poland* completely covers the centre and west of today's Poland.
@@
 
-`r a("Southern Germany")` \newline 
+`r a("Southern Germany")`:
 The artificial district region *Southern Germany* covers Bavaria completely, but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and Western Austria.
@@
 
-`r a("Northeastern France")` \newline 
+`r a("Northeastern France")`:
 *Northeastern France* is a district region comprising the French administrative regions Grand-Est and Bourgogne-Franche-Comté, the east of Baden-Württemberg, Rhineland-Palatinate and Saarland, as well as Luxembourg and Wallonia in Belgium.
@@
 
-`r a("Northern Germany")` \newline 
+`r a("Northern Germany")`:
 The artificial region *Northern Germany* includes all new federal states in East Germany, but also large parts of Lower Saxony and Schleswig-Holstein.
@@
 
-`r a("Southern Scandinavia")` \newline 
+`r a("Southern Scandinavia")`:
 The region *Southern Scandinavia* includes all of Denmark with Jutland, Funen, Zeeland and all smaller islands as well as large parts of the east of southern Sweden, and coastal regions in Schleswig-Holstein and Mecklenburg-Western Pomerania.
@@
 
-`r a("Benelux")` \newline 
+`r a("Benelux")`:
 The Benelux region includes the Netherlands, northern Belgium and large parts of Lower Saxony and North Rhine-Westphalia.
@@
 
-`r a("England")` \newline 
+`r a("England")`:
 The *England* region stretches across England to the heights of York and Liverpool, incorporating parts of Wales and leaving out Cornwall and Devon in the southwest.

### 2018-12-17 analysis/code/02_dataset_description_scripts/observations_preparation.R

@@ -130,5 +130,5 @@ regions_graves_amounts <- regions %>% sf::st_set_geometry(NULL) %>%
     region_name = factor(regions$NAME, levels = region_order)
   ) %>%
   dplyr::arrange(region_name) %$%
-  paste(paste0("*", region_name, "*"), paste0("(", number_of_graves, ")"), collapse = ", ", sep = " ")
+  paste(paste0("**", region_name, "**"), paste0("(", number_of_graves, ")"), collapse = ", ", sep = " ")
 txtstorage::store("regions graves amounts", regions_graves_amounts, storage_file)

### 2018-12-17 article/03_Results.Rmd
improvements of "Southeastern Central Europe"

@@
 
+General observations: **TODO**
+
 `r a("Southeastern Central Europe")`^[Amount of graves in Radon-B for this region.]:
-The artificial study region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, western Slovakia, Moravia and large parts of Bohemia.
-The density of ^14^C data on graves from radon B is low -- individual data are available from the various regions, but large areas must manage without evidence.
-Diachronic, the data volume is consistently low and reaches a low point in the Middle Bronze Age.
-While for the *burial type* variable a constant flow of information is given, the number of graves without information on *burial construction* increases in the Late Bronze Age to such an extent that no statement at all can be derived for them.
-To assess the relative development within the variables *burial type* and *burial construction*, as predicted with the Radon-B data set on the basis of ^14^C data, a résumé must first be drawn from the data in chapter \@ref(regions-archaeological-overview).
-This is complicated by the fact that the region intersects diverse geographical and cultural areas.
+The artificial study region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, Western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map} B).
+The density of ^14^C data on graves in Radon-B is low -- individual dates are available from the various regions, but large areas are without evidence.
+Diachronically the amount of data is consistently low and reaches a minimum in the Middle Bronze Age.
+While for the *burial type* variable a constant flow of information is given, the number of graves without information on *burial construction* increases in the Late Bronze Age to such an extent that no meaningful overview can be derived.
+The comparison with archaeological observations is complicated by the fact that the region intersects diverse geographical and cultural areas.
 In the southeast of Central Europe, cremation and bodily burial already existed side by side in the Early Bronze Age, and the local tradition remained stable in this respect also due to the expansion of the burial mound tradition in the Middle Bronze Age -- with a slight overall increase in cremation graves.
@@
 The information from ^14^C data on *burial construction* correctly traces the relevance of flat burials in the Early Bronze Age, but the latter are then overemphasized in the Middle Bronze Age compared to the construction of burial mounds.
-In the Late Bronze Age, burials on flat burial grounds were frequent, although hills of the burial mound culture were also further documented.
-This circumstance -- certainly due to the small amount of data -- cannot be correctly deduced from radon-B.
+In the Late Bronze Age, burials on flat burial grounds were frequent, although burial mounds of the Tumulus Culture were also further documented.
+This impact cannot be correctly deduced from Radon-B -- certainly due to the small amount of data.
 

### 2018-12-17 article/03_Results.Rmd
improvements of "Northeastern France"

@@
 General observations: **TODO**
+Literature! **TODO**
 
@@
 `r a("Poland")`:
-The region *Poland* completely covers the centre and west of today's Poland.
-However, the ^14^C data from this region come almost exclusively from the voivodships of Kajuwien-Pommern, Lodsch and Lower Silesia.
+The region *Poland* completely covers the centre and western part of today's Poland.
+However, the ^14^C data from this region are almost exclusively from the voivodships of Kajuwien-Pommern, Lodsch and Lower Silesia.
 The amount of data is small overall, with significantly more data from the Early Bronze Age up to about 1700calBC.
@@
 The relative development of *burial type* approximates the *real* conditions -- i.e.
-according to the classical, archaeological narrative --: body burials were actually significantly more frequent in western Poland until the Late Bronze Age, while burials did not occur until the Urnfield Period.
-Also before, cremation graves had been laid out again and again, but in principle the tendency recorded by Radon-B is correct.
-In contrast, *burial construction* is misrepresented insofar as the burial mounds also played an important role in the Middle Bronze Age, which the dataset does not appreciate.
+according to the classical, archaeological narrative: inhumations were actually significantly more frequent in western Poland until the Late Bronze Age, while cremations did not occur until the Urnfield Period.
+Also before, cremation graves had been created sometimes, but in general the tendency recorded by Radon-B is correct.
+In contrast, *burial construction* is misrepresented: burial mounds also played an important role in the Middle Bronze Age.
 
@@
 No ^14^C data are available from the latter regions.
-The majority of the available data comes from the Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative districts of Tübingen.
+The majority of the available data comes from the Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative district of Tübingen.
 For the Early Bronze Age a remarkable amount of data from Southern Germany are preserved in Radon-B, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves.
@@
 If one considers the relative development calculated from Radon-B and compares it with the archaeological literature, then a good agreement can be found for *burial type*.
-Body burial dominated for a long time, but was then replaced by cremation from the Middle Bronze Age onwards and massively in the Late Bronze Age.
-The calculated proportions for *burial construction* could easily be dismissed as completely worthless due to the negligibly small amount of data, but in fact the proxy indicates the right tendency for early and middle Bronze Age: The dominance of flat graves was replaced by hill burials.
-The data set does not allow any statements to be made about the Late Bronze Age.
+Inhumation dominated for a long time, but was then replaced by cremation from the Middle Bronze Age onwards and massively in the Late Bronze Age.
+The calculated proportions for *burial construction* could easily be dismissed as completely worthless due to the negligibly small amount of data, but in fact the proxy indicates the right tendency for early and middle Bronze Age: The dominance of flat graves was replaced by mounds.
+The data set does not allow any statements about the Late Bronze Age.
 
 `r a("Northeastern France")`:
-*Northeastern France* is a district region comprising the French administrative regions Grand-Est and Bourgogne-Franche-Comté, the east of Baden-Württemberg, Rhineland-Palatinate and Saarland, as well as Luxembourg and Wallonia in Belgium.
+*Northeastern France* comprises the French administrative regions Grand-Est and Bourgogne-Franche-Comté, the east of the German Baden-Württemberg, Rhineland-Palatinate and Saarland, as well as Luxembourg and Wallonia in Belgium.
 Accumulations of ^14^C data are only available from Württemberg and towards the Paris Basin.
@@
 The prediction of the relative development of these variables is correspondingly error-prone.
-After all, the parallel existence of cremation and body burial in early and middle Bronze Age is correctly reproduced.
-However, this already existed in the Early Bronze Age and certainly developed more nuanced than the proxy might suggest.
-Contrary to the derivation from the ^14^C ratios, flat graves were quite common in *Northeastern France*.
+Still the parallel existence of cremation and body burial in early and middle Bronze Age is correctly reflected.
+However, this situation already existed in the Early Bronze Age and certainly developed more nuanced than the proxy might suggest.
+Contrary to the derivation from the ^14^C ratios, flat graves were quite common in Northeastern France.
 

### 2018-12-17 article/03_Results.Rmd
improvements of "Northern Germany"

@@
 `r a("Northern Germany")`:
-The artificial region *Northern Germany* includes all new federal states in East Germany, but also large parts of Lower Saxony and Schleswig-Holstein.
+The artificial region *Northern Germany* includes all federal states in East Germany, but also large parts of Lower Saxony and Schleswig-Holstein.
 The ^14^C data for this region come mainly from Saxony-Anhalt and neighbouring regions in Thuringia, Saxony and Lower Saxony.
@@
 The database does not contain as many dated graves from any other region as from northern Germany.
-Most of them date back to the Early Bronze Age, while the Middle Bronze Age from about 1800calBC to 1400calBC is a striking period for some centuries.
-Only for the Late Bronze Age is a significant number of data documented again.
-The meta information on *burial type* and *burial construction* trace this development to about the same extent.
+Most of them date back to the Early Bronze Age, while the Middle Bronze Age from about 1800calBC to 1400calBC seems barren.
+Only for the Late Bronze Age there is a significant number of data documented again.
+The meta information on *burial type* and *burial construction* trace this development to the same extent.
 Since in Northern Germany influences from all four cardinal points come together and contribute significantly to the formation of local cultural phenomena, it is difficult to estimate the overall development of the primary variables here as well.
-The ^14^C proxy correctly represents *burial site* insofar as bodily burials dominated in the Early Bronze Age and are replaced by burials especially in the North from the Middle Bronze Age onwards.
-In the Late Bronze Age cremation was clearly more dominant than the data set suggests.
+The ^14^C proxy correctly represents *burial site* insofar as inhumations dominated in the Early Bronze Age and are replaced by cremations especially in the North from the Middle Bronze Age onwards.
+In the Late Bronze Age though, cremation was clearly more dominant than the data set suggests.
 With regard to *burial construction*, the current Radon-B stock records the basic dominance of flat graves well, but attributes more relevance to them in the Early and Middle Bronze Age than is usual in the archaeological assessment of the context.

### 2018-12-17 article/03_Results.Rmd
improvements of "Southern Scandinavia"

@@
 Since the Bronze Age in Scandinavia begins much later than in Central Europe and is measured by another chronological system, the diachronic developments cannot simply be compared by means of the tripartite structure of the Early, Middle and Late Bronze Ages.
+The period which is labeled Early Bronze Age due in Central Europe actually corresponds to the Scandinavian Late Neolithic.
 The majority of data from Scandinavia are from the period IB onwards.
 In this respect, the data density for almost the entire Bronze Age is good -- also with regard to the two variables *burial type* and *burial construction*.
-The Late Neolithic from 2200calBC is a desideratum -- although at a higher level than, for example, the Early Bronze Age in *Northeastern France*.
-The adjustment in Figure \@ref(fig:development-proportions-regions-pseudoquant) is difficult, since the period which is labeled with Early Bronze Age due to the situation in Central Europe actually corresponds to the local Late Neolithic.
-If one carries out the resulting mental shift -- the Central European Middle Bronze Age corresponds to the Nordic Early Bronze Age -- and then compares the proxy of the relative grave form development with the archaeological observations, then one finds a solid agreement: body graves were the rule in periods I and II (about 1800-1300calBC), but cremations also occurred sporadically.
+The Late Neolithic from 2200calBC is a desideratum.
+If one carries out the necessary mental shift -- the Central European Middle Bronze Age corresponds to the Nordic Early Bronze Age -- and then compares the proxy of the relative grave form development with the archaeological observations, then one finds a solid agreement: body graves were the rule in periods I and II (about 1800-1300calBC), but cremations also occurred sporadically.
 From period III onwards cremation burials gained the upper hand.
-In the Scandinavian Early Bronze Age (i.e.
-Middle Bronze Age in Figure \@ref(fig:development-proportions-regions-pseudoquant)) burials in burial mounds were the absolute rule.
-In the Late Bronze Age, shallow burial fields were added.
+In the Scandinavian Early Bronze Age burials in burial mounds were the absolute rule.
+In the Late Bronze Age, flat burial fields were added.
 

### 2018-12-17 article/03_Results.Rmd
improvements of "Benelux"

@@
 As in the *Southern Germany* region, there is a serious difference between the information density of both variables.
-If one considers the relative development of the variables as they can be derived from the grave data in comparison to the archaeological doctrine, both agreement and deviations result here in detail.
-Here, too, the study region includes different cultural phenomena with partially diametrically different burial customs, but a general assessment of the conditions is nevertheless possible.
-In the north and east of the Benelux region body burials were the rule until the Late Bronze Age, while burials on fire dominated earlier in the south of the Meuse.
+If one considers the relative development of the variables as they can be derived from the grave data in comparison to the archaeological state of research there is both deviation as well as agreement.
+Here, too, the study region includes different cultural phenomena with highly different burial customs: In the north and east of the Benelux region inhumations were the rule until the Late Bronze Age, while cremation dominated earlier south of the Meuse.
 The Radon-B proxy seems to overemphasise the dominance of cremation in the whole area.

### 2018-12-17 article/03_Results.Rmd
improvements of "England"

@@
 `r a("England")`:
-The *England* region stretches across England to the heights of York and Liverpool, incorporating parts of Wales and leaving out Cornwall and Devon in the southwest.
-In fact, it also touches a narrow coastal strip on the mainland, from which no ^14^C data are available.
+The *England* region stretches across England up to York and Liverpool, incorporating parts of Wales and leaving out Cornwall and Devon in the southwest.
+It also touches a narrow coastal strip on the mainland, from which no ^14^C data are documented.
 The available data are concentrated in the southeast of Britain, but are well distributed there.
@@
 *burial type* and *burial construction* behave accordingly.
-The relative development derived from the ^14^C data indicates the correct trend for *burial type*: Cremation graves became more and more important during the Bronze Age, while body burials were practiced for a long time in parallel.
+The relative development derived from the ^14^C data indicates the correct trend for *burial type*: Cremation graves became more and more important during the Bronze Age, while inhumations were practiced for a long time in parallel.
 However, the data set emphasizes the dominance of the cremation graves too much.

### 2018-12-17 article/index.Rmd
started to work with dynamic value integration
added conclusion segment
moved reference file

--- 2018-11-20 index.Rmd
@@
   - times
-bibliography: references
+bibliography: ../references.bib
 bibliographystyle: apalike
@@
     keep_tex: yes
+    includes:
+        in_header: preamble.tex
 ---
+
+```{r echo=FALSE}
+library(magrittr)
+
+# allow easy dynamic value integration
+sf_prep <- "../analysis/data/output_data/sf_prep.txt"
+sf_desc <- "../analysis/data/output_data/sf_desc.txt"
+r <- txtstorage::read
+a <- function(x) {r("regions graves amounts", sf_prep) %>% strsplit(", ") %>% unlist %>% magrittr::extract(grepl(x, .))}
+```
 
@@
 ```
+
+```{r child = '05_Conclusion.Rmd'}
+```

### 2018-12-17 article/03_Results.Rmd
added some literature for the comparison

@@
 General observations: **TODO**
-Literature! **TODO**
 
@@
 In the Late Bronze Age, burials on flat burial grounds were frequent, although burial mounds of the Tumulus Culture were also further documented.
-This impact cannot be correctly deduced from Radon-B -- certainly due to the small amount of data.
+This impact cannot be correctly deduced from Radon-B -- certainly due to the small amount of data [@lubosCzechLandsAustria2013; @markovaSlovakiaHungary2013].
+\newline
 
@@
 Also before, cremation graves had been created sometimes, but in general the tendency recorded by Radon-B is correct.
-In contrast, *burial construction* is misrepresented: burial mounds also played an important role in the Middle Bronze Age.
+In contrast, *burial construction* is misrepresented: burial mounds also played an important role in the Middle Bronze Age [@czebreszukBronzeAgePolish2013; @dabrowskiAeltereBronzezeitPolen2004].
+\newline
 
@@
 The calculated proportions for *burial construction* could easily be dismissed as completely worthless due to the negligibly small amount of data, but in fact the proxy indicates the right tendency for early and middle Bronze Age: The dominance of flat graves was replaced by mounds.
-The data set does not allow any statements about the Late Bronze Age.
+The data set does not allow any statements about the Late Bronze Age [@falkensteinWandelBestattungssittenHugelgraber2017; @jockenhovelGermanyBronzeAge2013].
+\newline
 
@@
 However, this situation already existed in the Early Bronze Age and certainly developed more nuanced than the proxy might suggest.
-Contrary to the derivation from the ^14^C ratios, flat graves were quite common in Northeastern France.
+Contrary to the derivation from the ^14^C ratios, flat graves were quite common in Northeastern France [@mordantBronzeAgeFrance2013].
+\newline
 
@@
 In the Late Bronze Age though, cremation was clearly more dominant than the data set suggests.
-With regard to *burial construction*, the current Radon-B stock records the basic dominance of flat graves well, but attributes more relevance to them in the Early and Middle Bronze Age than is usual in the archaeological assessment of the context.
+With regard to *burial construction*, the current Radon-B stock records the basic dominance of flat graves well, but attributes more relevance to them in the Early and Middle Bronze Age than is usual in the archaeological assessment of the context [@hofmannRituelleUmgangMit2008; @jockenhovelGermanyBronzeAge2013].
+\newline
 
@@
 In the Scandinavian Early Bronze Age burials in burial mounds were the absolute rule.
-In the Late Bronze Age, flat burial fields were added.
+In the Late Bronze Age, flat burial fields were added [@thraneScandinavia2013].
+\newline
 
@@
 The Radon-B proxy seems to overemphasise the dominance of cremation in the whole area.
-The relevance of hill burials in the Early and Middle Bronze Age, on the other hand, is correctly reflected, and the mixture between hill and flat graves in the context of urnfields, often resulting from the continued use of existing monumental sites, is also visible.
+The relevance of hill burials in the Early and Middle Bronze Age, on the other hand, is correctly reflected, and the mixture between hill and flat graves in the context of urnfields, often resulting from the continued use of existing monumental sites, is also visible [@arnoldussenBronzeAgeSettlements2008; @drenthMoundsDeadFuneray2005; @lohofTraditionChangeBurial1994].
+\newline
 
@@
 However, the data set emphasizes the dominance of the cremation graves too much.
-With regard to the variables *burial construction*, the basic dominance of burial mounds corresponds perfectly to the archaeological narrative, but nevertheless -- especially in the Late Bronze Age -- flat grave fields also occurred, which the data set wrongly excludes.
+With regard to the variables *burial construction*, the basic dominance of burial mounds corresponds perfectly to the archaeological narrative, but nevertheless -- especially in the Late Bronze Age -- flat grave fields also occurred, which the data set wrongly excludes [@robertsBritainIrelandBronze2013].
+\newline
 

### 2018-12-17 article/02_Materials_and_Methods.Rmd
decided to split the map figure

@@
 7.
-Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial date distribution (Figure \ref{map} A).
+Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial date distribution (Figure \ref{map_graves}).
 
@@
 The regions are circular and were delimited semi-automatically by creating a regular grid in accordance to main cultural-historical regions of the European Bronze Age (**citation**).
-The regions have a radius of 240km and overlap slightly (Figure \ref{map} B).
+The regions have a radius of 240km and overlap slightly (Figure \ref{map_regions} A).
 That means some graves are counted towards not just one but two regions.
@@
 \centering
-\includegraphics[width=\columnwidth]{../analysis/figures/map.jpeg}
-\caption{\label{map}This is a figure.}
+\includegraphics[width=\columnwidth]{../analysis/figures/map_graves.jpeg}
+\caption{\label{map_graves}This is a figure.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../analysis/figures/map_regions.jpeg}
+\caption{\label{map_regions}This is a figure.}
 \end{figure}
@@
 A continuous distance measure can't be employed, but it's possible to distinguish ordinal distance classes.
-For the eight regions four different categories of spatial distance were defined (Figure \ref{map} C).
+For the eight regions four different categories of spatial distance were defined (Figure \ref{map_regions} B).
 The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the difference in the data level of measurements.

### 2018-12-17 article/03_Results.Rmd
decided to split the map figure

@@
 `r a("Southeastern Central Europe")`^[Amount of graves in Radon-B for this region.]:
-The artificial study region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, Western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map} B).
+The artificial study region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, Western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map_regions} A).
 The density of ^14^C data on graves in Radon-B is low -- individual dates are available from the various regions, but large areas are without evidence.

### 2018-12-17 analysis/code/05_plot_scripts/map.R

@@ -46,6 +46,10 @@ map_A <- ggplot() +
     ),
     show.legend = "point"
   ) +
+  geom_sf(
+    data = research_area,
+    fill = NA, colour = "red", size = 0.8
+  ) +
   theme_bw() +
   coord_sf(
     xlim = xlimit_A, ylim = ylimit_A,
@@ -89,6 +93,18 @@ map_A <- ggplot() +
     shape = guide_legend(title = "Burial construction", override.aes = list(size = 10), nrow = 1, byrow = TRUE),
     size = FALSE
   )
+
+map_A %>%
+  ggsave(
+    "analysis/figures/map_graves.jpeg",
+    plot = .,
+    device = "jpeg",
+    scale = 1,
+    dpi = 300,
+    width = 330, height = 420, units = "mm",
+    limitsize = F
+  )
+
 
 #### map_B ####
 
@@ -277,12 +293,11 @@ map_C <- ggplot() +
 #### combine maps ####
 
 combined_map <- plot_grid(
-  map_A,
   map_B,
   map_C,
-  labels = c("A", "B", "C"),
-  rel_heights = c(1.73, 1, 1),
-  nrow = 3,
+  labels = c("A", "B"),
+  rel_heights = c(1, 1),
+  nrow = 2,
   align = "h",
   axis = "lr",
   label_size = 35
@@ -290,11 +305,11 @@ combined_map <- plot_grid(
 
 combined_map %>%
   ggsave(
-    "analysis/figures/map.jpeg",
+    "analysis/figures/map_regions.jpeg",
     plot = .,
     device = "jpeg",
     scale = 1,
     dpi = 300,
-    width = 330, height = 850, units = "mm",
+    width = 330, height = 440, units = "mm",
     limitsize = F
   )

### 2018-12-17 article/03_Results.Rmd
paragraph about cultural distance

@@
 Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region wise timeseries of trait occurrence.
-The relative proportion of each trait calculated based on ^14^C dates can be compared to the archaeological state of research: \newline
+The relative proportion of each trait calculated based on ^14^C dates can be compared to the archaeological state of research:
 
@@
 This impact cannot be correctly deduced from Radon-B -- certainly due to the small amount of data [@lubosCzechLandsAustria2013; @markovaSlovakiaHungary2013].
-\newline
 
@@
 In contrast, *burial construction* is misrepresented: burial mounds also played an important role in the Middle Bronze Age [@czebreszukBronzeAgePolish2013; @dabrowskiAeltereBronzezeitPolen2004].
-\newline
 
@@
 The data set does not allow any statements about the Late Bronze Age [@falkensteinWandelBestattungssittenHugelgraber2017; @jockenhovelGermanyBronzeAge2013].
-\newline
 
@@
 Contrary to the derivation from the ^14^C ratios, flat graves were quite common in Northeastern France [@mordantBronzeAgeFrance2013].
-\newline
 
@@
 With regard to *burial construction*, the current Radon-B stock records the basic dominance of flat graves well, but attributes more relevance to them in the Early and Middle Bronze Age than is usual in the archaeological assessment of the context [@hofmannRituelleUmgangMit2008; @jockenhovelGermanyBronzeAge2013].
-\newline
 
@@
 In the Late Bronze Age, flat burial fields were added [@thraneScandinavia2013].
-\newline
 
@@
 The relevance of hill burials in the Early and Middle Bronze Age, on the other hand, is correctly reflected, and the mixture between hill and flat graves in the context of urnfields, often resulting from the continued use of existing monumental sites, is also visible [@arnoldussenBronzeAgeSettlements2008; @drenthMoundsDeadFuneray2005; @lohofTraditionChangeBurial1994].
-\newline
 
@@
 With regard to the variables *burial construction*, the basic dominance of burial mounds corresponds perfectly to the archaeological narrative, but nevertheless -- especially in the Late Bronze Age -- flat grave fields also occurred, which the data set wrongly excludes [@robertsBritainIrelandBronze2013].
-\newline
 
-- Cultural distance space
+Beyond the simple comparison of archaeological knowledge and information derived from Radon-B, figure \ref{sed_region_matrix} shows the timeseries of squared euclidian distance for each region-region relationship.
+The majority of the curves in it's upper right half for *burial type* correspond to expectations: At the beginning of the Bronze Age, inhumation is widespread throughout the study area, so the Euclidean distance at 2200calBC is small in almost every regional relationship.
+Only the Benelux region, where the share of cremation is already high in the Early Bronze Age, is excluded.
+From the Middle Bronze Age onwards, but especially at the transition to the Late Bronze Age, the transition to the dominance of cremation took place all over Europe.
+The SED values increase in this period and show clear peaks or several successive maxima.
+This is a consequence of the fact that the transition process in each region as depicted in Radon-B is different and that it took place with a time lag compared to other regions.
+At the end of the Late Bronze Age, the average SED is approaching zero again, as burial by cremation dominates in almost all of Europe.
+The early and complete adoption of the cremation rite in the Benelux region, as depicted in Radon-B, clearly sets it apart from the other regions.
+England, on the other hand, with its slow, gradual transition from bodily burial to cremation, stands out due to its consistently moderate distances to all other regions.
+It remains without accentuated minima and maxima.
+The situation for *burial construction* is more complicated, as can be seen from the plots in the bottom right corner of figure \ref{sed_region_matrix}.
+No general trend can be seen here, but a multitude of local and temporally limited phenomena.
+This is also due to how correctly or incorrectly the proxy developed from radon-B depicts developments in the real world.
+In the Middle Bronze Age, the Tumulus culture should bring about a fundamental reduction of the SED for almost all regional relationships -- this effect is indeed weakly visible.
 

### 2018-12-17 article/03_Results.Rmd
more about cultural and spatial distance

@@
 
-- Correlation of cultural distances and spatial distance
+Table \ref{tab:distanceandcorrelationtable*} shows that the distance networks of *burial type* and *burial construction* are not equivalent.
+There is also no significant negative correlation.
+The degree of agreement between two regions in the question of body burial and cremation is therefore not suitable for predicting the similarity of developments in these regions with regard to the dominance of flat or mound graves.
+Only for the time window at the transition to the Middle Bronze Age from 1800 to 1600calBC is the 5% significance level of the mantle test almost undercut, thus suggesting a statistically significant relationship.
 
-- Simulation: What is to be expected in case of influence of spatial distance?
+For the relationship between the SED network based on *burial type* and the spatial distance several observations result from Table \ref{tab:distanceandcorrelationtable*}: Surprisingly, cultural distance does not seem to correlate with spatial distance.
+A significant correlation between the distance measures only reveals itself for the Early Bronze Age time window 2000 to 1800calBC.
+This means that only in the Early Bronze Age, long before the universal switch to the Urnfield burial, large regions directly adjacent to each other tend to undergo the same developments rather than the ones further away from each other.
+In the Middle Bronze Age, the SED developments of the *burial type* variable undergo the peaks already observed in Figure \ref{sed_region_matrix}.
+The correlation between cultural and spatial distance over these peaks is low and even tilts between 1400 and 1200calBC, at the transition to the Late Bronze Age, into the negative -- although not significantly.
+The cultural distance between several geographically distant regions is then smaller than between several directly adjacent regions.
+Looking at the SED network on the basis of *burial construction* and examining it for spatial correlation by means of the Table \ref{tab:distanceandcorrelationtable*}, similar conclusions can be drawn as in the *burial type* context.
+The distance values here are larger overall and the SED differences within the individual spatial distance classes are clearly more accentuated, but here, too, the early bronze age is by far the phase of the largest spatial correlation.
+This is only significant for the time window between 1800 and 1400calBC.
+Negative correlation is not indicative for *burial construction*, but here, too, the correlation coefficient approaches zero in the Late Bronze Age.
+
+Simulation: What is to be expected in case of influence of spatial distance?

### 2018-12-17 article/03_Results.Rmd
language and grammer

--- 2018-12-17 article/03_Results.Rmd
@@
 
-Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region wise timeseries of trait occurrence.
-The relative proportion of each trait calculated based on ^14^C dates can be compared to the archaeological state of research: \newline
+Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region-wise time series of trait occurrence.
+The relative proportion of each trait calculated based on ^14^C dates can be compared to the archaeological state of research:
 
@@
 The artificial study region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, Western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map_regions} A).
-The density of ^14^C data on graves in Radon-B is low -- individual dates are available from the various regions, but large areas are without evidence.
+The density of ^14^C data on graves in Radon-B is low -- dates are available from the various regions, but large areas are without evidence.
 Diachronically the amount of data is consistently low and reaches a minimum in the Middle Bronze Age.
-While for the *burial type* variable a constant flow of information is given, the number of graves without information on *burial construction* increases in the Late Bronze Age to such an extent that no meaningful overview can be derived.
-The comparison with archaeological observations is complicated by the fact that the region intersects diverse geographical and cultural areas.
-In the southeast of Central Europe, cremation and bodily burial already existed side by side in the Early Bronze Age, and the local tradition remained stable in this respect also due to the expansion of the burial mound tradition in the Middle Bronze Age -- with a slight overall increase in cremation graves.
+While there is a constant flow of information for the variable *burial type*, the number of graves without information on *burial construction* increases in the Late Bronze Age to such an extent that no meaningful overview can be derived.
+The region intersects diverse geographical and cultural areas complicating the comparison with archaeological observations.
+In the southeast of Central Europe, cremation and inhumation already existed side by side in the Early Bronze Age, and the local tradition remained stable in this respect also due to the expansion of the burial mound tradition in the Middle Bronze Age -- with a slight overall increase in cremation graves.
 The Urnfield period brought the transition to the absolute dominance of cremation burial.
 The development of the *burial type* variable in the Radon-B proxy shows this development correctly.
-The information from ^14^C data on *burial construction* correctly traces the relevance of flat burials in the Early Bronze Age, but the latter are then overemphasized in the Middle Bronze Age compared to the construction of burial mounds.
+The information from ^14^C data on *burial construction* correctly traces the relevance of flat burials in the Early Bronze Age but then overemphasises them in the Middle Bronze Age compared to the construction of burial mounds.
 In the Late Bronze Age, burials on flat burial grounds were frequent, although burial mounds of the Tumulus Culture were also further documented.
-This impact cannot be correctly deduced from Radon-B -- certainly due to the small amount of data [@lubosCzechLandsAustria2013; @markovaSlovakiaHungary2013].
-\newline
+This impact does not become clear in the Radon-B data -- indeed due to the small number of dates [@lubosCzechLandsAustria2013; @markovaSlovakiaHungary2013].
 
@@
 Almost no information is available from the Middle and Late Bronze Ages.
-This applies to both the *burial type* and the *burial construction* variables, the former being slightly better documented.
+This lack of data applies to both the *burial type* and the *burial construction* variables, the former being slightly better documented.
 The relative development of *burial type* approximates the *real* conditions -- i.e.
-according to the classical, archaeological narrative: inhumations were actually significantly more frequent in western Poland until the Late Bronze Age, while cremations did not occur until the Urnfield Period.
-Also before, cremation graves had been created sometimes, but in general the tendency recorded by Radon-B is correct.
-In contrast, *burial construction* is misrepresented: burial mounds also played an important role in the Middle Bronze Age [@czebreszukBronzeAgePolish2013; @dabrowskiAeltereBronzezeitPolen2004].
-\newline
+according to the classical, archaeological narrative: inhumations were significantly more frequent in western Poland until the Late Bronze Age, while cremations did rarely occur before the Urnfield Period.
+In contrast, it misrepresents *burial construction*: burial mounds also played an important role in Middle Bronze Age Poland [@czebreszukBronzeAgePolish2013; @dabrowskiAeltereBronzezeitPolen2004].
 
 `r a("Southern Germany")`:
-The artificial district region *Southern Germany* covers Bavaria completely, but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and Western Austria.
+The artificial district region *Southern Germany* covers Bavaria completely but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and Western Austria.
 No ^14^C data are available from the latter regions.
 The majority of the available data comes from the Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative district of Tübingen.
-For the Early Bronze Age a remarkable amount of data from Southern Germany are preserved in Radon-B, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves.
-The *burial type* variable is quite well understood -- although disproportionately much more of the early body graves were recorded than urnfield burials.
+For the Early Bronze Age, Radon-B preserves a remarkable amount of data from Southern Germany, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves.
+The *burial type* variable is quite well understood -- although disproportionately much more of the early inhumations were recorded than urnfield burials.
 This wealth of information contrasts sharply with the situation regarding the *burial construction* variable, for which almost no information from this area is available throughout the Bronze Age.
-If one considers the relative development calculated from Radon-B and compares it with the archaeological literature, then a good agreement can be found for *burial type*.
-Inhumation dominated for a long time, but was then replaced by cremation from the Middle Bronze Age onwards and massively in the Late Bronze Age.
-The calculated proportions for *burial construction* could easily be dismissed as completely worthless due to the negligibly small amount of data, but in fact the proxy indicates the right tendency for early and middle Bronze Age: The dominance of flat graves was replaced by mounds.
-The data set does not allow any statements about the Late Bronze Age [@falkensteinWandelBestattungssittenHugelgraber2017; @jockenhovelGermanyBronzeAge2013].
-\newline
+If one considers the relative development calculated from Radon-B and compares it with the archaeological literature, then a reasonable agreement can be found for *burial type*.
+Inhumation dominated for a long time but was then replaced by cremation from the Middle Bronze Age onwards and massively in the Late Bronze Age.
+The calculated proportions for *burial construction* are based on a negligibly small amount of data, but the proxy still indicates the right tendency for early and middle Bronze Age: Flat graves replaced mounds.
+The dataset does not allow any statements about the Late Bronze Age [@falkensteinWandelBestattungssittenHugelgraber2017; @jockenhovelGermanyBronzeAge2013].
 
@@
 The amount of data is low overall, but very low in the early and early Middle Bronze Age.
-For *burial type*, after all, limited statements are possible after this time window, *burial construction* on the other hand must remain obscure.
+For *burial type*, after all, limited statements are possible after this time window, *burial construction*, on the other hand, must remain obscure.
 The prediction of the relative development of these variables is correspondingly error-prone.
-Still the parallel existence of cremation and body burial in early and middle Bronze Age is correctly reflected.
-However, this situation already existed in the Early Bronze Age and certainly developed more nuanced than the proxy might suggest.
+Still, the parallel existence of cremation and inhumation in the Early and Middle Bronze Age is correctly reflected.
+However, this situation already existed in the Early Bronze Age and indeed developed more nuanced than the proxy might suggest.
 Contrary to the derivation from the ^14^C ratios, flat graves were quite common in Northeastern France [@mordantBronzeAgeFrance2013].
-\newline
 
@@
 Most of them date back to the Early Bronze Age, while the Middle Bronze Age from about 1800calBC to 1400calBC seems barren.
-Only for the Late Bronze Age there is a significant number of data documented again.
+Only for the Late Bronze Age, there is a significant number of data documented again.
 The meta information on *burial type* and *burial construction* trace this development to the same extent.
 Since in Northern Germany influences from all four cardinal points come together and contribute significantly to the formation of local cultural phenomena, it is difficult to estimate the overall development of the primary variables here as well.
-The ^14^C proxy correctly represents *burial site* insofar as inhumations dominated in the Early Bronze Age and are replaced by cremations especially in the North from the Middle Bronze Age onwards.
-In the Late Bronze Age though, cremation was clearly more dominant than the data set suggests.
-With regard to *burial construction*, the current Radon-B stock records the basic dominance of flat graves well, but attributes more relevance to them in the Early and Middle Bronze Age than is usual in the archaeological assessment of the context [@hofmannRituelleUmgangMit2008; @jockenhovelGermanyBronzeAge2013].
-\newline
+The ^14^C proxy correctly represents *burial site* insofar as inhumations dominated in the Early Bronze Age and were later on replaced by cremations especially in the North from the Middle Bronze Age onwards.
+In the Late Bronze Age though, cremation was more dominant than the dataset suggests.
+Concerning *burial construction*, the current Radon-B stock records the fundamental dominance of flat graves well but attributes more relevance to them in the Early and Middle Bronze Age than is usual in the archaeological assessment of the context [@hofmannRituelleUmgangMit2008; @jockenhovelGermanyBronzeAge2013].
 
@@
 The ^14^C data are very well distributed across Jutland, Funen and the Swedish Skåne region.
-Since the Bronze Age in Scandinavia begins much later than in Central Europe and is measured by another chronological system, the diachronic developments cannot simply be compared by means of the tripartite structure of the Early, Middle and Late Bronze Ages.
-The period which is labeled Early Bronze Age due in Central Europe actually corresponds to the Scandinavian Late Neolithic.
+Since the Bronze Age in Scandinavia begins much later than in Central Europe and is measured by another chronological system, the diachronic developments cannot merely be compared through the tripartite structure of the Early, Middle and Late Bronze Ages.
+The period which is labelled Early Bronze Age in Central Europe corresponds to the Scandinavian Late Neolithic.
 The majority of data from Scandinavia are from the period IB onwards.
@@
 The Late Neolithic from 2200calBC is a desideratum.
-If one carries out the necessary mental shift -- the Central European Middle Bronze Age corresponds to the Nordic Early Bronze Age -- and then compares the proxy of the relative grave form development with the archaeological observations, then one finds a solid agreement: body graves were the rule in periods I and II (about 1800-1300calBC), but cremations also occurred sporadically.
+If one carries out the necessary mental shift -- the Central European Middle Bronze Age corresponds to the Nordic Early Bronze Age -- and then compares the proxy of the relative grave form development with the archaeological observations, then one finds a substantial agreement: body graves were the rule in periods I and II (about 1800-1300calBC), but cremations also occurred sporadically.
 From period III onwards cremation burials gained the upper hand.
@@
 As in the *Southern Germany* region, there is a serious difference between the information density of both variables.
-If one considers the relative development of the variables as they can be derived from the grave data in comparison to the archaeological state of research there is both deviation as well as agreement.
-Here, too, the study region includes different cultural phenomena with highly different burial customs: In the north and east of the Benelux region inhumations were the rule until the Late Bronze Age, while cremation dominated earlier south of the Meuse.
+If one considers the relative development of the variables as they can be derived from the burial data in comparison to the archaeological state of research, there is both deviations as well as agreement.
+Here, too, the study region includes different cultural phenomena with highly different burial customs: In the north and east of the Benelux region, inhumations were the rule until the Late Bronze Age, while cremation dominated earlier south of the Meuse.
 The Radon-B proxy seems to overemphasise the dominance of cremation in the whole area.
-The relevance of hill burials in the Early and Middle Bronze Age, on the other hand, is correctly reflected, and the mixture between hill and flat graves in the context of urnfields, often resulting from the continued use of existing monumental sites, is also visible [@arnoldussenBronzeAgeSettlements2008; @drenthMoundsDeadFuneray2005; @lohofTraditionChangeBurial1994].
-\newline
+The relevance of burial mounds in the Early and Middle Bronze Age, on the other hand, is correctly reflected, and the mixture between mounds and flat graves in the context of urnfields, often resulting from the continued use of existing grave monuments, is also visible [@arnoldussenBronzeAgeSettlements2008; @drenthMoundsDeadFuneray2005; @lohofTraditionChangeBurial1994].
 
@@
 It also touches a narrow coastal strip on the mainland, from which no ^14^C data are documented.
-The available data are concentrated in the southeast of Britain, but are well distributed there.
-The diachronic data availability is relatively good in the early and middle Bronze Age, but then decreases in the late Bronze Age.
+The available data are concentrated in the southeast of Britain but are well distributed there.
+The diachronic data availability is relatively good in the early and middle Bronze Age but then decreases in the late Bronze Age.
 *burial type* and *burial construction* behave accordingly.
-The relative development derived from the ^14^C data indicates the correct trend for *burial type*: Cremation graves became more and more important during the Bronze Age, while inhumations were practiced for a long time in parallel.
-However, the data set emphasizes the dominance of the cremation graves too much.
-With regard to the variables *burial construction*, the basic dominance of burial mounds corresponds perfectly to the archaeological narrative, but nevertheless -- especially in the Late Bronze Age -- flat grave fields also occurred, which the data set wrongly excludes [@robertsBritainIrelandBronze2013].
-\newline
+The relative development derived from the ^14^C data indicates the correct trend for *burial type*: Cremation graves became more and more important during the Bronze Age, while inhumations appeared for a long time in parallel.
+However, the dataset emphasises the dominance of the cremation graves too much.
+Concerning the variables *burial construction*, the dominance of burial mounds corresponds perfectly to the archaeological narrative, but -- especially in the Late Bronze Age -- flat grave fields also occurred, which the data set wrongly excludes [@robertsBritainIrelandBronze2013].
 
-- Cultural distance space
+Beyond the simple comparison of archaeological knowledge and information derived from Radon-B, figure \ref{sed_region_matrix} shows the time series of squared euclidian distance for each region-region relationship.
+The majority of the curves in its upper right half for *burial type* correspond to expectations: At the beginning of the Bronze Age, inhumation is widespread throughout the study area, so the Euclidean distance at 2200calBC is small in almost every region-region relationship.
+Only the Benelux region, where the share of cremation is already high in the Early Bronze Age, is excluded.
+From the Middle Bronze Age onwards, but especially at the transition to the Late Bronze Age, the transition to the dominance of cremation took place all over Europe.
+The SED values increase in this period and show clear peaks or several successive maxima.
+This rise is a consequence of the fact that the transition process in each region as depicted in Radon-B is different and that it took place with a time lag compared to other regions.
+At the end of the Late Bronze Age, the average SED is approaching zero again, as burial by cremation dominates in almost all of Europe.
+The early and complete adoption of the cremation rite in the Benelux region, as depicted in Radon-B, clearly sets it apart from the other regions.
+England, on the other hand, with its slow, gradual transition from inhumation to cremation, stands out due to its consistently moderate distances to all other regions.
+It remains without accentuated minima and maxima.
+The situation for *burial construction* is more complicated, as can be seen from the plots in the bottom right corner of figure \ref{sed_region_matrix}.
+The distance does not seem to follow a general trend, but a multitude of local and temporally limited phenomena.
+This heterogeneity is also due to how correctly or incorrectly the proxy developed from radon-B depicts developments in the real world.
+In the Middle Bronze Age, the Tumulus culture should bring about a fundamental reduction of the SED for almost all region-region relationships -- this effect is indeed weakly visible.
 
-- Correlation of cultural distances and spatial distance
+Table \ref{tab:distanceandcorrelationtable*} shows that the distance networks of *burial type* and *burial construction* are not equivalent.
+There is also no significant negative correlation.
+The degree of agreement between two regions in the question of inhumation and cremation is therefore not suitable for predicting the similarity of developments in these regions concerning the dominance of flat or mound graves.
+Only for the time window at the transition to the Middle Bronze Age from 1800 to 1600calBC is the 5% significance level of the mantle test almost undercut, thus suggesting a statistically significant relationship.
 
-- Simulation: What is to be expected in case of influence of spatial distance?
+For the relationship between the SED network based on *burial type* and the spatial distance several observations result from Table \ref{tab:distanceandcorrelationtable*}: Surprisingly, cultural distance does not seem to correlate with spatial distance.
+A significant correlation between the distance measures only reveals itself for the Early Bronze Age time window 2000 to 1800calBC: Only in the Early Bronze Age, long before the universal switch to the Urnfield burial, large regions directly adjacent to each other tend to undergo the same developments rather than the ones further away from each other.
+In the Middle Bronze Age, the SED developments of the *burial type* variable undergo the peaks already observed in Figure \ref{sed_region_matrix}.
+The correlation between cultural and spatial distance over these peaks is low and even tilts between 1400 and 1200calBC, at the transition to the Late Bronze Age, into the negative -- although not significantly.
+The cultural distance between several geographically distant regions is then smaller than between several directly adjacent regions.
+Looking at the SED network of *burial construction* (Table \ref{tab:distanceandcorrelationtable*}), similar conclusions can be drawn as in the *burial type* context: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated, but here, too, the early bronze age is by far the phase of the most substantial spatial correlation.
+It is only significant for the time window between 1800 and 1400calBC.
+There is no time period with negative mean correlation for *burial construction* cultural distance and spatial distance, but also here the correlation coefficient approaches zero in the Late Bronze Age.
+
+Simulation: What is to be expected in case of influence of spatial distance?

### 2018-12-18 article/02_Materials_and_Methods.Rmd
made correlation table more nice with highlighting

@@
 load("../analysis/data/output_data/distance_and_correlation_table.RData")
+distance_and_correlation_table %>%
+  dplyr::mutate_at(
+    dplyr::vars(dplyr::matches("Mean|Correlation")),
+    dplyr::funs(kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black"))) 
+  ) %>%
+  dplyr::mutate_at(
+    dplyr::vars(dplyr::matches("p-Value")),
+    dplyr::funs(kableExtra::cell_spec(., "latex", bold = . < 0.1)) 
+  ) %>%
 knitr::kable(
-  distance_and_correlation_table, 
   "latex", 
@@
   booktabs = T,
-  linesep = "\\addlinespace \\hline \\addlinespace"
+  linesep = "\\addlinespace \\hline \\addlinespace",
+  escape = F
 ) %>%

### 2018-12-18 article/01_Introduction.Rmd
language and grammer

@@
 
-This articles employs some of the methods that arose within archaeology's contemplation of *Cultural transmission theory* to trace the development of four major Bronze Age burial traditions in Europe.
-The cultural-historical evolution of archaeological cultures and ethnic groups during the European Bronze Age is complex and it's reconstruction an ongoing endeavour of archaeology.
+This article employs some of the methods that arose within archaeology's contemplation of *Cultural transmission theory* to trace the development of four major Bronze Age burial traditions in Europe.
+The cultural-historical evolution of archaeological cultures and ethnic groups during the European Bronze Age is complex, and its reconstruction an ongoing endeavour of archaeology.
 Within this history of constant emergence, displacement and demise many different forms of burial traditions were practised.
@@
 In the Middle Bronze Age, barrows gained importance in Western and Central Europe.
-This is highly linked to the emergence of the *Tumulus culture* in Central Europe.
+This observation is highly linked to the emergence of the *Tumulus culture* in Central Europe.
 In the late Bronze Age, the Urnfield period, cremation became the most common burial custom in the context of *Urnfield culture*.
 
-Hardings narrative is, even when taking into account it's enormous simplification, partially misleading.
+Harding's narrative is, even when taking into account its enormous simplification, partially misleading.
 Cremation, for example, was a lot more common already in the early and middle Bronze Age, though not in the context of the later Urnfield culture (e.g.
 in the Benelux area [@fokkensBronzeAgeLow2013] or Great-Britain [@robertsBritainIrelandBronze2013]).
-Beyond simple research progress, one additional reason why major misconceptions like this can appear and thrive might be a lack of unified data repositories and universally acknowledged proxies for the distribution of cultural phenomena in European archaeology: Instead of directly querying quantitative data the typical process for the creation of large scale overviews is to excerpt the literature of all relevant local research traditions.
-The present article in contrast explores whether openly available bulk radiocarbon data could allow to reconstruct the trends within the distribution of burial rites.
-This could set the precedent for for a more reproducible and statistically sound observation of the main shifts.
+Beyond simple research progress, one additional reason why major misconceptions like this can appear and thrive might be a lack of unified data repositories and universally acknowledged proxies for the distribution of cultural phenomena in European archaeology.
+Instead of directly querying quantitative data the typical process for the creation of large-scale overviews is to excerpt the literature of all relevant local research traditions.
+The present article, in contrast, explores whether openly available bulk radiocarbon data could allow reconstructing the trends within the distribution of burial rites.
+This approach could set a precedent for a more reproducible and statistically sound observation of the main shifts.
 The quantitative comparison of macroregional trends also paves the way for an analysis of patterns of interaction intensity on a continental scale.
-A core question is how cultural and spatial distance are related over the course of the European Bronze Age.
+A core question is how cultural and spatial distance are related throughout the European Bronze Age.
 
@@
 The spread of ideas in human populations over space and time is a major topic of *Cultural evolution theory* [@creanzaCulturalEvolutionaryTheory2017].
-This line of thought is as least as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* and back to *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004].
-In the late seventies the the discussion took another turn away from social structure to the more general observation of human behaviour.
+This line of thought is at least as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* and back to *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004].
+In the late seventies, the discussion took another turn away from the political and social structure to the more general observation of human behaviour.
 Three major theoretical movements emerged in this context: *Evolutionary psychology*, *Human behavioural ecology* and *Dual inheritance theory* [@smithThreeStylesEvolutionary2000].
-The latter -- heavily popularized by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007].
+The latter -- heavily popularised by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007].
 Just like in biological evolution there might be selection and mutation within *Social learning* and therefore methods of *Population genetics* could be employed within cultural anthropology -- of course with the much faster pace of cultural exchange in mind.
@@
 
-Archaeology has a special perspective to offer for Cultural evolution research: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018].
+Archaeology has a unique perspective to offer for Cultural evolution research: It is concerned with questions that can embrace continent-size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018].
 Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous by-products of living.
 We rightfully expect humans to learn handcraft as well as building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them.
-The similarity of historically and spatially close material remains can therefore be explained as a result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
-In this context a distinction of culture in function and style became an important question for *Evolutionary archaeology*, explicitly elaborated already by @dunnell1978style.
-Some types of human behaviour obviously increase well-being and simplify survival, like certain hunting techniques, superior tools or sowing more fruitful crops.
-Without resorting to the inappropriate simplifications of Evolutionism their diffusion can be expected to increase over time.
-On the other hand there are behaviours that are not obviously related to survival and reproductive capacity, like e.g.
+The similarity of historically and spatially close material remains can, therefore, be explained as a result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
+In this context a distinction of culture in function and style became an essential question for *Evolutionary archaeology*, explicitly elaborated already by @dunnell1978style.
+
+Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
+Without resorting to the inappropriate simplifications of Evolutionism, their diffusion can be expected to increase over time.
+On the other hand, some behaviours are not related to survival and reproductive capacity, like e.g.
 ceramic decorations, traditional costume fashion or jewellery design.
-Despite (or because?) their seemingly secondary importance these characteristics often pass through a breathtakingly fast development and serve archaeologists as proxy for relative chronology.
+Despite (or because?) their seemingly secondary importance these characteristics often pass through a breathtakingly fast development and serve archaeologists as a proxy for relative chronology.
 Dunnell postulates that the distribution and emergence of these stylistic cultural traits behave randomly on a population scale level.
@@
 @neimanStylisticVariationEvolutionary1995 takes a closer look at this processes and develops a simple framework for the simulation of non-functional idea propagation.
-Neiman identifies the essential phenomena *drift* (random dominance of equivalent traits) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.
-Furthermore he implements a measure of cultural distance to apply his theoretical insights to real world data.
+Neiman identifies the essential phenomena *drift* (random dominance of similar traits) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.
+Furthermore, he implements a measure of cultural distance to apply his theoretical insights to real-world data.
 The workflow suggested by Neiman can be partially applied to the burial rites data examined in this article and was an important methodological inspiration: As burial traditions are ritualised behaviour and most of the time heavily linked to religious and spiritual beliefs they can be assumed to be stylistic traits as defined by @dunnell1978style.
-Besides @neimanStylisticVariationEvolutionary1995, another important template for the comparison of cultural and spatial was provided by @porcicSimulatingCulturalTransmission2014.
+Besides @neimanStylisticVariationEvolutionary1995, another vital template for the comparison of cultural and spatial was provided by @porcicSimulatingCulturalTransmission2014.

### 2018-12-18 article/02_Materials_and_Methods.Rmd
language and grammer

@@
 
-The analysis code of this paper is written in the statistics programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible.
+The analysis code of this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible.
 A more in-depth description and discussion of the data preparation can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
@@
 It contains ^14^C dates from the European Bronze and Early Iron Age (2300 BC to 500 BC).
-Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B is a collection of more than 11000 published dates from around 2800 sites from all over Europe (as of December 2018).
-Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the the type and cultural attribution of the sampled context.
-The variable *Feature type* systematically distinguishes for example between settlements, rockshelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, etc..
-This information allows to accurately and precisely determine a moment in space and time, when a grave of a certain type was created.
-In theory the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe.
-Unfortunately the Radon-B dataset has some quality issues mostly due to inconsistent manual data input: Incompleteness and faultiness of meta information, no supervised or well thought-out variable classification for *sample material* and *cultural attribution* and inconsistent multilingualism even within individual entries.
+Radiocarbon dating is a crucial absolute dating technique for this period [@ramseyRadiocarbonDatingRevolutions2008], and Radon-B is a collection of more than 11000 published dates from around 2800 sites from all over Europe (as of December 2018).
+Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the type and cultural attribution of the sampled context.
+The variable *Feature type* systematically distinguishes for example between settlements, rock shelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, i.a..
+This information allows to accurately and precisely determine a moment in space and time when a grave of a particular type was created.
+In theory, the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe.
+Unfortunately, the Radon-B dataset has some quality issues mostly due to inconsistent manual data input: Incompleteness and faultiness of meta information, no supervised or well-thought-out variable classification for *sample material* and *cultural attribution* and inconsistent multilingualism even within individual entries.
 These issues though do not inhibit the general data analysis strategy of this paper.
 
-To create a tidy and useful subset of Radon-B, the following major steps of data preparation were applied.
+To create a tidy and useful subset of Radon-B, the following steps of data preparation were applied.
 Please see the code for details of implementation.
@@
 Transforming data from the meaning *^14^C dates on graves* to *burial type at a certain time*.
-For `r r("bronze15 multi dates one grave", sf_prep)` entries in the thus far assembled subset the attribution to *site* and *feature* is not unique.
+For `r r("bronze15 multi dates one grave", sf_prep)` entries in the thus far assembled subset, the attribution to *site* and *feature* is not unique.
 That means that for many graves multiple (up to 8) ^14^C dates have been created and entered in Radon-B.
-Dates can't be treated as semantically equal to graves.
-There's a variety of phenomena that cause this effect: In some cases safety measurements were taken -- multiple dates for one single grave that all fall within the margin of errors of each other.
-In other cases there is a set of strongly diverging dates for a collective grave that was used for hundreds of years.
+Dates cannot be treated as semantically equal to graves.
+There's a variety of phenomena that cause this effect: In some cases, safety measurements were taken -- multiple dates for one single grave that all fall within the margin of errors of each other.
+In other cases, there is a set of strongly diverging dates for a collective grave that was used for hundreds of years.
 There are also cases where the *feature* information in the database is not sufficient to determine whether two dates are really from the same grave despite the identical entry.
@@
 
-    1. If for an individual ^14^C date no numbers appear in it's *feature* variable value, than it is treated as an individual grave.
+    1. If for an individual ^14^C date no numbers appear in its *feature* variable value, then it is treated as an individual grave.
 
@@
 For the remaining `r r("bronze16 multi dates one grave with numbers", sf_prep)` problematic cases, groups were created by *site* and *feature*.
-The cumulated timespan covered by the $2\sigma$ age range of all these dates where taken as the relevant timespan for the respective grave.
+The cumulated timespan covered by the $2\sigma$ age range of all these dates were taken as the relevant timespan for the respective grave.
 
 11.
-Defining macro regions to observe the development of *burial type* and *burial construction* within sensible geographical and cultural units that are appropriate to the amount of available data.
+Defining macro-regions to observe the development of *burial type* and *burial construction* within sensible geographical and cultural units that are appropriate to the amount of available data.
 The regions are circular and were delimited semi-automatically by creating a regular grid in accordance to main cultural-historical regions of the European Bronze Age (**citation**).
@@
 
-12. Reducing the date selection to the space covered by the macro regions and attributing the dates to the respective units. (`r r("gpr size", sf_desc)` graves)
+12. Reducing the date selection to the space covered by the macro-regions and attributing the dates to the respective units. (`r r("gpr size", sf_desc)` graves)
 
@@
 This data for individual graves attributed to a macro region can be transformed to a time series of trait prevalence within the respective regions.
-If a certain kind of grave can be attributed to a fuzzy point in time with radiocarbon dating, than we also know that most likely a ceremony was held around this time were the respective burial tradition was practised.
+If a certain kind of grave can be attributed to a fuzzy point in time with radiocarbon dating, then we also know that most likely a ceremony was held around this time where the particular burial tradition was practised.
 That means that the cultural trait behind this behaviour must also have been present in one form or the other.
 An example: The idea *inhumation* must have been present in 1447 calBC in Great Britain since this year lies in the $2\sigma$ age range of the ^14^C date *NZA-32497* which was made for the Amesbury Down inhumation burial *I2639/25217*.
-For each year in the time frame 2200 -- 800 calBC we can count the amount of graves with dates that cover this year with their $2\sigma$ age range.
+For each year in the time frame 2200 -- 800 calBC we can count the number of graves with dates that cover this year with their $2\sigma$ age range.
 That allows the construction of region wise time series of absolute occurrence numbers for the variables *burial type* and *burial construction* (Figure \ref{development_burial_type} and \ref{development_burial_construction}).
-Although the number of dates is very small for some regions and time periods we still can attempt to calculate relative proportions of the mutually exclusive traits (inhumation vs.
-cremation and mound grave vs.
-flat grave).
+Although the number of dates is tiny for some regions and periods, we still can attempt to calculate relative proportions of the mutually exclusive traits (inhumation vs cremation and mound grave vs flat grave).
 That leads to relative time series of trait prevalence, which could be a powerful proxy for the development of the distribution of cultural behaviour.
@@
 Several different measures of cultural distance could be used here following the discussion of these matters in the relevant literature (**citation**).
-In this context the Squared Euclidian Distance (SED) seems to be a sensible statistic, because it is fairly simple, works with only few assumptions and emphasizes differences through squaring.
+In this context the Squared Euclidian Distance (SED) seems to be a sensible statistic, because it is relatively simple, works with only a few assumptions and emphasises differences through squaring.
 
@@
 
-To test for correlation between the distances based on the variables *burial type* and *burial construction*, the time frame of interest was divided into sections of 200 years.
+To test for correlation between the distances based on the variables *burial type* and *burial construction*, the timeframe of interest was divided into sections of 200 years.
 For each section of each the two variables, a matrix of the mean SED over time was calculated.
@@
 
-Defining a measure of spatial distance for the above established main units of observation, the macro regions, is only possible in a highly simplified way.
-A continuous distance measure can't be employed, but it's possible to distinguish ordinal distance classes.
-For the eight regions four different categories of spatial distance were defined (Figure \ref{map_regions} B).
+Defining a measure of spatial distance for the above established main units of observation, the macro-regions, is only possible in a highly simplified way.
+A continuous distance measure cannot be employed, but it is possible to distinguish ordinal distance classes.
+For the eight regions, four different categories of spatial distance were defined (Figure \ref{map_regions} B).
 The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the difference in the data level of measurements.
-The test result are listed in Table \ref{tab:distanceandcorrelationtable*}.
+The test results are listed in Table \ref{tab:distanceandcorrelationtable*}.
 
@@
 
-To get a better understanding how the correlation coefficient should develop if there were strong correlation between spatial and cultural distance as measured by *burial type* and *burial construction*, a simple simulation was implemented based on the model proposed by @neimanStylisticVariationEvolutionary1995.
+To get a better understanding how the correlation coefficient should develop if there was a strong correlation between spatial and cultural distance as measured by *burial type* and *burial construction*, a simple simulation was implemented based on the model proposed by @neimanStylisticVariationEvolutionary1995.
 This simulation assumes a population separated into $g$ groups with a group population of $N_g$ each.
-At the beginning of the simulation there are $k$ ideas randomly, but equally distributed in this population.
-With every timestep -- $t$ defines the total number of time steps -- all individuals learn from an other randomly chosen individual.
-The probability that this other individual is from an other group equals $m_i$, for the opposite therefore $1-m_i$.
-The partner group for each individual learning instance is selected randomly following a group interaction probability defined in the matrix $I$.
-The distribution of the $k$ variants can be observed for every timestep and documented in a time series like the one for real world data.
+At the beginning of the simulation, there are $k$ ideas randomly but equally distributed in this population.
+With every timestep -- $t$ defines the total number of time steps -- all individuals learn from another randomly chosen individual.
+The probability that this other individual is from another group equals $m_i$, for the opposite therefore $1-m_i$.
+The partner group for each learning instance is selected randomly following a group interaction probability defined in the matrix $I$.
+The distribution of the $k$ variants can be observed for every timestep and documented in a time series like the one for real-world data.
 

### 2018-12-18 article/04_Discussion.Rmd
language and grammer

--- 2018-12-17 article/04_Discussion.Rmd
@@
 
-Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region wise timeseries of trait occurrence.
-The figure part A with absolute numbers indicates areas and timeperiods with a lot, and others with very few points of data.
-For some there is not enough information to make any decision about the relative dominance and for many the general relevance and accuracy of the resulting timeseries can be questioned.
+The A sections of the figures \ref{development_burial_type} and \ref{development_burial_construction} with absolute grave numbers per region and year indicate areas and periods with a lot and others with very few points of data.
+For some, there is not enough information to make any decision about the relative dominance and for many, the general relevance and accuracy of the resulting time series can be questioned.
 There's a variety of reasons for this:
 
-1. The data in Radon-B depends on the projects and interests of the involved researchers: The data was collected as needed for individual regions rather then systematically for all of europe.
+1. The data in Radon-B depends on the projects and interests of the involved researchers: The data was collected as needed for individual regions rather than systematically for all of Europe.
 
@@
 The availability of ^14^C dates for a region and period in the literature depends on whether a modern research project took place within this context.
-Besides the research interets of researchers and institutions, there are also political reasons why an area might be better or worse explored.
+Besides the research interests of researchers and institutions, there are also political reasons why an area might be better or worse explored.
 
@@
 In various Metal Age contexts, 14C data are inferior to the accuracy of relative chronological classifications - e.g.
-on the basis of primer typology - and are therefore used only sporadically to create absolute chronological fix points.
+by primer typology - and are therefore used only sporadically to create absolute chronological fixpoints.
 Dendrochronological dating is preferred in moist soil contexts and when the necessary woods are available.
-The available ^14^C calibration curves form plateaus in some periods, which significantly limit the achievable dating accuracy of the data.
+The available ^14^C calibration curves form plateaus in some periods, which significantly limit the possible dating accuracy of the data.
 If such an effect is known, then no data is ordered at all.
@@
 The majority of ^14^C data for this work comes from bones and teeth.
-However, their preservation is subject to great differences depending on the soil [@brockPrescreeningTechniquesIdentification2010; @gordonSoilPHBone1981; @nielsen-marshBoneDiagenesisEuropean2007].
+However, their preservation is subject to huge differences depending on the soil [@brockPrescreeningTechniquesIdentification2010; @gordonSoilPHBone1981; @nielsen-marshBoneDiagenesisEuropean2007].
 In many mineral soil contexts no sufficiently good, dateable material is available.
-This can affect entire regions, if, for example, poor bone preservation prevails in general as a result of low-calcification soils.
+This situation can affect entire regions, if, for example, poor bone preservation prevails in general as a result of low-calcification soils.
 
@@
 From very few regions and periods of the European Bronze Age so many graves are documented that it can be assumed that a significant sample of the burial culture of the entire population has been researched.
-Instead, it is evident in several contexts that only the burials of individual social groups, gender or ethnic group were recorded: The quantity and nature of the known graves simply cannot be representative of all the deceased.
-In this case it can be assumed that further, deviant burial rituals were practiced, which left no remains or remains that were difficult to grasp archaeologically.
+Instead, it is evident in several contexts that only the burials of individual social groups, gender or ethnic group were recorded: The quantity and nature of the known graves cannot be representative of all the deceased.
+In this case, it can be assumed that further, deviant burial rituals were practised, which left no remains or remains that were difficult to grasp archaeologically.
 These rituals are not shown in the ^14^C data from Radon-B.
+
+Concerning the primary variables *burial type* and *burial construction*, further special effects must be taken into account: The disturbance and robbery of graves was a widespread phenomenon in history and prehistory.
+Burial mounds were usually attacked more severely than flat graves due to their better visibility and could, therefore, be underrepresented in Radon-B.
+In many places, they were also the subject of archaeological expeditions in the early days of archaeology at the end of the 19th and beginning of the 20th century.
+Of course, these excavations did not yield ^14^C-dates.
+Irrespective of this, it can be assumed that a large part of the burial mounds erected in the Bronze Age was destroyed by agricultural activity.
+Ploughing could also result in an under-representation of the archaeological findings [@hardingEuropeanSocietiesBronze2000a].
+On the other hand, if a burial mound is still present, it is more readily visible and recognised by archaeologists today: Mounds are often clearly identifiable in aerial photographs or digital terrain models, while flat graves are much more difficult to identify for today's remote sensing methods.
+It is difficult to estimate the impact of these effects in detail, but an overrepresentation of mounds seems more likely than an overrepresentation of flat graves.
+The dating of cremation graves on bone remains methodically differs from the dating of body graves because it is based on the extraction of the mineral, not the organic, fraction in the bone [@lantingDatingCrematedBones2001].
+The underlying chemical method was developed later, has only been used more frequently in the past 15 years and is associated with some particular difficulties [@olsenCharacterisationBlindTesting2008; @strydonck14CDatingCremated2009].
+Burnt bone is generally more susceptible to weathering than unburned bone.
+An under-representation of cremation graves in current data collections is, therefore, more plausible.
 
@@
 
-- General and concluding observations
-
-The data representativity and semantic consequences will be discussed below.
-
 In reality, beyond Dunnels distinction, most behaviour is not unambiguously stylistic or functional.
-For the sake of this case study burial rites are assumed to be stylistic, but in fact their creation causes costs in terms of working time and material to a varying extent.
-This cost might have been influenced or had their own effect on the human-nature relation: Pyres for example require a significant amount of wood and the erection of barrows several days of manpower.
-Still it can be expected that these functional disadvantages weren't critical most of the time for the general decision which kind of burial tradition is applied, but maybe for the magnitude of invested energy.
+For the sake of this case study burial rites are assumed to be stylistic, but in fact, their creation causes costs regarding working time and material to a varying extent.
+This cost might have been influenced or had their effect on the human-nature relation: Pyres, for example, require a significant amount of wood and the erection of barrows much workforce.
+Still, it can be expected that these functional disadvantages were not critical most of the time for the general decision which kind of burial tradition is applied, but maybe for the magnitude of invested energy.
 Other factors to consider here are the level of grief [@macdonaldGriefBurialAmerican2001] or the representational intentions of the burying relatives [@henrichEvolutionCostlyDisplays2009].

### 2018-12-18 article/02_Materials_and_Methods.Rmd
made tables more similar

@@
     escape = F,
-    linesep = "\\addlinespace"
+    linesep = "\\addlinespace \\hline \\addlinespace"
   ) %>%
@@
     full_width = T, 
-    font_size = 8,
+    font_size = 9,
     latex_options = "hold_position"
@@
   kableExtra::kable_styling(
-    latex_options = c("hold_position")
+    latex_options = c("hold_position"),
+    font_size = 9,
+    full_width = T
   ) %>%

### 2018-12-18 analysis/code/04_simulation_scripts/050_general_neiman_simulation.R
copied necessary code fragments for simulation

@@ -1,0 +1,197 @@
+#### simulation function ####
+
+#' neiman_simulation
+#'
+#' @param k Integer. Number of ideas at t = 0
+#' @param N_g Integer. Population per region
+#' @param t_final Integer. Final timestep
+#' @param mu Double. Innovation rate
+#' @param g Integer. Number of regions
+#' @param mi Double. Degree of interregion interaction
+#' @param I Doublematrix. Interregion interaction matrix. NA means equal interaction
+#' 
+neiman_simulation <- function(k, N_g, t_final, mu, g, mi, I) {
+
+  # define variables
+  regions <- 1:g
+  population <- 1:N_g
+  ideas <- 1:k
+  timesteps <- 2:t_final
+  if (is.na(I)) {
+    I <- matrix(
+      rep(1, g*g), g, g
+    )
+    diag(I) <- 0
+  }
+  
+  # create starting populations
+  pop0 <- lapply(
+    regions, function(region, N, k) {
+      tibble::tibble(
+        timestep = as.integer(0),
+        individual = population,
+        idea = rep_len(ideas, max(population)),
+        region = region
+      )
+    },
+    k, population
+  )
+  
+  # create development list
+  pop_devel <- list()
+  pop_devel[[1]] <- pop0
+  
+  # determine number of ideas
+  last_idea <- max(do.call(rbind, pop_devel[[1]])$idea)
+  
+  # simulation loop
+  for (p1 in timesteps) {
+    
+    # new timestep list
+    pop_old <- pop_devel[[p1 - 1]]
+    pop_new <- pop_old
+    
+    # adjust time in new timestep list
+    pop_new <- lapply(
+      pop_new, function(x, p1) {
+        x$timestep <- p1 - 1
+        return(x)
+      },
+      p1
+    )
+    
+    # intraregion learning
+    pop_new <- lapply(
+      pop_new, function(x) {
+        x$idea <- sample(x$idea, length(x$idea), replace = T)
+        return(x)
+      }
+    )
+    
+    # interregion learning
+    pop_new <- lapply(
+      regions, function(i, pop_new, pop_old, mi, I, regions) {
+        exchange_where <- which(sample(c(TRUE, FALSE), nrow(pop_new[[i]]), prob = c(mi, 1 - mi), replace = T))
[…]

### 2018-12-18 analysis/code/04_simulation_scripts/100_multiple_simulation_sed.R

@@ -1,0 +1,69 @@
+load("data_analysis/region_order.RData")
+
+##### read simulation output data #####
+
+models <- pbapply::pblapply(
+  list.files("../simulationdata/sed_simulation", full.names = TRUE),
+  function(y) {
+    read.csv(y) %>% tibble::as.tibble()
+  }
+)
+
+prop <- do.call(rbind, models)
+
+#### load sed function ####
+
+load("data_analysis/sed_function.RData")
+
+#### data preparation ####
+
+long_prop <- prop %>%
+  tidyr::spread(
+    idea, proportion
+  )
+
+regions_grid <- 
+  expand.grid(
+    regionA = prop$region %>% unique(), 
+    regionB = prop$region %>% unique(), 
+    time = prop$timestep %>% unique(), 
+    model_id =  prop$model_id %>% unique(),
+    stringsAsFactors = FALSE
+  ) %>%
+  tibble::as.tibble() %>%
+  dplyr::left_join(
+    long_prop,
+    by = c(
+      "regionA" = "region", 
+      "time" = "timestep", 
+      "model_id" = "model_id"
+    )
+  ) %>%
+  dplyr::left_join(
+    subset(long_prop, select=-c(model_group)),
+    by = c(
+      "regionB" = "region", 
+      "time" = "timestep", 
+      "model_id" = "model_id"
+    ),
+    suffix = c("_regionA", "_regionB")
+  )
+
+region_A <- mapply(c, regions_grid$idea_1_regionA, regions_grid$idea_2_regionA, SIMPLIFY = FALSE)
+region_B <- mapply(c, regions_grid$idea_1_regionB, regions_grid$idea_2_regionB, SIMPLIFY = FALSE)
+region_A_region_B <- mapply(list, region_A, region_B, SIMPLIFY = FALSE)
+
+regions_grid$sed <- unlist(pbapply::pblapply(
+  region_A_region_B, function(x) {
+    sed(x[[1]], x[[2]])
+  }
+))
+
+regions_grid %<>% dplyr::select(
+    regionA, regionB, time, sed, model_id, model_group
+  )
+
+regions_grid$regionA <- factor(regions_grid$regionA, levels = region_order)
+regions_grid$regionB <- factor(regions_grid$regionB, levels = region_order)
+
+save(regions_grid, file = "data_simulation/sed_simulation_regions_grid.RData")

### 2018-12-18 analysis/code/04_simulation_scripts/200_multiple_simulation_sed_create_matritzes.R

@@ -1,0 +1,44 @@
+load("data_simulation/sed_simulation_regions_grid.RData")
+
+##### cut sed regions grid into timeslices ####
+
+modelid_time_regions_grid <- regions_grid %>% dplyr::mutate(
+  time = base::cut(
+    time, 
+    seq(-2200, -800, 200), labels = paste(seq(-2200, -1000, 200), seq(-2000, -800, 200), sep = " - "),
+    include.lowest = TRUE, 
+    right = FALSE)
+) %>%
+  dplyr::group_by(
+    model_group, model_id, time, regionA, regionB
+  ) %>%
+  dplyr::summarise(
+    mean_sed = mean(sed, na.rm = TRUE)
+  ) %>%
+  dplyr::ungroup() %>%
+  # that's dangerous...
+  dplyr::mutate(
+    mean_sed = tidyr::replace_na(mean_sed, 0)
+  )
+
+#### create list of matrizes from timesliced regions grid ####
+
+distance_matrizes_sed <- lapply(
+  base::split(modelid_time_regions_grid, modelid_time_regions_grid$model_id), function(time_regions_grid) {
+    lapply(
+      base::split(time_regions_grid, time_regions_grid$time), function(x){
+        x %>%
+          dplyr::select(
+            -model_group, -model_id, -time
+          ) %>%
+          tidyr::spread(regionA, mean_sed) %>%
+          dplyr::select(
+            -regionB
+          ) %>%
+          as.matrix()
+      }
+    )
+  }
+)
+
+save(distance_matrizes_sed, file = "data_simulation/sed_simulation_regions_timeslices_matrizes.RData")

### 2018-12-18 analysis/code/04_simulation_scripts/300_multiple_simulation_sed_vs_spatial_distance_create_data.R

@@ -1,0 +1,61 @@
+load("data_analysis/region_order.RData")
+load("data_simulation/sed_simulation_regions_grid.RData")
+load("data_analysis/distance_matrix_spatial_long.RData")
+
+#### half regions_grid -- removal of double entries ####
+
+regions_grid$regionA <- as.character(regions_grid$regionA)
+regions_grid$regionB <- as.character(regions_grid$regionB)
+
+regions_grid_half <- pbapply::pblapply(
+  base::split(regions_grid, f = regions_grid$model_group), function(z) { 
+    lapply(
+      base::split(z, f = regions_grid$model_id), function(y) { 
+        lapply(
+          base::split(y, f = y$time), function(x) {
+            mn <- pmin(x$regionA, x$regionB)
+            mx <- pmax(x$regionA, x$regionB)
+            int <- as.numeric(interaction(mn, mx))
+            x <- x[match(unique(int), int),]
+            return(x)
+          }) %>%
+          do.call(rbind, .)
+      }) %>%
+      do.call(rbind, .)    
+  }) %>%
+  do.call(rbind, .)
+
+regions_grid_half$regionA <- factor(regions_grid_half$regionA, levels = region_order)
+regions_grid_half$regionB <- factor(regions_grid_half$regionB, levels = region_order)
+
+#### combine spatial and cultural distance into one data.frame ####
+
+sed_spatial_distance <- regions_grid_half %>% dplyr::left_join(
+  distance_matrix_spatial_long, by = c("regionA", "regionB")
+) %>% 
+  dplyr::filter(
+    distance != 0
+  ) %>%
+  dplyr::mutate(
+    relation = paste(regionA, "+", regionB),
+    time = base::cut(
+      time, 
+      seq(-2200, -800, 200), labels = paste(seq(-2200, -1000, 200), seq(-2000, -800, 200), sep = " - "),
+      include.lowest = TRUE, 
+      right = FALSE)
+  ) %>%
+  dplyr::group_by(
+    model_id, model_group, time, regionA, regionB, distance
+  ) %>%
+  dplyr::summarise(
+    mean_sed = mean(sed, na.rm = TRUE)
+  ) %>%
+  dplyr::ungroup() %>%
+  dplyr::filter(
+    !is.na(mean_sed)
+  )
+
+save(
+  sed_spatial_distance, 
+  file = "data_simulation/sed_simulation_regions_timeslices_spatial_distance.RData"
+)

### 2018-12-18 analysis/code/04_simulation_scripts/400_multiple_simulation_sed_vs_spatial_distance_mantel_test.R

@@ -1,0 +1,30 @@
+load("data_analysis/distance_matrix_spatial.RData")
+load("data_simulation/sed_simulation_regions_timeslices_matrizes.RData")
+
+mantel_test_results <- pbapply::pblapply(
+  1:length(distance_matrizes_sed), function(i, x, y, model_id) {
+      lapply(
+        1:length(x[[i]]), function(i, x, y, time, model_id) {
+          mantel_result <- vegan::mantel(x[[i]], y, method = "spear", permutations=999)
+          data.frame(
+            model_id = model_id,
+            time = time[[i]],
+            statistic = mantel_result$statistic, 
+            signif = mantel_result$signif
+          )
+        },
+        x = x[[i]],
+        y = y,
+        time = names(x[[i]]),
+        model_id = model_id[[i]]
+      ) %>% do.call(rbind, .)
+    },
+    x = distance_matrizes_sed,
+    y = distance_matrix_spatial,
+    model_id = names(distance_matrizes_sed)
+  ) %>% do.call(rbind, .) %>% tibble::as.tibble()
+
+save(
+  mantel_test_results, 
+  file = "data_simulation/sed_simulation_mantel_sed_spatial.RData"
+)

### 2018-12-18 analysis/code/04_simulation_scripts/500_multiple_simulation_sed_vs_spatial_distance.R

@@ -1,0 +1,73 @@
+library(magrittr)
+
+load("data_simulation/sed_simulation_regions_timeslices_spatial_distance.RData")
+sed_simulation <- sed_spatial_distance
+load("data_analysis/squared_euclidian_distance_over_timeblocks_burial_type.RData")
+sed_burial_type <- sed_spatial_distance %>% dplyr::mutate(model_group = "burial_type")
+load("data_analysis/squared_euclidian_distance_over_timeblocks_burial_construction.RData")
+sed_burial_construction <- sed_spatial_distance %>% dplyr::mutate(model_group = "burial_construction")
+
+sed_all <- sed_simulation %>%
+  dplyr::select(-model_id) %>%
+  rbind(sed_burial_type) %>%
+  rbind(sed_burial_construction)
+
+sed_all <- sed_all %>%
+  dplyr::filter(mean_sed != 0)
+
+#print(sed_all)
+
+library(ggplot2)
+plu <- ggplot(sed_all) +
+  geom_boxplot(
+    mapping = aes(
+      x = as.factor(distance),
+      y = mean_sed,
+      fill = model_group
+    ),
+    width = 0.9
+  ) +
+  facet_wrap(
+    nrow = 2,
+    ~time
+  ) +
+  scale_fill_manual(
+    name = "Real world context",
+    values = c(
+      "burial_type" = "#0072B2",
+      "burial_construction" = "#009E73",
+      "low equal interaction" = "#fb9a99",
+      "low spatial interaction" = "#ffff33",
+      "high equal interaction" = "#e31a1c",
+      "high spatial interaction" = "#ff7f00"
+    )
+  ) +
+  theme_bw() +
+  theme(
+    plot.title = element_text(size = 30, face = "bold"),
+    legend.position = "bottom",
+    legend.title = element_text(size = 20, face = "bold"),
+    legend.text = element_text(size = 20),
+    strip.text.x = element_text(size = 20),
+    axis.text = element_text(size = 20),
+    axis.title = element_text(size = 20)
+  ) +
+  guides(
+    fill = guide_legend(nrow = 3, byrow = TRUE)
+  ) +
+  xlab("Spatial Distance Classes") +
+  ylab("Squared Euclidian Distance") +
+  ylim(0, 2) +
+  NULL
+
+plu %>%
+  ggsave(
[…]

### 2018-12-18 analysis/code/04_simulation_scripts/600_multiple_simulation_sed_vs_spatial_distance_spearman.R

@@ -1,0 +1,158 @@
+library(magrittr)
+
+load("data_simulation/sed_simulation_mantel_sed_spatial.RData")
+mantel_simulations <- mantel_test_results %>% dplyr::mutate(model_id = as.integer(model_id))
+load("data_simulation/sed_simulation_model_grid.RData")
+mantel_simulations %<>% dplyr::left_join(models_grid[, c("model_id", "model_group")], by = "model_id")
+load("data_analysis/mantel_sed_spatial_burial_type.RData")
+mantel_burial_type <- mantel_test_results
+load("data_analysis/mantel_sed_spatial_burial_construction.RData")
+mantel_burial_construction <- mantel_test_results
+
+mantel_burial_type %<>%
+  dplyr::mutate(
+    context = "burial_type"
+  )
+
+mantel_burial_construction %<>%
+  dplyr::mutate(
+    context = "burial_construction"
+  )
+
+mantel_real_world <- rbind(mantel_burial_type, mantel_burial_construction)
+
+library(ggplot2)
+
+plot_mantel <- function(title, mantel_simulations, mantel_real_world) {
+  ju <- ggplot() +
+    geom_hline(
+      yintercept = 0,
+      colour = "red",
+      size = 2
+    ) + 
+    geom_line(
+      data = mantel_simulations,
+      mapping = aes(
+        x = time,
+        y = statistic,
+        group = model_id
+      ),
+      position = position_nudge(x = -0.25),
+      size = 0.2,
+      alpha = 0.2
+    ) +
+    geom_point(
+      data = mantel_simulations,
+      mapping = aes(
+        x = time,
+        y = statistic
+      ),
+      position = position_nudge(x = -0.25),
+      size = 1
+    ) +
+    geom_point(
+      data = mantel_real_world,
+      mapping = aes(
+        x = time,
+        y = statistic,
+        colour = context
+      ),
+      position = position_nudge(x = -0.25),
+      size = 6
+    ) +
+    geom_line(
+      data = mantel_real_world,
+      mapping = aes(
+        x = time,
+        y = statistic,
+        colour = context,
+        group = context
+      ),
+      position = position_nudge(x = -0.25),
+      size = 0.7
+    ) +
+    geom_boxplot(
+      data = mantel_simulations,
+      mapping = aes(
+        x = time,
+        y = statistic
+      ),
+      width = 0.2
+    ) +
+    geom_dotplot(
[…]

### 2018-12-18 analysis/code/04_simulation_scripts/605_multiple_simulation_sed_vs_spatial_distance_spearman_stepwise_plot.R

@@ -1,0 +1,506 @@
+library(magrittr)
+
+load("data_simulation/sed_simulation_mantel_sed_spatial.RData")
+mantel_simulations <- mantel_test_results %>% dplyr::mutate(model_id = as.integer(model_id))
+load("data_simulation/sed_simulation_model_grid.RData")
+mantel_simulations %<>% dplyr::left_join(models_grid[, c("model_id", "model_group")], by = "model_id")
+load("data_analysis/mantel_sed_spatial_burial_type.RData")
+mantel_burial_type <- mantel_test_results
+load("data_analysis/mantel_sed_spatial_burial_construction.RData")
+mantel_burial_construction <- mantel_test_results
+
+mantel_burial_type %<>%
+  dplyr::mutate(
+    context = "burial_type"
+  )
+
+mantel_burial_construction %<>%
+  dplyr::mutate(
+    context = "burial_construction"
+  )
+
+mantel_real_world <- rbind(mantel_burial_type, mantel_burial_construction)
+
+library(ggplot2)
+
+variant <- "high equal interaction"
+title <- variant %>% gsub(" ", "_", .)
+mantel_simulations <- mantel_simulations %>% dplyr::filter(model_group == variant)
+mantel_real_world <- mantel_real_world
+
+#### 1 ####
+
+ju1 <- ggplot() +
+  geom_hline(
+    yintercept = 0,
+    colour = "red",
+    size = 2
+  ) + 
+  geom_line(
+    data = mantel_simulations,
+    mapping = aes(
+      x = time,
+      y = statistic,
+      group = model_id
+    ),
+   position = position_nudge(x = -0.25),
+    size = 0.2,
+    alpha = 0
+  ) +
+  geom_point(
+    data = mantel_simulations,
+    mapping = aes(
+      x = time,
+      y = statistic
+    ),
+    position = position_nudge(x = -0.25),
+    size = 1,
+    alpha = 0
+  ) +
+  geom_point(
+    data = mantel_real_world,
+    mapping = aes(
+      x = time,
+      y = statistic,
+      colour = context
+    ),
+    position = position_nudge(x = -0.25),
+    size = 6
+  ) +
+  geom_line(
+    data = mantel_real_world,
+    mapping = aes(
+      x = time,
+      y = statistic,
+      colour = context,
+      group = context
+    ),
[…]

### 2018-12-18 analysis/code/05_plot_scripts/sed_simulation.R
preliminary version of simulation plot

@@ -122,22 +122,15 @@ plot_mantel <- function(title, mantel_simulations, mantel_real_world) {
       strip.text.x = element_text(size = 13),
       legend.title = element_text(size = 15, face = "bold"),
       legend.text = element_text(size = 15),
-      legend.box = "vertical"
+      legend.box = "vertical",
+      legend.text.align = 0
     ) +
-    ylab("Spearman's rank correlation coefficient") +
-    xlab("time blocks calBC") +
-    ylim(c(-0.6, 0.75))
+    #ylab("Spearman's rank correlation coefficient") +
+    #xlab("time blocks calBC") +
+    ylim(c(-0.5, 0.7))
 
-ju %>%
-  ggsave(
-    paste0("analysis/figures/", title, ".jpeg"),
-    plot = .,
-    device = "jpeg",
-    scale = 1,
-    dpi = 300,
-    width = 300, height = 300, units = "mm",
-    limitsize = F
-  )
+  ju
+
 }
 
 variants <- c(1,2,3,4,5,6) %>% lapply(
@@ -149,5 +142,35 @@ variants <- c(1,2,3,4,5,6) %>% lapply(
     )
   }
 )
+
+top <- cowplot::plot_grid(
+  plotlist = lapply(variants, function(x){ x + theme(legend.position = "none") }),
+  labels = LETTERS[1:6],
+  rel_heights = c(1, 1, 1),
+  nrow = 3,
+  ncol = 2,
+  align = "h",
+  axis = "lr",
+  label_size = 25
+)
+
+total <- cowplot::plot_grid(
+  top,
+  cowplot::get_legend(variants[[1]]),
+  labels = c('', ''),
+  ncol = 1,
+  rel_heights = c(1, 0.1)
+)
+
+total %>%
+  ggsave(
+    paste0("analysis/figures/simulation.jpeg"),
+    plot = .,
+    device = "jpeg",
+    scale = 1,
+    dpi = 300,
+    width = 330, height = 400, units = "mm",
+    limitsize = F
+  )
 
 

### 2018-12-18 analysis/code/run_all_scripts.R
simulation scripts should work now

@@ -5,7 +5,7 @@ purrr::walk(
     "analysis/code/01_data_preparation_scripts",
     "analysis/code/02_dataset_description_scripts",
     "analysis/code/03_cultural_distance_scripts",
-    #"analysis/code/04_simulation_scripts",
+    "analysis/code/04_simulation_scripts",
     "analysis/code/05_plot_scripts"
   ),
   function(x) {

### 2019-01-10 article/02_Materials_and_Methods.Rmd
added preliminary simulation figure to text
some work on figure captions

--- 2018-12-18 article/02_Materials_and_Methods.Rmd
@@
 # Materials and Methods {#materials-and-methods}
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../analysis/figures/map_graves.jpeg}
+\caption{\label{map_graves}All graves dating between 2200 and 800calBC as extracted from the Radon-B database. The red frame indicates the research area.}
+\end{figure}
 
@@
 \centering
-\includegraphics[width=\columnwidth]{../analysis/figures/map_graves.jpeg}
-\caption{\label{map_graves}This is a figure.}
-\end{figure}
-
-\begin{figure}
-\centering
 \includegraphics[width=\columnwidth]{../analysis/figures/map_regions.jpeg}
-\caption{\label{map_regions}This is a figure.}
+\caption{\label{map_regions}Circular, artificial regions (\textbf{A}) and spatial distance network with classified distance values (\textbf{B}).}
 \end{figure}
@@
   knitr::kable(
-    caption = "Table of burial rite distribution within the active dataset",
+    caption = "Burial rite distribution in the active dataset.",
     format = "latex",
@@
 \includegraphics[width=\columnwidth]{../analysis/figures/development_burial_type.jpeg}
-\caption{\label{development_burial_type}This is a figure.}
+\caption{\label{development_burial_type}Regional distribution time series of the variable \textit{burial type} with the absolute amount (\textbf{A}) and proportion (\textbf{B}) of grave types.}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{../analysis/figures/development_burial_construction.jpeg}
-\caption{\label{development_burial_construction}This is a figure.}
+\caption{\label{development_burial_construction}Like Figure \ref{development_burial_type}, but for the variable \textit{burial construction}.}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{../analysis/figures/sed_region_matrix.jpeg}
-\caption{\label{sed_region_matrix}This is a figure.}
+\caption{\label{sed_region_matrix}Plot matrix of the region-region squared euclidian distance time series.
+In the top right half for the variable \textit{burial type} and in the bottom left for the variable \textit{burial construction}.}
 \end{figure*}
@@
   "latex", 
-  caption = "Mean cultural distance, correlation and mantel test results", 
+  caption = "Mean distance, correlation and mantel test results for cultural and spatial distance.", 
   booktabs = T,
@@
 
-For the actual simulation runs, the parameters were chosen as follows: **TODO**
+For the actual simulation runs, the parameters were chosen as follows: **TODO** Figure \ref{simulation*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{../analysis/figures/simulation.jpeg}
+\caption{\label{simulation}Simulation results in comparison with observed data.}
+\end{figure*}

### 2019-01-10 article/01_Introduction.Rmd
improvements inn the introduction section

@@
 Within this history of constant emergence, displacement and demise many different forms of burial traditions were practised.
-Most of them can be distinguished by the criteria *burial construction* and *treatment of the body*.
+Most of them can be distinguished by the criteria *construction of the grave* and *treatment of the body*.
 Bronze Age burials are either flat graves or part of a burial mound, and they contain either untreated or cremated human remains.
+
 In a nutshell, the classic narrative of the development of these rites in Europe presented e.g.
@@
 In the late Bronze Age, the Urnfield period, cremation became the most common burial custom in the context of *Urnfield culture*.
-
 Harding's narrative is, even when taking into account its enormous simplification, partially misleading.
@@
 in the Benelux area [@fokkensBronzeAgeLow2013] or Great-Britain [@robertsBritainIrelandBronze2013]).
-Beyond simple research progress, one additional reason why major misconceptions like this can appear and thrive might be a lack of unified data repositories and universally acknowledged proxies for the distribution of cultural phenomena in European archaeology.
-Instead of directly querying quantitative data the typical process for the creation of large-scale overviews is to excerpt the literature of all relevant local research traditions.
-The present article, in contrast, explores whether openly available bulk radiocarbon data could allow reconstructing the trends within the distribution of burial rites.
-This approach could set a precedent for a more reproducible and statistically sound observation of the main shifts.
-The quantitative comparison of macroregional trends also paves the way for an analysis of patterns of interaction intensity on a continental scale.
-A core question is how cultural and spatial distance are related throughout the European Bronze Age.
+
+How can major misconceptions like this arise and persist in European archaeology?
+Beyond missing data due to preservation and research history, an additional reason might be a lack of unified data repositories and universally acknowledged methods to measure the spatial and temporal distribution of cultural phenomena.
+Instead of directly querying archives with quantitative information, the typical process for the creation of large-scale syntheses is still to manually excerpt the literature of all relevant, local research traditions.
+The present article, in contrast, explores the former strategy: Does openly available, bulk radiocarbon data allow reconstructing the main trends within the distribution of burial rites in Bronze Age Europe?
+This approach could set a precedent for a more reproducible and statistically sound research practice.
+If so, then the quantitative comparison of macroregional trends could also pave the way for an analysis of interaction intensity on a long-term continental scale: Do time series constructed from radiocarbon data allow meaningful observations concerning cultural and spatial distance?
 
@@
 Just like in biological evolution there might be selection and mutation within *Social learning* and therefore methods of *Population genetics* could be employed within cultural anthropology -- of course with the much faster pace of cultural exchange in mind.
-A lot of theoretical and quantitative research arose from this assumptions: e.g.
+A lot of theoretical and quantitative research arose from these assumptions: e.g.
 @cavalli-sforzaCulturalTransmissionEvolution1981, @boydCultureEvolutionaryProcess1985, @teltserEvolutionaryArchaeologyMethodological1995 or @obrienEvolutionaryArchaeologyTheory1996.

### 2019-01-10 article/01_Introduction.Rmd
and even more love for the introduction

@@
 
-Already the assumption that the cremation aspect of Early Bronze Age burial rituals in Britain can be directly compared to the same aspect in Hungary requires a theoretical foundation.
-*Cultural transmission theory* provides a framework where this claim can be explored as a hypothesis for further research.
-The spread of ideas in human populations over space and time is a major topic of *Cultural evolution theory* [@creanzaCulturalEvolutionaryTheory2017].
-This line of thought is at least as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* and back to *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004].
-In the late seventies, the discussion took another turn away from the political and social structure to the more general observation of human behaviour.
-Three major theoretical movements emerged in this context: *Evolutionary psychology*, *Human behavioural ecology* and *Dual inheritance theory* [@smithThreeStylesEvolutionary2000].
-The latter -- heavily popularised by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007].
-Just like in biological evolution there might be selection and mutation within *Social learning* and therefore methods of *Population genetics* could be employed within cultural anthropology -- of course with the much faster pace of cultural exchange in mind.
-A lot of theoretical and quantitative research arose from these assumptions: e.g.
+To answer these questions, a theoretical framework is required that allows to compare human behaviour across spatial, temporal and cultural boundaries: The assumption that the cremation aspect of Early Bronze Age burial rituals in Britain can be directly compared to the same aspect in Hungary is not self-evident.
+*Cultural evolution theory* provides such a framework [@creanzaCulturalEvolutionaryTheory2017].
+The spread of ideas in human populations over space and time is one of it's major topics: *Cultural transmission*.
+
+*Cultural evolution theory* is at least as old as theories of biological evolution and underwent a long and versatile research history.
+Its beginnings are in *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, matured in conflict with *Cultural relativism* into *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004].
+In the late seventies three major theoretical movements emerged: *Evolutionary psychology*, *Human behavioural ecology* and *Dual inheritance theory* [@smithThreeStylesEvolutionary2000].
+The latter -- heavily popularised by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007]: Just like in biological evolution also *Social learning* might be prone to selection and mutation.
+Under this assumption methods of *Population genetics* could be employed within cultural anthropology to analyse the spread and development of ideas -- of course with the much faster pace of cultural exchange in mind.
+A lot of theoretical and quantitative research follows this line of thought: e.g.
 @cavalli-sforzaCulturalTransmissionEvolution1981, @boydCultureEvolutionaryProcess1985, @teltserEvolutionaryArchaeologyMethodological1995 or @obrienEvolutionaryArchaeologyTheory1996.

### 2019-01-10 article/01_Introduction.Rmd
archaeology and cultural evolution

@@
 
-Archaeology has a unique perspective to offer for Cultural evolution research: It is concerned with questions that can embrace continent-size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018].
-Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous by-products of living.
+Archaeology has a unique perspective on *Cultural evolution* [@garveyCurrentPotentialRoles2018]: It is concerned with human behaviour over big areas and long timescales.
+It also works with solid proxies of cultural expression: material remains -- individual lost or discarded artefacts, the ruins of past architecture and all more inconspicuous by-products of living.
 We rightfully expect humans to learn handcraft as well as building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them.
-The similarity of historically and spatially close material remains can, therefore, be explained as a result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
+The similarity of historically and spatially close material remains can, therefore, be explained as a direct result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
+Questions of *Cultural Transmission* form the core of archaeology.
+
 In this context a distinction of culture in function and style became an essential question for *Evolutionary archaeology*, explicitly elaborated already by @dunnell1978style.
-
 Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.

### 2019-01-10 article/01_Introduction.Rmd
better integration of evolutionary archaeology

@@
 
-Archaeology has a unique perspective on *Cultural evolution* [@garveyCurrentPotentialRoles2018]: It is concerned with human behaviour over big areas and long timescales.
+Archaeology has a unique perspective on Cultural evolution [@garveyCurrentPotentialRoles2018]: It is concerned with human behaviour over big areas and long timescales.
 It also works with solid proxies of cultural expression: material remains -- individual lost or discarded artefacts, the ruins of past architecture and all more inconspicuous by-products of living.
@@
 
-In this context a distinction of culture in function and style became an essential question for *Evolutionary archaeology*, explicitly elaborated already by @dunnell1978style.
+*Evolutionary archaeology*, the branch of archaeology that explicitly adopted the terminology of Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* (@dunnell1978style).
 Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
-Without resorting to the inappropriate simplifications of Evolutionism, their diffusion can be expected to increase over time.
+Without resorting to the inappropriate simplifications of Evolutionism, their diffusion can be expected to increase over time if we assume reasonable human actors.
 On the other hand, some behaviours are not related to survival and reproductive capacity, like e.g.
 ceramic decorations, traditional costume fashion or jewellery design.
-Despite (or because?) their seemingly secondary importance these characteristics often pass through a breathtakingly fast development and serve archaeologists as a proxy for relative chronology.
+Despite or because their seemingly subordinate importance, these characteristics often pass through a breathtakingly fast development.
+They often serve archaeologists as a proxy for precise relative chronology.
 Dunnell postulates that the distribution and emergence of these stylistic cultural traits behave randomly on a population scale level.
 The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @porcicExploringEffectsAssemblage2015; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001].
+@neimanStylisticVariationEvolutionary1995 takes a closer look at this processes and develops a simple framework for the simulation of non-functional idea propagation.
+He identifies the essential phenomena *drift* (random dominance of similar traits) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.
+Furthermore, he implements a measure of cultural distance to apply his theoretical insights to real-world data.
 
-@neimanStylisticVariationEvolutionary1995 takes a closer look at this processes and develops a simple framework for the simulation of non-functional idea propagation.
-Neiman identifies the essential phenomena *drift* (random dominance of similar traits) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.
-Furthermore, he implements a measure of cultural distance to apply his theoretical insights to real-world data.
 The workflow suggested by Neiman can be partially applied to the burial rites data examined in this article and was an important methodological inspiration: As burial traditions are ritualised behaviour and most of the time heavily linked to religious and spiritual beliefs they can be assumed to be stylistic traits as defined by @dunnell1978style.
-Besides @neimanStylisticVariationEvolutionary1995, another vital template for the comparison of cultural and spatial was provided by @porcicSimulatingCulturalTransmission2014.
+Besides @neimanStylisticVariationEvolutionary1995, another vital template for the comparison of cultural and spatial distance was provided by @porcicSimulatingCulturalTransmission2014.

### 2019-01-11 article/01_Introduction.Rmd
more love for the introduction to make it more well-rounded

@@
 Instead of directly querying archives with quantitative information, the typical process for the creation of large-scale syntheses is still to manually excerpt the literature of all relevant, local research traditions.
-The present article, in contrast, explores the former strategy: Does openly available, bulk radiocarbon data allow reconstructing the main trends within the distribution of burial rites in Bronze Age Europe?
+The present article, in contrast, explores the former strategy.
+The first major question therefore is: Does openly available, bulk radiocarbon data allow reconstructing the main trends within the distribution of burial rites in Bronze Age Europe?
 This approach could set a precedent for a more reproducible and statistically sound research practice.
-If so, then the quantitative comparison of macroregional trends could also pave the way for an analysis of interaction intensity on a long-term continental scale: Do time series constructed from radiocarbon data allow meaningful observations concerning cultural and spatial distance?
+If so, then the quantitative comparison of macroregional trends could also pave the way for an analysis of interaction intensity on a long-term continental scale: Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?
 
-To answer these questions, a theoretical framework is required that allows to compare human behaviour across spatial, temporal and cultural boundaries: The assumption that the cremation aspect of Early Bronze Age burial rituals in Britain can be directly compared to the same aspect in Hungary is not self-evident.
+To answer these questions, a theoretical framework is required that permits the comparison of human behaviour across spatial, temporal and cultural boundaries: The assumption that the cremation aspect of Early Bronze Age burial rituals in Britain could be related to the same aspect in Hungary hundreds of years later is not self-evident.
 *Cultural evolution theory* provides such a framework [@creanzaCulturalEvolutionaryTheory2017].
@@
 
-*Cultural evolution theory* is at least as old as theories of biological evolution and underwent a long and versatile research history.
-Its beginnings are in *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, matured in conflict with *Cultural relativism* into *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004].
+Cultural evolution theory is at least as old as theories of biological evolution and underwent a long and versatile research history.
+Its beginnings are in *Evolutionism* and *Social darwinism* in the 19th and early 20th century, matured in conflict with *Cultural relativism* into *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004].
 In the late seventies three major theoretical movements emerged: *Evolutionary psychology*, *Human behavioural ecology* and *Dual inheritance theory* [@smithThreeStylesEvolutionary2000].
-The latter -- heavily popularised by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007]: Just like in biological evolution also *Social learning* might be prone to selection and mutation.
+The latter -- heavily popularised by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007]: Just like in biological evolution, also *Social learning* might be prone to selection and mutation.
 Under this assumption methods of *Population genetics* could be employed within cultural anthropology to analyse the spread and development of ideas -- of course with the much faster pace of cultural exchange in mind.
@@
 The similarity of historically and spatially close material remains can, therefore, be explained as a direct result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
-Questions of *Cultural Transmission* form the core of archaeology.
+Questions of Cultural Transmission form the core of archaeology.
 
-*Evolutionary archaeology*, the branch of archaeology that explicitly adopted the terminology of Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* (@dunnell1978style).
+*Evolutionary archaeology*, the branch of archaeology that explicitly adopted the terminology of Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style].
 Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
@@
 Dunnell postulates that the distribution and emergence of these stylistic cultural traits behave randomly on a population scale level.
-The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @porcicExploringEffectsAssemblage2015; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001].
+The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @porcicSimulatingCulturalTransmission2014; @porcicExploringEffectsAssemblage2015; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001].
+
 @neimanStylisticVariationEvolutionary1995 takes a closer look at this processes and develops a simple framework for the simulation of non-functional idea propagation.
@@
 
-The workflow suggested by Neiman can be partially applied to the burial rites data examined in this article and was an important methodological inspiration: As burial traditions are ritualised behaviour and most of the time heavily linked to religious and spiritual beliefs they can be assumed to be stylistic traits as defined by @dunnell1978style.
-Besides @neimanStylisticVariationEvolutionary1995, another vital template for the comparison of cultural and spatial distance was provided by @porcicSimulatingCulturalTransmission2014.
+The workflow suggested by Neiman is relevant to the burial rites data examined in this paper and was an important methodological inspiration: As burial traditions are ritualised behaviour and most of the time heavily linked to religious and spiritual beliefs they can be assumed to be stylistic traits.
+Of course in reality, beyond the distinction of @dunnell1978style, most behaviour is not unambiguously stylistic or functional.
+Burial rites cause costs regarding working time and material to a varying extent.
+This cost might have been influenced or had their effect on the human-nature relation: Pyres, for example, require a significant amount of wood and the erection of barrows much workforce.
+Still, it can be expected that these functional disadvantages were not critical most of the time -- at least regarding the general decision which kind of burial tradition is applied.
+Most probably they influence the magnitude of invested energy for a burial, just like the level of grief [@macdonaldGriefBurialAmerican2001] or the representational intentions of the burying relatives [@henrichEvolutionCostlyDisplays2009].
+
+In the following sections I will (1) explain how radiocarbon data with relevant meta information can be transformed to macro regional time series of burial traditions, (2) explore how these timeseries compare to traditional archaeological narratives and (3) construct and contextualise cultural distances between macro regions over time in relation to spatial distance.
+For the latter a simulation based on the @neimanStylisticVariationEvolutionary1995 template was employed.

### 2019-01-11 article/04_Discussion.Rmd
more love for the introduction to make it more well-rounded

@@
 - Surprising result of no correlation: Meaning? Construction of an own story?
-
-In reality, beyond Dunnels distinction, most behaviour is not unambiguously stylistic or functional.
-For the sake of this case study burial rites are assumed to be stylistic, but in fact, their creation causes costs regarding working time and material to a varying extent.
-This cost might have been influenced or had their effect on the human-nature relation: Pyres, for example, require a significant amount of wood and the erection of barrows much workforce.
-Still, it can be expected that these functional disadvantages were not critical most of the time for the general decision which kind of burial tradition is applied, but maybe for the magnitude of invested energy.
-Other factors to consider here are the level of grief [@macdonaldGriefBurialAmerican2001] or the representational intentions of the burying relatives [@henrichEvolutionCostlyDisplays2009].

### 2019-01-11 article/02_Materials_and_Methods.Rmd
small changes in the materials and methods section

@@
 The analysis code of this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible.
-A more in-depth description and discussion of the data preparation can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
+A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
 
@@
 Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the type and cultural attribution of the sampled context.
-The variable *Feature type* systematically distinguishes for example between settlements, rock shelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, i.a..
+The variable *feature type* systematically distinguishes between settlements, rock shelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, i.a..
 This information allows to accurately and precisely determine a moment in space and time when a grave of a particular type was created.
@@
 
-3. Calibrating the ^14^C dates with the R package Bchron [@haslettSimpleMonotoneProcess2008], which implements the calibration algorithm with numerical integration.
+3. Calibrating the ^14^C dates with the R package Bchron [@haslettSimpleMonotoneProcess2008], which implements a calibration algorithm based on numerical integration.
 
@@
 
-5. Reducing the date selection to all dates taken from graves as indicated by the *Feature type* variable. (`r r("size bronze05", sf_prep)` dates)
+5. Reducing the date selection to all dates taken from graves as indicated by the *feature type* variable. (`r r("size bronze05", sf_prep)` dates)
 
-6. Transforming the variable *Feature type* via pattern matching to two new variables *burial type* (inhumation/cremation/unknown) and *burial construction* (mound/flat/unknown).
+6. Transforming the variable *feature type* via pattern matching to two new variables *burial type* (inhumation/cremation/unknown) and *burial construction* (mound/flat/unknown).
 
 7.
-Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial date distribution (Figure \ref{map_graves}).
+Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial data distribution (Figure \ref{map_graves}).
 
@@
 Defining macro-regions to observe the development of *burial type* and *burial construction* within sensible geographical and cultural units that are appropriate to the amount of available data.
-The regions are circular and were delimited semi-automatically by creating a regular grid in accordance to main cultural-historical regions of the European Bronze Age (**citation**).
+The regions are circular and were delimited semi-automatically by creating a regular grid in accordance to main cultural-historical regions of the European Bronze Age [@fokkensIntroductionBronzeAge2013].
 The regions have a radius of 240km and overlap slightly (Figure \ref{map_regions} A).
@@
     escape = F,
-    linesep = "\\addlinespace \\hline \\addlinespace"
+    linesep = "\\addlinespace \\hline \\addlinespace",
+    align = "l"
   ) %>%
@@
     latex_options = "hold_position"
-  )
+  ) %>%
+  kableExtra::add_header_above(c("burial type" = 1, "burial construction" = 3))
 ```

### 2019-01-11 article/02_Materials_and_Methods.Rmd
more improvements of the materials and methods chapter

@@
 
-This data for individual graves attributed to a macro region can be transformed to a time series of trait prevalence within the respective regions.
+This data for individual graves can be transformed to a time series of trait prevalence within the respective regions.
 If a certain kind of grave can be attributed to a fuzzy point in time with radiocarbon dating, then we also know that most likely a ceremony was held around this time where the particular burial tradition was practised.
 That means that the cultural trait behind this behaviour must also have been present in one form or the other.
-An example: The idea *inhumation* must have been present in 1447 calBC in Great Britain since this year lies in the $2\sigma$ age range of the ^14^C date *NZA-32497* which was made for the Amesbury Down inhumation burial *I2639/25217*.
+An example: The idea *inhumation* must have been present in 1447 calBC in Great Britain since this year lies in the $2\sigma$ age range of the ^14^C date *NZA-32497* taken from the Amesbury Down inhumation burial *I2639/25217*.
 For each year in the time frame 2200 -- 800 calBC we can count the number of graves with dates that cover this year with their $2\sigma$ age range.
@@
 
-The continuous time series of year wise burial rite proportions for each region allow to calculate time series of a measure of cultural distance for each region-region relation (Figure \ref{sed_region_matrix}).
-Several different measures of cultural distance could be used here following the discussion of these matters in the relevant literature (**citation**).
-In this context the Squared Euclidian Distance (SED) seems to be a sensible statistic, because it is relatively simple, works with only a few assumptions and emphasises differences through squaring.
+These time series also allow to derive another set of time series with a measure of cultural distance for each region-region relationship (Figure \ref{sed_region_matrix}).
+Cultural distance is defined here as the Squared Euclidian Distance (SED) between the trait proportions of two different regions at a certain point in time [@neimanStylisticVariationEvolutionary1995].
+SED seems to be a sensible statistic here, because it is relatively simple, works with only a few assumptions and emphasises differences through squaring.
 
@@
 
-The SED is defined here as:
-
 $$d_{ij}^2 = \sum_{k = 1}^{n} (p_{ik} - p_{jk})^2$$
 
-- $d_{ij}^2$: SED between tqo groups $i$ and $j$
+- $d_{ij}^2$: SED between two groups $i$ and $j$
 
@@
 The two matrices of each section can be tested for correlation with the Mantel test, which uses permutations of allegedly correlating matrices to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967].
-Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test.
+Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test (Table \ref{tab:distanceandcorrelationtable*} C).
 
@@
 For the eight regions, four different categories of spatial distance were defined (Figure \ref{map_regions} B).
-The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the difference in the data level of measurements.
-The test results are listed in Table \ref{tab:distanceandcorrelationtable*}.
+The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the difference in the data level of measurements (Table \ref{tab:distanceandcorrelationtable*} A&B).
 
@@
   kableExtra::add_header_above(
-    c(" ", "burial type" = 3, "burial construction" = 3, "variable correlation" = 2)
+    c(" ", "burial type & spatial distance" = 3, "burial construction & spatial distance" = 3, "variable correlation" = 2)
+  ) %>%
+  kableExtra::add_header_above(
+    c(" ", "A" = 3, "B" = 3, "C" = 2),
+    bold = T, line = F
   ) %>%

### 2019-01-11 analysis/code/05_plot_scripts/development_burial_construction.R
improved time series plots

@@ -40,17 +40,16 @@ development_burial_construction_A <- ggplot() +
     alpha = 0.8
   ) +
   facet_wrap(~region_name, nrow = 8) +
-  xlab("") +
-  ylab("Amount of 14C dates from burials") +
-  labs(fill = "Ideas (mutually exclusive)") +
+  xlab("Time in years calBC") +
+  ylab("Amount of graves") +
+  labs(fill = "Burial customs") +
   theme_bw() +
   theme(
     legend.position = "bottom",
     panel.grid.major.x = element_line(colour = "black", size = 0.3),
-    axis.text.y = element_text(size = 25),
-    axis.title.y = element_text(size = 25),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank(),
+    axis.text = element_text(size = 25),
+    axis.title = element_text(size = 25),
+    axis.text.x = element_text(angle = 45, hjust = 1),
     strip.text.x = element_text(size = 25),
     legend.title = element_text(size = 30, face = "bold"),
     legend.text = element_text(size = 30),
@@ -72,9 +71,10 @@ development_burial_construction_A <- ggplot() +
     breaks = seq(-2200, -800, 200),
     limits = c(-2500, -800)
   ) +
-  guides(
-    fill = FALSE
-  )
+  # guides(
+  #   fill = FALSE
+  # ) +
+  NULL
 
 development_burial_construction_A <- development_burial_construction_A +
   geom_custom(
@@ -96,13 +96,14 @@ development_burial_construction_B <- ggplot() +
   scale_alpha_continuous(range = c(0.0, 0.7)) +
   facet_wrap(~region_name, nrow = 8) +
   xlab("Time in years calBC") +
-  ylab("Proportion of 14C dates from burials") +
-  labs(fill = "Ideas (mutually exclusive)") +
+  ylab("Proportion of burial customs") +
+  labs(fill = "Burial customs") +
   theme_bw() +
   theme(
     legend.position = "bottom",
     panel.grid.major.x = element_line(colour = "black", size = 0.3),
     axis.text = element_text(size = 25),
+    axis.text.x = element_text(angle = 45, hjust = 1),
     axis.title = element_text(size = 25),
     strip.text.x = element_text(size = 25),
     legend.title = element_text(size = 30, face = "bold"),
@@ -142,7 +143,9 @@ combined_plot <- plot_grid(
[…]

### 2019-01-11 analysis/code/05_plot_scripts/development_burial_type.R

@@ -40,17 +40,16 @@ development_burial_type_A <- ggplot() +
     alpha = 0.8
   ) +
   facet_wrap(~region_name, nrow = 8) +
-  xlab("") +
-  ylab("Amount of 14C dates from burials") +
-  labs(fill = "Ideas (mutually exclusive)") +
+  xlab("Time in years calBC") +
+  ylab("Amount of graves") +
+  labs(fill = "Burial customs") +
   theme_bw() +
   theme(
     legend.position = "bottom",
     panel.grid.major.x = element_line(colour = "black", size = 0.3),
-    axis.text.y = element_text(size = 25),
-    axis.title.y = element_text(size = 25),
-    axis.text.x = element_blank(),
-    axis.ticks.x = element_blank(),
+    axis.text = element_text(size = 25),
+    axis.title = element_text(size = 25),
+    axis.text.x = element_text(angle = 45, hjust = 1),
     strip.text.x = element_text(size = 25),
     legend.title = element_text(size = 30, face = "bold"),
     legend.text = element_text(size = 30),
@@ -72,9 +71,10 @@ development_burial_type_A <- ggplot() +
     breaks = seq(-2200, -800, 200),
     limits = c(-2500, -800)
   ) +
-  guides(
-    fill = FALSE
-  )
+  # guides(
+  #   fill = FALSE
+  # ) +
+  NULL
 
 development_burial_type_A <- development_burial_type_A +
   geom_custom(
@@ -97,13 +97,14 @@ development_burial_type_B <- ggplot() +
   scale_alpha_continuous(range = c(0.0, 0.7)) +
   facet_wrap(~region_name, nrow = 8) +
   xlab("Time in years calBC") +
-  ylab("Proportion of 14C dates from burials") +
-  labs(fill = "Ideas (mutually exclusive)") +
+  ylab("Proportion of burial customs") +
+  labs(fill = "Burial customs") +
   theme_bw() +
   theme(
     legend.position = "bottom",
     panel.grid.major.x = element_line(colour = "black", size = 0.3),
     axis.text = element_text(size = 25),
+    axis.text.x = element_text(angle = 45, hjust = 1),
     axis.title = element_text(size = 25),
     strip.text.x = element_text(size = 25),
     legend.title = element_text(size = 30, face = "bold"),
@@ -143,7 +144,9 @@ combined_plot <- plot_grid(
   labels = c("A", "B"),
[…]

### 2019-01-11 article/02_Materials_and_Methods.Rmd
adjusted figure caption row count

@@
 \includegraphics[width=\columnwidth]{../analysis/figures/development_burial_construction.jpeg}
-\caption{\label{development_burial_construction}Like Figure \ref{development_burial_type}, but for the variable \textit{burial construction}.}
+\caption{\label{development_burial_construction}Like Figure \ref{development_burial_type}, but for the variable \textit{burial construction}.\newline\newline}
 \end{figure}

### 2019-01-12 article/02_Materials_and_Methods.Rmd
started to work on simulation parametrisation

@@
 
-$$d_{ij}^2 = \sum_{k = 1}^{n} (p_{ik} - p_{jk})^2$$
+$$
+d_{ij}^2 = \sum_{k = 1}^{n} (p_{ik} - p_{jk})^2
+$$
 
@@
 For the eight regions, four different categories of spatial distance were defined (Figure \ref{map_regions} B).
-The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the difference in the data level of measurements (Table \ref{tab:distanceandcorrelationtable*} A&B).
+The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the difference in the data level of measurements (Table \ref{tab:distanceandcorrelationtable*} AB).
 
@@
 To get a better understanding how the correlation coefficient should develop if there was a strong correlation between spatial and cultural distance as measured by *burial type* and *burial construction*, a simple simulation was implemented based on the model proposed by @neimanStylisticVariationEvolutionary1995.
-This simulation assumes a population separated into $g$ groups with a group population of $N_g$ each.
-At the beginning of the simulation, there are $k$ ideas randomly but equally distributed in this population.
-With every timestep -- $t$ defines the total number of time steps -- all individuals learn from another randomly chosen individual.
-The probability that this other individual is from another group equals $m_i$, for the opposite therefore $1-m_i$.
-The partner group for each learning instance is selected randomly following a group interaction probability defined in the matrix $I$.
+This simulation assumes a population separated into $g = 8$ groups with a group population of $N_g = 100$ individuals each.
+$k = 2$ mutually exclusive ideas are randomly but equally distributed in this population.
+With every timestep all individuals learn from another randomly chosen individual and the total number of time steps is $t = 1400$ -- one for each year in the observed Bronze Age context.
+The probability that the other agent the individual is learning from comes from another group equals
+
+$$
+m_i = 
+  \begin{dcases*} 
+    0 & Figure \ref{simulation} AB, no intergroup exchange \\ 
+    0.01 & Figure \ref{simulation} CD \\
+    0.1 & Figure \ref{simulation} EF \\
+    0.5 & Figure \ref{simulation} GH, very high exchange
+  \end{dcases*}
+$$
+
+The partner group for each learning instance is selected randomly following a group interaction probability defined in the matrix
+
+$$
+I =
+  \begin{dcases*}
+    \begin{matrix}
+      0 & 1 & 1 & ... \\
+      1 & 0 & 1 & ...\\
+      1 & 1 & 0 & ... \\
+      ... & ... & ... & ...
+    \end{matrix} \\
+    \text{Figure \ref{simulation} ACEG, equal distance between groups} \\
+    \begin{matrix}
+      0 & |b-a| & |c-a| & ... \\
+      |a-b| & 0 & |c-b| & ... \\
+      |a-c| & |b-c| & 0 & ... \\
+      ... & ... & ... & ...
+    \end{matrix} \\
+    \text{Figure \ref{simulation} BDFH, spatial distance}
+  \end{dcases*}
+$$.
+
 The distribution of the $k$ variants can be observed for every timestep and documented in a time series like the one for real-world data.

### 2019-01-12 article/02_Materials_and_Methods.Rmd
minimal description of the steps that lead to fig 6

@@
 $k = 2$ mutually exclusive ideas are randomly but equally distributed in this population.
-With every timestep all individuals learn from another randomly chosen individual and the total number of time steps is $t = 1400$ -- one for each year in the observed Bronze Age context.
-The probability that the other agent the individual is learning from comes from another group equals
+With every timestep $t$ (with $t_{final} = 1400$) all individuals learn from another randomly chosen individual.
+Several different values where tested for the probability $m_i$ that the other agent the individual is learning from comes not from the same, but from another group:
 
@@
 
-The partner group for each learning instance is selected randomly following a group interaction probability defined in the matrix
+The partner group for each learning instance is selected randomly following a group interaction probability matrix $I$. Two different matrizes where tested:
 
@@
   \end{dcases*}
-$$.
+$$
 
-The distribution of the $k$ variants can be observed for every timestep and documented in a time series like the one for real-world data.
-
-For the actual simulation runs, the parameters were chosen as follows: **TODO** Figure \ref{simulation*}
+The distribution of the $k$ variants within the groups can be observed for every timestep and documented in time series just like the ones for the real-world regions data (Figure \ref{development_burial_type} B and \ref{development_burial_construction} B).
+These time series can as well be used to calculate time series of region-region distance, which again can be split up into time blocks of 200 years to calculate mean distance matrices.
+The latter can be checked for correlation with the spatial distance matrix (cf.
+Table \ref{tab:distanceandcorrelationtable*} AB) and compared with other simulation results and real-world observations (Figure \ref{simulation}).
 

### 2019-01-13 article/02_Materials_and_Methods.Rmd
improvements on the description of the simulation

@@
 To get a better understanding how the correlation coefficient should develop if there was a strong correlation between spatial and cultural distance as measured by *burial type* and *burial construction*, a simple simulation was implemented based on the model proposed by @neimanStylisticVariationEvolutionary1995.
-This simulation assumes a population separated into $g = 8$ groups with a group population of $N_g = 100$ individuals each.
-$k = 2$ mutually exclusive ideas are randomly but equally distributed in this population.
-With every timestep $t$ (with $t_{final} = 1400$) all individuals learn from another randomly chosen individual.
+This simulation assumes a population separated into $g = 8$ groups (regions) with $N_g = 100$ individuals each.
+$k = 2$ mutually exclusive ideas are randomly but equally distributed in this population at $t_{0}$.
+With every timestep ($t_{total} = (2200 - 800)/20 = 70$) all individuals learn from another randomly chosen individual.
 Several different values where tested for the probability $m_i$ that the other agent the individual is learning from comes not from the same, but from another group:
@@
     \begin{matrix}
-      0 & 1 & 1 & ... \\
-      1 & 0 & 1 & ...\\
-      1 & 1 & 0 & ... \\
-      ... & ... & ... & ...
+        & a & b & c & ... \\
+      a & 0 & 1 & 1 & ... \\
+      b & 1 & 0 & 1 & ...\\
+      c & 1 & 1 & 0 & ... \\
+      ... & ... & ... & ... & ...
     \end{matrix} \\
-    \text{Figure \ref{simulation} ACEG, equal distance between groups} \\
+    \text{Figure \ref{simulation} ACEG, equal distance between groups a:h} \\
     \begin{matrix}
-      0 & |b-a| & |c-a| & ... \\
-      |a-b| & 0 & |c-b| & ... \\
-      |a-c| & |b-c| & 0 & ... \\
-      ... & ... & ... & ...
+        & a & b       & c       & ... \\
+      a & 0 & |b-a|^2 & |c-a|^2 & ... \\
+      b & |a-b|^2 & 0 & |c-b|^2 & ... \\
+      c & |a-c|^2 & |b-c|^2 & 0 & ... \\
+      ... & ... & ... & ... & ...
     \end{matrix} \\
-    \text{Figure \ref{simulation} BDFH, spatial distance}
+    \text{Figure \ref{simulation} BDFH, spatial distance (Figure \ref{map_regions} B)}
   \end{dcases*}
@@
 \includegraphics[width=\textwidth]{../analysis/figures/simulation.jpeg}
-\caption{\label{simulation}Simulation results in comparison with observed data.}
+\caption{\label{simulation}Development of correlation between cultural and spatial distance for real world data and simulation results.
+The plots in the left column of the plot matrix (\textbf{ACE}) contain simulation results for an equal group distance matrix $I$, the ones on the right (\textbf{BDF}) for a distance matrix based on the macro unit distances defined for this paper (Figure \ref{map_regions} B).
+The rows of the plot matrix (\textbf{AB}, \textbf{CD}, \textbf{EF}) represent different degrees of intergroup transmission $m_i$.
+Each plot shows the results of 50 simulation runs.
+The x-axis displays the time distiguished into 200 year timeblocks, the y-axis Spearman's rank correlation coefficient between the mean squared euclidian distance matrix for a 200 year timeblock and the static spatial distance matrix.
+This information is visualized for each timeblock with (from left to right) a dotplot coloured according to the $p$-value of the Mantel test, a scatter plot line and a boxplot.
+The scatter plot points of one simulation run are connected with a faint line from one timeblock to the next.
+To compare the runs with the real world observations, the results for the burial customs data (Table \ref{tab:distanceandcorrelationtable*} AB) is added here with with a stronger green and blue line.
+The strong horizontal red line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}
 \end{figure*}

### 2019-01-13 analysis/code/04_simulation_scripts/01_run_simulation.R

@@ -12,7 +12,7 @@ models_grid <- expand.grid(
   N_g = 100,
   t_start = -2200,
   t_end = -800,
-  t_steps = 10,
+  t_steps = 20,
   mu = 0,
   g = 8,
   I = list(NA, interaction_matrix_spatial),

### 2019-01-13 analysis/code/05_plot_scripts/sed_simulation.R

@@ -190,7 +190,7 @@ total %>%
     device = "jpeg",
     scale = 1,
     dpi = 300,
-    width = 660, height = 930, units = "mm",
+    width = 660, height = 840, units = "mm",
     limitsize = F
   )
 

### 2019-01-13 article/02_Materials_and_Methods.Rmd
better caption for table 2

@@
   "latex", 
-  caption = "Mean distance, correlation and mantel test results for cultural and spatial distance.", 
+  caption = "Mean distance, correlation and mantel test results for cultural and spatial distance.
+Listed are the the overall mean Squared Euclidian Distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bt)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B).
+(\\textbf{C}) shows the Mantel test results for the two variables if compared directly with each other.",
   booktabs = T,
@@
 Each plot shows the results of 50 simulation runs.
-The x-axis displays the time distiguished into 200 year timeblocks, the y-axis Spearman's rank correlation coefficient between the mean squared euclidian distance matrix for a 200 year timeblock and the static spatial distance matrix.
-This information is visualized for each timeblock with (from left to right) a dotplot coloured according to the $p$-value of the Mantel test, a scatter plot line and a boxplot.
-The scatter plot points of one simulation run are connected with a faint line from one timeblock to the next.
+The x-axis displays the time distiguished into 200 year time slots, the y-axis Spearman's rank correlation coefficient between the mean squared euclidian distance matrix for a 200 year time slot and the static spatial distance matrix.
+This information is visualized for each time slot with (from left to right) a dotplot coloured according to the $p$-value of the Mantel test, a scatter plot line and a boxplot.
+The scatter plot points of one simulation run are connected with a faint line from one time slot to the next.
 To compare the runs with the real world observations, the results for the burial customs data (Table \ref{tab:distanceandcorrelationtable*} AB) is added here with with a stronger green and blue line.

### 2019-01-13 analysis/code/05_plot_scripts/sed_region_matrix.R
plot now with axis labels

@@ -83,8 +83,6 @@ for (row_region in region_order) {
           limits = c(0, 2)
         ) +
         theme_bw() +
-        ylab("Squared Euclidian Distance") +
-        xlab("Time in years calBC") +
         scale_colour_manual(
           name = "Burial customs",
           values = c(
@@ -106,7 +104,9 @@ sed_matrix <- ggmatrix(
   yAxisLabels = region_order,
   byrow = FALSE,
   switch = "y",
-  labeller = label_wrap_gen()
+  labeller = label_wrap_gen(),
+  ylab ="Squared Euclidian Distance",
+  xlab = "Time in years calBC"
 ) + theme_bw() +
   theme(
     axis.text = element_text(size = 15),

### 2019-01-13 article/02_Materials_and_Methods.Rmd
better figure captions

@@
 These time series also allow to derive another set of time series with a measure of cultural distance for each region-region relationship (Figure \ref{sed_region_matrix}).
-Cultural distance is defined here as the Squared Euclidian Distance (SED) between the trait proportions of two different regions at a certain point in time [@neimanStylisticVariationEvolutionary1995].
+Cultural distance is defined here as the squared euclidian distance (SED) between the trait proportions of two different regions at a certain point in time [@neimanStylisticVariationEvolutionary1995].
 SED seems to be a sensible statistic here, because it is relatively simple, works with only a few assumptions and emphasises differences through squaring.
@@
 \caption{\label{sed_region_matrix}Plot matrix of the region-region squared euclidian distance time series.
-In the top right half for the variable \textit{burial type} and in the bottom left for the variable \textit{burial construction}.}
+In the top right half for the variable \textit{burial type} and in the bottom left for the variable \textit{burial construction}.
+On the x-axis the time is displayed, on the y-axis the squared euclidian distance between the two relevant regions.
+The faint, grey line shows the actual data, the stronger, coloured line an approximation (LOESS with $span = 0.3$).}
 \end{figure*}
@@
   caption = "Mean distance, correlation and mantel test results for cultural and spatial distance.
-Listed are the the overall mean Squared Euclidian Distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bt)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B).
+Listed are the the overall mean squared euclidian distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bt)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B).
 (\\textbf{C}) shows the Mantel test results for the two variables if compared directly with each other.",

### 2019-01-13 article/02_Materials_and_Methods.Rmd
better figure captions

@@
 \includegraphics[width=\columnwidth]{../analysis/figures/map_graves.jpeg}
-\caption{\label{map_graves}All graves dating between 2200 and 800calBC as extracted from the Radon-B database. The red frame indicates the research area.}
+\caption{\label{map_graves}All graves dating between 2200 and 800calBC as extracted from the Radon-B database.
+The classes of the variable \textit{burial type} are distinguished by colour, the ones of \textit{burial construction} by shape.
+The red frame indicates the research area.
+The Map projection is EPSG:102013 and the base layer (land, rivers, lakes) data is taken from the Natural Earth project (\url{www.naturalearthdata.com}).}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{../analysis/figures/map_regions.jpeg}
-\caption{\label{map_regions}Circular, artificial regions (\textbf{A}) and spatial distance network with classified distance values (\textbf{B}).}
+\caption{\label{map_regions}Circular, artificial regions (\textbf{A}) and spatial distance network with classified distance values (\textbf{B}).
+The map layout is the same as in Figure \ref{map_graves}.}
 \end{figure}

### 2019-01-13 article/02_Materials_and_Methods.Rmd
and more better figure captions

@@
 \includegraphics[width=\columnwidth]{../analysis/figures/development_burial_type.jpeg}
-\caption{\label{development_burial_type}Regional distribution time series of the variable \textit{burial type} with the absolute amount (\textbf{A}) and proportion (\textbf{B}) of grave types.}
+\caption{\label{development_burial_type}Regional distribution time series of the variable \textit{burial type} with the absolute amount (\textbf{A}) and proportion (\textbf{B}) of grave types.
+The x-axis shows the time, the y-axis the amount or the percentage shares of graves.
+The individual plots are separated by region, and each region is featured by a small map according to Figure \ref{map_regions} A.}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{../analysis/figures/development_burial_construction.jpeg}
-\caption{\label{development_burial_construction}Like Figure \ref{development_burial_type}, but for the variable \textit{burial construction}.\newline\newline}
+\caption{\label{development_burial_construction}Like Figure \ref{development_burial_type}, but for the variable \textit{burial construction}.\newline\newline\newline\newline\newline}
 \end{figure}

### 2019-01-13 analysis/code/05_plot_scripts/development_burial_construction.R

@@ -155,7 +155,7 @@ combined_plot %>%
     device = "jpeg",
     scale = 1,
     dpi = 300,
-    width = 330, height = 930, units = "mm",
+    width = 330, height = 910, units = "mm",
     limitsize = F
   )
 

### 2019-01-13 analysis/code/05_plot_scripts/development_burial_type.R

@@ -156,7 +156,7 @@ combined_plot %>%
     device = "jpeg",
     scale = 1,
     dpi = 300,
-    width = 330, height = 930, units = "mm",
+    width = 330, height = 910, units = "mm",
     limitsize = F
   )
 

### 2019-01-13 article/03_Results.Rmd
added text for an own narrative

@@
 Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region-wise time series of trait occurrence.
-The relative proportion of each trait calculated based on ^14^C dates can be compared to the archaeological state of research:
+If we ignore all other archaeological knowledge, we can construct the following narrative from them:
 
-General observations: **TODO**
+In the course of the Bronze Age in Europe there was a trend away from body burial towards cremation.
+Around 2200calBC cremation burials were almost unknown in Poland, southern Germany, northeastern France and northern Germany.
+In the northwest, in Great Britain and the Benelux area, and in the southeast, in Austria and the Czech Republic, however, there were early cremation traditions.
+These contexts could be discussed as areas of origin of the later omnipresent phenomenon.
+While cremation burials lost importance in Austria and the Czech Republic as well as in the Benelux region during the Early Bronze Age and only regained their importance in the Middle Bronze Age, their share increased steadily in England.
+Inhumation and cremation were balanced here for a long time.
+The situation was similar in southern Scandinavia, where the proportion of cremation burials was much lower until the Late Bronze Age.
+In Poland and northern Germany, cremation became relatively suddenly and completely established with the beginning of the Late Bronze Age.
+In northeastern France and northern Germany this change took place already in the Middle Bronze Age.
+In both contexts, inhumation continued to play an important role.
+
+With regard to the question of grave mound construction, the picture is considerably more heterogeneous.
+In Austria and the Czech Republic, flat graves were the dominant form of burial until the Late Bronze Age, but mounds became more relevant from the Middle Bronze Age onwards.
+In Poland, flat and hill graves remained in balance until the Middle Bronze Age, when flat graves prevailed.
+The amount of data from southern Germany and northeastern France is extraordinarily small: If one believes the sample, then a sudden, radical change from burial in flat graves to hill graves took place in southern Germany at the beginning of the Middle Bronze Age.
+In northeastern France there were no flat graves in the Bronze Age.
+Northern Germany underwent a development from the dominance of flat graves in the Early Bronze Age, a short phase of increasing mound construction in the Middle Bronze Age followed by renewed dominance of the flat grave custom in the Late Bronze Age.
+A similar development can be seen in southern Scandinavia: Flat graves clearly predominated, but were increasingly -- almost completely -- replaced by hills, until burial in flat graves became more frequent again at the end of the Bronze Age.
+Burial mounds were dominant throughout the Benelux region, but in the Early and Late Bronze Ages flat graves also occurred to a significant extent.
+In England, flat graves were a short phenomenon in the Early Bronze Age, which did not recur later.
+
+This narrative is wrong in many aspects. The following section contains a short comparison with the archaeological state of research for each region:
 
 `r a("Southeastern Central Europe")`^[Amount of graves in Radon-B for this region.]:
-The artificial study region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, Western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map_regions} A).
+The artificial circular region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, Western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map_regions} A).
 The density of ^14^C data on graves in Radon-B is low -- dates are available from the various regions, but large areas are without evidence.
@@
 
-Beyond the simple comparison of archaeological knowledge and information derived from Radon-B, figure \ref{sed_region_matrix} shows the time series of squared euclidian distance for each region-region relationship.
+Beyond the simple comparison of archaeological knowledge and information derived from Radon-B, Figure \ref{sed_region_matrix} shows the time series of squared euclidian distance for each region-region relationship.
 The majority of the curves in its upper right half for *burial type* correspond to expectations: At the beginning of the Bronze Age, inhumation is widespread throughout the study area, so the Euclidean distance at 2200calBC is small in almost every region-region relationship.
@@
 It remains without accentuated minima and maxima.
-The situation for *burial construction* is more complicated, as can be seen from the plots in the bottom right corner of figure \ref{sed_region_matrix}.
+The situation for *burial construction* is more complicated, as can be seen from the plots in the bottom right corner of Figure \ref{sed_region_matrix}.
 The distance does not seem to follow a general trend, but a multitude of local and temporally limited phenomena.

### 2019-01-13 article/03_Results.Rmd
improvements in the language of the detailed comparison of archaeology and radon-b

@@
 `r a("Southeastern Central Europe")`^[Amount of graves in Radon-B for this region.]:
-The artificial circular region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, Western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map_regions} A).
+The artificial circular region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map_regions} A).
 The density of ^14^C data on graves in Radon-B is low -- dates are available from the various regions, but large areas are without evidence.
@@
 `r a("Southern Germany")`:
-The artificial district region *Southern Germany* covers Bavaria completely but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and Western Austria.
-No ^14^C data are available from the latter regions.
-The majority of the available data comes from the Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative district of Tübingen.
-For the Early Bronze Age, Radon-B preserves a remarkable amount of data from Southern Germany, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves.
+The artificial district region *Southern Germany* covers Bavaria completely but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and western Austria.
+No ^14^C data are available from the latter regions, the majority of the available data comes from Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative district of Tübingen.
+For the Early Bronze Age, Radon-B preserves a remarkable amount of data from southern Germany, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves.
 The *burial type* variable is quite well understood -- although disproportionately much more of the early inhumations were recorded than urnfield burials.
@@
 The amount of data is low overall, but very low in the early and early Middle Bronze Age.
-For *burial type*, after all, limited statements are possible after this time window, *burial construction*, on the other hand, must remain obscure.
+For *burial type* limited statements are possible after this time window, *burial construction*, on the other hand, must remain obscure.
 The prediction of the relative development of these variables is correspondingly error-prone.
@@
 However, this situation already existed in the Early Bronze Age and indeed developed more nuanced than the proxy might suggest.
-Contrary to the derivation from the ^14^C ratios, flat graves were quite common in Northeastern France [@mordantBronzeAgeFrance2013].
+Contrary to the derivation from the ^14^C ratios, flat graves were quite common in northeastern France [@mordantBronzeAgeFrance2013].
 
@@
 The meta information on *burial type* and *burial construction* trace this development to the same extent.
-Since in Northern Germany influences from all four cardinal points come together and contribute significantly to the formation of local cultural phenomena, it is difficult to estimate the overall development of the primary variables here as well.
+Since in Northern Germany influences from all four cardinal directions come together and contribute significantly to the formation of local cultural phenomena, it is difficult to estimate the overall development of the primary variables here as well.
 The ^14^C proxy correctly represents *burial site* insofar as inhumations dominated in the Early Bronze Age and were later on replaced by cremations especially in the North from the Middle Bronze Age onwards.
@@
 `r a("Southern Scandinavia")`:
-The region *Southern Scandinavia* includes all of Denmark with Jutland, Funen, Zeeland and all smaller islands as well as large parts of the east of southern Sweden, and coastal regions in Schleswig-Holstein and Mecklenburg-Western Pomerania.
+The region *Southern Scandinavia* includes all of Denmark with Jutland, Funen, Zeeland and all smaller islands as well as large parts of east of southern Sweden, and coastal regions in Schleswig-Holstein and Mecklenburg-Western Pomerania.
 The ^14^C data are very well distributed across Jutland, Funen and the Swedish Skåne region.
@@
 The Late Neolithic from 2200calBC is a desideratum.
-If one carries out the necessary mental shift -- the Central European Middle Bronze Age corresponds to the Nordic Early Bronze Age -- and then compares the proxy of the relative grave form development with the archaeological observations, then one finds a substantial agreement: body graves were the rule in periods I and II (about 1800-1300calBC), but cremations also occurred sporadically.
+If one carries out the necessary mental shift -- the Central European Middle Bronze Age corresponds to the Nordic Early Bronze Age -- and compares the proxy of the relative grave form development with the archaeological observations, then one finds a substantial agreement: inhumations were the rule in periods I and II (about 1800-1300calBC), but cremations also occurred sporadically.
 From period III onwards cremation burials gained the upper hand.
-In the Scandinavian Early Bronze Age burials in burial mounds were the absolute rule.
-In the Late Bronze Age, flat burial fields were added [@thraneScandinavia2013].
+In the Scandinavian Early Bronze Age burials in burial mounds dominated.
+In the Late Bronze Age, flat burial fields began to occure [@thraneScandinavia2013].
 
@@
 The relative development derived from the ^14^C data indicates the correct trend for *burial type*: Cremation graves became more and more important during the Bronze Age, while inhumations appeared for a long time in parallel.
-However, the dataset emphasises the dominance of the cremation graves too much.
+However, the dataset emphasises the occurence of cremation graves too much.
 Concerning the variables *burial construction*, the dominance of burial mounds corresponds perfectly to the archaeological narrative, but -- especially in the Late Bronze Age -- flat grave fields also occurred, which the data set wrongly excludes [@robertsBritainIrelandBronze2013].

### 2019-01-13 article/02_Materials_and_Methods.Rmd
added subsections

@@
 \includegraphics[width=\columnwidth]{../analysis/figures/map_graves.jpeg}
-\caption{\label{map_graves}All graves dating between 2200 and 800calBC as extracted from the Radon-B database.
+\caption{\label{map_graves}All graves dating between 2200 and 800calBC as extracted from the Radon-B database (\url{radon-b.ufg.uni-kiel.de}).
 The classes of the variable \textit{burial type} are distinguished by colour, the ones of \textit{burial construction} by shape.
@@
 
-The analysis code of this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible.
+The code of this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible.
 A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
+
+## Data
 
@@
 These issues though do not inhibit the general data analysis strategy of this paper.
+
+## Data preparation
 
@@
 
+## Time series
+
 This data for individual graves can be transformed to a time series of trait prevalence within the respective regions.
@@
 \end{figure}
+
+## Cultural distance
 
@@
 Listed are the the overall mean squared euclidian distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bt)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B).
-(\\textbf{C}) shows the Mantel test results for the two variables if compared directly with each other.",
+(\\textbf{C}) shows the Mantel test results for the two variables if compared directly with each other.
+Negative correlation coefficients are printed red and $p$-values below $0.1$ bold.",
   booktabs = T,

### 2019-01-13 article/03_Results.Rmd
added subsections

@@
 
+## Timeseries
+
 Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region-wise time series of trait occurrence.
-If we ignore all other archaeological knowledge, we can construct the following narrative from them:
+If we ignore all other archaeological knowledge, we can construct the following narrative from them.
 
@@
 
-This narrative is wrong in many aspects. The following section contains a short comparison with the archaeological state of research for each region:
+This narrative is wrong in many aspects. The following section contains a short comparison with the archaeological state of research for each region.
 
@@
 
+## Cultural distance
+
 Beyond the simple comparison of archaeological knowledge and information derived from Radon-B, Figure \ref{sed_region_matrix} shows the time series of squared euclidian distance for each region-region relationship.

### 2019-01-13 article/03_Results.Rmd
moved detailed discussion of the timeseries to the discussion section

@@
 Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region-wise time series of trait occurrence.
-If we ignore all other archaeological knowledge, we can construct the following narrative from them.
+If we ignore all other archaeological knowledge, we can construct the following narrative from them: \newline
 
@@
 In England, flat graves were a short phenomenon in the Early Bronze Age, which did not recur later.
+\newline
 
-This narrative is wrong in many aspects. The following section contains a short comparison with the archaeological state of research for each region.
-
-`r a("Southeastern Central Europe")`^[Amount of graves in Radon-B for this region.]:
-The artificial circular region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map_regions} A).
-The density of ^14^C data on graves in Radon-B is low -- dates are available from the various regions, but large areas are without evidence.
-Diachronically the amount of data is consistently low and reaches a minimum in the Middle Bronze Age.
-While there is a constant flow of information for the variable *burial type*, the number of graves without information on *burial construction* increases in the Late Bronze Age to such an extent that no meaningful overview can be derived.
-The region intersects diverse geographical and cultural areas complicating the comparison with archaeological observations.
-In the southeast of Central Europe, cremation and inhumation already existed side by side in the Early Bronze Age, and the local tradition remained stable in this respect also due to the expansion of the burial mound tradition in the Middle Bronze Age -- with a slight overall increase in cremation graves.
-The Urnfield period brought the transition to the absolute dominance of cremation burial.
-The development of the *burial type* variable in the Radon-B proxy shows this development correctly.
-The information from ^14^C data on *burial construction* correctly traces the relevance of flat burials in the Early Bronze Age but then overemphasises them in the Middle Bronze Age compared to the construction of burial mounds.
-In the Late Bronze Age, burials on flat burial grounds were frequent, although burial mounds of the Tumulus Culture were also further documented.
-This impact does not become clear in the Radon-B data -- indeed due to the small number of dates [@lubosCzechLandsAustria2013; @markovaSlovakiaHungary2013].
-
-`r a("Poland")`:
-The region *Poland* completely covers the centre and western part of today's Poland.
-However, the ^14^C data from this region are almost exclusively from the voivodships of Kajuwien-Pommern, Lodsch and Lower Silesia.
-The amount of data is small overall, with significantly more data from the Early Bronze Age up to about 1700calBC.
-Almost no information is available from the Middle and Late Bronze Ages.
-This lack of data applies to both the *burial type* and the *burial construction* variables, the former being slightly better documented.
-The relative development of *burial type* approximates the *real* conditions -- i.e.
-according to the classical, archaeological narrative: inhumations were significantly more frequent in western Poland until the Late Bronze Age, while cremations did rarely occur before the Urnfield Period.
-In contrast, it misrepresents *burial construction*: burial mounds also played an important role in Middle Bronze Age Poland [@czebreszukBronzeAgePolish2013; @dabrowskiAeltereBronzezeitPolen2004].
-
-`r a("Southern Germany")`:
-The artificial district region *Southern Germany* covers Bavaria completely but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and western Austria.
-No ^14^C data are available from the latter regions, the majority of the available data comes from Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative district of Tübingen.
-For the Early Bronze Age, Radon-B preserves a remarkable amount of data from southern Germany, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves.
-The *burial type* variable is quite well understood -- although disproportionately much more of the early inhumations were recorded than urnfield burials.
-This wealth of information contrasts sharply with the situation regarding the *burial construction* variable, for which almost no information from this area is available throughout the Bronze Age.
-If one considers the relative development calculated from Radon-B and compares it with the archaeological literature, then a reasonable agreement can be found for *burial type*.
-Inhumation dominated for a long time but was then replaced by cremation from the Middle Bronze Age onwards and massively in the Late Bronze Age.
-The calculated proportions for *burial construction* are based on a negligibly small amount of data, but the proxy still indicates the right tendency for early and middle Bronze Age: Flat graves replaced mounds.
-The dataset does not allow any statements about the Late Bronze Age [@falkensteinWandelBestattungssittenHugelgraber2017; @jockenhovelGermanyBronzeAge2013].
-
-`r a("Northeastern France")`:
-*Northeastern France* comprises the French administrative regions Grand-Est and Bourgogne-Franche-Comté, the east of the German Baden-Württemberg, Rhineland-Palatinate and Saarland, as well as Luxembourg and Wallonia in Belgium.
-Accumulations of ^14^C data are only available from Württemberg and towards the Paris Basin.
-The amount of data is low overall, but very low in the early and early Middle Bronze Age.
-For *burial type* limited statements are possible after this time window, *burial construction*, on the other hand, must remain obscure.
-The prediction of the relative development of these variables is correspondingly error-prone.
-Still, the parallel existence of cremation and inhumation in the Early and Middle Bronze Age is correctly reflected.
-However, this situation already existed in the Early Bronze Age and indeed developed more nuanced than the proxy might suggest.
-Contrary to the derivation from the ^14^C ratios, flat graves were quite common in northeastern France [@mordantBronzeAgeFrance2013].
-
-`r a("Northern Germany")`:
-The artificial region *Northern Germany* includes all federal states in East Germany, but also large parts of Lower Saxony and Schleswig-Holstein.
-The ^14^C data for this region come mainly from Saxony-Anhalt and neighbouring regions in Thuringia, Saxony and Lower Saxony.
-Also from the area north of Bremen some data are documented in Radon-B.
-The database does not contain as many dated graves from any other region as from northern Germany.
-Most of them date back to the Early Bronze Age, while the Middle Bronze Age from about 1800calBC to 1400calBC seems barren.
-Only for the Late Bronze Age, there is a significant number of data documented again.
-The meta information on *burial type* and *burial construction* trace this development to the same extent.
-Since in Northern Germany influences from all four cardinal directions come together and contribute significantly to the formation of local cultural phenomena, it is difficult to estimate the overall development of the primary variables here as well.
-The ^14^C proxy correctly represents *burial site* insofar as inhumations dominated in the Early Bronze Age and were later on replaced by cremations especially in the North from the Middle Bronze Age onwards.
-In the Late Bronze Age though, cremation was more dominant than the dataset suggests.
-Concerning *burial construction*, the current Radon-B stock records the fundamental dominance of flat graves well but attributes more relevance to them in the Early and Middle Bronze Age than is usual in the archaeological assessment of the context [@hofmannRituelleUmgangMit2008; @jockenhovelGermanyBronzeAge2013].
-
-`r a("Southern Scandinavia")`:
-The region *Southern Scandinavia* includes all of Denmark with Jutland, Funen, Zeeland and all smaller islands as well as large parts of east of southern Sweden, and coastal regions in Schleswig-Holstein and Mecklenburg-Western Pomerania.
-The ^14^C data are very well distributed across Jutland, Funen and the Swedish Skåne region.
-Since the Bronze Age in Scandinavia begins much later than in Central Europe and is measured by another chronological system, the diachronic developments cannot merely be compared through the tripartite structure of the Early, Middle and Late Bronze Ages.
-The period which is labelled Early Bronze Age in Central Europe corresponds to the Scandinavian Late Neolithic.
-The majority of data from Scandinavia are from the period IB onwards.
-In this respect, the data density for almost the entire Bronze Age is good -- also with regard to the two variables *burial type* and *burial construction*.
-The Late Neolithic from 2200calBC is a desideratum.
-If one carries out the necessary mental shift -- the Central European Middle Bronze Age corresponds to the Nordic Early Bronze Age -- and compares the proxy of the relative grave form development with the archaeological observations, then one finds a substantial agreement: inhumations were the rule in periods I and II (about 1800-1300calBC), but cremations also occurred sporadically.
-From period III onwards cremation burials gained the upper hand.
-In the Scandinavian Early Bronze Age burials in burial mounds dominated.
-In the Late Bronze Age, flat burial fields began to occure [@thraneScandinavia2013].
-
-`r a("Benelux")`:
-The Benelux region includes the Netherlands, northern Belgium and large parts of Lower Saxony and North Rhine-Westphalia.
-All these regions have ^14^C data on graves in Radon-B, but the data density in Belgium and the Netherlands is much higher.
-As in Southern Scandinavia, the Early Bronze Age is under-represented in the Benelux region as opposed to the Middle and Late Bronze Age, although in a slightly weaker form.
-While the information density of the *burial type* variable follows this course, *burial construction* is almost undocumented in the Late Bronze Age.
-As in the *Southern Germany* region, there is a serious difference between the information density of both variables.
-If one considers the relative development of the variables as they can be derived from the burial data in comparison to the archaeological state of research, there is both deviations as well as agreement.
-Here, too, the study region includes different cultural phenomena with highly different burial customs: In the north and east of the Benelux region, inhumations were the rule until the Late Bronze Age, while cremation dominated earlier south of the Meuse.
-The Radon-B proxy seems to overemphasise the dominance of cremation in the whole area.
-The relevance of burial mounds in the Early and Middle Bronze Age, on the other hand, is correctly reflected, and the mixture between mounds and flat graves in the context of urnfields, often resulting from the continued use of existing grave monuments, is also visible [@arnoldussenBronzeAgeSettlements2008; @drenthMoundsDeadFuneray2005; @lohofTraditionChangeBurial1994].
-
-`r a("England")`:
-The *England* region stretches across England up to York and Liverpool, incorporating parts of Wales and leaving out Cornwall and Devon in the southwest.
-It also touches a narrow coastal strip on the mainland, from which no ^14^C data are documented.
-The available data are concentrated in the southeast of Britain but are well distributed there.
-The diachronic data availability is relatively good in the early and middle Bronze Age but then decreases in the late Bronze Age.
-*burial type* and *burial construction* behave accordingly.
-The relative development derived from the ^14^C data indicates the correct trend for *burial type*: Cremation graves became more and more important during the Bronze Age, while inhumations appeared for a long time in parallel.
-However, the dataset emphasises the occurence of cremation graves too much.
-Concerning the variables *burial construction*, the dominance of burial mounds corresponds perfectly to the archaeological narrative, but -- especially in the Late Bronze Age -- flat grave fields also occurred, which the data set wrongly excludes [@robertsBritainIrelandBronze2013].
+This narrative is wrong or inaccurate in some aspects. Please see the analysis in Section \ref{time-series-discussion}.
 

### 2019-01-13 article/04_Discussion.Rmd
added subsections
moved detailed discussion of the timeseries to the discussion section

@@
 # Discussion {#discussion}
+
+## Time series {#time-series-discussion}
+
+The following section contains a short evaluation of the time series in Figure \ref{development_burial_type} and \ref{development_burial_construction} in the light of the archaeological state of research for each region: \newline
+
+`r a("Southeastern Central Europe")`^[Amount of graves in Radon-B for this region.]:
+The artificial circular region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map_regions} A).
+The density of ^14^C data on graves in Radon-B is low -- dates are available from the various regions, but large areas are without evidence.
+Diachronically the amount of data is consistently low and reaches a minimum in the Middle Bronze Age.
+While there is a constant flow of information for the variable *burial type*, the number of graves without information on *burial construction* increases in the Late Bronze Age to such an extent that no meaningful overview can be derived.
+The region intersects diverse geographical and cultural areas complicating the comparison with archaeological observations.
+In the southeast of Central Europe, cremation and inhumation already existed side by side in the Early Bronze Age, and the local tradition remained stable in this respect also due to the expansion of the burial mound tradition in the Middle Bronze Age -- with a slight overall increase in cremation graves.
+The Urnfield period brought the transition to the absolute dominance of cremation burial.
+The development of the *burial type* variable in the Radon-B proxy shows this development correctly.
+The information from ^14^C data on *burial construction* correctly traces the relevance of flat burials in the Early Bronze Age but then overemphasises them in the Middle Bronze Age compared to the construction of burial mounds.
+In the Late Bronze Age, burials on flat burial grounds were frequent, although burial mounds of the Tumulus Culture were also further documented.
+This impact does not become clear in the Radon-B data -- indeed due to the small number of dates [@lubosCzechLandsAustria2013; @markovaSlovakiaHungary2013].
+
+`r a("Poland")`:
+The region *Poland* completely covers the centre and western part of today's Poland.
+However, the ^14^C data from this region are almost exclusively from the voivodships of Kajuwien-Pommern, Lodsch and Lower Silesia.
+The amount of data is small overall, with significantly more data from the Early Bronze Age up to about 1700calBC.
+Almost no information is available from the Middle and Late Bronze Ages.
+This lack of data applies to both the *burial type* and the *burial construction* variables, the former being slightly better documented.
+The relative development of *burial type* approximates the *real* conditions -- i.e.
+according to the classical, archaeological narrative: inhumations were significantly more frequent in western Poland until the Late Bronze Age, while cremations did rarely occur before the Urnfield Period.
+In contrast, it misrepresents *burial construction*: burial mounds also played an important role in Middle Bronze Age Poland [@czebreszukBronzeAgePolish2013; @dabrowskiAeltereBronzezeitPolen2004].
+
+`r a("Southern Germany")`:
+The artificial district region *Southern Germany* covers Bavaria completely but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and western Austria.
+No ^14^C data are available from the latter regions, the majority of the available data comes from Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative district of Tübingen.
+For the Early Bronze Age, Radon-B preserves a remarkable amount of data from southern Germany, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves.
+The *burial type* variable is quite well understood -- although disproportionately much more of the early inhumations were recorded than urnfield burials.
+This wealth of information contrasts sharply with the situation regarding the *burial construction* variable, for which almost no information from this area is available throughout the Bronze Age.
+If one considers the relative development calculated from Radon-B and compares it with the archaeological literature, then a reasonable agreement can be found for *burial type*.
+Inhumation dominated for a long time but was then replaced by cremation from the Middle Bronze Age onwards and massively in the Late Bronze Age.
+The calculated proportions for *burial construction* are based on a negligibly small amount of data, but the proxy still indicates the right tendency for early and middle Bronze Age: Flat graves replaced mounds.
+The dataset does not allow any statements about the Late Bronze Age [@falkensteinWandelBestattungssittenHugelgraber2017; @jockenhovelGermanyBronzeAge2013].
+
+`r a("Northeastern France")`:
+*Northeastern France* comprises the French administrative regions Grand-Est and Bourgogne-Franche-Comté, the east of the German Baden-Württemberg, Rhineland-Palatinate and Saarland, as well as Luxembourg and Wallonia in Belgium.
+Accumulations of ^14^C data are only available from Württemberg and towards the Paris Basin.
+The amount of data is low overall, but very low in the early and early Middle Bronze Age.
+For *burial type* limited statements are possible after this time window, *burial construction*, on the other hand, must remain obscure.
+The prediction of the relative development of these variables is correspondingly error-prone.
+Still, the parallel existence of cremation and inhumation in the Early and Middle Bronze Age is correctly reflected.
+However, this situation already existed in the Early Bronze Age and indeed developed more nuanced than the proxy might suggest.
+Contrary to the derivation from the ^14^C ratios, flat graves were quite common in northeastern France [@mordantBronzeAgeFrance2013].
+
+`r a("Northern Germany")`:
+The artificial region *Northern Germany* includes all federal states in East Germany, but also large parts of Lower Saxony and Schleswig-Holstein.
+The ^14^C data for this region come mainly from Saxony-Anhalt and neighbouring regions in Thuringia, Saxony and Lower Saxony.
+Also from the area north of Bremen some data are documented in Radon-B.
+The database does not contain as many dated graves from any other region as from northern Germany.
+Most of them date back to the Early Bronze Age, while the Middle Bronze Age from about 1800calBC to 1400calBC seems barren.
+Only for the Late Bronze Age, there is a significant number of data documented again.
+The meta information on *burial type* and *burial construction* trace this development to the same extent.
+Since in Northern Germany influences from all four cardinal directions come together and contribute significantly to the formation of local cultural phenomena, it is difficult to estimate the overall development of the primary variables here as well.
+The ^14^C proxy correctly represents *burial site* insofar as inhumations dominated in the Early Bronze Age and were later on replaced by cremations especially in the North from the Middle Bronze Age onwards.
+In the Late Bronze Age though, cremation was more dominant than the dataset suggests.
+Concerning *burial construction*, the current Radon-B stock records the fundamental dominance of flat graves well but attributes more relevance to them in the Early and Middle Bronze Age than is usual in the archaeological assessment of the context [@hofmannRituelleUmgangMit2008; @jockenhovelGermanyBronzeAge2013].
+
+`r a("Southern Scandinavia")`:
+The region *Southern Scandinavia* includes all of Denmark with Jutland, Funen, Zeeland and all smaller islands as well as large parts of east of southern Sweden, and coastal regions in Schleswig-Holstein and Mecklenburg-Western Pomerania.
+The ^14^C data are very well distributed across Jutland, Funen and the Swedish Skåne region.
+Since the Bronze Age in Scandinavia begins much later than in Central Europe and is measured by another chronological system, the diachronic developments cannot merely be compared through the tripartite structure of the Early, Middle and Late Bronze Ages.
+The period which is labelled Early Bronze Age in Central Europe corresponds to the Scandinavian Late Neolithic.
+The majority of data from Scandinavia are from the period IB onwards.
+In this respect, the data density for almost the entire Bronze Age is good -- also with regard to the two variables *burial type* and *burial construction*.
+The Late Neolithic from 2200calBC is a desideratum.
+If one carries out the necessary mental shift -- the Central European Middle Bronze Age corresponds to the Nordic Early Bronze Age -- and compares the proxy of the relative grave form development with the archaeological observations, then one finds a substantial agreement: inhumations were the rule in periods I and II (about 1800-1300calBC), but cremations also occurred sporadically.
+From period III onwards cremation burials gained the upper hand.
+In the Scandinavian Early Bronze Age burials in burial mounds dominated.
+In the Late Bronze Age, flat burial fields began to occure [@thraneScandinavia2013].
+
+`r a("Benelux")`:
+The Benelux region includes the Netherlands, northern Belgium and large parts of Lower Saxony and North Rhine-Westphalia.
+All these regions have ^14^C data on graves in Radon-B, but the data density in Belgium and the Netherlands is much higher.
+As in Southern Scandinavia, the Early Bronze Age is under-represented in the Benelux region as opposed to the Middle and Late Bronze Age, although in a slightly weaker form.
+While the information density of the *burial type* variable follows this course, *burial construction* is almost undocumented in the Late Bronze Age.
+As in the *Southern Germany* region, there is a serious difference between the information density of both variables.
+If one considers the relative development of the variables as they can be derived from the burial data in comparison to the archaeological state of research, there is both deviations as well as agreement.
+Here, too, the study region includes different cultural phenomena with highly different burial customs: In the north and east of the Benelux region, inhumations were the rule until the Late Bronze Age, while cremation dominated earlier south of the Meuse.
+The Radon-B proxy seems to overemphasise the dominance of cremation in the whole area.
+The relevance of burial mounds in the Early and Middle Bronze Age, on the other hand, is correctly reflected, and the mixture between mounds and flat graves in the context of urnfields, often resulting from the continued use of existing grave monuments, is also visible [@arnoldussenBronzeAgeSettlements2008; @drenthMoundsDeadFuneray2005; @lohofTraditionChangeBurial1994].
+
+`r a("England")`:
+The *England* region stretches across England up to York and Liverpool, incorporating parts of Wales and leaving out Cornwall and Devon in the southwest.
+It also touches a narrow coastal strip on the mainland, from which no ^14^C data are documented.
+The available data are concentrated in the southeast of Britain but are well distributed there.
+The diachronic data availability is relatively good in the early and middle Bronze Age but then decreases in the late Bronze Age.
+*burial type* and *burial construction* behave accordingly.
+The relative development derived from the ^14^C data indicates the correct trend for *burial type*: Cremation graves became more and more important during the Bronze Age, while inhumations appeared for a long time in parallel.
+However, the dataset emphasises the occurence of cremation graves too much.
+Concerning the variables *burial construction*, the dominance of burial mounds corresponds perfectly to the archaeological narrative, but -- especially in the Late Bronze Age -- flat grave fields also occurred, which the data set wrongly excludes [@robertsBritainIrelandBronze2013].
+
+## Cultural distance
+
+Surprising result of no correlation: Meaning? Construction of an own story?
+
+## Data quality
 
@@
 An under-representation of cremation graves in current data collections is, therefore, more plausible.
-
-- Comparison archaeological observations and reconstruction based on data
-
-- Surprising result of no correlation: Meaning? Construction of an own story?

### 2019-01-13 article/03_Results.Rmd
language in the results chapter

@@
 
-This narrative is wrong or inaccurate in some aspects. Please see the analysis in Section \ref{time-series-discussion}.
+This narrative is wrong or inaccurate in some aspects. Please see the analysis in section \ref{time-series-discussion}.
 
@@
 Beyond the simple comparison of archaeological knowledge and information derived from Radon-B, Figure \ref{sed_region_matrix} shows the time series of squared euclidian distance for each region-region relationship.
-The majority of the curves in its upper right half for *burial type* correspond to expectations: At the beginning of the Bronze Age, inhumation is widespread throughout the study area, so the Euclidean distance at 2200calBC is small in almost every region-region relationship.
+The majority of the curves in its top right half for *burial type* correspond to expectations: At the beginning of the Bronze Age, inhumation is widespread throughout the study area, so the Euclidean distance at 2200calBC is small in almost every region-region relationship.
 Only the Benelux region, where the share of cremation is already high in the Early Bronze Age, is excluded.
@@
 The SED values increase in this period and show clear peaks or several successive maxima.
-This rise is a consequence of the fact that the transition process in each region as depicted in Radon-B is different and that it took place with a time lag compared to other regions.
+This rise is a consequence of the time lag in the transition processes as depicted in Radon-B: It is different in every region.
 At the end of the Late Bronze Age, the average SED is approaching zero again, as burial by cremation dominates in almost all of Europe.
-The early and complete adoption of the cremation rite in the Benelux region, as depicted in Radon-B, clearly sets it apart from the other regions.
+The early and complete adoption of the cremation rite in the Benelux area, as depicted in Radon-B, clearly sets it apart from the other regions.
 England, on the other hand, with its slow, gradual transition from inhumation to cremation, stands out due to its consistently moderate distances to all other regions.
@@
 The distance does not seem to follow a general trend, but a multitude of local and temporally limited phenomena.
-This heterogeneity is also due to how correctly or incorrectly the proxy developed from radon-B depicts developments in the real world.
+This heterogeneity is of course also due to how correctly or incorrectly the proxy developed from Radon-B depicts developments in the real world.
 In the Middle Bronze Age, the Tumulus culture should bring about a fundamental reduction of the SED for almost all region-region relationships -- this effect is indeed weakly visible.
@@
 The degree of agreement between two regions in the question of inhumation and cremation is therefore not suitable for predicting the similarity of developments in these regions concerning the dominance of flat or mound graves.
-Only for the time window at the transition to the Middle Bronze Age from 1800 to 1600calBC is the 5% significance level of the mantle test almost undercut, thus suggesting a statistically significant relationship.
+Only for the time window at the transition to the Middle Bronze Age from 1800 to 1600calBC is the 5% significance level of the Mantle test almost undercut, thus suggesting a statistically significant relationship.
 
-For the relationship between the SED network based on *burial type* and the spatial distance several observations result from Table \ref{tab:distanceandcorrelationtable*}: Surprisingly, cultural distance does not seem to correlate with spatial distance.
-A significant correlation between the distance measures only reveals itself for the Early Bronze Age time window 2000 to 1800calBC: Only in the Early Bronze Age, long before the universal switch to the Urnfield burial, large regions directly adjacent to each other tend to undergo the same developments rather than the ones further away from each other.
-In the Middle Bronze Age, the SED developments of the *burial type* variable undergo the peaks already observed in Figure \ref{sed_region_matrix}.
+For the relationship between the SED network based on *burial type* and the spatial distance several observations result from Table \ref{tab:distanceandcorrelationtable*}: Cultural distance calculated from these to variables does not seem to correlate with spatial distance.
+A significant correlation between the distance measures only reveals itself for the Early Bronze Age time slot 2000 to 1800calBC: Only in the Early Bronze Age, long before the universal switch to the Urnfield burial tradition, large regions directly adjacent to each other tend to undergo the same developments rather than the ones further away from each other.
+In the Middle Bronze Age the SED developments of the *burial type* variable undergo the peaks already observed in Figure \ref{sed_region_matrix}.
 The correlation between cultural and spatial distance over these peaks is low and even tilts between 1400 and 1200calBC, at the transition to the Late Bronze Age, into the negative -- although not significantly.
@@
 Looking at the SED network of *burial construction* (Table \ref{tab:distanceandcorrelationtable*}), similar conclusions can be drawn as in the *burial type* context: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated, but here, too, the early bronze age is by far the phase of the most substantial spatial correlation.
-It is only significant for the time window between 1800 and 1400calBC.
+It is only significant for the time slot between 1800 and 1400calBC.
 There is no time period with negative mean correlation for *burial construction* cultural distance and spatial distance, but also here the correlation coefficient approaches zero in the Late Bronze Age.

### 2019-01-13 article/03_Results.Rmd
some description of the simulation results

@@
 
-Simulation: What is to be expected in case of influence of spatial distance?
+The results of the simulation as presented in Figure \ref{simulation} confirm the expectation: In case of equal interaction intensity $I$ of separate populations, the correlation of cultural and spatial distance is weak (Figure \ref{simulation} A vs.
+C vs.
+E).
+It appears to be normally distributed around zero for many simulation runs, although individual runs can still yield remarkably high positive and negative correlation for certain time slots.
+The fluctuation of the correlation within one run can be strong and seems to increase for bigger values of $m_i$.
+If the interaction matrix $I$ is predefined to represent the spatial distance matrix (Figure \ref{simulation} BDF), then the mean correlation per time slot increases.
+The correlation increases as well with the degree of interaction $m_i$ (Figure \ref{simulation} D vs.
+F).
+Counterintuitively the fluctuation of the correlation is also here very high (Figure \ref{simulation} E vs.
+F).
+The correlation development observed for the variables *burial type* and *burial construction* fits to simulation runs in all presets, but becomes less likely if we assume $I$ to represent spatial distance and a high degree of interaction $m_i$ (Figure \ref{simulation} A vs.
+F).
+That's mostly because of the relatively strong negative correlation of cultural and spatial distance in the Middle and Late Bronze Age (1400 to 1000calBC).

### 2019-01-13 article/04_Discussion.Rmd
introuction to the time series discussion

@@
 
-The following section contains a short evaluation of the time series in Figure \ref{development_burial_type} and \ref{development_burial_construction} in the light of the archaeological state of research for each region: \newline
+One major question asked in this paper was, if openly available, bulk radiocarbon data does allow the reconstruction of main trends within the distribution of burial rites in Bronze Age Europe.
+This question can be answered with Yes.
+The narrative in section \ref{time-series-results} is surprisingly similar to the one presented by @hardingEuropeanSocietiesBronze2000a.
+The following paragraphs contain a short evaluation of the time series in Figure \ref{development_burial_type} and \ref{development_burial_construction} in the light of the archaeological state of research for each region.
+Despite the general success of this methodological apporach, unfortunately the amount and quality of data is not sufficient yet for a comprehensive picture.
+Please see the discussion of data quality in section \ref{data-quality-discussion}.
 
@@
 
-## Data quality
+## Data quality {#data-quality-discussion}
 

### 2019-01-13 article/02_Materials_and_Methods.Rmd
changes in simulation paramaters

@@
     0 & Figure \ref{simulation} AB, no intergroup exchange \\ 
-    0.01 & Figure \ref{simulation} CD \\
-    0.1 & Figure \ref{simulation} EF \\
-    0.5 & Figure \ref{simulation} GH, very high exchange
+    0.05 & Figure \ref{simulation} CD \\
+    0.5 & Figure \ref{simulation} EF, very high exchange
   \end{dcases*}

### 2019-01-13 analysis/code/04_simulation_scripts/01_run_simulation.R

@@ -16,7 +16,7 @@ models_grid <- expand.grid(
   mu = 0,
   g = 8,
   I = list(NA, interaction_matrix_spatial),
-  mi = c(0, 0.05, 0.2)
+  mi = c(0, 0.05, 0.5)
 ) %>%
   tibble::as.tibble() %>%
   dplyr::mutate(

### 2019-01-13 article/04_Discussion.Rmd
layout

@@
 The following paragraphs contain a short evaluation of the time series in Figure \ref{development_burial_type} and \ref{development_burial_construction} in the light of the archaeological state of research for each region.
-Despite the general success of this methodological apporach, unfortunately the amount and quality of data is not sufficient yet for a comprehensive picture.
+Despite the general success of this methodological apporach, unfortunately the amount and quality of data is still not sufficient for a comprehensive picture.
 Please see the discussion of data quality in section \ref{data-quality-discussion}.
+\newline
 

### 2019-01-13 analysis/code/05_plot_scripts/sed_simulation.R
wrong data caused wrong plot

@@ -7,7 +7,7 @@ models_grid %<>% dplyr::mutate(I_type = ifelse(sapply(I, length) == 1, "equal di
 mantel_simulations %<>% dplyr::left_join(models_grid[, c("model_id", "model_group", "mi", "I_type")], by = "model_id")
 load("analysis/data/tmp_data/mantel_sed_spatial_burial_type.RData")
 mantel_burial_type <- mantel_test_results
-load("analysis/data/tmp_data/mantel_sed_spatial_burial_type_burial_construction.RData")
+load("analysis/data/tmp_data/mantel_sed_spatial_burial_construction.RData")
 mantel_burial_construction <- mantel_test_results
 
 mantel_burial_type %<>%

### 2019-01-13 article/03_Results.Rmd
introuction to the time series discussion
some discussion text: cultural distance

@@
 
-## Timeseries
+## Timeseries {#time-series-results}
 
@@
 The cultural distance between several geographically distant regions is then smaller than between several directly adjacent regions.
-Looking at the SED network of *burial construction* (Table \ref{tab:distanceandcorrelationtable*}), similar conclusions can be drawn as in the *burial type* context: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated, but here, too, the early bronze age is by far the phase of the most substantial spatial correlation.
+Looking at the SED network of *burial construction* (Table \ref{tab:distanceandcorrelationtable*}), similar conclusions can be drawn as in the *burial type* context: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated.
+As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation.
 It is only significant for the time slot between 1800 and 1400calBC.

### 2019-01-13 article/04_Discussion.Rmd
some discussion text: cultural distance

@@
 
-Surprising result of no correlation: Meaning? Construction of an own story?
+The second important question posed for this paper was, if these time series constructed from radiocarbon data (Figure \ref{development_burial_type} and \ref{development_burial_construction}) allow further analysis concerning cultural and spatial distance?
+While the idea behind the first question is rather technical -- doing something that's already available in a quantitative and reproducible way -- the second one has a larger scope.
+A quantitative representation of an archaeological narrative should path the way for a more diverse set of methods to be applied to this data.
+Cultural distance is one of the first topics that comes to mind.
+
+The expectation for this question was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* as well in between these variables and the spatial distance.
+The simulation, though with this implementation and parametrisation for sure rather qualitativly than quantitativly, visualises how a high degree of interaction should translate to a significantly higher degree of cultural similarity across human populations (Figure \ref{simulation}).
+And as the European Bronze Age is usually depicted as a period of intensive trade, travel and exchange [e.g.
+@kristiansenRiseBronzeAge2005; @sherrattWhatWouldBronzeAge1993] there was no reason to expect something else than Tobler's famous First Law of Geography: "everything is related to everything else, but near things are more related than distant things." [@toblerComputerMovieSimulating1970].
+
+Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:distanceandcorrelationtable*}).
+Only for small periods of time this correlation can be assumed based on the Radon-B data.
+The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems behave in a opposing manner: negative correlation.
+If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000calBC) and a decline (after 2000) of cultural-spatial-distance correlation.
+One interpretation of this oberservation might be, that the Early Bronze Age was a time of rising cultural exchange and religious homogenization, while in later periods different domains secluded themselves more strongly.
+There are two main reasons why such interpretations are difficult so far:
+
+1. Data amount, distributiona and quality:
+
+2. Data meaning
+
+Only two variables
+
+Keeping this in mind, the surprising result is still valid and has to be considered in future hypotheses building.
 
@@
 The A sections of the figures \ref{development_burial_type} and \ref{development_burial_construction} with absolute grave numbers per region and year indicate areas and periods with a lot and others with very few points of data.
-For some, there is not enough information to make any decision about the relative dominance and for many, the general relevance and accuracy of the resulting time series can be questioned.
-There's a variety of reasons for this:
+For some there is not enough information to make any decision about the relative dominance and for many the general relevance and accuracy of the resulting time series can be questioned.
+There's a variety of reasons for this which should be considered for future data acquisition in Radon-B and beyond:
 

### 2019-01-20 article/04_Discussion.Rmd
a section about data meaning

@@
 Only for small periods of time this correlation can be assumed based on the Radon-B data.
-The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems behave in a opposing manner: negative correlation.
+The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation.
 If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000calBC) and a decline (after 2000) of cultural-spatial-distance correlation.
@@
 
-1. Data amount, distributiona and quality:
+1. The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} queries the credibility of the cultural distance calculation.
 
-2. Data meaning
+2. Data meaning \ref{data-meaning-discussion}
 
@@
 An under-representation of cremation graves in current data collections is, therefore, more plausible.
+
+## Data meaning {#data-meaning-discussion}
+
+This paper is based on ^14^C dates extracted from the Radon-B data set which were prepared for the dating of grave findings.
+^14^C data on graves are used to chronologically attach a feature to the absolute chronological context or to better describe the relative development within a finding complex.
+The questions underlying the creation of a ^14^C date are therefore generally not equivalent to those imposed in this paper.
+While here the presence of an idea is to be measured abstractly at a certain place at a certain time, graves are usually dated exactly the other way round to contextualize and better understand them as concrete object.
+By definition, field research is oriented towards small-scale questions that can be answered to their respective extent.
+It must put a lot of energy into doing justice to all findings that are inevitably destroyed by excavation.
+There is a fundamental break between that great, theoretical archaeology, which wants to develop comprehensive narratives and only pay quantitative attention to the individual case, and the individual research project or emergency excavation, which has to dedicate a measure of respect and sometimes a lot of working time to each grave.
+In this thesis, ^14^C data are converted into a proxy of calibrated density distributions within artificial working regions, which is completely detached from the individual context.
+
+The superordinate, large perspective also tries to simplify and categorize the parameters associated with a single burial.
+Radon-B anticipated this step by providing the variables *burial type* and *burial construction* that can be reduced to three categories each.
+This is extremely useful on the one hand, but also questionable on the other.
+The respective contexts can be multifarious: similar rituals in different cultural contexts can have different meanings.
+For this work, it is assumed -- with reference to *Cultural Evolution Theory* -- that despite this restriction, actions such as burning the dead or burying in burial mounds can be passed on across cultural and ethnic group boundaries through imitation and teaching.
+To put it provocatively: cremation burial in the Hungarian Early Bronze Age is a manifestation of the same idea as cremation in the Northern Late Bronze Age.
+This is a working hypothesis, but may be questioned.
+
+The meaning of the data in Radon-B is thus manipulated in at least three ways for the purpose of this work: By concentrating on the dated grave in contrast to the ^14^C date, this work detaches itself from the actual content of the database.
+By creating artificial time steps and investigation regions, the cultural-historical context of the grave is dissolved.
+By creating the relative development proxies, the individual grave loses its context.
+The present paper is therefore a clear secondary use of Radon-B.
+A better dataset should be grave-related, not date-related, and should contain much more meta-information to each grave.

### 2019-01-20 article/04_Discussion.Rmd
some possible interpretation of the results

@@
 If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000calBC) and a decline (after 2000) of cultural-spatial-distance correlation.
-One interpretation of this oberservation might be, that the Early Bronze Age was a time of rising cultural exchange and religious homogenization, while in later periods different domains secluded themselves more strongly.
-There are two main reasons why such interpretations are difficult so far:
+There are several reasons why any interpretation of this observation is difficult so far:
 
-1. The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} queries the credibility of the cultural distance calculation.
+1.
+The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} queries the credibility of the cultural distance calculation.
+The macroregions have to be very big due to a lack of data and don't allow for a precise picture.
 
-2. Data meaning \ref{data-meaning-discussion}
+2.
+The data used for this analysis was collected for another purpose and its application in this context requires a problematic shift in data meaning as explained in section \ref{data-meaning-discussion}.
 
-Only two variables
+3.
+All observations of cultural distance in this paper are based on only the two variables *burial type* and *burial construction*.
+This is by far not enough for any overarching conclusions about the general state of society in the Bronze Age.
 
-Keeping this in mind, the surprising result is still valid and has to be considered in future hypotheses building.
+Keeping this in mind, the surprising result is still valid and has to be explained.
+One interpretation might be, that the Early Bronze Age was a time of rising cultural exchange and religious homogenization, while in later periods different domains secluded themselves more strongly.
+The negative correlation of *burial type* in the Late Bronze Age could indicate long distance contacts that explicitly excluded certain intermediate regions of Central Europe.
+The rise of the Tumulus Culture in the Middle Bronze Age might have cushioned this effect for the variable *burial construction* and at the same time introduced more variability for a less important aspect *burial type*.
+The latter may have been of subordinate relevance in the Tumulus Culture context.
+With the rise of the Urnfield Culture this phenomonon could have been reversed as *burial type* became an expression of something that could have been one of the first formalized religions in Europe: The spread of the Urnfield Culture was an extraordinary development in early European history and was possibly accompanied by a structured, religious dogma that could have built up around the symbolic content of cult wagons, sunbarks and waterfowl [@bilicSwanChariotSolar2016; @kimmig_seevolkerbewegung_1964; @molloy_european_2018].
+Of course these explanations could, if accurate, only be considered on a really big scale.
+For individual regions and sites specific ideological, social or even economic considerations might have been much more critical for the adopted burial tradition [@barrett_monumentality_1990; @fokkens_genesis_1997-1; @muller-scheesel_variabilitat_2009].
 

### 2019-01-20 article/06_Acknowledgements.Rmd
Acknowledgements

# Acknowledgements

I would like to offer my special thanks to my supervisors PD Dr. Oliver Nakoinz and Dr. Martin Hinz for developing this research idea with me and for ongoing support and advice before, while and after the creation of the master thesis this article is based on. With Dr. Jutta Kneisel, Dr. Christoph Rinne, Dr. Daniel Knitter, Dr. Nils Müller-Scheeßel and David Matzig, B.A. I had helpful discussions at different stages of the research process. Beyond these colleagues at Kiel University I am particularly grateful for the feedback, encouragement and literature recommendations by Prof. Dr. Ben Marwick and the valuable C++ code review by Benny Baumann.

### 2019-01-21 article/01_Introduction.Rmd
some discussion text: cultural distance
spelling

@@
 This approach could set a precedent for a more reproducible and statistically sound research practice.
-If so, then the quantitative comparison of macroregional trends could also pave the way for an analysis of interaction intensity on a long-term continental scale: Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?
+If so, then the quantitative comparison of macro regional trends could also pave the way for an analysis of interaction intensity on a long-term continental scale: Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?
 
@@
 
-Archaeology has a unique perspective on Cultural evolution [@garveyCurrentPotentialRoles2018]: It is concerned with human behaviour over big areas and long timescales.
+Archaeology has a unique perspective on Cultural evolution [@garveyCurrentPotentialRoles2018]: It is concerned with human behaviour over big areas and long time scales.
 It also works with solid proxies of cultural expression: material remains -- individual lost or discarded artefacts, the ruins of past architecture and all more inconspicuous by-products of living.
@@
 
-In the following sections I will (1) explain how radiocarbon data with relevant meta information can be transformed to macro regional time series of burial traditions, (2) explore how these timeseries compare to traditional archaeological narratives and (3) construct and contextualise cultural distances between macro regions over time in relation to spatial distance.
+In the following sections I will (1) explain how radiocarbon data with relevant meta information can be transformed to macro regional time series of burial traditions, (2) explore how these time series compare to traditional archaeological narratives and (3) construct and contextualise cultural distances between macro regions over time in relation to spatial distance.
 For the latter a simulation based on the @neimanStylisticVariationEvolutionary1995 template was employed.

### 2019-01-21 article/02_Materials_and_Methods.Rmd
spelling

@@
 These time series also allow to derive another set of time series with a measure of cultural distance for each region-region relationship (Figure \ref{sed_region_matrix}).
-Cultural distance is defined here as the squared euclidian distance (SED) between the trait proportions of two different regions at a certain point in time [@neimanStylisticVariationEvolutionary1995].
+Cultural distance is defined here as the squared Euclidean distance (SED) between the trait proportions of two different regions at a certain point in time [@neimanStylisticVariationEvolutionary1995].
 SED seems to be a sensible statistic here, because it is relatively simple, works with only a few assumptions and emphasises differences through squaring.
@@
 \includegraphics[width=\textwidth]{../analysis/figures/sed_region_matrix.jpeg}
-\caption{\label{sed_region_matrix}Plot matrix of the region-region squared euclidian distance time series.
+\caption{\label{sed_region_matrix}Plot matrix of the region-region squared Euclidean distance time series.
 In the top right half for the variable \textit{burial type} and in the bottom left for the variable \textit{burial construction}.
-On the x-axis the time is displayed, on the y-axis the squared euclidian distance between the two relevant regions.
+On the x-axis the time is displayed, on the y-axis the squared Euclidean distance between the two relevant regions.
 The faint, grey line shows the actual data, the stronger, coloured line an approximation (LOESS with $span = 0.3$).}
@@
 
-To test for correlation between the distances based on the variables *burial type* and *burial construction*, the timeframe of interest was divided into sections of 200 years.
+To test for correlation between the distances based on the variables *burial type* and *burial construction*, the time frame of interest was divided into sections of 200 years.
 For each section of each the two variables, a matrix of the mean SED over time was calculated.
@@
 $k = 2$ mutually exclusive ideas are randomly but equally distributed in this population at $t_{0}$.
-With every timestep ($t_{total} = (2200 - 800)/20 = 70$) all individuals learn from another randomly chosen individual.
+With every time step ($t_{total} = (2200 - 800)/20 = 70$) all individuals learn from another randomly chosen individual.
 Several different values where tested for the probability $m_i$ that the other agent the individual is learning from comes not from the same, but from another group:
@@
 
-The partner group for each learning instance is selected randomly following a group interaction probability matrix $I$. Two different matrizes where tested:
+The partner group for each learning instance is selected randomly following a group interaction probability matrix $I$. Two different matrices where tested:
 
@@
 
-The distribution of the $k$ variants within the groups can be observed for every timestep and documented in time series just like the ones for the real-world regions data (Figure \ref{development_burial_type} B and \ref{development_burial_construction} B).
+The distribution of the $k$ variants within the groups can be observed for every time step and documented in time series just like the ones for the real-world regions data (Figure \ref{development_burial_type} B and \ref{development_burial_construction} B).
 These time series can as well be used to calculate time series of region-region distance, which again can be split up into time blocks of 200 years to calculate mean distance matrices.
@@
 Each plot shows the results of 50 simulation runs.
-The x-axis displays the time distiguished into 200 year time slots, the y-axis Spearman's rank correlation coefficient between the mean squared euclidian distance matrix for a 200 year time slot and the static spatial distance matrix.
+The x-axis displays the time distinguished into 200 year time slots, the y-axis Spearman's rank correlation coefficient between the mean squared Euclidean distance matrix for a 200 year time slot and the static spatial distance matrix.
 This information is visualized for each time slot with (from left to right) a dotplot coloured according to the $p$-value of the Mantel test, a scatter plot line and a boxplot.

### 2019-01-21 article/03_Results.Rmd
spelling

@@
 Around 2200calBC cremation burials were almost unknown in Poland, southern Germany, northeastern France and northern Germany.
-In the northwest, in Great Britain and the Benelux area, and in the southeast, in Austria and the Czech Republic, however, there were early cremation traditions.
+In the Northwest, in Great Britain and the Benelux area, and in the Southeast, in Austria and the Czech Republic, however, there were early cremation traditions.
 These contexts could be discussed as areas of origin of the later omnipresent phenomenon.
@@
 
-Beyond the simple comparison of archaeological knowledge and information derived from Radon-B, Figure \ref{sed_region_matrix} shows the time series of squared euclidian distance for each region-region relationship.
+Beyond the simple comparison of archaeological knowledge and information derived from Radon-B, Figure \ref{sed_region_matrix} shows the time series of squared Euclidean distance for each region-region relationship.
 The majority of the curves in its top right half for *burial type* correspond to expectations: At the beginning of the Bronze Age, inhumation is widespread throughout the study area, so the Euclidean distance at 2200calBC is small in almost every region-region relationship.

### 2019-01-21 article/04_Discussion.Rmd
spelling

@@
 The following paragraphs contain a short evaluation of the time series in Figure \ref{development_burial_type} and \ref{development_burial_construction} in the light of the archaeological state of research for each region.
-Despite the general success of this methodological apporach, unfortunately the amount and quality of data is still not sufficient for a comprehensive picture.
+Despite the general success of this methodological approach, unfortunately the amount and quality of data is still not sufficient for a comprehensive picture.
 Please see the discussion of data quality in section \ref{data-quality-discussion}.
@@
 The region intersects diverse geographical and cultural areas complicating the comparison with archaeological observations.
-In the southeast of Central Europe, cremation and inhumation already existed side by side in the Early Bronze Age, and the local tradition remained stable in this respect also due to the expansion of the burial mound tradition in the Middle Bronze Age -- with a slight overall increase in cremation graves.
+In the Southeast of Central Europe, cremation and inhumation already existed side by side in the Early Bronze Age, and the local tradition remained stable in this respect also due to the expansion of the burial mound tradition in the Middle Bronze Age -- with a slight overall increase in cremation graves.
 The Urnfield period brought the transition to the absolute dominance of cremation burial.
@@
 For the Early Bronze Age, Radon-B preserves a remarkable amount of data from southern Germany, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves.
-The *burial type* variable is quite well understood -- although disproportionately much more of the early inhumations were recorded than urnfield burials.
+The *burial type* variable is quite well understood -- although disproportionately much more of the early inhumations were recorded than Urnfield burials.
 This wealth of information contrasts sharply with the situation regarding the *burial construction* variable, for which almost no information from this area is available throughout the Bronze Age.
@@
 In the Scandinavian Early Bronze Age burials in burial mounds dominated.
-In the Late Bronze Age, flat burial fields began to occure [@thraneScandinavia2013].
+In the Late Bronze Age, flat burial fields began to occur [@thraneScandinavia2013].
 
@@
 `r a("England")`:
-The *England* region stretches across England up to York and Liverpool, incorporating parts of Wales and leaving out Cornwall and Devon in the southwest.
+The *England* region stretches across England up to York and Liverpool, incorporating parts of Wales and leaving out Cornwall and Devon in the Southwest.
 It also touches a narrow coastal strip on the mainland, from which no ^14^C data are documented.
-The available data are concentrated in the southeast of Britain but are well distributed there.
+The available data are concentrated in the Southeast of Britain but are well distributed there.
 The diachronic data availability is relatively good in the early and middle Bronze Age but then decreases in the late Bronze Age.
@@
 The relative development derived from the ^14^C data indicates the correct trend for *burial type*: Cremation graves became more and more important during the Bronze Age, while inhumations appeared for a long time in parallel.
-However, the dataset emphasises the occurence of cremation graves too much.
+However, the dataset emphasises the occurrence of cremation graves too much.
 Concerning the variables *burial construction*, the dominance of burial mounds corresponds perfectly to the archaeological narrative, but -- especially in the Late Bronze Age -- flat grave fields also occurred, which the data set wrongly excludes [@robertsBritainIrelandBronze2013].
@@
 The expectation for this question was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* as well in between these variables and the spatial distance.
-The simulation, though with this implementation and parametrisation for sure rather qualitativly than quantitativly, visualises how a high degree of interaction should translate to a significantly higher degree of cultural similarity across human populations (Figure \ref{simulation}).
+The simulation, though with this implementation and parametrisation for sure rather qualitatively than quantitatively, visualises how a high degree of interaction should translate to a significantly higher degree of cultural similarity across human populations (Figure \ref{simulation}).
 And as the European Bronze Age is usually depicted as a period of intensive trade, travel and exchange [e.g.
@@
 The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} queries the credibility of the cultural distance calculation.
-The macroregions have to be very big due to a lack of data and don't allow for a precise picture.
+The macro regions have to be very big due to a lack of data and don't allow for a precise picture.
 
@@
 The latter may have been of subordinate relevance in the Tumulus Culture context.
-With the rise of the Urnfield Culture this phenomonon could have been reversed as *burial type* became an expression of something that could have been one of the first formalized religions in Europe: The spread of the Urnfield Culture was an extraordinary development in early European history and was possibly accompanied by a structured, religious dogma that could have built up around the symbolic content of cult wagons, sunbarks and waterfowl [@bilicSwanChariotSolar2016; @kimmig_seevolkerbewegung_1964; @molloy_european_2018].
+With the rise of the Urnfield Culture this phenomenon could have been reversed as *burial type* became an expression of something that could have been one of the first formalized religions in Europe: The spread of the Urnfield Culture was an extraordinary development in early European history and was possibly accompanied by a structured, religious dogma that could have built up around the symbolic content of cult wagons, sun barks and waterfowl [@bilicSwanChariotSolar2016; @kimmig_seevolkerbewegung_1964; @molloy_european_2018].
 Of course these explanations could, if accurate, only be considered on a really big scale.
@@
 In various Metal Age contexts, 14C data are inferior to the accuracy of relative chronological classifications - e.g.
-by primer typology - and are therefore used only sporadically to create absolute chronological fixpoints.
+by primer typology - and are therefore used only sporadically to create absolute chronological fix points.
 Dendrochronological dating is preferred in moist soil contexts and when the necessary woods are available.
@@
 4.
-^14^C dating requires a small but sufficiently contextualizable amount of organic finds.
+^14^C dating requires a small but sufficiently contextualised amount of organic finds.
 Due to the old wood and old water effect, short-lived sample materials such as bones of terrestrial organisms or charred cereal grains are preferred.

### 2019-01-21 article/02_Materials_and_Methods.Rmd
removed colour terms from figure and table captions

@@
 The classes of the variable \textit{burial type} are distinguished by colour, the ones of \textit{burial construction} by shape.
-The red frame indicates the research area.
+The rectangular frame indicates the research area.
 The Map projection is EPSG:102013 and the base layer (land, rivers, lakes) data is taken from the Natural Earth project (\url{www.naturalearthdata.com}).}
@@
 On the x-axis the time is displayed, on the y-axis the squared Euclidean distance between the two relevant regions.
-The faint, grey line shows the actual data, the stronger, coloured line an approximation (LOESS with $span = 0.3$).}
+The faint line shows the actual data, the stronger line an approximation (LOESS with $span = 0.3$).}
 \end{figure*}
@@
 (\\textbf{C}) shows the Mantel test results for the two variables if compared directly with each other.
-Negative correlation coefficients are printed red and $p$-values below $0.1$ bold.",
+$p$-values below $0.1$ are printed bold.",
   booktabs = T,
@@
 The scatter plot points of one simulation run are connected with a faint line from one time slot to the next.
-To compare the runs with the real world observations, the results for the burial customs data (Table \ref{tab:distanceandcorrelationtable*} AB) is added here with with a stronger green and blue line.
-The strong horizontal red line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}
+To compare the runs with the real world observations, the results for the burial customs data (Table \ref{tab:distanceandcorrelationtable*} AB) is added here with with a stronger solid (\textit{burial type}) and dashed (\textit{burial construction}) line.
+The strong horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}
 \end{figure*}

### 2019-01-26 article/01_Introduction.Rmd
work on the language of the introduction

--- 2019-01-11 article/01_Introduction.Rmd
@@
 Within this history of constant emergence, displacement and demise many different forms of burial traditions were practised.
-Most of them can be distinguished by the criteria *construction of the grave* and *treatment of the body*.
-Bronze Age burials are either flat graves or part of a burial mound, and they contain either untreated or cremated human remains.
+Most of them can be distinguished by the criteria *construction of the grave* and *treatment of the body*: Bronze Age burials are either flat graves or part of a burial mound, and they contain either untreated or cremated human remains.
 
-In a nutshell, the classic narrative of the development of these rites in Europe presented e.g.
+The classic narrative of the development of these rites in Europe presented e.g.
 by @hardingEuropeanSocietiesBronze2000a (following @hauslerBestattungssittenFruhenBronzezeit1977; @hauslerGrabUndBestattungssitten1994; @hauslerTotenorientierungUndGeographischer1996) can be summarised as follows: In the early and middle Bronze Age inhumations dominated in different variations.
 Cremation burials in this time only occur in the Hungarian lowlands.
-Burial mounds are concentrated in parts of the Balkans as well as in eastern, western and northern Europe, while in central and southern Europe flat graves -- mostly burials -- predominate.
+Burial mounds are concentrated in parts of the Balkans as well as in Eastern, Western and Northern Europe, while in Central and Southern Europe flat graves -- mostly inhumations -- predominate.
 In the Middle Bronze Age, barrows gained importance in Western and Central Europe.
@@
 How can major misconceptions like this arise and persist in European archaeology?
-Beyond missing data due to preservation and research history, an additional reason might be a lack of unified data repositories and universally acknowledged methods to measure the spatial and temporal distribution of cultural phenomena.
+Beyond missing data due to preservation and research history, another reason might be a lack of unified data repositories and universally acknowledged methods to measure the spatial and temporal distribution of cultural phenomena.
 Instead of directly querying archives with quantitative information, the typical process for the creation of large-scale syntheses is still to manually excerpt the literature of all relevant, local research traditions.
 The present article, in contrast, explores the former strategy.
-The first major question therefore is: Does openly available, bulk radiocarbon data allow reconstructing the main trends within the distribution of burial rites in Bronze Age Europe?
-This approach could set a precedent for a more reproducible and statistically sound research practice.
-If so, then the quantitative comparison of macroregional trends could also pave the way for an analysis of interaction intensity on a long-term continental scale: Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?
+The first major question therefore is: *Does openly available, bulk radiocarbon data allow reconstructing the main trends within the distribution of burial rites in Bronze Age Europe?* This approach could set a precedent for a more reproducible and statistically sound research practice.
+If we could answer this first question with *yes*, then the quantitative comparison of macro regional trends could pave the way for an analysis of interaction intensity on a long-term continental scale: *Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?*
 
@@
 Cultural evolution theory is at least as old as theories of biological evolution and underwent a long and versatile research history.
-Its beginnings are in *Evolutionism* and *Social darwinism* in the 19th and early 20th century, matured in conflict with *Cultural relativism* into *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004].
+Its beginnings are in *Evolutionism* and *Social darwinism* in the 19th and early 20th century and until the nineteen sixties and seventies it matured in conflict with *Cultural relativism* into *Neoevolutionism* [@petermannGeschichteEthnologie2004].
 In the late seventies three major theoretical movements emerged: *Evolutionary psychology*, *Human behavioural ecology* and *Dual inheritance theory* [@smithThreeStylesEvolutionary2000].
-The latter -- heavily popularised by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007]: Just like in biological evolution, also *Social learning* might be prone to selection and mutation.
+The latter -- heavily popularised by Richard Dawkins's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007]: Just like in biological evolution, also *Social learning* might be prone to selection and mutation.
 Under this assumption methods of *Population genetics* could be employed within cultural anthropology to analyse the spread and development of ideas -- of course with the much faster pace of cultural exchange in mind.
-A lot of theoretical and quantitative research follows this line of thought: e.g.
+A lot of theoretical and quantitative research followed this line of thought: e.g.
 @cavalli-sforzaCulturalTransmissionEvolution1981, @boydCultureEvolutionaryProcess1985, @teltserEvolutionaryArchaeologyMethodological1995 or @obrienEvolutionaryArchaeologyTheory1996.
 
-Archaeology has a unique perspective on Cultural evolution [@garveyCurrentPotentialRoles2018]: It is concerned with human behaviour over big areas and long timescales.
-It also works with solid proxies of cultural expression: material remains -- individual lost or discarded artefacts, the ruins of past architecture and all more inconspicuous by-products of living.
+Archaeology has a unique perspective on Cultural evolution as it is concerned with human behaviour over big areas and long time scales [@garveyCurrentPotentialRoles2018].
+It focuses on an abundant and highly diagnostic proxy of cultural expression: material remains.
+These include individual lost or discarded artefacts, the ruins of past architecture and all more inconspicuous by-products of living.
 We rightfully expect humans to learn handcraft as well as building techniques and traditions -- both functionally meaningful and seemingly pointless -- from others close to them.
@@
 
-The workflow suggested by Neiman is relevant to the burial rites data examined in this paper and was an important methodological inspiration: As burial traditions are ritualised behaviour and most of the time heavily linked to religious and spiritual beliefs they can be assumed to be stylistic traits.
+The workflow suggested by Neiman is relevant to the burial rites data examined in this paper and was an important methodological inspiration: As burial traditions are ritualised behaviour which is most of the time heavily linked to religious and spiritual beliefs they can be assumed to be stylistic traits.
 Of course in reality, beyond the distinction of @dunnell1978style, most behaviour is not unambiguously stylistic or functional.
 Burial rites cause costs regarding working time and material to a varying extent.
-This cost might have been influenced or had their effect on the human-nature relation: Pyres, for example, require a significant amount of wood and the erection of barrows much workforce.
+These expenses might have been influenced by or had their effect on the human-nature relation: Pyres, for example, require a significant amount of wood and the erection of barrows much workforce.
 Still, it can be expected that these functional disadvantages were not critical most of the time -- at least regarding the general decision which kind of burial tradition is applied.
@@
 
-In the following sections I will (1) explain how radiocarbon data with relevant meta information can be transformed to macro regional time series of burial traditions, (2) explore how these timeseries compare to traditional archaeological narratives and (3) construct and contextualise cultural distances between macro regions over time in relation to spatial distance.
+In the following sections I will (1) *explain how radiocarbon data with relevant meta information can be transformed to macro regional time series of burial traditions*, (2) *explore how these time series compare to traditional archaeological narratives* and (3) *construct and contextualise cultural distances between macro regions over time in relation to spatial distance*.
 For the latter a simulation based on the @neimanStylisticVariationEvolutionary1995 template was employed.

### 2019-01-26 article/02_Materials_and_Methods.Rmd
language in materials&methods

@@
 The rectangular frame indicates the research area.
-The Map projection is EPSG:102013 and the base layer (land, rivers, lakes) data is taken from the Natural Earth project (\url{www.naturalearthdata.com}).}
+The Map projection is EPSG:102013 and the base layer (land, rivers, lakes) data is taken from Natural Earth project (\url{www.naturalearthdata.com}).}
 \end{figure}
@@
 The dataset used for the analysis in this paper is a subset of the openly available database Radon-B [@kneiselRadonB2013].
-It contains ^14^C dates from the European Bronze and Early Iron Age (2300 BC to 500 BC).
-Radiocarbon dating is a crucial absolute dating technique for this period [@ramseyRadiocarbonDatingRevolutions2008], and Radon-B is a collection of more than 11000 published dates from around 2800 sites from all over Europe (as of December 2018).
+It contains ^14^C dates from the European Bronze and Early Iron Age (approximately 2300 BC to 500 BC).
+Radiocarbon dating is a crucial absolute dating technique for this period [@ramseyRadiocarbonDatingRevolutions2008], and Radon-B is a collection of more than 11000 published dates from around 2800 sites located all over Europe (as of December 2018).
 Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the type and cultural attribution of the sampled context.

### 2019-01-26 article/02_Materials_and_Methods.Rmd
language in materials&methods

@@
 Please see the code for details of implementation.
-The value in brackets indicates the number of dates still available after the respective filtering operation.
+The value in brackets indicates the number of dates or graves still available after the respective filtering operation.
 
@@
 7.
-Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial data distribution (Figure \ref{map_graves}).
+Manually defining a rectangular (in EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial data distribution (Figure \ref{map_graves}).
 
@@
 Transforming data from the meaning *^14^C dates on graves* to *burial type at a certain time*.
-For `r r("bronze15 multi dates one grave", sf_prep)` entries in the thus far assembled subset, the attribution to *site* and *feature* is not unique.
-That means that for many graves multiple (up to 8) ^14^C dates have been created and entered in Radon-B.
-Dates cannot be treated as semantically equal to graves.
-There's a variety of phenomena that cause this effect: In some cases, safety measurements were taken -- multiple dates for one single grave that all fall within the margin of errors of each other.
-In other cases, there is a set of strongly diverging dates for a collective grave that was used for hundreds of years.
-There are also cases where the *feature* information in the database is not sufficient to determine whether two dates are really from the same grave despite the identical entry.
-The following steps were applied to solve the problem automatically.
-This approach is compatible with the intention to create a cultural trait time series as explained below.
+For `r r("bronze15 multi dates one grave", sf_prep)` entries in the thus far assembled subset the attribution to *site* and *feature* is not unique.
+That means that for many graves multiple (up to 8) ^14^C dates have been created (and entered in Radon-B) and dates therefore cannot be treated as semantically equal to graves.
+There is a variety of phenomena that cause this effect: In some cases, safety measurements were taken so that there are multiple dates for one single grave that all fall within the margin of errors of each other.
+In other cases, the set of dates for one grave strongly diverge because they cover the life cycle of a collective grave that was used for hundreds of years.
+There are also cases where the *feature* information in Radon-B is not sufficient to determine whether two dates are really from the same grave despite the identical entry.
+The following steps were applied to solve this problem algorithmically.
+The approach is compatible with the intention to create a cultural trait time series as explained below.
 (`r r("size bronze17", sf_prep)` graves)
@@
 The regions have a radius of 240km and overlap slightly (Figure \ref{map_regions} A).
-That means some graves are counted towards not just one but two regions.
+That means some graves are counted towards not only one but two regions.
 
@@
 \caption{\label{map_regions}Circular, artificial regions (\textbf{A}) and spatial distance network with classified distance values (\textbf{B}).
-The map layout is the same as in Figure \ref{map_graves}.}
+The map layout is similar to Figure \ref{map_graves} but instead of rivers modern country borders are drawn.}
 \end{figure}
@@
 The graves are linked to `r r("dpr period amount", sf_desc)` different *period* terms and `r r("dpr culture amount", sf_desc)` different archaeological *cultures* as they are distinguished in Radon-B.
-`r r("dpr material bone amount", sf_desc)` dates were measured on sample material of bones and teeth (at least `r r("dpr material cremated bones amount", sf_desc)` of them burnt), `r r("dpr material charcoal wood amount", sf_desc)` on wood and charcoal, a small part (`r r("dpr material other amount", sf_desc)`) of other materials such as nuts, resin or pitch.
+`r r("dpr material bone amount", sf_desc)` dates were measured on sample material of bones and teeth (at least `r r("dpr material cremated bones amount", sf_desc)` of them burnt), `r r("dpr material charcoal wood amount", sf_desc)` on wood and charcoal, a small part (`r r("dpr material other amount", sf_desc)`) on other materials such as nuts, resin or pitch.
 For the remaining `r r("dpr material unknown amount", sf_desc)` dates, no material information is stored.

### 2019-01-26 article/02_Materials_and_Methods.Rmd
language in materials&methods

@@
 This data for individual graves can be transformed to a time series of trait prevalence within the respective regions.
-If a certain kind of grave can be attributed to a fuzzy point in time with radiocarbon dating, then we also know that most likely a ceremony was held around this time where the particular burial tradition was practised.
+If a certain kind of grave can be attributed to a (fuzzy) point in time with radiocarbon dating, then we also know that most likely a ceremony was held around this time where the particular burial tradition was practised.
 That means that the cultural trait behind this behaviour must also have been present in one form or the other.

### 2019-01-26 article/02_Materials_and_Methods.Rmd
language in materials&methods

@@
 \caption{\label{sed_region_matrix}Plot matrix of the region-region squared Euclidean distance time series.
-In the top right half for the variable \textit{burial type} and in the bottom left for the variable \textit{burial construction}.
-On the x-axis the time is displayed, on the y-axis the squared Euclidean distance between the two relevant regions.
+In the top right half with solid lines for the variable \textit{burial type} and in the bottom left with dashed lines for the variable \textit{burial construction}.
+The x-axis displays time, the y-axis the squared Euclidean distance between two relevant regions.
 The faint line shows the actual data, the stronger line an approximation (LOESS with $span = 0.3$).}
@@
 To test for correlation between the distances based on the variables *burial type* and *burial construction*, the time frame of interest was divided into sections of 200 years.
-For each section of each the two variables, a matrix of the mean SED over time was calculated.
+For each of the $1400/200 = 7$ section and each of the two variables, a matrix of the mean SED over time was calculated.
 The two matrices of each section can be tested for correlation with the Mantel test, which uses permutations of allegedly correlating matrices to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967].
@@
 For the eight regions, four different categories of spatial distance were defined (Figure \ref{map_regions} B).
-The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the difference in the data level of measurements (Table \ref{tab:distanceandcorrelationtable*} AB).
+The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the different data levels of the measurements (Table \ref{tab:distanceandcorrelationtable*} AB).
 
@@
   caption = "Mean distance, correlation and mantel test results for cultural and spatial distance.
-Listed are the the overall mean squared euclidian distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bt)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B).
-(\\textbf{C}) shows the Mantel test results for the two variables if compared directly with each other.
+Listed are the the overall mean squared Euclidian distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bc)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B).
+\\textbf{C} shows the Mantel test results for the two variables if compared directly.
 $p$-values below $0.1$ are printed bold.",

### 2019-01-26 article/02_Materials_and_Methods.Rmd
language in materials&methods

@@
 
-The distribution of the $k$ variants within the groups can be observed for every time step and documented in time series just like the ones for the real-world regions data (Figure \ref{development_burial_type} B and \ref{development_burial_construction} B).
+The distribution of the $k$ variants within the groups can be observed for every time step and documented in time series like the ones for the real-world regions data (Figure \ref{development_burial_type} B and \ref{development_burial_construction} B).
 These time series can as well be used to calculate time series of region-region distance, which again can be split up into time blocks of 200 years to calculate mean distance matrices.
@@
 The scatter plot points of one simulation run are connected with a faint line from one time slot to the next.
-To compare the runs with the real world observations, the results for the burial customs data (Table \ref{tab:distanceandcorrelationtable*} AB) is added here with with a stronger solid (\textit{burial type}) and dashed (\textit{burial construction}) line.
+To compare the runs with the real world observations, the results for the burial customs data (Table \ref{tab:distanceandcorrelationtable*} AB) is added here with a stronger solid (\textit{burial type}) and dashed (\textit{burial construction}) line.
 The strong horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}

### 2019-01-26 article/03_Results.Rmd
language in results

@@
 In the course of the Bronze Age in Europe there was a trend away from body burial towards cremation.
-Around 2200calBC cremation burials were almost unknown in Poland, southern Germany, northeastern France and northern Germany.
+Around 2200calBC cremation burials were almost unknown in Poland, Southern Germany, Northeastern France and Northern Germany.
 In the Northwest, in Great Britain and the Benelux area, and in the Southeast, in Austria and the Czech Republic, however, there were early cremation traditions.
 These contexts could be discussed as areas of origin of the later omnipresent phenomenon.
-While cremation burials lost importance in Austria and the Czech Republic as well as in the Benelux region during the Early Bronze Age and only regained their importance in the Middle Bronze Age, their share increased steadily in England.
+While cremation burials lost importance in Austria and the Czech Republic as well as in the Benelux region during the Early Bronze Age and only regained relevance there in the Middle Bronze Age, their share increased steadily in England.
 Inhumation and cremation were balanced here for a long time.
-The situation was similar in southern Scandinavia, where the proportion of cremation burials was much lower until the Late Bronze Age.
+The situation was similar in Southern Scandinavia, where the proportion of cremation burials was much lower until the Late Bronze Age.
 In Poland and northern Germany, cremation became relatively suddenly and completely established with the beginning of the Late Bronze Age.
-In northeastern France and northern Germany this change took place already in the Middle Bronze Age.
-In both contexts, inhumation continued to play an important role.
+In Northeastern France and Northern Germany this change took place already in the Middle Bronze Age.
+In both contexts inhumation continued to play an important role.
 
-With regard to the question of grave mound construction, the picture is considerably more heterogeneous.
-In Austria and the Czech Republic, flat graves were the dominant form of burial until the Late Bronze Age, but mounds became more relevant from the Middle Bronze Age onwards.
-In Poland, flat and hill graves remained in balance until the Middle Bronze Age, when flat graves prevailed.
-The amount of data from southern Germany and northeastern France is extraordinarily small: If one believes the sample, then a sudden, radical change from burial in flat graves to hill graves took place in southern Germany at the beginning of the Middle Bronze Age.
-In northeastern France there were no flat graves in the Bronze Age.
+With regard to the question of grave mound construction the picture is considerably more heterogeneous.
+In Austria and the Czech Republic flat graves were the dominant form of burial until the Late Bronze Age, but mounds became more relevant from the Middle Bronze Age onwards.
+In Poland mounds and flat graves remained in balance until the Middle Bronze Age, when the latter finally prevailed.
+The amount of data from Southern Germany and Northeastern France is extraordinarily small: If one believes the sample, then a sudden, radical change from burial in flat graves to hill graves took place in Southern Germany at the beginning of the Middle Bronze Age.
+In Northeastern France there were no flat graves in the Bronze Age.
 Northern Germany underwent a development from the dominance of flat graves in the Early Bronze Age, a short phase of increasing mound construction in the Middle Bronze Age followed by renewed dominance of the flat grave custom in the Late Bronze Age.
-A similar development can be seen in southern Scandinavia: Flat graves clearly predominated, but were increasingly -- almost completely -- replaced by hills, until burial in flat graves became more frequent again at the end of the Bronze Age.
+A similar development can be observed in Southern Scandinavia: Flat graves clearly predominated, but were increasingly -- almost completely -- replaced by mounds, until burial in flat graves became more frequent again at the end of the Bronze Age.
 Burial mounds were dominant throughout the Benelux region, but in the Early and Late Bronze Ages flat graves also occurred to a significant extent.
-In England, flat graves were a short phenomenon in the Early Bronze Age, which did not recur later.
+In England flat graves were a short phenomenon in the Early Bronze Age, which did not recur later.
 \newline
@@
 F).
-Counterintuitively the fluctuation of the correlation is also here very high (Figure \ref{simulation} E vs.
+Counter-intuitively the fluctuation of the correlation is also here very high (Figure \ref{simulation} E vs.
 F).

### 2019-01-26 article/03_Results.Rmd
language in discussion

@@
 The majority of the curves in its top right half for *burial type* correspond to expectations: At the beginning of the Bronze Age, inhumation is widespread throughout the study area, so the Euclidean distance at 2200calBC is small in almost every region-region relationship.
-Only the Benelux region, where the share of cremation is already high in the Early Bronze Age, is excluded.
+Only the Benelux countries, where the share of cremation is already high in the Early Bronze Age, are excluded.
 From the Middle Bronze Age onwards, but especially at the transition to the Late Bronze Age, the transition to the dominance of cremation took place all over Europe.
@@
 The degree of agreement between two regions in the question of inhumation and cremation is therefore not suitable for predicting the similarity of developments in these regions concerning the dominance of flat or mound graves.
-Only for the time window at the transition to the Middle Bronze Age from 1800 to 1600calBC is the 5% significance level of the Mantle test almost undercut, thus suggesting a statistically significant relationship.
+Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600calBC the 5% significance level of the Mantle test is almost undercut, thus suggesting a statistically significant relationship.
 
@@
 Looking at the SED network of *burial construction* (Table \ref{tab:distanceandcorrelationtable*}), similar conclusions can be drawn as in the *burial type* context: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated.
-As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation.
-It is only significant for the time slot between 1800 and 1400calBC.
+As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation, though it is only significant for the time slot between 1800 and 1400calBC.
 There is no time period with negative mean correlation for *burial construction* cultural distance and spatial distance, but also here the correlation coefficient approaches zero in the Late Bronze Age.
@@
 F).
-Counter-intuitively the fluctuation of the correlation is also here very high (Figure \ref{simulation} E vs.
+Counter-intuitively the fluctuation of the correlation is also very high here (Figure \ref{simulation} E vs.
 F).

### 2019-01-26 article/04_Discussion.Rmd
language in discussion

@@
 One major question asked in this paper was, if openly available, bulk radiocarbon data does allow the reconstruction of main trends within the distribution of burial rites in Bronze Age Europe.
-This question can be answered with Yes.
+This is the case.
 The narrative in section \ref{time-series-results} is surprisingly similar to the one presented by @hardingEuropeanSocietiesBronze2000a.
@@
 Despite the general success of this methodological approach, unfortunately the amount and quality of data is still not sufficient for a comprehensive picture.
-Please see the discussion of data quality in section \ref{data-quality-discussion}.
+The reasons for this lack of data quality are discussed in section \ref{data-quality-discussion}.
 \newline
@@
 `r a("Southeastern Central Europe")`^[Amount of graves in Radon-B for this region.]:
-The artificial circular region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map_regions} A).
+The artificial circular region *Southeastern Central Europe* comprises the eastern part of Austria (Upper Austria, Lower Austria, Styria, Burgenland), the northwest of Hungary, Western Slovakia, Moravia and large parts of Bohemia (Figure \ref{map_regions} A).
 The density of ^14^C data on graves in Radon-B is low -- dates are available from the various regions, but large areas are without evidence.
@@
 The region intersects diverse geographical and cultural areas complicating the comparison with archaeological observations.
-In the Southeast of Central Europe, cremation and inhumation already existed side by side in the Early Bronze Age, and the local tradition remained stable in this respect also due to the expansion of the burial mound tradition in the Middle Bronze Age -- with a slight overall increase in cremation graves.
+In the southeast of Central Europe, cremation and inhumation already existed side by side in the Early Bronze Age, and the local tradition remained stable in this respect also due to the expansion of the burial mound tradition in the Middle Bronze Age -- with a slight overall increase in cremation graves.
 The Urnfield period brought the transition to the absolute dominance of cremation burial.
 The development of the *burial type* variable in the Radon-B proxy shows this development correctly.
-The information from ^14^C data on *burial construction* correctly traces the relevance of flat burials in the Early Bronze Age but then overemphasises them in the Middle Bronze Age compared to the construction of burial mounds.
-In the Late Bronze Age, burials on flat burial grounds were frequent, although burial mounds of the Tumulus Culture were also further documented.
+The information from ^14^C data on *burial construction* also correctly traces the relevance of flat burials in the Early Bronze Age but then overemphasises them in the Middle Bronze Age compared to the construction of burial mounds.
+In the Late Bronze Age, burials on flat burial grounds were frequent, although burial mounds of the Tumulus Culture were still present and in use.
 This impact does not become clear in the Radon-B data -- indeed due to the small number of dates [@lubosCzechLandsAustria2013; @markovaSlovakiaHungary2013].
@@
 However, the ^14^C data from this region are almost exclusively from the voivodships of Kajuwien-Pommern, Lodsch and Lower Silesia.
-The amount of data is small overall, with significantly more data from the Early Bronze Age up to about 1700calBC.
+The amount of data is small, with significantly more data from the Early Bronze Age up to about 1700calBC.
 Almost no information is available from the Middle and Late Bronze Ages.
-This lack of data applies to both the *burial type* and the *burial construction* variables, the former being slightly better documented.
+This lack of dates applies to both the *burial type* and the *burial construction* variables, the former being slightly better documented.
 The relative development of *burial type* approximates the *real* conditions -- i.e.
@@
 `r a("Southern Germany")`:
-The artificial district region *Southern Germany* covers Bavaria completely but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and western Austria.
-No ^14^C data are available from the latter regions, the majority of the available data comes from Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative district of Tübingen.
-For the Early Bronze Age, Radon-B preserves a remarkable amount of data from southern Germany, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves.
-The *burial type* variable is quite well understood -- although disproportionately much more of the early inhumations were recorded than Urnfield burials.
+The artificial region *Southern Germany* covers Bavaria but also includes most of Baden-Württemberg and Thuringia as well as extensive areas in Moravia and western Austria.
+No ^14^C dates are available from the latter regions, the majority of the available data comes from Upper Palatinate, Lower Bavaria and Swabia as well as (in Württemberg) the administrative district of Tübingen.
+For the Early Bronze Age, Radon-B preserves a remarkable amount of data from Southern Germany, but this amount decreases in the course of the Middle Bronze Age and leaves the Late Bronze Age almost without dated graves.
+The *burial type* variable is well understood -- although disproportionately much more of the early inhumations were recorded than Urnfield burials.
 This wealth of information contrasts sharply with the situation regarding the *burial construction* variable, for which almost no information from this area is available throughout the Bronze Age.
@@
 Inhumation dominated for a long time but was then replaced by cremation from the Middle Bronze Age onwards and massively in the Late Bronze Age.
-The calculated proportions for *burial construction* are based on a negligibly small amount of data, but the proxy still indicates the right tendency for early and middle Bronze Age: Flat graves replaced mounds.
+The calculated proportions for *burial construction* are based on a negligibly small amount of data, but the proxy still indicates the right tendency for the Early and Middle Bronze Age: Flat graves replaced mounds.
 The dataset does not allow any statements about the Late Bronze Age [@falkensteinWandelBestattungssittenHugelgraber2017; @jockenhovelGermanyBronzeAge2013].
@@
 Accumulations of ^14^C data are only available from Württemberg and towards the Paris Basin.
-The amount of data is low overall, but very low in the early and early Middle Bronze Age.
+The amount of data is low overall, but very low in the Early and Middle Bronze Age.
 For *burial type* limited statements are possible after this time window, *burial construction*, on the other hand, must remain obscure.
-The prediction of the relative development of these variables is correspondingly error-prone.
-Still, the parallel existence of cremation and inhumation in the Early and Middle Bronze Age is correctly reflected.
-However, this situation already existed in the Early Bronze Age and indeed developed more nuanced than the proxy might suggest.
-Contrary to the derivation from the ^14^C ratios, flat graves were quite common in northeastern France [@mordantBronzeAgeFrance2013].
+The prediction of the relative development of these variables is error-prone: The parallel existence of cremation and inhumation in the Early and Middle Bronze Age is correctly reflected, but this situation already existed in the Early Bronze Age and indeed developed more nuanced than the proxy might suggest.
+Contrary to the derivation from the ^14^C ratios, flat graves were quite common in Northeastern France [@mordantBronzeAgeFrance2013].
 
@@
 Also from the area north of Bremen some data are documented in Radon-B.
-The database does not contain as many dated graves from any other region as from northern Germany.
+The database does not contain as many dated graves from any other region as from Northern Germany.
 Most of them date back to the Early Bronze Age, while the Middle Bronze Age from about 1800calBC to 1400calBC seems barren.
@@
 The meta information on *burial type* and *burial construction* trace this development to the same extent.
-Since in Northern Germany influences from all four cardinal directions come together and contribute significantly to the formation of local cultural phenomena, it is difficult to estimate the overall development of the primary variables here as well.
+Since in Northern Germany influences from all four cardinal directions come together and contribute significantly to the formation of local cultural phenomena, it is difficult to estimate the overall development of the primary variables.
 The ^14^C proxy correctly represents *burial site* insofar as inhumations dominated in the Early Bronze Age and were later on replaced by cremations especially in the North from the Middle Bronze Age onwards.
 In the Late Bronze Age though, cremation was more dominant than the dataset suggests.
-Concerning *burial construction*, the current Radon-B stock records the fundamental dominance of flat graves well but attributes more relevance to them in the Early and Middle Bronze Age than is usual in the archaeological assessment of the context [@hofmannRituelleUmgangMit2008; @jockenhovelGermanyBronzeAge2013].
+Concerning *burial construction*, the current Radon-B stock records the fundamental dominance of flat graves well, but attributes more relevance to them in the Early and Middle Bronze Age than it is usual in the archaeological assessment of the context [@hofmannRituelleUmgangMit2008; @jockenhovelGermanyBronzeAge2013].
 
 `r a("Southern Scandinavia")`:
-The region *Southern Scandinavia* includes all of Denmark with Jutland, Funen, Zeeland and all smaller islands as well as large parts of east of southern Sweden, and coastal regions in Schleswig-Holstein and Mecklenburg-Western Pomerania.
+The region *Southern Scandinavia* includes all of Denmark with Jutland, Funen, Zeeland and all smaller islands as well as large parts of East of Southern Sweden, and coastal regions in Schleswig-Holstein and Mecklenburg-Western Pomerania.
 The ^14^C data are very well distributed across Jutland, Funen and the Swedish Skåne region.
-Since the Bronze Age in Scandinavia begins much later than in Central Europe and is measured by another chronological system, the diachronic developments cannot merely be compared through the tripartite structure of the Early, Middle and Late Bronze Ages.
+Since the Bronze Age in Scandinavia begins later than in Central Europe and is measured by another chronological system, the diachronic developments cannot merely be compared through the tripartite structure of the Early, Middle and Late Bronze Ages.
 The period which is labelled Early Bronze Age in Central Europe corresponds to the Scandinavian Late Neolithic.
@@
 In the Scandinavian Early Bronze Age burials in burial mounds dominated.
-In the Late Bronze Age, flat burial fields began to occur [@thraneScandinavia2013].
+In the Late Bronze Age, flat grave fields began to occur [@thraneScandinavia2013].
 
 `r a("Benelux")`:
-The Benelux region includes the Netherlands, northern Belgium and large parts of Lower Saxony and North Rhine-Westphalia.
+The Benelux region includes the Netherlands, Northern Belgium and large parts of Lower Saxony and North Rhine-Westphalia.
 All these regions have ^14^C data on graves in Radon-B, but the data density in Belgium and the Netherlands is much higher.
@@
 As in the *Southern Germany* region, there is a serious difference between the information density of both variables.
-If one considers the relative development of the variables as they can be derived from the burial data in comparison to the archaeological state of research, there is both deviations as well as agreement.
-Here, too, the study region includes different cultural phenomena with highly different burial customs: In the north and east of the Benelux region, inhumations were the rule until the Late Bronze Age, while cremation dominated earlier south of the Meuse.
+If one considers the relative development of the variables as they can be derived from the burial data in comparison to the archaeological state of research, there is both deviation as well as agreement.
+Here, too, the study region includes different cultural phenomena with highly different burial customs: In the north and east of the Benelux region inhumations were the rule until the Late Bronze Age, while cremation dominated earlier south of the Meuse.
 The Radon-B proxy seems to overemphasise the dominance of cremation in the whole area.
@@
 It also touches a narrow coastal strip on the mainland, from which no ^14^C data are documented.
-The available data are concentrated in the Southeast of Britain but are well distributed there.
-The diachronic data availability is relatively good in the early and middle Bronze Age but then decreases in the late Bronze Age.
+The available data are concentrated in the southeast of Britain but are well distributed there.
+The diachronic data availability is relatively good in the Early and Middle Bronze Age but then decreases in the Late Bronze Age.
 *burial type* and *burial construction* behave accordingly.
@@
 However, the dataset emphasises the occurrence of cremation graves too much.
-Concerning the variables *burial construction*, the dominance of burial mounds corresponds perfectly to the archaeological narrative, but -- especially in the Late Bronze Age -- flat grave fields also occurred, which the data set wrongly excludes [@robertsBritainIrelandBronze2013].
+Concerning the variables *burial construction*, the dominance of burial mounds corresponds perfectly to the archaeological narrative, but -- especially in the Late Bronze Age -- flat cemeteries also occurred, which the data set wrongly excludes [@robertsBritainIrelandBronze2013].
 

### 2019-01-27 article/04_Discussion.Rmd
language in discussion

@@
 
-One major question asked in this paper was, if openly available, bulk radiocarbon data does allow the reconstruction of main trends within the distribution of burial rites in Bronze Age Europe.
+One major question asked in this paper was, *if openly available, bulk radiocarbon data does allow the reconstruction of main trends within the distribution of burial rites in Bronze Age Europe*.
 This is the case.
@@
 
-The second important question posed for this paper was, if these time series constructed from radiocarbon data (Figure \ref{development_burial_type} and \ref{development_burial_construction}) allow further analysis concerning cultural and spatial distance?
-While the idea behind the first question is rather technical -- doing something that's already available in a quantitative and reproducible way -- the second one has a larger scope.
-A quantitative representation of an archaeological narrative should path the way for a more diverse set of methods to be applied to this data.
-Cultural distance is one of the first topics that comes to mind.
+The second important question posed for this paper was, *if these time series constructed from radiocarbon data* (Figure \ref{development_burial_type} and \ref{development_burial_construction}) *allow further analysis concerning cultural and spatial distance*?
+While the idea behind the first question is rather technical -- reproducing already available information in a quantitative way -- the second one has a larger scope.
+A quantitative representation of an archaeological narrative should path the way for a more diverse set of methods to be applied to this data and cultural distance is an integral topic of archaeological research.
 
-The expectation for this question was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* as well in between these variables and the spatial distance.
+The expectation for this question was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* as well in between these variables as with spatial distance.
 The simulation, though with this implementation and parametrisation for sure rather qualitatively than quantitatively, visualises how a high degree of interaction should translate to a significantly higher degree of cultural similarity across human populations (Figure \ref{simulation}).
-And as the European Bronze Age is usually depicted as a period of intensive trade, travel and exchange [e.g.
+And as the European Bronze Age is usually depicted as a period of trade, travel and exchange [e.g.
 @kristiansenRiseBronzeAge2005; @sherrattWhatWouldBronzeAge1993] there was no reason to expect something else than Tobler's famous First Law of Geography: "everything is related to everything else, but near things are more related than distant things." [@toblerComputerMovieSimulating1970].
@@
 1.
-The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} queries the credibility of the cultural distance calculation.
-The macro regions have to be very big due to a lack of data and don't allow for a precise picture.
+The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} question the credibility of the cultural distance calculation.
+The macro regions have to be very big due to a lack of data and don't allow for a more precise picture.
 
@@
 The negative correlation of *burial type* in the Late Bronze Age could indicate long distance contacts that explicitly excluded certain intermediate regions of Central Europe.
-The rise of the Tumulus Culture in the Middle Bronze Age might have cushioned this effect for the variable *burial construction* and at the same time introduced more variability for a less important aspect *burial type*.
+The rise of the Tumulus Culture in the Middle Bronze Age might have cushioned this effect for the variable *burial construction* and introduced more variability for a -- at that time -- less important aspect *burial type*.
 The latter may have been of subordinate relevance in the Tumulus Culture context.
 With the rise of the Urnfield Culture this phenomenon could have been reversed as *burial type* became an expression of something that could have been one of the first formalized religions in Europe: The spread of the Urnfield Culture was an extraordinary development in early European history and was possibly accompanied by a structured, religious dogma that could have built up around the symbolic content of cult wagons, sun barks and waterfowl [@bilicSwanChariotSolar2016; @kimmig_seevolkerbewegung_1964; @molloy_european_2018].
-Of course these explanations could, if accurate, only be considered on a really big scale.
-For individual regions and sites specific ideological, social or even economic considerations might have been much more critical for the adopted burial tradition [@barrett_monumentality_1990; @fokkens_genesis_1997-1; @muller-scheesel_variabilitat_2009].
+Of course these explanations could, if correct, only be considered on a big scale.
+For individual regions and sites specific ideological, social or even economic considerations might have been much more critical for the decision which burial tradition to adopt [@barrett_monumentality_1990; @fokkens_genesis_1997-1; @muller-scheesel_variabilitat_2009].
 

### 2019-01-27 article/04_Discussion.Rmd
language in discussion

@@
 
-The A sections of the figures \ref{development_burial_type} and \ref{development_burial_construction} with absolute grave numbers per region and year indicate areas and periods with a lot and others with very few points of data.
+The **A** sections of the figures \ref{development_burial_type} and \ref{development_burial_construction} with absolute grave numbers per region and year indicate areas and periods with a lot and others with few points of data.
 For some there is not enough information to make any decision about the relative dominance and for many the general relevance and accuracy of the resulting time series can be questioned.
-There's a variety of reasons for this which should be considered for future data acquisition in Radon-B and beyond:
+There is a variety of reasons for this.
+These should also be considered for future data acquisition in Radon-B and beyond.
 
-1. The data in Radon-B depends on the projects and interests of the involved researchers: The data was collected as needed for individual regions rather than systematically for all of Europe.
+1. The data in Radon-B depend on the projects and interests of the participating researchers: It was collected as needed for individual regions rather than systematically for all of Europe.
 
@@
 3.
-In the field of investigation of this work different research traditions meet, which in the past considered ^14^C data more or less essential for the construction of an absolute chronological assessment.
-In various Metal Age contexts, 14C data are inferior to the accuracy of relative chronological classifications - e.g.
-by primer typology - and are therefore used only sporadically to create absolute chronological fix points.
-Dendrochronological dating is preferred in moist soil contexts and when the necessary woods are available.
-The available ^14^C calibration curves form plateaus in some periods, which significantly limit the possible dating accuracy of the data.
-If such an effect is known, then no data is ordered at all.
+Different research traditions in archaeology considered ^14^C data in the past more or less essential for the reconstruction of absolute chronology.
+In various Metal Age contexts, ^14^C data provide inferior precision in comparison to relative chronological classifications - e.g.
+by fibula typology - and are therefore used only sporadically to create absolute chronological fix points.
+Dendrochronological dating is preferred in waterlogged deposits and whenever the necessary woods are available.
+^14^C calibration curves form plateaus in some periods, which significantly limit the possible dating precision.
+If such an effect is known, then often no dates are commissioned at all.
 
@@
 In many mineral soil contexts no sufficiently good, dateable material is available.
-This situation can affect entire regions, if, for example, poor bone preservation prevails in general as a result of low-calcification soils.
+This situation can affect entire regions, if, for example, poor bone preservation prevails as a result of lime-deficient soils.
 
 5.
-From very few regions and periods of the European Bronze Age so many graves are documented that it can be assumed that a significant sample of the burial culture of the entire population has been researched.
-Instead, it is evident in several contexts that only the burials of individual social groups, gender or ethnic group were recorded: The quantity and nature of the known graves cannot be representative of all the deceased.
-In this case, it can be assumed that further, deviant burial rituals were practised, which left no remains or remains that were difficult to grasp archaeologically.
-These rituals are not shown in the ^14^C data from Radon-B.
+In very few regions and periods of the European Bronze Age the number of excavated graves is sufficiently big to assume that a statistically relevant sample of the burial culture of the entire population is documented.
+Instead, it is evident in most contexts that only the burials of certain social groups, genders or ethnic groups were recorded.
+The quantity and nature of the known graves cannot be representative.
+In this case it can be expected that further, deviant burial rituals were practised, which left no archaeologically visible remains.
+These rituals are invisible in the ^14^C data from Radon-B.
 
 Concerning the primary variables *burial type* and *burial construction*, further special effects must be taken into account: The disturbance and robbery of graves was a widespread phenomenon in history and prehistory.
-Burial mounds were usually attacked more severely than flat graves due to their better visibility and could, therefore, be underrepresented in Radon-B.
-In many places, they were also the subject of archaeological expeditions in the early days of archaeology at the end of the 19th and beginning of the 20th century.
+Burial mounds were usually attacked more severely than flat graves due to their better visibility and could, therefore, be under-represented in Radon-B.
+In many places, they were also the subject of expeditions in the early days of archaeology at the end of the 19th and beginning of the 20th century.
 Of course, these excavations did not yield ^14^C-dates.
@@
 On the other hand, if a burial mound is still present, it is more readily visible and recognised by archaeologists today: Mounds are often clearly identifiable in aerial photographs or digital terrain models, while flat graves are much more difficult to identify for today's remote sensing methods.
-It is difficult to estimate the impact of these effects in detail, but an overrepresentation of mounds seems more likely than an overrepresentation of flat graves.
-The dating of cremation graves on bone remains methodically differs from the dating of body graves because it is based on the extraction of the mineral, not the organic, fraction in the bone [@lantingDatingCrematedBones2001].
+It is difficult to estimate the impact of these effects in detail, but an over-representation of mounds seems more likely than an over-representation of flat graves.
+The dating of cremation graves on bone remains methodically differs from the dating of inhumations because it is based on the extraction of the mineral, not the organic, fraction in the bone [@lantingDatingCrematedBones2001].
 The underlying chemical method was developed later, has only been used more frequently in the past 15 years and is associated with some particular difficulties [@olsenCharacterisationBlindTesting2008; @strydonck14CDatingCremated2009].

### 2019-01-27 article/04_Discussion.Rmd
language in discussion

@@
 
-This paper is based on ^14^C dates extracted from the Radon-B data set which were prepared for the dating of grave findings.
-^14^C data on graves are used to chronologically attach a feature to the absolute chronological context or to better describe the relative development within a finding complex.
-The questions underlying the creation of a ^14^C date are therefore generally not equivalent to those imposed in this paper.
-While here the presence of an idea is to be measured abstractly at a certain place at a certain time, graves are usually dated exactly the other way round to contextualize and better understand them as concrete object.
-By definition, field research is oriented towards small-scale questions that can be answered to their respective extent.
-It must put a lot of energy into doing justice to all findings that are inevitably destroyed by excavation.
-There is a fundamental break between that great, theoretical archaeology, which wants to develop comprehensive narratives and only pay quantitative attention to the individual case, and the individual research project or emergency excavation, which has to dedicate a measure of respect and sometimes a lot of working time to each grave.
+This paper is based on ^14^C dates which were prepared for the dating of graves.
+There is a fundamental break between that great, theoretical archaeology, which wants to develop comprehensive narratives and only pay quantitative attention to the individual case, and the on-site research or emergency excavation, which has to dedicate a measure of respect and sometimes a lot of time to each grave.
 In this thesis, ^14^C data are converted into a proxy of calibrated density distributions within artificial working regions, which is completely detached from the individual context.
+^14^C data on graves are used to attribute a feature to the absolute chronological context or to better describe the relative development within a site.
+The questions underlying the creation of a ^14^C date are therefore not equivalent to those imposed in this paper.
 
@@
 By creating the relative development proxies, the individual grave loses its context.
-The present paper is therefore a clear secondary use of Radon-B.
-A better dataset should be grave-related, not date-related, and should contain much more meta-information to each grave.
+The present paper is a secondary use of Radon-B.
+A better dataset should be grave-related, not date-related, and should contain much more meta-information for each grave.

### 2019-01-27 article/06_Acknowledgements.Rmd
update of Acknowledgements

@@
 I would like to offer my special thanks to my supervisors PD Dr.
-Oliver Nakoinz and Dr.
-Martin Hinz for developing this research idea with me and for ongoing support and advice before, while and after the creation of the master thesis this article is based on.
-With Dr.
+Oliver Nakoinz (University of Kiel) and Dr.
+Martin Hinz (University of Bern) for developing this research idea with me and for ongoing support and advice before, while and after the creation of the master thesis this article is based on.
+I had helpful discussions at different stages of the research process with Dr.
 Jutta Kneisel, Dr.
@@
 Nils Müller-Scheeßel and David Matzig, B.A.
-I had helpful discussions at different stages of the research process.
+(all University of Kiel).
 Beyond these colleagues at Kiel University I am particularly grateful for the feedback, encouragement and literature recommendations by Prof.
 Dr.
-Ben Marwick and the valuable C++ code review by Benny Baumann.
+Ben Marwick (University of Washington) and the valuable C++ code review by Benny Baumann.
+
+Finally I would like to thank Sergi Valverde (Pompeu Fabra University Barcelona) and Simon Carrignon (Barcelona Supercomputing Center) for the organisation of the session *Evolution of Cultural Complexity III* at the Conference on Complex Systems 2018 in Thessaloniki where this paper was initially presented.

### 2019-01-27 article/06_Remarks.Rmd
reorganized the technical stuff at the end

--- 2019-01-27 article/06_Acknowledgements.Rmd
@@
-# Acknowledgements
+# Remarks
+
+## Acknowledgements
 
@@
 
-Finally I would like to thank Sergi Valverde (Pompeu Fabra University Barcelona) and Simon Carrignon (Barcelona Supercomputing Center) for the organisation of the session *Evolution of Cultural Complexity III* at the Conference on Complex Systems 2018 in Thessaloniki where this paper was initially presented.
+Finally I would like to thank Sergi Valverde (Pompeu Fabra University Barcelona) and Simon Carrignon (Barcelona Supercomputing Center) for the organisation of the session *Evolution of Cultural Complexity III* at the Conference on Complex Systems 2018 in Thessaloniki (Greece) where this paper was initially presented.
+
+## Declaration of Conflicting Interests
+
+The author declared no potential conflicts of interest with respect to the research, authorship, and/or publication of this article.
+
+## Funding
+
+This paper was prepared without financial support, except for a 300Euro travel grant to attend the Conference on Complex Systems 2018 by the International Center of Kiel University.

### 2019-01-27 article/05_Conclusion.Rmd
work on a conclusion

# Conclusion {#conclusion}

The major questions posed in the introduction of this paper (section \ref{introduction}) have been answered (section \ref{discussion}): *Bulk radiocarbon data is a powerful source to trace general trends in prehistoric Europe* and *allows long-term observations concerning cultural distance*. Radiocarbon dates can be transformed to meaningful regional time series, if they can be linked to the relevant meta information. Although only `r r("gpr size", sf_desc)` graves of a six-figure statistical population that can be estimated for all burials in the European Bronze Age [@nikulkaArchaologischeDemographieMethoden2016] have been available for this analysis, the resulting reconstruction of general burial tradition trends in wast areas of Central, Northern and North-western Europe comes close to the one achieved by literature research. The time series can be employed for cultural distance calculation and they yield the surprising observation, that cultural distance and spatial distance do not significantly correlate. Although this interpretation must be treated with caution, it seems as if the spread of burial traditions are not simple *Diffusion of Innovation* events as introduced by @rogersDiffusionInnovations1962, but the result of more complex interactions between competing ideas.

Now it is time to ask

For this paper burial traditions were the main object of research, but the same methodological approach could be applied for other aspects of cultural behaviour.

### 2019-01-27 article/05_Conclusion.Rmd
more love for the conclusion
reproducibility is key

@@
 
-Now it is time to ask
+Beyond these intrinsic results, it is important now to ask what this paper means for the study of human behaviour in general. Two possible answers:
 
-For this paper burial traditions were the main object of research, but the same methodological approach could be applied for other aspects of cultural behaviour.
+1.
+For this paper burial traditions were the main object of research, but the same methodological approach could be applied for other aspects of cultural behaviour reconstruction in archaeology.
+The idea to systematically collect data points that combine spatial and temporal location with simplified meta information about cultural phenomena might be a way to achieve better reproducibility of overarching narratives.
+Fast and easy *reproducibility* is of crucial importance to achieve scientific progress and it includes the notion of rather technical *preproducibility* [@starkReproducibilityMustCome2018]: Data (and code) must be shared to fuel advances archaeological theory building [@kitzesPracticeReproducibleResearch2017; @marwickComputationalReproducibilityArchaeological2017].
+
+2. test

### 2019-01-27 article/05_Conclusion.Rmd
more work on the conclusion

@@
 
-The major questions posed in the introduction of this paper (section \ref{introduction}) have been answered (section \ref{discussion}): *Bulk radiocarbon data is a powerful source to trace general trends in prehistoric Europe* and *allows long-term observations concerning cultural distance*.
+The major questions posed in the introduction of this paper have been answered : *Bulk radiocarbon data is a powerful source to trace general trends in prehistoric Europe* and *allows long-term observations concerning cultural distance*.
 Radiocarbon dates can be transformed to meaningful regional time series, if they can be linked to the relevant meta information.
-Although only `r r("gpr size", sf_desc)` graves of a six-figure statistical population that can be estimated for all burials in the European Bronze Age [@nikulkaArchaologischeDemographieMethoden2016] have been available for this analysis, the resulting reconstruction of general burial tradition trends in wast areas of Central, Northern and North-western Europe comes close to the one achieved by literature research.
-The time series can be employed for cultural distance calculation and they yield the surprising observation, that cultural distance and spatial distance do not significantly correlate.
+Although only `r r("gpr size", sf_desc)` graves of a maybe six-figure statistical population that can be estimated for all burials in the European Bronze Age [@nikulkaArchaologischeDemographieMethoden2016] have been available for this analysis, the resulting reconstruction of general burial tradition trends in wast areas of Central, Northern and North-western Europe comes close to the one achieved by literature research.
+The time series can be employed for cultural distance calculation and they yield the surprising observation that cultural distance and spatial distance do not significantly correlate.
 Although this interpretation must be treated with caution, it seems as if the spread of burial traditions are not simple *Diffusion of Innovation* events as introduced by @rogersDiffusionInnovations1962, but the result of more complex interactions between competing ideas.
+
+Future research is required to de-tangle these interactions -- of course on the basis of more precise data.
+This endeavour will not end with the simple description of social learning cascades, but with the analysis of the initial reasons for diversification.
+From the point of view of Cultural Complexity -- the question, which factors accelerate or decelerate the variability of human behaviour [@henrichDemographyCulturalEvolution2004; @mesoudiWhatCumulativeCultural2018; @vaesenPopulationSizeDoes2016] -- the long term fluctuations of dominance of one burial rite over the other are remarkable.
+Its standard explanation models (among others population density and environmental risk) can not be applied to stylistic variables.
 
@@
 The idea to systematically collect data points that combine spatial and temporal location with simplified meta information about cultural phenomena might be a way to achieve better reproducibility of overarching narratives.
-Fast and easy *reproducibility* is of crucial importance to achieve scientific progress and it includes the notion of rather technical *preproducibility* [@starkReproducibilityMustCome2018]: Data (and code) must be shared to fuel advances archaeological theory building [@kitzesPracticeReproducibleResearch2017; @marwickComputationalReproducibilityArchaeological2017].
+Fast and easy *reproducibility* is of crucial importance to achieve scientific progress and it includes the notion of rather technical *preproducibility* [@starkReproducibilityMustCome2018]: Data (and code) must be shared to fuel advances in anthropological theory building [@kitzesPracticeReproducibleResearch2017; @marwickComputationalReproducibilityArchaeological2017].
 
-2. test
+2.
+This paper is built upon Cultural Evolution theory (see section \ref{introduction}).
+It not just relies on terminology and methodology developed in this field, the process of abstraction necessary to go from radiocarbon dates to cultural distance was also heavily influenced by Evolutionary archaeology.
+This theoretical framework might also be a source of inspiration and methods for other projects.
+The adoption and adaption of statistical tools and models from population genetics into anthropology was and is a feature of Cultural Evolution theory [@creanzaCulturalEvolutionaryTheory2017].

### 2019-01-27 article/index.Rmd
Acknowledgements
reorganized the technical stuff at the end
update of affiliation

@@
 - num: 1
-  org: Institute of Pre- and Protohistoric Archaeology, Kiel University, Kiel, Germany
-corrauth: "Clemens Schmid, Institut für Ur- und Frühgeschichte, Christian-Albrechts-Universität, D - 24098 Kiel."
+  org: Römisch-Germanisches Zentralmuseum, Leibniz-Forschungsinstitut für Archäologie, Mainz, Germany
+corrauth: Clemens Schmid, Römisch-Germanisches Zentralmuseum, Leibniz-Forschungsinstitut für Archäologie, D - 55116 Mainz.
 email: clemens@nevrome.de
-abstract: "Petersilie."
+abstract: Petersilie.
 keywords: Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission
@@
 ```
+
+```{r child = '06_Remarks.Rmd'}
+```

### 2019-01-27 article/index.Rmd
abstract primer

@@
 email: clemens@nevrome.de
-abstract: Petersilie.
+abstract: European Bronze age archaeology traditionally focusses on two major dimensions to categorise burials - although there's an immense variability of attendant phenomena within this spectrum -- Flat graves versus burial mounds and cremation versus inhumation.
+These traits are an indispensable ingredient for common narratives of sociocultural interaction in the Bronze age.
+This complex system of ideological affiliation and exchange can be described in the terms of Cultural Evolution theory.
+Burial rites are extraordinary cultural traits -- Following Dunnels [1] distinction between function and style based on relevance for selection, they can be handled as neutral variants.
+As demonstrated by Neiman [2], drift and intergroup transmission as opposed to natural selection should therefore be the decisive processes for their expansion.
+On the other hand funerals touch upon personal loss and profound religious beliefs -- They are not short-lived fashion and most probably well interlinked with other - many archaeologically inaccessible - traits.
+This paper will present the results of my currently ongoing master thesis.
+To study the diffusion of burial rites, I employ the dataset RADON-B [3] which contains more then two thousand Bronze age 14C dates of graves from Central, Northern and North-western Europe.
+Based on this information I construct regional time series that document how rituals change.
+For a better understanding of the real-world phenomena I implement a computational model in R and C++.
+It simulates the expansion of mutually exclusive ideas in an artificial population graph and provides an environment to explore the effects of parameters like group size or the degree of intergroup idea transmission.
 keywords: Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission

### 2019-01-28 article/05_Conclusion.Rmd
small changes in the conclusion

@@
 
-The major questions posed in the introduction of this paper have been answered : *Bulk radiocarbon data is a powerful source to trace general trends in prehistoric Europe* and *allows long-term observations concerning cultural distance*.
+The major questions posed in the introduction to this paper have been answered : *Bulk radiocarbon data is a powerful source to trace general trends in prehistoric Europe* and *allows long-term observations concerning cultural distance*.
 Radiocarbon dates can be transformed to meaningful regional time series, if they can be linked to the relevant meta information.
 Although only `r r("gpr size", sf_desc)` graves of a maybe six-figure statistical population that can be estimated for all burials in the European Bronze Age [@nikulkaArchaologischeDemographieMethoden2016] have been available for this analysis, the resulting reconstruction of general burial tradition trends in wast areas of Central, Northern and North-western Europe comes close to the one achieved by literature research.
-The time series can be employed for cultural distance calculation and they yield the surprising observation that cultural distance and spatial distance do not significantly correlate.
+The time series can be employed for cultural distance calculation.
+They yield the surprising observation that cultural distance and spatial distance do not significantly correlate.
 Although this interpretation must be treated with caution, it seems as if the spread of burial traditions are not simple *Diffusion of Innovation* events as introduced by @rogersDiffusionInnovations1962, but the result of more complex interactions between competing ideas.
@@
 This endeavour will not end with the simple description of social learning cascades, but with the analysis of the initial reasons for diversification.
-From the point of view of Cultural Complexity -- the question, which factors accelerate or decelerate the variability of human behaviour [@henrichDemographyCulturalEvolution2004; @mesoudiWhatCumulativeCultural2018; @vaesenPopulationSizeDoes2016] -- the long term fluctuations of dominance of one burial rite over the other are remarkable.
-Its standard explanation models (among others population density and environmental risk) can not be applied to stylistic variables.
+From the point of view of Cultural Complexity -- the question, which factors accelerate or decelerate the variability of human behaviour [@henrichDemographyCulturalEvolution2004; @mesoudiWhatCumulativeCultural2018; @vaesenPopulationSizeDoes2016] -- the long term dominance fluctuations of one burial rite over the other are remarkable.
+Its standard explanation models (among others population density and environmental risk) can hardly be applied to stylistic variables.
 
-Beyond these intrinsic results, it is important now to ask what this paper means for the study of human behaviour in general. Two possible answers:
+Beyond these intrinsic results, it is also important to ask what this paper means for the study of human behaviour in general. Two possible answers:
 
 1.
-For this paper burial traditions were the main object of research, but the same methodological approach could be applied for other aspects of cultural behaviour reconstruction in archaeology.
+For this paper burial traditions were the main object of research, but the same methodological approach could be applied to reconstruct other aspects of cultural behaviour.
 The idea to systematically collect data points that combine spatial and temporal location with simplified meta information about cultural phenomena might be a way to achieve better reproducibility of overarching narratives.
@@
 It not just relies on terminology and methodology developed in this field, the process of abstraction necessary to go from radiocarbon dates to cultural distance was also heavily influenced by Evolutionary archaeology.
-This theoretical framework might also be a source of inspiration and methods for other projects.
-The adoption and adaption of statistical tools and models from population genetics into anthropology was and is a feature of Cultural Evolution theory [@creanzaCulturalEvolutionaryTheory2017].
+This theoretical framework might be a source of inspiration and methods for other projects.
+The adoption and adaption of statistical tools and models from population genetics into anthropology was and is a unique characteristic of Cultural Evolution theory [@creanzaCulturalEvolutionaryTheory2017].

### 2019-02-02 article/03_Results.Rmd
more consistent spacing before "calBC" (@yesdavid)

@@
 In the course of the Bronze Age in Europe there was a trend away from body burial towards cremation.
-Around 2200calBC cremation burials were almost unknown in Poland, Southern Germany, Northeastern France and Northern Germany.
+Around 2200 calBC cremation burials were almost unknown in Poland, Southern Germany, Northeastern France and Northern Germany.
 In the Northwest, in Great Britain and the Benelux area, and in the Southeast, in Austria and the Czech Republic, however, there were early cremation traditions.
@@
 Beyond the simple comparison of archaeological knowledge and information derived from Radon-B, Figure \ref{sed_region_matrix} shows the time series of squared Euclidean distance for each region-region relationship.
-The majority of the curves in its top right half for *burial type* correspond to expectations: At the beginning of the Bronze Age, inhumation is widespread throughout the study area, so the Euclidean distance at 2200calBC is small in almost every region-region relationship.
+The majority of the curves in its top right half for *burial type* correspond to expectations: At the beginning of the Bronze Age, inhumation is widespread throughout the study area, so the Euclidean distance at 2200 calBC is small in almost every region-region relationship.
 Only the Benelux countries, where the share of cremation is already high in the Early Bronze Age, are excluded.
@@
 The degree of agreement between two regions in the question of inhumation and cremation is therefore not suitable for predicting the similarity of developments in these regions concerning the dominance of flat or mound graves.
-Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600calBC the 5% significance level of the Mantle test is almost undercut, thus suggesting a statistically significant relationship.
+Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600 calBC the 5% significance level of the Mantle test is almost undercut, thus suggesting a statistically significant relationship.
 
 For the relationship between the SED network based on *burial type* and the spatial distance several observations result from Table \ref{tab:distanceandcorrelationtable*}: Cultural distance calculated from these to variables does not seem to correlate with spatial distance.
-A significant correlation between the distance measures only reveals itself for the Early Bronze Age time slot 2000 to 1800calBC: Only in the Early Bronze Age, long before the universal switch to the Urnfield burial tradition, large regions directly adjacent to each other tend to undergo the same developments rather than the ones further away from each other.
+A significant correlation between the distance measures only reveals itself for the Early Bronze Age time slot 2000 to 1800 calBC: Only in the Early Bronze Age, long before the universal switch to the Urnfield burial tradition, large regions directly adjacent to each other tend to undergo the same developments rather than the ones further away from each other.
 In the Middle Bronze Age the SED developments of the *burial type* variable undergo the peaks already observed in Figure \ref{sed_region_matrix}.
-The correlation between cultural and spatial distance over these peaks is low and even tilts between 1400 and 1200calBC, at the transition to the Late Bronze Age, into the negative -- although not significantly.
+The correlation between cultural and spatial distance over these peaks is low and even tilts between 1400 and 1200 calBC, at the transition to the Late Bronze Age, into the negative -- although not significantly.
 The cultural distance between several geographically distant regions is then smaller than between several directly adjacent regions.
 Looking at the SED network of *burial construction* (Table \ref{tab:distanceandcorrelationtable*}), similar conclusions can be drawn as in the *burial type* context: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated.
-As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation, though it is only significant for the time slot between 1800 and 1400calBC.
+As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation, though it is only significant for the time slot between 1800 and 1400 calBC.
 There is no time period with negative mean correlation for *burial construction* cultural distance and spatial distance, but also here the correlation coefficient approaches zero in the Late Bronze Age.
@@
 F).
-That's mostly because of the relatively strong negative correlation of cultural and spatial distance in the Middle and Late Bronze Age (1400 to 1000calBC).
+That's mostly because of the relatively strong negative correlation of cultural and spatial distance in the Middle and Late Bronze Age (1400 to 1000 calBC).

### 2019-02-02 article/04_Discussion.Rmd
more consistent spacing before "calBC" (@yesdavid)

@@
 However, the ^14^C data from this region are almost exclusively from the voivodships of Kajuwien-Pommern, Lodsch and Lower Silesia.
-The amount of data is small, with significantly more data from the Early Bronze Age up to about 1700calBC.
+The amount of data is small, with significantly more data from the Early Bronze Age up to about 1700 calBC.
 Almost no information is available from the Middle and Late Bronze Ages.
@@
 The database does not contain as many dated graves from any other region as from Northern Germany.
-Most of them date back to the Early Bronze Age, while the Middle Bronze Age from about 1800calBC to 1400calBC seems barren.
+Most of them date back to the Early Bronze Age, while the Middle Bronze Age from about 1800 to 1400 calBC seems barren.
 Only for the Late Bronze Age, there is a significant number of data documented again.
@@
 In this respect, the data density for almost the entire Bronze Age is good -- also with regard to the two variables *burial type* and *burial construction*.
-The Late Neolithic from 2200calBC is a desideratum.
-If one carries out the necessary mental shift -- the Central European Middle Bronze Age corresponds to the Nordic Early Bronze Age -- and compares the proxy of the relative grave form development with the archaeological observations, then one finds a substantial agreement: inhumations were the rule in periods I and II (about 1800-1300calBC), but cremations also occurred sporadically.
+The Late Neolithic from 2200 calBC is a desideratum.
+If one carries out the necessary mental shift -- the Central European Middle Bronze Age corresponds to the Nordic Early Bronze Age -- and compares the proxy of the relative grave form development with the archaeological observations, then one finds a substantial agreement: inhumations were the rule in periods I and II (about 1800-1300 calBC), but cremations also occurred sporadically.
 From period III onwards cremation burials gained the upper hand.
@@
 The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation.
-If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000calBC) and a decline (after 2000) of cultural-spatial-distance correlation.
+If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000 calBC) and a decline (after 2000 calBC) of cultural-spatial-distance correlation.
 There are several reasons why any interpretation of this observation is difficult so far:

### 2019-02-02 article/index.Rmd

--- 2019-01-27 article/index.Rmd
@@
 email: clemens@nevrome.de
-abstract: Petersilie.
+abstract: European Bronze Age archaeology traditionally focusses on two major dimensions to categorise burials -- although there is an immense variability of attendant phenomena within this spectrum -- flat graves versus burial mounds and cremation versus inhumation.
+These traits are an indispensable ingredient for common archaeological narratives of sociocultural interaction and cultural evolution.
+This paper presents a quantitative reconstruction of the general trends in the distribution of Bronze Age burial traditions based on bulk radiocarbon data and employs the resulting time series for the  estimation of macro-regional cultural distance.
+Despite the relatively small amount of input data the trend reconstruction fits to established archaeological observations for prehistoric Europe.
+The comparison of cultural and spatial distance leads to the remarkable result of no significant permanent correlation, which indicates that the spread of the relevant aspects of burial traditions can not be explained with simple diffusion models.
+Instead a more complex process of cultural transmission has to be considered.
 keywords: Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission

### 2019-02-05 article/index.Rmd
change of title -- the following changes are based on suggestions by @benmarwick

@@
 ---
-title: A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe
+title: Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data
 runninghead: Schmid.

### 2019-02-05 article/01_Introduction.Rmd
restructuring the introduction

@@
 
-This article employs some of the methods that arose within archaeology's contemplation of *Cultural transmission theory* to trace the development of four major Bronze Age burial traditions in Europe.
 The cultural-historical evolution of archaeological cultures and ethnic groups during the European Bronze Age is complex, and its reconstruction an ongoing endeavour of archaeology.
@@
 The present article, in contrast, explores the former strategy.
-The first major question therefore is: *Does openly available, bulk radiocarbon data allow reconstructing the main trends within the distribution of burial rites in Bronze Age Europe?* This approach could set a precedent for a more reproducible and statistically sound research practice.
+The first major question therefore is: *Can the main trends in the distribution of burial rites in Bronze Age Europe be detected in openly available bulk radiocarbon data?* This approach could set a precedent for a more reproducible and statistically sound research practice.
 If we could answer this first question with *yes*, then the quantitative comparison of macro regional trends could pave the way for an analysis of interaction intensity on a long-term continental scale: *Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?*
@@
 *Cultural evolution theory* provides such a framework [@creanzaCulturalEvolutionaryTheory2017].
-The spread of ideas in human populations over space and time is one of it's major topics: *Cultural transmission*.
+The spread of ideas in human populations over space and time is one of its major topics: *Cultural transmission*.
 

### 2019-02-05 article/01_Introduction.Rmd
restructuring the introduction

@@
 
-To answer these questions, a theoretical framework is required that permits the comparison of human behaviour across spatial, temporal and cultural boundaries: The assumption that the cremation aspect of Early Bronze Age burial rituals in Britain could be related to the same aspect in Hungary hundreds of years later is not self-evident.
-*Cultural evolution theory* provides such a framework [@creanzaCulturalEvolutionaryTheory2017].
-The spread of ideas in human populations over space and time is one of its major topics: *Cultural transmission*.
+Burial traditions are a ritual behaviour that is transmitted between generations.
+They are therefore suitable for study using methods of *Cultural transmission theory*, a special aspect of *Cultural evolution* concerned with the spread of ideas in human populations [@creanzaCulturalEvolutionaryTheory2017].
+This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries.
+*Evolutionary archaeology*, the branch of archaeology that explicitly adopted the terminology of Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is very relevant for burial traditions: Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
+Without resorting to the inappropriate simplifications of Evolutionism, their diffusion can be expected to increase over time if we assume reasonable human actors.
+On the other hand, some behaviours are not related to survival and reproductive capacity, like e.g.
+ceramic decorations, traditional costume fashion or jewellery design.
+Despite or because their seemingly subordinate importance, these characteristics often pass through a breathtakingly fast development.
+They serve archaeologists as a proxy for precise relative chronology.
+Burial traditions can be descriped as a special form of *stylistic traits*.
+Special because of their strong dependence on religious and spiritual beliefs and their connection with the emotionally stirring topic of death, but still subject to seemingly random fashion changes.
+
+Of course in reality most behaviour is not unambiguously stylistic or functional.
+Burial rites cause costs regarding working time and material to a varying extent.
+These expenses might have been influenced by or had their effect on the human-nature relation: Pyres, for example, require a significant amount of wood and the erection of barrows much workforce.
+Still, it can be expected that these functional disadvantages were not critical most of the time -- at least regarding the general decision which kind of burial tradition is applied.
+Most probably they influence the magnitude of invested energy for a burial, just like the level of grief [@macdonaldGriefBurialAmerican2001] or the representational intentions of the burying relatives [@henrichEvolutionCostlyDisplays2009].
+
+Dunnell postulates that the distribution and emergence of stylistic cultural traits behave randomly on a population scale level.
+The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @porcicSimulatingCulturalTransmission2014; @porcicExploringEffectsAssemblage2015; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001].
+@neimanStylisticVariationEvolutionary1995 takes a closer look at these processes and develops a simple framework for the simulation of non-functional idea propagation.
+He identifies the essential phenomena *drift* (random dominance of similar traits) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.
+Furthermore, he implements a measure of cultural distance to apply his theoretical insights to real-world data.
+This workflow was an important methodological inspiration for the burial rites data examination in this paper.
+
+In the following sections I will (1) *explain how radiocarbon data with relevant meta information can be transformed to macro regional time series of burial traditions*, (2) *explore how these time series compare to traditional archaeological narratives* and (3) *construct and contextualise cultural distances between macro regions over time in relation to spatial distance*.
+For the latter a simulation based on the @neimanStylisticVariationEvolutionary1995 template was employed.
 
@@
 Questions of Cultural Transmission form the core of archaeology.
-
-*Evolutionary archaeology*, the branch of archaeology that explicitly adopted the terminology of Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style].
-Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
-Without resorting to the inappropriate simplifications of Evolutionism, their diffusion can be expected to increase over time if we assume reasonable human actors.
-On the other hand, some behaviours are not related to survival and reproductive capacity, like e.g.
-ceramic decorations, traditional costume fashion or jewellery design.
-Despite or because their seemingly subordinate importance, these characteristics often pass through a breathtakingly fast development.
-They often serve archaeologists as a proxy for precise relative chronology.
-Dunnell postulates that the distribution and emergence of these stylistic cultural traits behave randomly on a population scale level.
-The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @porcicSimulatingCulturalTransmission2014; @porcicExploringEffectsAssemblage2015; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001].
-
-@neimanStylisticVariationEvolutionary1995 takes a closer look at this processes and develops a simple framework for the simulation of non-functional idea propagation.
-He identifies the essential phenomena *drift* (random dominance of similar traits) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.
-Furthermore, he implements a measure of cultural distance to apply his theoretical insights to real-world data.
-
-The workflow suggested by Neiman is relevant to the burial rites data examined in this paper and was an important methodological inspiration: As burial traditions are ritualised behaviour which is most of the time heavily linked to religious and spiritual beliefs they can be assumed to be stylistic traits.
-Of course in reality, beyond the distinction of @dunnell1978style, most behaviour is not unambiguously stylistic or functional.
-Burial rites cause costs regarding working time and material to a varying extent.
-These expenses might have been influenced by or had their effect on the human-nature relation: Pyres, for example, require a significant amount of wood and the erection of barrows much workforce.
-Still, it can be expected that these functional disadvantages were not critical most of the time -- at least regarding the general decision which kind of burial tradition is applied.
-Most probably they influence the magnitude of invested energy for a burial, just like the level of grief [@macdonaldGriefBurialAmerican2001] or the representational intentions of the burying relatives [@henrichEvolutionCostlyDisplays2009].
-
-In the following sections I will (1) *explain how radiocarbon data with relevant meta information can be transformed to macro regional time series of burial traditions*, (2) *explore how these time series compare to traditional archaeological narratives* and (3) *construct and contextualise cultural distances between macro regions over time in relation to spatial distance*.
-For the latter a simulation based on the @neimanStylisticVariationEvolutionary1995 template was employed.

### 2019-02-05 article/01_Introduction.Rmd
even more restructuring of the introduction -- some aspects where moved to the conclusion

@@
 Burial traditions are a ritual behaviour that is transmitted between generations.
-They are therefore suitable for study using methods of *Cultural transmission theory*, a special aspect of *Cultural evolution* concerned with the spread of ideas in human populations [@creanzaCulturalEvolutionaryTheory2017].
-This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries.
-*Evolutionary archaeology*, the branch of archaeology that explicitly adopted the terminology of Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is very relevant for burial traditions: Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
-Without resorting to the inappropriate simplifications of Evolutionism, their diffusion can be expected to increase over time if we assume reasonable human actors.
+They are therefore suitable for study using methods of *Cultural transmission theory*, a special aspect of *Cultural evolution theory* concerned with the spread of ideas in human populations [@creanzaCulturalEvolutionaryTheory2017].
+This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries, as it assumes *Social learning* to produce adaptive behaviour through transmission, selection and mutation -- a secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007; @lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
+
+*Evolutionary archaeology*, the branch of archaeology that explicitly adopted the terminology of Cultural evolution, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is very relevant for burial traditions: Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
+Without resorting to the inappropriate simplifications of Evolutionism, their prevalence generally happens to increase over time.
 On the other hand, some behaviours are not related to survival and reproductive capacity, like e.g.
@@
 Despite or because their seemingly subordinate importance, these characteristics often pass through a breathtakingly fast development.
-They serve archaeologists as a proxy for precise relative chronology.
-Burial traditions can be descriped as a special form of *stylistic traits*.
-Special because of their strong dependence on religious and spiritual beliefs and their connection with the emotionally stirring topic of death, but still subject to seemingly random fashion changes.
+Burial traditions can be descriped as a special form of these *stylistic traits*.
+Special because of their strong dependence on religious and spiritual beliefs and their connection with the emotionally stirring topic of death, but still subject to sometimes rapid and seemingly random fashion changes.
 
@@
 In the following sections I will (1) *explain how radiocarbon data with relevant meta information can be transformed to macro regional time series of burial traditions*, (2) *explore how these time series compare to traditional archaeological narratives* and (3) *construct and contextualise cultural distances between macro regions over time in relation to spatial distance*.
-For the latter a simulation based on the @neimanStylisticVariationEvolutionary1995 template was employed.
-
-Cultural evolution theory is at least as old as theories of biological evolution and underwent a long and versatile research history.
-Its beginnings are in *Evolutionism* and *Social darwinism* in the 19th and early 20th century and until the nineteen sixties and seventies it matured in conflict with *Cultural relativism* into *Neoevolutionism* [@petermannGeschichteEthnologie2004].
-In the late seventies three major theoretical movements emerged: *Evolutionary psychology*, *Human behavioural ecology* and *Dual inheritance theory* [@smithThreeStylesEvolutionary2000].
-The latter -- heavily popularised by Richard Dawkins's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007]: Just like in biological evolution, also *Social learning* might be prone to selection and mutation.
-Under this assumption methods of *Population genetics* could be employed within cultural anthropology to analyse the spread and development of ideas -- of course with the much faster pace of cultural exchange in mind.
-A lot of theoretical and quantitative research followed this line of thought: e.g.
-@cavalli-sforzaCulturalTransmissionEvolution1981, @boydCultureEvolutionaryProcess1985, @teltserEvolutionaryArchaeologyMethodological1995 or @obrienEvolutionaryArchaeologyTheory1996.
-
-Archaeology has a unique perspective on Cultural evolution as it is concerned with human behaviour over big areas and long time scales [@garveyCurrentPotentialRoles2018].
-It focuses on an abundant and highly diagnostic proxy of cultural expression: material remains.
-These include individual lost or discarded artefacts, the ruins of past architecture and all more inconspicuous by-products of living.
-We rightfully expect humans to learn handcraft as well as building techniques and traditions -- both functionally meaningful and seemingly pointless -- from others close to them.
-The similarity of historically and spatially close material remains can, therefore, be explained as a direct result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
-Questions of Cultural Transmission form the core of archaeology.

### 2019-02-05 article/05_Conclusion.Rmd
even more restructuring of the introduction -- some aspects where moved to the conclusion

@@
 2.
-This paper is built upon Cultural Evolution theory (see section \ref{introduction}).
+This paper is built upon Cultural evolution theory (see section \ref{introduction}).
 It not just relies on terminology and methodology developed in this field, the process of abstraction necessary to go from radiocarbon dates to cultural distance was also heavily influenced by Evolutionary archaeology.
 This theoretical framework might be a source of inspiration and methods for other projects.
-The adoption and adaption of statistical tools and models from population genetics into anthropology was and is a unique characteristic of Cultural Evolution theory [@creanzaCulturalEvolutionaryTheory2017].
+The adoption and adaption of statistical tools and models from population genetics into anthropology was and is a unique characteristic of Cultural Evolution theory [@cavalli-sforzaCulturalTransmissionEvolution1981; @boydCultureEvolutionaryProcess1985; @teltserEvolutionaryArchaeologyMethodological1995; @obrienEvolutionaryArchaeologyTheory1996].
+On the other hand, archaeology is explicitly concerned with human behaviour over big areas and long time scales and can offer a unique perspective to other anthropological fields [@garveyCurrentPotentialRoles2018].

### 2019-02-05 article/02_Materials_and_Methods.Rmd
more consistent spacing before "calBC" (@yesdavid)
improving text in materials&methods

@@
 \includegraphics[width=\columnwidth]{../analysis/figures/map_graves.jpeg}
-\caption{\label{map_graves}All graves dating between 2200 and 800calBC as extracted from the Radon-B database (\url{radon-b.ufg.uni-kiel.de}).
+\caption{\label{map_graves}All graves dating between 2200 and 800 calBC as extracted from the Radon-B database (\url{radon-b.ufg.uni-kiel.de}).
 The classes of the variable \textit{burial type} are distinguished by colour, the ones of \textit{burial construction} by shape.
@@
 
-The code of this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible.
+The code for this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible [@marwickPackagingData2018].
 A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
@@
 
-The dataset used for the analysis in this paper is a subset of the openly available database Radon-B [@kneiselRadonB2013].
-It contains ^14^C dates from the European Bronze and Early Iron Age (approximately 2300 BC to 500 BC).
-Radiocarbon dating is a crucial absolute dating technique for this period [@ramseyRadiocarbonDatingRevolutions2008], and Radon-B is a collection of more than 11000 published dates from around 2800 sites located all over Europe (as of December 2018).
+The dataset used for the analysis in this paper is a subset of the openly available database Radon-B [@kneiselRadonB2013], which contains more than 11000 ^14^C dates from the European Bronze and Early Iron Age (approximately 2300 BC to 500 BC).
+Radiocarbon dating is a crucial absolute dating technique for this period [@ramseyRadiocarbonDatingRevolutions2008], and Radon-B is a collection of published dates from around 2800 sites located all over Europe (as of February 2019).
 Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the type and cultural attribution of the sampled context.
@@
 In theory, the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe.
-Unfortunately, the Radon-B dataset has some quality issues mostly due to inconsistent manual data input: Incompleteness and faultiness of meta information, no supervised or well-thought-out variable classification for *sample material* and *cultural attribution* and inconsistent multilingualism even within individual entries.
-These issues though do not inhibit the general data analysis strategy of this paper.
+Unfortunately, the Radon-B dataset has some quality issues mostly due to inconsistent manual data input: Incompleteness and faultiness of meta information, no well-thought-out variable classification for *sample material* and *cultural attribution* and inconsistent multilingualism even within individual entries.
+These issues though do not inhibit the general data analysis strategy of this paper, as the information about temporal and spatial position as well as *feature type* attribution is mostly complete.
 

### 2019-02-05 article/05_Conclusion.Rmd
improving text in materials&methods

@@
 The idea to systematically collect data points that combine spatial and temporal location with simplified meta information about cultural phenomena might be a way to achieve better reproducibility of overarching narratives.
-Fast and easy *reproducibility* is of crucial importance to achieve scientific progress and it includes the notion of rather technical *preproducibility* [@starkReproducibilityMustCome2018]: Data (and code) must be shared to fuel advances in anthropological theory building [@kitzesPracticeReproducibleResearch2017; @marwickComputationalReproducibilityArchaeological2017].
+Fast and easy *reproducibility* is of crucial importance to achieve scientific progress and it includes the notion of rather technical *preproducibility* [@starkReproducibilityMustCome2018]: Data (and code) must be shared to fuel advances in anthropological theory building [@kitzesPracticeReproducibleResearch2017; @marwickComputationalReproducibilityArchaeological2017; @marwickPackagingData2018].
 

### 2019-02-05 article/06_Remarks.Rmd
more work on the conclusion
used the right sentence for funding declaration

@@
 Dr.
-Ben Marwick (University of Washington) and the valuable C++ code review by Benny Baumann.
+Ben Marwick (University of Washington) and the valuable C++ code review by Benny Baumann (Kiel).
 
@@
 
-This paper was prepared without financial support, except for a 300Euro travel grant to attend the Conference on Complex Systems 2018 by the International Center of Kiel University.
+This research received no specific grant from any funding agency in the public, commercial, or not-for-profit sectors, except for a 300Euro travel grant to attend the Conference on Complex Systems 2018 by the International Center of Kiel University.

### 2019-02-05 article/02_Materials_and_Methods.Rmd
some more minor changes in the materials&methods section

@@
 The x-axis shows the time, the y-axis the amount or the percentage shares of graves.
-The individual plots are separated by region, and each region is featured by a small map according to Figure \ref{map_regions} A.}
+The individual plots are separated by region, and each region is featured by a small map according to Figure \ref{map_regions} A.
+The order of the regions follows a southeast-\\northwest direction.}
 \end{figure}
@@
 
-These time series also allow to derive another set of time series with a measure of cultural distance for each region-region relationship (Figure \ref{sed_region_matrix}).
+From these time series another set of time series with a measure of cultural distance for each region-region relationship can be derived (Figure \ref{sed_region_matrix}).
 Cultural distance is defined here as the squared Euclidean distance (SED) between the trait proportions of two different regions at a certain point in time [@neimanStylisticVariationEvolutionary1995].
-SED seems to be a sensible statistic here, because it is relatively simple, works with only a few assumptions and emphasises differences through squaring.
+SED is a sensible metric here, because it is relatively simple, works with only a few assumptions and emphasises differences through squaring.
 
@@
 For each of the $1400/200 = 7$ section and each of the two variables, a matrix of the mean SED over time was calculated.
-The two matrices of each section can be tested for correlation with the Mantel test, which uses permutations of allegedly correlating matrices to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967].
+The two matrices of each section were tested for correlation with the Mantel test, which uses permutations of potentially correlating matrices to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967].
 Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test (Table \ref{tab:distanceandcorrelationtable*} C).
@@
 Defining a measure of spatial distance for the above established main units of observation, the macro-regions, is only possible in a highly simplified way.
-A continuous distance measure cannot be employed, but it is possible to distinguish ordinal distance classes.
+A continuous distance metric such as least-cost path distance cannot be employed on this scale, but it is possible to distinguish ordinal distance classes.
 For the eight regions, four different categories of spatial distance were defined (Figure \ref{map_regions} B).

### 2019-02-05 article/02_Materials_and_Methods.Rmd
tried to make it really clear what's the purpose of the simulation

@@
 
-To get a better understanding how the correlation coefficient should develop if there was a strong correlation between spatial and cultural distance as measured by *burial type* and *burial construction*, a simple simulation was implemented based on the model proposed by @neimanStylisticVariationEvolutionary1995.
+To understand the results in Table \ref{tab:distanceandcorrelationtable*} we need a frame of reference: *How should the the correlation coefficient develop if there was a strong correlation between spatial and cultural distance?
+How is the degree of cultural transmission related to correlation and similariy of cultural traits and therefore cultural distance?*
+
+For this questions a toy model simulation was implemented based on the model proposed by @neimanStylisticVariationEvolutionary1995.
 This simulation assumes a population separated into $g = 8$ groups (regions) with $N_g = 100$ individuals each.
@@
 
-The partner group for each learning instance is selected randomly following a group interaction probability matrix $I$. Two different matrices where tested:
+The partner group for each learning instance is selected randomly following a group interaction probability matrix $I$. Two different matrices were tested:
 
@@
 
+The main purpose of this comparison is to decide, if the real-world events -- the development of *burial type* and *burial construction* -- are really as incompatible with the assumption of spatial correlation as the results in Table \ref{tab:distanceandcorrelationtable*} suggest: If the toy model simulation yields no runs with weak or negative correlation for later time steps given the assumption of spatial correlation, then that would indicate that interaction with neighbouring communities can indeed be ruled out as a decisive factor for the distribution of the variable *burial type* in the Middle and Late Bronze Age.
+For this question Figure \ref{simulation} DF are most important, while Figure \ref{simulation} ABCE only show control runs.
+
 \begin{figure*}

### 2019-02-05 article/03_Results.Rmd
more work on the description of results

@@
 Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region-wise time series of trait occurrence.
-If we ignore all other archaeological knowledge, we can construct the following narrative from them: \newline
+If we consider these results in isolation, we can construct the following narrative from them: \newline
 
-In the course of the Bronze Age in Europe there was a trend away from body burial towards cremation.
+The overall picture is that during the Bronze Age in Europe there was a trend away from body burial towards cremation.
 Around 2200 calBC cremation burials were almost unknown in Poland, Southern Germany, Northeastern France and Northern Germany.
@@
 
-This narrative is wrong or inaccurate in some aspects. Please see the analysis in section \ref{time-series-discussion}.
-
 ## Cultural distance
@@
 
-Table \ref{tab:distanceandcorrelationtable*} shows that the distance networks of *burial type* and *burial construction* are not equivalent.
+Table \ref{tab:distanceandcorrelationtable*} C shows that the distance networks of *burial type* and *burial construction* are not equivalent.
 There is also no significant negative correlation.
@@
 
-For the relationship between the SED network based on *burial type* and the spatial distance several observations result from Table \ref{tab:distanceandcorrelationtable*}: Cultural distance calculated from these to variables does not seem to correlate with spatial distance.
-A significant correlation between the distance measures only reveals itself for the Early Bronze Age time slot 2000 to 1800 calBC: Only in the Early Bronze Age, long before the universal switch to the Urnfield burial tradition, large regions directly adjacent to each other tend to undergo the same developments rather than the ones further away from each other.
-In the Middle Bronze Age the SED developments of the *burial type* variable undergo the peaks already observed in Figure \ref{sed_region_matrix}.
-The correlation between cultural and spatial distance over these peaks is low and even tilts between 1400 and 1200 calBC, at the transition to the Late Bronze Age, into the negative -- although not significantly.
+For the relationship between the SED networks and the spatial distance network several observations result from Table \ref{tab:distanceandcorrelationtable*} AB: Cultural distance generally does not seem to correlate with spatial distance.
+For the variable *burial type* an exception with significant correlation between the distance measures only reveals itself for the Early Bronze Age time slot 2000 to 1800 calBC.
+In the Early Bronze Age, long before the universal switch to the Urnfield burial tradition, large regions directly adjacent to each other tend to see the same developments.
+In the Middle Bronze Age the SED developments of the *burial type* variable undergo the peaks already observed.
+The correlation between cultural and spatial distance during these peaks is low and even tilts between 1400 and 1200 calBC, at the transition to the Late Bronze Age, into the negative -- although not significantly.
 The cultural distance between several geographically distant regions is then smaller than between several directly adjacent regions.
-Looking at the SED network of *burial construction* (Table \ref{tab:distanceandcorrelationtable*}), similar conclusions can be drawn as in the *burial type* context: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated.
+Looking at the SED network of *burial construction*, similar conclusions can be drawn: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated.
 As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation, though it is only significant for the time slot between 1800 and 1400 calBC.
@@
 
-The results of the simulation as presented in Figure \ref{simulation} confirm the expectation: In case of equal interaction intensity $I$ of separate populations, the correlation of cultural and spatial distance is weak (Figure \ref{simulation} A vs.
+The simulation as presented in Figure \ref{simulation} was implemented to contextualize this observation of low correlation between spatial and cultural distance.
+Some general remarks: In case of equal interaction intensity $I$ of separate populations, the correlation of cultural and spatial distance is weak (Figure \ref{simulation} A vs.
 C vs.
 E).
-It appears to be normally distributed around zero for many simulation runs, although individual runs can still yield remarkably high positive and negative correlation for certain time slots.
-The fluctuation of the correlation within one run can be strong and seems to increase for bigger values of $m_i$.
+It appears to be normally distributed around zero for many simulation runs, although individual runs can still yield high positive and negative correlation for certain time slots.
+The fluctuation of the correlation within one run can be strong and seems to increase for greater values of $m_i$.
 If the interaction matrix $I$ is predefined to represent the spatial distance matrix (Figure \ref{simulation} BDF), then the mean correlation per time slot increases.
@@
 That's mostly because of the relatively strong negative correlation of cultural and spatial distance in the Middle and Late Bronze Age (1400 to 1000 calBC).
+Despite this tendency, the real-world correlation time series are still within the margins of potential toy model events.
+Therefore it is not possible to rule out the influence of spatial proximity as an important factor for the Urnfield burial rite diffusion.
+On the contrary, the simulation results challenge an overly simplistic interpretation of Table \ref{tab:distanceandcorrelationtable*}.

### 2019-02-05 article/04_Discussion.Rmd
changes and improvements in the discussion section

@@
 
-One major question asked in this paper was, *if openly available, bulk radiocarbon data does allow the reconstruction of main trends within the distribution of burial rites in Bronze Age Europe*.
+One major question asked in this paper was, *if the main trends in the distribution of burial rites in Bronze Age Europe can be detected in openly available bulk radiocarbon data?*.
 This is the case.
@@
 The following paragraphs contain a short evaluation of the time series in Figure \ref{development_burial_type} and \ref{development_burial_construction} in the light of the archaeological state of research for each region.
-Despite the general success of this methodological approach, unfortunately the amount and quality of data is still not sufficient for a comprehensive picture.
-The reasons for this lack of data quality are discussed in section \ref{data-quality-discussion}.
-\newline
 
@@
 The second important question posed for this paper was, *if these time series constructed from radiocarbon data* (Figure \ref{development_burial_type} and \ref{development_burial_construction}) *allow further analysis concerning cultural and spatial distance*?
-While the idea behind the first question is rather technical -- reproducing already available information in a quantitative way -- the second one has a larger scope.
-A quantitative representation of an archaeological narrative should path the way for a more diverse set of methods to be applied to this data and cultural distance is an integral topic of archaeological research.
+That can also be affirmed.
+While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one had a larger scope.
+The quantitative representation of an archaeological narrative paths the way for a diverse set of methods to be applied and the analysis undertaken here is merely the beginning.
 
-The expectation for this question was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* as well in between these variables as with spatial distance.
-The simulation, though with this implementation and parametrisation for sure rather qualitatively than quantitatively, visualises how a high degree of interaction should translate to a significantly higher degree of cultural similarity across human populations (Figure \ref{simulation}).
+Concerning the question of cultural distance, the expectation based on Cultural transmission theory was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* with spatial distance.
+The simulation, though with this implementation and parametrisation for sure rather qualitatively than quantitatively, visualises how a high degree of interaction should translate to a higher degree of cultural similarity across human populations (Figure \ref{simulation}).
 And as the European Bronze Age is usually depicted as a period of trade, travel and exchange [e.g.

### 2019-02-05 article/04_Discussion.Rmd
some love for the conclusion

@@
 For some there is not enough information to make any decision about the relative dominance and for many the general relevance and accuracy of the resulting time series can be questioned.
-There is a variety of reasons for this.
+There are a variety of reasons for this.
 These should also be considered for future data acquisition in Radon-B and beyond.
@@
 There is a fundamental break between that great, theoretical archaeology, which wants to develop comprehensive narratives and only pay quantitative attention to the individual case, and the on-site research or emergency excavation, which has to dedicate a measure of respect and sometimes a lot of time to each grave.
-In this thesis, ^14^C data are converted into a proxy of calibrated density distributions within artificial working regions, which is completely detached from the individual context.
+In this paper, ^14^C data are converted into a proxy of calibrated density distributions within artificial working regions, which is completely detached from the individual context.
 ^14^C data on graves are used to attribute a feature to the absolute chronological context or to better describe the relative development within a site.

### 2019-02-05 article/05_Conclusion.Rmd
some love for the conclusion

@@
 
-The major questions posed in the introduction to this paper have been answered : *Bulk radiocarbon data is a powerful source to trace general trends in prehistoric Europe* and *allows long-term observations concerning cultural distance*.
+*The main trends in the distribution of burial rites in Bronze Age Europe can be detected in bulk radiocarbon data* and indeed *they allow long-term observations concerning cultural distance*.
 Radiocarbon dates can be transformed to meaningful regional time series, if they can be linked to the relevant meta information.
-Although only `r r("gpr size", sf_desc)` graves of a maybe six-figure statistical population that can be estimated for all burials in the European Bronze Age [@nikulkaArchaologischeDemographieMethoden2016] have been available for this analysis, the resulting reconstruction of general burial tradition trends in wast areas of Central, Northern and North-western Europe comes close to the one achieved by literature research.
+Although only `r r("gpr size", sf_desc)` graves -- a small fraction of maybe hundreds of thousands burials that can be estimated for the European Bronze Age [@nikulkaArchaologischeDemographieMethoden2016] -- have been available for this analysis, the resulting reconstruction of general burial tradition trends in wast areas of Central, Northern and North-western Europe comes close to the one achieved by literature research.
 The time series can be employed for cultural distance calculation.
@@
 
-Future research is required to de-tangle these interactions -- of course on the basis of more precise data.
+Future research is required to untangle these interactions -- of course on the basis of more precise data.
 This endeavour will not end with the simple description of social learning cascades, but with the analysis of the initial reasons for diversification.
-From the point of view of Cultural Complexity -- the question, which factors accelerate or decelerate the variability of human behaviour [@henrichDemographyCulturalEvolution2004; @mesoudiWhatCumulativeCultural2018; @vaesenPopulationSizeDoes2016] -- the long term dominance fluctuations of one burial rite over the other are remarkable.
+From the point of view of Cultural complexity -- the question, which factors accelerate or decelerate the variability of human behaviour [@henrichDemographyCulturalEvolution2004; @mesoudiWhatCumulativeCultural2018; @vaesenPopulationSizeDoes2016] -- the long term dominance fluctuations of one burial rite over the other are remarkable.
 Its standard explanation models (among others population density and environmental risk) can hardly be applied to stylistic variables.
@@
 1.
-For this paper burial traditions were the main object of research, but the same methodological approach could be applied to reconstruct other aspects of cultural behaviour.
+For this paper burial traditions were the main object of research, but the same approach could be applied to reconstruct other aspects of cultural behaviour.
 The idea to systematically collect data points that combine spatial and temporal location with simplified meta information about cultural phenomena might be a way to achieve better reproducibility of overarching narratives.
@@
 2.
-This paper is built upon Cultural evolution theory (see section \ref{introduction}).
+This paper is built upon Cultural transmission and thereby Cultural evolution theory (see section \ref{introduction}).
 It not just relies on terminology and methodology developed in this field, the process of abstraction necessary to go from radiocarbon dates to cultural distance was also heavily influenced by Evolutionary archaeology.
@@
 The adoption and adaption of statistical tools and models from population genetics into anthropology was and is a unique characteristic of Cultural Evolution theory [@cavalli-sforzaCulturalTransmissionEvolution1981; @boydCultureEvolutionaryProcess1985; @teltserEvolutionaryArchaeologyMethodological1995; @obrienEvolutionaryArchaeologyTheory1996].
-On the other hand, archaeology is explicitly concerned with human behaviour over big areas and long time scales and can offer a unique perspective to other anthropological fields [@garveyCurrentPotentialRoles2018].
+On the other hand, archaeology is explicitly concerned with human behaviour over big areas and long time scales and can reciprocate with a unique perspective [@garveyCurrentPotentialRoles2018].

### 2019-02-05 article/04_Discussion.Rmd
detached boring parts of the discussion and moved them to a new main chapter "supplementary remarks"

@@
 For individual regions and sites specific ideological, social or even economic considerations might have been much more critical for the decision which burial tradition to adopt [@barrett_monumentality_1990; @fokkens_genesis_1997-1; @muller-scheesel_variabilitat_2009].
-
-## Data quality {#data-quality-discussion}
-
-The **A** sections of the figures \ref{development_burial_type} and \ref{development_burial_construction} with absolute grave numbers per region and year indicate areas and periods with a lot and others with few points of data.
-For some there is not enough information to make any decision about the relative dominance and for many the general relevance and accuracy of the resulting time series can be questioned.
-There are a variety of reasons for this.
-These should also be considered for future data acquisition in Radon-B and beyond.
-
-1. The data in Radon-B depend on the projects and interests of the participating researchers: It was collected as needed for individual regions rather than systematically for all of Europe.
-
-2.
-The availability of ^14^C dates for a region and period in the literature depends on whether a modern research project took place within this context.
-Besides the research interests of researchers and institutions, there are also political reasons why an area might be better or worse explored.
-
-3.
-Different research traditions in archaeology considered ^14^C data in the past more or less essential for the reconstruction of absolute chronology.
-In various Metal Age contexts, ^14^C data provide inferior precision in comparison to relative chronological classifications - e.g.
-by fibula typology - and are therefore used only sporadically to create absolute chronological fix points.
-Dendrochronological dating is preferred in waterlogged deposits and whenever the necessary woods are available.
-^14^C calibration curves form plateaus in some periods, which significantly limit the possible dating precision.
-If such an effect is known, then often no dates are commissioned at all.
-
-4.
-^14^C dating requires a small but sufficiently contextualised amount of organic finds.
-Due to the old wood and old water effect, short-lived sample materials such as bones of terrestrial organisms or charred cereal grains are preferred.
-The majority of ^14^C data for this work comes from bones and teeth.
-However, their preservation is subject to huge differences depending on the soil [@brockPrescreeningTechniquesIdentification2010; @gordonSoilPHBone1981; @nielsen-marshBoneDiagenesisEuropean2007].
-In many mineral soil contexts no sufficiently good, dateable material is available.
-This situation can affect entire regions, if, for example, poor bone preservation prevails as a result of lime-deficient soils.
-
-5.
-In very few regions and periods of the European Bronze Age the number of excavated graves is sufficiently big to assume that a statistically relevant sample of the burial culture of the entire population is documented.
-Instead, it is evident in most contexts that only the burials of certain social groups, genders or ethnic groups were recorded.
-The quantity and nature of the known graves cannot be representative.
-In this case it can be expected that further, deviant burial rituals were practised, which left no archaeologically visible remains.
-These rituals are invisible in the ^14^C data from Radon-B.
-
-Concerning the primary variables *burial type* and *burial construction*, further special effects must be taken into account: The disturbance and robbery of graves was a widespread phenomenon in history and prehistory.
-Burial mounds were usually attacked more severely than flat graves due to their better visibility and could, therefore, be under-represented in Radon-B.
-In many places, they were also the subject of expeditions in the early days of archaeology at the end of the 19th and beginning of the 20th century.
-Of course, these excavations did not yield ^14^C-dates.
-Irrespective of this, it can be assumed that a large part of the burial mounds erected in the Bronze Age was destroyed by agricultural activity.
-Ploughing could also result in an under-representation of the archaeological findings [@hardingEuropeanSocietiesBronze2000a].
-On the other hand, if a burial mound is still present, it is more readily visible and recognised by archaeologists today: Mounds are often clearly identifiable in aerial photographs or digital terrain models, while flat graves are much more difficult to identify for today's remote sensing methods.
-It is difficult to estimate the impact of these effects in detail, but an over-representation of mounds seems more likely than an over-representation of flat graves.
-The dating of cremation graves on bone remains methodically differs from the dating of inhumations because it is based on the extraction of the mineral, not the organic, fraction in the bone [@lantingDatingCrematedBones2001].
-The underlying chemical method was developed later, has only been used more frequently in the past 15 years and is associated with some particular difficulties [@olsenCharacterisationBlindTesting2008; @strydonck14CDatingCremated2009].
-Burnt bone is generally more susceptible to weathering than unburned bone.
-An under-representation of cremation graves in current data collections is, therefore, more plausible.
-
-## Data meaning {#data-meaning-discussion}
-
-This paper is based on ^14^C dates which were prepared for the dating of graves.
-There is a fundamental break between that great, theoretical archaeology, which wants to develop comprehensive narratives and only pay quantitative attention to the individual case, and the on-site research or emergency excavation, which has to dedicate a measure of respect and sometimes a lot of time to each grave.
-In this paper, ^14^C data are converted into a proxy of calibrated density distributions within artificial working regions, which is completely detached from the individual context.
-^14^C data on graves are used to attribute a feature to the absolute chronological context or to better describe the relative development within a site.
-The questions underlying the creation of a ^14^C date are therefore not equivalent to those imposed in this paper.
-
-The superordinate, large perspective also tries to simplify and categorize the parameters associated with a single burial.
-Radon-B anticipated this step by providing the variables *burial type* and *burial construction* that can be reduced to three categories each.
-This is extremely useful on the one hand, but also questionable on the other.
-The respective contexts can be multifarious: similar rituals in different cultural contexts can have different meanings.
-For this work, it is assumed -- with reference to *Cultural Evolution Theory* -- that despite this restriction, actions such as burning the dead or burying in burial mounds can be passed on across cultural and ethnic group boundaries through imitation and teaching.
-To put it provocatively: cremation burial in the Hungarian Early Bronze Age is a manifestation of the same idea as cremation in the Northern Late Bronze Age.
-This is a working hypothesis, but may be questioned.
-
-The meaning of the data in Radon-B is thus manipulated in at least three ways for the purpose of this work: By concentrating on the dated grave in contrast to the ^14^C date, this work detaches itself from the actual content of the database.
-By creating artificial time steps and investigation regions, the cultural-historical context of the grave is dissolved.
-By creating the relative development proxies, the individual grave loses its context.
-The present paper is a secondary use of Radon-B.
-A better dataset should be grave-related, not date-related, and should contain much more meta-information for each grave.

### 2019-02-05 article/06_Supplementary_Remarks.Rmd
detached boring parts of the discussion and moved them to a new main chapter "supplementary remarks"

# Supplementary Remarks

## Data quality {#data-quality-discussion}

The **A** sections of the figures \ref{development_burial_type} and \ref{development_burial_construction} with absolute grave numbers per region and year indicate areas and periods with a lot and others with few points of data. For some there is not enough information to make any decision about the relative dominance and for many the general relevance and accuracy of the resulting time series can be questioned. There are a variety of reasons for this. These should also be considered for future data acquisition in Radon-B and beyond.

1. The data in Radon-B depend on the projects and interests of the participating researchers: It was collected as needed for individual regions rather than systematically for all of Europe.

2. The availability of ^14^C dates for a region and period in the literature depends on whether a modern research project took place within this context. Besides the research interests of researchers and institutions, there are also political reasons why an area might be better or worse explored.

3. Different research traditions in archaeology considered ^14^C data in the past more or less essential for the reconstruction of absolute chronology. In various Metal Age contexts, ^14^C data provide inferior precision in comparison to relative chronological classifications - e.g. by fibula typology - and are therefore used only sporadically to create absolute chronological fix points. Dendrochronological dating is preferred in waterlogged deposits and whenever the necessary woods are available. ^14^C calibration curves form plateaus in some periods, which significantly limit the possible dating precision. If such an effect is known, then often no dates are commissioned at all.

4. ^14^C dating requires a small but sufficiently contextualised amount of organic finds. Due to the old wood and old water effect, short-lived sample materials such as bones of terrestrial organisms or charred cereal grains are preferred. The majority of ^14^C data for this paper comes from bones and teeth. However, their preservation is subject to huge differences depending on the soil [@brockPrescreeningTechniquesIdentification2010; @gordonSoilPHBone1981; @nielsen-marshBoneDiagenesisEuropean2007]. In many mineral soil contexts no sufficiently good, dateable material is available. This situation can affect entire regions, if, for example, poor bone preservation prevails as a result of lime-deficient soils.

5. In very few regions and periods of the European Bronze Age the number of excavated graves is sufficiently big to assume that a statistically relevant sample of the burial culture of the entire population is documented. Instead, it is evident in most contexts that only the burials of certain social groups, genders or ethnic groups were recorded. The quantity and nature of the known graves cannot be representative. In this case it can be expected that further, deviant burial rituals were practised, which left no archaeologically visible remains. These rituals are invisible in the ^14^C data from Radon-B.

Concerning the primary variables *burial type* and *burial construction*, further special effects must be taken into account: The disturbance and robbery of graves was a widespread phenomenon in history and prehistory. Burial mounds were usually attacked more severely than flat graves due to their better visibility and could, therefore, be under-represented in Radon-B. In many places, they were also the subject of expeditions in the early days of archaeology at the end of the 19th and beginning of the 20th century. Of course, these excavations did not yield ^14^C-dates. Irrespective of this, it can be assumed that a large part of the burial mounds erected in the Bronze Age was destroyed by agricultural activity. Ploughing could also result in an under-representation of the archaeological findings [@hardingEuropeanSocietiesBronze2000a]. On the other hand, if a burial mound is still present, it is more readily visible and recognised by archaeologists today: Mounds are often clearly identifiable in aerial photographs or digital terrain models, while flat graves are much more difficult to identify for today's remote sensing methods. It is difficult to estimate the impact of these effects in detail, but an over-representation of mounds seems more likely than an over-representation of flat graves. The dating of cremation graves on bone remains methodically differs from the dating of inhumations because it is based on the extraction of the mineral, not the organic, fraction in the bone [@lantingDatingCrematedBones2001]. The underlying chemical method was developed later, has only been used more frequently in the past 15 years and is associated with some particular difficulties [@olsenCharacterisationBlindTesting2008; @strydonck14CDatingCremated2009]. Burnt bone is generally more susceptible to weathering than unburned bone. An under-representation of cremation graves in current data collections is, therefore, more plausible.

## Data meaning {#data-meaning-discussion}

This paper is based on ^14^C dates which were prepared for the dating of graves. There is a fundamental break between that great, theoretical archaeology, which wants to develop comprehensive narratives and only pay quantitative attention to the individual case, and the on-site research or emergency excavation, which has to dedicate a measure of respect and sometimes a lot of time to each grave. In this paper, ^14^C data are converted into a proxy of calibrated density distributions within artificial working regions, which is completely detached from the individual context. ^14^C data on graves are used to attribute a feature to the absolute chronological context or to better describe the relative development within a site. The questions underlying the creation of a ^14^C date are therefore not equivalent to those imposed in this paper.

The superordinate, large perspective also tries to simplify and categorize the parameters associated with a single burial. Radon-B anticipated this step by providing the variables *burial type* and *burial construction* that can be reduced to three categories each. This is extremely useful on the one hand, but also questionable on the other. The respective contexts can be multifarious: similar rituals in different cultural contexts can have different meanings. For this paper, it is assumed -- with reference to *Cultural Evolution Theory* -- that despite this restriction, actions such as burning the dead or burying in burial mounds can be passed on across cultural and ethnic group boundaries through imitation and teaching. To put it provocatively: cremation burial in the Hungarian Early Bronze Age is a manifestation of the same idea as cremation in the Northern Late Bronze Age. This is a working hypothesis, but may be questioned.

The meaning of the data in Radon-B is thus manipulated in at least three ways for the purpose of this paper: By concentrating on the dated grave in contrast to the ^14^C date, it detaches itself from the actual content of the database. By creating artificial time steps and investigation regions, the cultural-historical context of the grave is dissolved. By creating the relative development proxies, the individual grave loses its context. The present paper is a secondary use of Radon-B. A better dataset should be grave-related, not date-related, and should contain much more meta-information for each grave.

### 2019-02-05 article/04_Discussion.Rmd
another important point concerning the simulation

@@
 Concerning the question of cultural distance, the expectation based on Cultural transmission theory was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* with spatial distance.
-The simulation, though with this implementation and parametrisation for sure rather qualitatively than quantitatively, visualises how a high degree of interaction should translate to a higher degree of cultural similarity across human populations (Figure \ref{simulation}).
+The simulation, though with this implementation and parametrisation for sure rather qualitatively than quantitatively, shows how an increase in the degree of interaction should translate to a higher degree of cultural similarity across human populations (Figure \ref{simulation}).
 And as the European Bronze Age is usually depicted as a period of trade, travel and exchange [e.g.
@@
 Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:distanceandcorrelationtable*}).
-Only for small periods of time this correlation can be assumed based on the Radon-B data.
+Only for small periods of time this correlation is visible in the Radon-B data.
 The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation.
@@
 3.
+The toy simulation based on @neimanStylisticVariationEvolutionary1995 suggests that even under the assumption of strong transmission between directly neighbouring groups, random and emergent effects can still cause 	fluctuation and temporary lack of correlation in human populations.
+
+4.
 All observations of cultural distance in this paper are based on only the two variables *burial type* and *burial construction*.
@@
 The rise of the Tumulus Culture in the Middle Bronze Age might have cushioned this effect for the variable *burial construction* and introduced more variability for a -- at that time -- less important aspect *burial type*.
-The latter may have been of subordinate relevance in the Tumulus Culture context.
-With the rise of the Urnfield Culture this phenomenon could have been reversed as *burial type* became an expression of something that could have been one of the first formalized religions in Europe: The spread of the Urnfield Culture was an extraordinary development in early European history and was possibly accompanied by a structured, religious dogma that could have built up around the symbolic content of cult wagons, sun barks and waterfowl [@bilicSwanChariotSolar2016; @kimmig_seevolkerbewegung_1964; @molloy_european_2018].
+The latter may have been of subordinate relevance in the Tumulus culture context.
+With the rise of the Urnfield culture this phenomenon could have been reversed as *burial type* became an expression of something that could have been one of the first formalized religions in Europe: The spread of the Urnfield Culture was an extraordinary development and possibly accompanied by a structured, religious dogma that could have built up around the symbolic content of cult wagons, sun barks and waterfowl [@bilicSwanChariotSolar2016; @kimmig_seevolkerbewegung_1964; @molloy_european_2018].
 Of course these explanations could, if correct, only be considered on a big scale.

### 2019-02-05 README.Rmd
first draft of updated README

@@
 ---
-output: github_document
+output:
+  md_document:
+    variant: gfm
 ---
@@
 ```{r, echo = FALSE} 
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>",
-  fig.path = "README-"
-)
+knitr::opts_chunk$set( collapse = TRUE, comment = "#>", fig.path = "README-" )
 ```
 
-# workingtitle
+[![Build Status](https://travis-ci.com/nevrome/cultrans.bronzeageburials.article2019.svg?token=vxsQ9RjxoGASGtX4Q8jc&branch=master)](https://travis-ci.com/nevrome/cultrans.bronzeageburials.article2019) [![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/nevrome/nevrome/cultrans.bronzeageburials.article2019/commits/master)
+[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.5.0-brightgreen.svg)](https://cran.r-project.org/)
+[![ORCiD](https://img.shields.io/badge/ORCiD-0000--0003--3448--5715-green.svg)](http://orcid.org/0000-0003-3448-5715)
 
-This repository contains the data and code for our paper:
+## Research compendium for ‘Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data’
 
-> Authors, (YYYY). _Title of paper_. Name of journal/book <https://doi.org/xxx/xxx>
+### Compendium DOI:
 
-Our pre-print is online here:
+<http://dx.doi.org/>…
 
-> Authors, (YYYY). _Title of paper_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+The files at the URL above will generate the results as found in the publication.
+The files hosted at <https://github.com/nevrome/cultrans.bronzeageburials.article2019> are the development versions and may have changed since the paper was published
 
-### How to cite
+### Author of this repository:
 
-Please cite this compendium as:
+Clemens Schmid (<clemens@nevrome.de>)
 
-> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for 'Title of paper'_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+### Published in:
 
-### How to download or install
+Not yet published.
 
-You can download the compendium as a zip from from this URL: <https://github.com/nevrome/workingtitle/archive/master.zip>
+### Abstract:
 
-Or you can install this compendium as an R package, workingtitle, from GitHub with:
-```{r gh-installation, eval = FALSE}
-# install.packages("devtools")
-devtools::install_github("nevrome/workingtitle")
-```
-### Licenses
+### Overview of contents:
 
-**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)
+This repository contains text, code and data for the paper. Further
+context information can be found there.
 
-**Code :** See the [DESCRIPTION](DESCRIPTION) file
+The rendered document can be downloaded
+[here](https://github.com/nevrome/cultrans.bronzeageburials.article2019/blob/document/article.pdf).
 
-**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse
+### How to reproduce:
 
-### Contributions
+As the data and code in this repository are complete and self-contained, it can be reproduced with any R environment (\> version 3.5.0).
+The necessary package dependencies are documented in the `DESCRIPTION` file and can be installed manually or automatically with `devtools::install()`.
+If it’s not possible any more to construct a working environment with these methods due to technological progress, one can use the Docker image.
 
-We welcome contributions from everyone.
-Before you get started, please see our [contributor guidelines](CONTRIBUTING.md).
-Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md).
-By participating in this project you agree to abide by its terms.
+A Docker image is a lightweight GNU/Linux virtual computer that can be run as a piece of software on Windows, Linux, and OSX.
+To capture the complete computational environment used for this project we have a Dockerfile that specifies how to make the Docker image that we developed this project in.
+The Docker image includes all of the software dependencies needed to run the code in this project, including the data and code itself.
+To launch the Docker image for this project, first, [install Docker](https://docs.docker.com/installation/) on your computer and download the `.tar` file with the paper image [here](...) (not yet available).
+At the Docker prompt, you can load and run the image with:
+
+    docker load -i cultransbronze19_docker_image.tar
+    docker run -e PASSWORD=cultransbronze19 -dp 8787:8787 --name cultransbronze19 cultransbronze19
+
+This will start a server instance of RStudio.
+Then open your web browser at localhost:8787 or run `docker-machine ip default` in the shell to find the correct IP address, and log in with rstudio/cultransbronze19.
+Once logged in, use the Files pane (bottom right) to navigate to the script files.
+More information about using RStudio in Docker is available at the [Rocker](https://github.com/rocker-org) [wiki](https://github.com/rocker-org/rocker/wiki/Using-the-RStudio-image) pages.
+
+We developed and tested the package on this Docker container, so this is the only platform that we’re confident it works on. It was built and stored with:
+
+    docker build -t cultransbronze19 .
+    docker save -o cultransbronze19_docker_image.tar cultransbronze19
+
+### Licenses:
+
+Code: MIT <http://opensource.org/licenses/MIT> year: 2019, copyright holder: Clemens Schmid
+
+Data: Please see the license agreements of Radon-B and Natural Earth data
+
+Text: Please see the license agreements of the Adaptive Behavior journal

### 2019-02-05 README.Rmd
added abstract to README

@@
 ### Abstract:
+
+European Bronze Age archaeology traditionally focusses on two major dimensions to categorise burials -- although there is an immense variability of attendant phenomena within this spectrum -- flat graves versus burial mounds and cremation versus inhumation.
+These traits are an indispensable ingredient for common archaeological narratives of sociocultural interaction and cultural evolution.
+This paper presents a quantitative reconstruction of the general trends in the distribution of Bronze Age burial traditions based on bulk radiocarbon data and employs the resulting time series for the estimation of macro-regional cultural distance.
+Despite the relatively small amount of input data the trend reconstruction fits to established archaeological observations for prehistoric Europe.
+The comparison of cultural and spatial distance leads to the remarkable result of no significant permanent correlation, which indicates that the spread of the relevant aspects of burial traditions can not be explained with simple diffusion models.
+Instead a more complex process of cultural transmission has to be considered.
+
+### Keywords:
+
+Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission
 

### 2019-02-05 README.Rmd
some more info about the directory structure in the readme

@@
 
-This repository contains text, code and data for the paper. Further
-context information can be found there.
+This repository contains text, code and data for the paper.
+The paper is rendered via continuous integration and can be downloaded from the [document branch](https://github.com/nevrome/cultrans.bronzeageburials.article2019/tree/document), so
+[here](https://github.com/nevrome/cultrans.bronzeageburials.article2019/blob/document/article.pdf).
 
-The rendered document can be downloaded
-[here](https://github.com/nevrome/cultrans.bronzeageburials.article2019/blob/document/article.pdf).
+The `analysis` directory contains `code` and `data` to reproduce the preparations, calculations, simulations and figure renderings in this paper.
+The `article` directory contains the tex template and the text for the paper in *.Rmd* format.
 

### 2019-02-05 analysis/code/02_cultural_distance_scripts/02b_burial_type_sed.R
a little code cleaning

@@ -3,12 +3,6 @@ load("analysis/data/tmp_data/development_proportions_burial_type.RData")
 
 prop <- proportion_development_burial_type
 
-# proportion_development_burial_type %>%
-#   tidyr::complete(
-#     region_name, timestep, idea,
-#     fill = list(proportion = as.integer(0))
-#   )
-
 long_prop <- prop %>%
   tidyr::spread(
     idea, proportion

### 2019-02-05 analysis/code/02_cultural_distance_scripts/05a_burial_construction_mantel_test.R

@@ -1,8 +1,6 @@
 load("analysis/data/tmp_data/distance_matrix_spatial.RData")
 load("analysis/data/tmp_data/distance_matrizes_sed_burial_construction.RData")
 
-####
-
 mantel_test_results <- lapply(
   1:length(distance_matrizes_sed), function(i, x, y, z) {
     mantel_result <- vegan::mantel(x[[i]], y, method = "spear", permutations=999)

### 2019-02-05 analysis/code/02_cultural_distance_scripts/05b_burial_type_mantel_test.R

@@ -1,8 +1,6 @@
 load("analysis/data/tmp_data/distance_matrix_spatial.RData")
 load("analysis/data/tmp_data/distance_matrizes_sed_burial_type.RData")
 
-####
-
 mantel_test_results <- lapply(
   1:length(distance_matrizes_sed), function(i, x, y, z) {
     mantel_result <- vegan::mantel(x[[i]], y, method = "spear", permutations=999)

### 2019-02-06 article/index.Rmd
detached boring parts of the discussion and moved them to a new main chapter "supplementary remarks"
added abstract to README
new key word

@@
 Instead a more complex process of cultural transmission has to be considered.
-keywords: Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission
+keywords: Bronze Age, Burial Traditions, Cultural Transmission, Radiocarbon Dating 
 classoption:
@@
 
-```{r child = '06_Remarks.Rmd'}
+```{r child = '06_Supplementary_Remarks.Rmd'}
 ```
+
+```{r child = '07_Technical_Remarks.Rmd'}
+```

### 2019-02-06 article/02_Materials_and_Methods.Rmd
citation for euclidian distance

@@
 Cultural distance is defined here as the squared Euclidean distance (SED) between the trait proportions of two different regions at a certain point in time [@neimanStylisticVariationEvolutionary1995].
-SED is a sensible metric here, because it is relatively simple, works with only a few assumptions and emphasises differences through squaring.
+SED is a sensible metric here, because it is relatively simple, works with only a few assumptions and emphasises differences through squaring [@nakoinzFingerprintingIronAge2014].
 

### 2019-03-17 review/answer_to_review.md
preparation to answer the review

# Reviewer: 1

## Comments to the Author:

> A lot of work has been already done in quantifying changes in funerary practices during the Bronze Age in central and south-western Europe through the analysis of radiocarbon-dated archaeological contexts. I recommend citing these works in the paper, since the adopted approach is similar.

> -	Capuzzo G, Barceló JA. 2015. Cultural changes in the 2nd millennium BC:  a Bayesian examination of radiocarbon evidence from Switzerland and Catalonia. World Archaeology 47 (4) Special Issue: Bayesian approaches to Prehistoric Chronologies:622-641.
> -	Capuzzo G, López Cachero FJ. 2017. De la inhumación a la cremación en el nordeste peninsular: cronología y sociedad. In: Barceló JA, Bogdanovic I, Morell B, editors. Iber-Crono 2016. Cronometrías Para la Historia de la Península Ibérica (Chronometry for the History of the Iberian Peninsula), Barcelona (Spain), 17-19 October 2016:192-208.
> -	Capuzzo G, Barceló JA (in press). Cremation burials in Central and Western Europe: quantifying an adoption of innovation in the 2nd millennium BC. In: Kneisel J, Nakoinz O, Barcelo J, editors. Turning Points and Change in Bronze Age Europe (2400 – 800 BC). Modes of change – inhumation versus cremation in Bronze Age burial rites. Proceedings of the International Workshop "Socio-Environmental Dynamics over the Last 12,000 Years: The Creation of Landscapes IV (24th-27th March 2015) in Kiel. Universitätsforschungen zur Prähistorischen Archäologie 
> -	Capuzzo G. 2014. Space-temporal Analysis of Radiocarbon Evidence and Associated Archaeological Record: from Danube to Ebro Rivers and from Bronze to Iron Ages. Ph.D. thesis, Autonomous University of Barcelona.

> In the paper, it’s very high the amount of data, for which the archaeological information regarding the type of burial and funerary rite is lacking. At table 1, page 4, we can clearly detect that for 60% of the dates associated to cremation burials we have no information regarding the type of funerary structure. For 37% of the inhumations, no information about the funerary structure is available. Moreover, for 32% of the dataset (559 radiocarbon dates out of 1701) we do not have any kind of information regarding neither the funerary rite nor the burial structure. This lack of data obliges the author to consider carefully the results reached in the paper. It’s relevant to highlight all the biases of the prosed research and how the conclusion can be misleading due to the quality of available data. In particular, when the statistical analyses carried out on data aim to explore in the details the cultural distance among the different regions.

## There are some typing error in the manuscript:

> Line 28 page 2: can be described

> Line 58 page 4: How should the correlation

> Line 2 page 5: Listed are the overall

> I advise the author to review carefully the text.

> At lines 48-50, page 2, the bibliographic reference to Dunnell is missing.

> Regarding the images, it would be preferable not to use negative number in order to indicate years BC.

# Reviewer: 2

## Comments to the Author:

> I had the chance to read the manuscript entitled “Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data” authored by C. Schimd. This is an interesting paper that combines empirical analysis of Bronze age burial sites across Europe with a computer simulation of intergroup cultural transmission. The manuscript is well written, I enjoyed reading it, and think that the core findings are definitely of interest for archaeologist working on Bronze age Europe and for those interested in the application of cultural evolutionary framework. I am thus keen to recommend its publication, although I think several issues needs to be addressed before this.

> Firstly, my impression is that the author is trying to squeeze perhaps a bit too much in a single paper. I felt this was particularly the case of the computer simulation model which is almost thrown in in the mix, without much discussion on model parameters (see below). I would be tempted to suggest removing this entirely and include this into another paper.

> Secondly, I feel like the empirical analyses have too much preprocessing. The raw data provides spatial and temporal coordinates of individual burials, yet these are combined spatially and chronologically first and analysed as aggregates. As a consequence: 1) there is a substantial loss of spatial information; 2) sampling error is effectively disregarded (i.e. mantel tests are based on proportion values that are calculated for regions with very different samples sizes) and strictly speaking the samples are also non-independent (i.e. a large site with multiple burials can strongly bias the proportions of a particular region); and 3) there is an increased risk of ecological fallacy/Simpson’s paradox. I take this aggregation provides a direct mapping to existing literature, but I am not convinced that this is worth sacrificing so much.

> Thirdly, the author implies that the scripts used for his analysis are supplied but there are no links on the article to any repository and I did not find any electronic supplementary material. As the author is well aware, we need to do our best to make our work fully reproducible.

## Minor points:

> Page 1, Column 2, Lines 52-43: “Do time series ... spatial distance”. I think this question should be phrased better. Eg. What  does “meaningful” mean in this context?

> Page 2, Column 1, Line 9: “[...] adaptive behaviour”. Not necessarily.

> Page 2, Column 1, Line 15: “explicitly adopted the terminology of Cultural evolution”. This reads as if only the terminology is adopted? Rephrase?

> Page 2, Column 1, Lines 58. Not sure what “random dominance of similar traits” means - should be rephrased.

> Page 3, Column 1, Line 33. Please provide a link to the repository with the code.

> Page 3, Column 2, Lines 40-44. The definition of these regions seems rather arbitrary and necessary. Why create such an artificial unit when the spatial coordinates are available?

> Page 3, Column 2, Line 60. “For the remaining 154 dates...”. I assume these dates were excluded as we don’t know whether they are terrestrial or marine?  Also, the paragraph mentions “bones and teeth”, are there any risks of aquatic diet determining a reservoir effect?

> Page 4, Column 1, Lines 57-59. Why not use directly the summed probabilities?

> Page 4, Columns 1-2, Lines 58-60; 1-8. This is a major issue. Perhaps it is worth bootstrap confidence interval for these proportions? This way both sample size and sample independence could be tackled. Also, please provide the sample size for each region.

> Page 4, Column 2, Lines 43-56. I think it is worth computing partial mantel tests so that the correlation between burial type and construction can take into account geography.

> Page 5, Column 1, Line 24. Perhaps worth making this a subsection?

> Page 5, Column 1, Lines 24-60. I have several issues with this model. The idea here is to emulate the observed data so we can build expectations that can be directly compared against empirical patterns. But I do find several issues: 1) How is Ng=100 justified? This seems a somewhat arbitrary number for a parameter that is likely going to play a significant role in the amount of drift in the system? 2) I understand each time-step is 20 years. Why? I assume this is related to generation length but this ought to be discussed; 3) Is the model supposed to reach an equilibrium state? If this is the case I don’t think 70 time-steps and Ng=100 is sufficient for this. If the answer is no, the initial condition of the system shouldn’t be random and perhaps be informed from the empirical data.

> Figure 3. These are very nice plots - I would add the sample sizes for each region though.

> Figure 6. I find these plots a bit messy and with too much information. I also don’t see much the point of testing for significance the results of each simulation run - they are effectively samples from the same generative process so I think testing them does not make much sense?

### 2019-03-17 article/01_Introduction.Rmd
fixed spelling mistakes

@@
 Despite or because their seemingly subordinate importance, these characteristics often pass through a breathtakingly fast development.
-Burial traditions can be descriped as a special form of these *stylistic traits*.
+Burial traditions can be described as a special form of these *stylistic traits*.
 Special because of their strong dependence on religious and spiritual beliefs and their connection with the emotionally stirring topic of death, but still subject to sometimes rapid and seemingly random fashion changes.
@@
 
-Dunnell postulates that the distribution and emergence of stylistic cultural traits behave randomly on a population scale level.
+@dunnell1978style postulates that the distribution and emergence of stylistic cultural traits behave randomly on a population scale level.
 The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @porcicSimulatingCulturalTransmission2014; @porcicExploringEffectsAssemblage2015; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001].

### 2019-03-17 article/02_Materials_and_Methods.Rmd
fixed spelling mistakes

@@
   caption = "Mean distance, correlation and mantel test results for cultural and spatial distance.
-Listed are the the overall mean squared Euclidian distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bc)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B).
+Listed are the overall mean squared Euclidian distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bc)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B).
 \\textbf{C} shows the Mantel test results for the two variables if compared directly.
@@
 
-To understand the results in Table \ref{tab:distanceandcorrelationtable*} we need a frame of reference: *How should the the correlation coefficient develop if there was a strong correlation between spatial and cultural distance?
+To understand the results in Table \ref{tab:distanceandcorrelationtable*} we need a frame of reference: *How should the correlation coefficient develop if there was a strong correlation between spatial and cultural distance?
 How is the degree of cultural transmission related to correlation and similariy of cultural traits and therefore cultural distance?*

### 2019-03-17 review/answer_to_review.md
fixed spelling mistakes
decided to remove simulation from this paper

@@
 
+Fixed.
+
 > Line 58 page 4: How should the correlation
 
+Fixed.
+
 > Line 2 page 5: Listed are the overall
+
+Fixed.
 
@@
 > At lines 48-50, page 2, the bibliographic reference to Dunnell is missing.
+
+Fixed.
 
@@
 I would be tempted to suggest removing this entirely and include this into another paper.
+
+After careful deliberation, I have decided to follow this advice.
+The simulation (1) does not add information that would be absolutely essential for the core topics, (2) reduces the overall comprehensibility due to the terminology that has to be introduced to describe it and (3) cannot be adequately presented and discussed here without leaving the initial scope of the paper.
+I will prepare another paper in the future that will present this simulation and other simulation approaches.
 

### 2019-03-17 README.Rmd
replaced key word
updated README accordingly

@@
 
-Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission
+Bronze Age, Burial Traditions, Cultural Transmission, Radiocarbon Dating
 
@@
 
-The `analysis` directory contains `code` and `data` to reproduce the preparations, calculations, simulations and figure renderings in this paper.
+The `analysis` directory contains `code` and `data` to reproduce the preparations, calculations and figure renderings in this paper.
 The `article` directory contains the tex template and the text for the paper in *.Rmd* format.

### 2019-03-17 article/02_Materials_and_Methods.Rmd
removed simulation and everything related to it from the text: careful rereading will be necessary

@@
 ```
-
-To understand the results in Table \ref{tab:distanceandcorrelationtable*} we need a frame of reference: *How should the correlation coefficient develop if there was a strong correlation between spatial and cultural distance?
-How is the degree of cultural transmission related to correlation and similariy of cultural traits and therefore cultural distance?*
-
-For this questions a toy model simulation was implemented based on the model proposed by @neimanStylisticVariationEvolutionary1995.
-This simulation assumes a population separated into $g = 8$ groups (regions) with $N_g = 100$ individuals each.
-$k = 2$ mutually exclusive ideas are randomly but equally distributed in this population at $t_{0}$.
-With every time step ($t_{total} = (2200 - 800)/20 = 70$) all individuals learn from another randomly chosen individual.
-Several different values where tested for the probability $m_i$ that the other agent the individual is learning from comes not from the same, but from another group:
-
-$$
-m_i = 
-  \begin{dcases*} 
-    0 & Figure \ref{simulation} AB, no intergroup exchange \\ 
-    0.05 & Figure \ref{simulation} CD \\
-    0.5 & Figure \ref{simulation} EF, very high exchange
-  \end{dcases*}
-$$
-
-The partner group for each learning instance is selected randomly following a group interaction probability matrix $I$. Two different matrices were tested:
-
-$$
-I =
-  \begin{dcases*}
-    \begin{matrix}
-        & a & b & c & ... \\
-      a & 0 & 1 & 1 & ... \\
-      b & 1 & 0 & 1 & ...\\
-      c & 1 & 1 & 0 & ... \\
-      ... & ... & ... & ... & ...
-    \end{matrix} \\
-    \text{Figure \ref{simulation} ACEG, equal distance between groups a:h} \\
-    \begin{matrix}
-        & a & b       & c       & ... \\
-      a & 0 & |b-a|^2 & |c-a|^2 & ... \\
-      b & |a-b|^2 & 0 & |c-b|^2 & ... \\
-      c & |a-c|^2 & |b-c|^2 & 0 & ... \\
-      ... & ... & ... & ... & ...
-    \end{matrix} \\
-    \text{Figure \ref{simulation} BDFH, spatial distance (Figure \ref{map_regions} B)}
-  \end{dcases*}
-$$
-
-The distribution of the $k$ variants within the groups can be observed for every time step and documented in time series like the ones for the real-world regions data (Figure \ref{development_burial_type} B and \ref{development_burial_construction} B).
-These time series can as well be used to calculate time series of region-region distance, which again can be split up into time blocks of 200 years to calculate mean distance matrices.
-The latter can be checked for correlation with the spatial distance matrix (cf.
-Table \ref{tab:distanceandcorrelationtable*} AB) and compared with other simulation results and real-world observations (Figure \ref{simulation}).
-
-The main purpose of this comparison is to decide, if the real-world events -- the development of *burial type* and *burial construction* -- are really as incompatible with the assumption of spatial correlation as the results in Table \ref{tab:distanceandcorrelationtable*} suggest: If the toy model simulation yields no runs with weak or negative correlation for later time steps given the assumption of spatial correlation, then that would indicate that interaction with neighbouring communities can indeed be ruled out as a decisive factor for the distribution of the variable *burial type* in the Middle and Late Bronze Age.
-For this question Figure \ref{simulation} DF are most important, while Figure \ref{simulation} ABCE only show control runs.
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{../analysis/figures/simulation.jpeg}
-\caption{\label{simulation}Development of correlation between cultural and spatial distance for real world data and simulation results.
-The plots in the left column of the plot matrix (\textbf{ACE}) contain simulation results for an equal group distance matrix $I$, the ones on the right (\textbf{BDF}) for a distance matrix based on the macro unit distances defined for this paper (Figure \ref{map_regions} B).
-The rows of the plot matrix (\textbf{AB}, \textbf{CD}, \textbf{EF}) represent different degrees of intergroup transmission $m_i$.
-Each plot shows the results of 50 simulation runs.
-The x-axis displays the time distinguished into 200 year time slots, the y-axis Spearman's rank correlation coefficient between the mean squared Euclidean distance matrix for a 200 year time slot and the static spatial distance matrix.
-This information is visualized for each time slot with (from left to right) a dotplot coloured according to the $p$-value of the Mantel test, a scatter plot line and a boxplot.
-The scatter plot points of one simulation run are connected with a faint line from one time slot to the next.
-To compare the runs with the real world observations, the results for the burial customs data (Table \ref{tab:distanceandcorrelationtable*} AB) is added here with a stronger solid (\textit{burial type}) and dashed (\textit{burial construction}) line.
-The strong horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}
-\end{figure*}

### 2019-03-17 article/03_Results.Rmd
detached boring parts of the discussion and moved them to a new main chapter "supplementary remarks"
removed simulation and everything related to it from the text: careful rereading will be necessary

@@
 
-## Timeseries {#time-series-results}
+## Time series {#time-series-results}
 
@@
 There is no time period with negative mean correlation for *burial construction* cultural distance and spatial distance, but also here the correlation coefficient approaches zero in the Late Bronze Age.
-
-The simulation as presented in Figure \ref{simulation} was implemented to contextualize this observation of low correlation between spatial and cultural distance.
-Some general remarks: In case of equal interaction intensity $I$ of separate populations, the correlation of cultural and spatial distance is weak (Figure \ref{simulation} A vs.
-C vs.
-E).
-It appears to be normally distributed around zero for many simulation runs, although individual runs can still yield high positive and negative correlation for certain time slots.
-The fluctuation of the correlation within one run can be strong and seems to increase for greater values of $m_i$.
-If the interaction matrix $I$ is predefined to represent the spatial distance matrix (Figure \ref{simulation} BDF), then the mean correlation per time slot increases.
-The correlation increases as well with the degree of interaction $m_i$ (Figure \ref{simulation} D vs.
-F).
-Counter-intuitively the fluctuation of the correlation is also very high here (Figure \ref{simulation} E vs.
-F).
-The correlation development observed for the variables *burial type* and *burial construction* fits to simulation runs in all presets, but becomes less likely if we assume $I$ to represent spatial distance and a high degree of interaction $m_i$ (Figure \ref{simulation} A vs.
-F).
-That's mostly because of the relatively strong negative correlation of cultural and spatial distance in the Middle and Late Bronze Age (1400 to 1000 calBC).
-Despite this tendency, the real-world correlation time series are still within the margins of potential toy model events.
-Therefore it is not possible to rule out the influence of spatial proximity as an important factor for the Urnfield burial rite diffusion.
-On the contrary, the simulation results challenge an overly simplistic interpretation of Table \ref{tab:distanceandcorrelationtable*}.

### 2019-03-17 article/04_Discussion.Rmd
removed simulation and everything related to it from the text: careful rereading will be necessary

@@
 Concerning the question of cultural distance, the expectation based on Cultural transmission theory was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* with spatial distance.
-The simulation, though with this implementation and parametrisation for sure rather qualitatively than quantitatively, shows how an increase in the degree of interaction should translate to a higher degree of cultural similarity across human populations (Figure \ref{simulation}).
-And as the European Bronze Age is usually depicted as a period of trade, travel and exchange [e.g.
+As the European Bronze Age is usually depicted as a period of trade, travel and exchange [e.g.
 @kristiansenRiseBronzeAge2005; @sherrattWhatWouldBronzeAge1993] there was no reason to expect something else than Tobler's famous First Law of Geography: "everything is related to everything else, but near things are more related than distant things." [@toblerComputerMovieSimulating1970].
@@
 3.
-The toy simulation based on @neimanStylisticVariationEvolutionary1995 suggests that even under the assumption of strong transmission between directly neighbouring groups, random and emergent effects can still cause 	fluctuation and temporary lack of correlation in human populations.
-
-4.
 All observations of cultural distance in this paper are based on only the two variables *burial type* and *burial construction*.

### 2019-03-17 analysis/code/03_simulation_scripts/01_run_simulation.R
removed simulation code

@@ -1,64 +1,0 @@
-library(magrittr)
-source("analysis/code/helper_functions/neiman_simulation.R")
-
-load("analysis/data/tmp_data/region_order.RData")
-load("analysis/data/tmp_data/distance_matrix_spatial.RData")
-interaction_matrix_spatial <- (1/distance_matrix_spatial)^2
-diag(interaction_matrix_spatial) <- 0
-
-#### setup settings grid ####
-
-models_grid <- expand.grid(
-  k = 2,
-  N_g = 100,
-  t_start = -2200,
-  t_end = -800,
-  t_steps = 20,
-  mu = 0,
-  g = 8,
-  I = list(NA, interaction_matrix_spatial),
-  mi = c(0, 0.05, 0.5)
-) %>%
-  tibble::as.tibble() %>%
-  dplyr::mutate(
-    model_group = 1:nrow(.)
-  ) %>%
-  tidyr::uncount(50) %>%
-  dplyr::mutate(
-    model_id = 1:nrow(.)
-  )
-
-save(models_grid, file = "analysis/data/tmp_data/sed_simulation_model_grid.RData")
-
-
-#### run simulation ####
-
-models <- pbapply::pblapply(
-  1:nrow(models_grid),
-  function(i, models_grid) {
-    neiman_simulation(
-      models_grid$k[i],
-      models_grid$N_g[i],
-      models_grid$t_start[i],
-      models_grid$t_end[i],
-      models_grid$t_steps[i],
-      models_grid$mu[i],
-      models_grid$g[i],
-      models_grid$mi[i],
-      models_grid$I[[i]]
-    ) %>% standardize_neiman_output(., region_order) %>%
-      dplyr::mutate(
-        model_id = models_grid$model_id[i],
-        model_group = models_grid$model_group[i],
-        region_population_size = models_grid$N_g[i],
-        degree_interregion_interaction = models_grid$mi[i]
-      )
-  },
-  models_grid,
-  cl = 2
-)
-
-#### store results in file system ####
-
-save(models, file = "analysis/data/tmp_data/simulation_models.RData")
-

### 2019-03-17 analysis/code/03_simulation_scripts/02_create_regions_grid_from_simulation_results.R

@@ -1,64 +1,0 @@
-load("analysis/data/tmp_data/region_order.RData")
-
-##### read simulation output data #####
-
-load("analysis/data/tmp_data/simulation_models.RData")
-
-prop <- do.call(rbind, models)
-
-#### load sed function ####
-
-load("analysis/data/tmp_data/sed_function.RData")
-
-#### data preparation ####
-
-long_prop <- prop %>%
-  tidyr::spread(
-    idea, proportion
-  )
-
-regions_grid <-
-  expand.grid(
-    regionA = prop$region %>% unique(),
-    regionB = prop$region %>% unique(),
-    time = prop$timestep %>% unique(),
-    model_id =  prop$model_id %>% unique(),
-    stringsAsFactors = FALSE
-  ) %>%
-  tibble::as.tibble() %>%
-  dplyr::left_join(
-    long_prop,
-    by = c(
-      "regionA" = "region",
-      "time" = "timestep",
-      "model_id" = "model_id"
-    )
-  ) %>%
-  dplyr::left_join(
-    subset(long_prop, select=-c(model_group)),
-    by = c(
-      "regionB" = "region",
-      "time" = "timestep",
-      "model_id" = "model_id"
-    ),
-    suffix = c("_regionA", "_regionB")
-  )
-
-region_A <- mapply(c, regions_grid$idea_1_regionA, regions_grid$idea_2_regionA, SIMPLIFY = FALSE)
-region_B <- mapply(c, regions_grid$idea_1_regionB, regions_grid$idea_2_regionB, SIMPLIFY = FALSE)
-region_A_region_B <- mapply(list, region_A, region_B, SIMPLIFY = FALSE)
-
-regions_grid$sed <- unlist(pbapply::pblapply(
-  region_A_region_B, function(x) {
-    sed(x[[1]], x[[2]])
-  }
-))
-
-regions_grid %<>% dplyr::select(
-    regionA, regionB, time, sed, model_id, model_group
-  )
-
-regions_grid$regionA <- factor(regions_grid$regionA, levels = region_order)
-regions_grid$regionB <- factor(regions_grid$regionB, levels = region_order)
-
-save(regions_grid, file = "analysis/data/tmp_data/sed_simulation_regions_grid.RData")

### 2019-03-17 analysis/code/03_simulation_scripts/03_cut_grid_into_timeslices.R

@@ -1,44 +1,0 @@
-load("analysis/data/tmp_data/sed_simulation_regions_grid.RData")
-
-##### cut sed regions grid into timeslices ####
-
-modelid_time_regions_grid <- regions_grid %>% dplyr::mutate(
-  time = base::cut(
-    time,
-    seq(-2200, -800, 200), labels = paste(seq(-2200, -1000, 200), seq(-2000, -800, 200), sep = " - "),
-    include.lowest = TRUE,
-    right = FALSE)
-) %>%
-  dplyr::group_by(
-    model_group, model_id, time, regionA, regionB
-  ) %>%
-  dplyr::summarise(
-    mean_sed = mean(sed, na.rm = TRUE)
-  ) %>%
-  dplyr::ungroup() %>%
-  # that's dangerous...
-  dplyr::mutate(
-    mean_sed = tidyr::replace_na(mean_sed, 0)
-  )
-
-#### create list of matrizes from timesliced regions grid ####
-
-distance_matrizes_sed <- lapply(
-  base::split(modelid_time_regions_grid, modelid_time_regions_grid$model_id), function(time_regions_grid) {
-    lapply(
-      base::split(time_regions_grid, time_regions_grid$time), function(x){
-        x %>%
-          dplyr::select(
-            -model_group, -model_id, -time
-          ) %>%
-          tidyr::spread(regionA, mean_sed) %>%
-          dplyr::select(
-            -regionB
-          ) %>%
-          as.matrix()
-      }
-    )
-  }
-)
-
-save(distance_matrizes_sed, file = "analysis/data/tmp_data/sed_simulation_regions_timeslices_matrizes.RData")

### 2019-03-17 analysis/code/03_simulation_scripts/04_combine_spatial_and_cultural_distance.R

@@ -1,61 +1,0 @@
-load("analysis/data/tmp_data/region_order.RData")
-load("analysis/data/tmp_data/sed_simulation_regions_grid.RData")
-load("analysis/data/tmp_data/distance_matrix_spatial_long.RData")
-
-#### half regions_grid -- removal of double entries ####
-
-regions_grid$regionA <- as.character(regions_grid$regionA)
-regions_grid$regionB <- as.character(regions_grid$regionB)
-
-regions_grid_half <- pbapply::pblapply(
-  base::split(regions_grid, f = regions_grid$model_group), function(z) {
-    lapply(
-      base::split(z, f = z$model_id), function(y) {
-        lapply(
-          base::split(y, f = y$time), function(x) {
-            mn <- pmin(x$regionA, x$regionB)
-            mx <- pmax(x$regionA, x$regionB)
-            int <- as.numeric(interaction(mn, mx))
-            x <- x[match(unique(int), int),]
-            return(x)
-          }) %>%
-          do.call(rbind, .)
-      }) %>%
-      do.call(rbind, .)
-  }) %>%
-  do.call(rbind, .)
-
-regions_grid_half$regionA <- factor(regions_grid_half$regionA, levels = region_order)
-regions_grid_half$regionB <- factor(regions_grid_half$regionB, levels = region_order)
-
-#### combine spatial and cultural distance into one data.frame ####
-
-sed_spatial_distance <- regions_grid_half %>% dplyr::left_join(
-  distance_matrix_spatial_long, by = c("regionA", "regionB")
-) %>%
-  dplyr::filter(
-    distance != 0
-  ) %>%
-  dplyr::mutate(
-    relation = paste(regionA, "+", regionB),
-    time = base::cut(
-      time,
-      seq(-2200, -800, 200), labels = paste(seq(-2200, -1000, 200), seq(-2000, -800, 200), sep = " - "),
-      include.lowest = TRUE,
-      right = FALSE)
-  ) %>%
-  dplyr::group_by(
-    model_id, model_group, time, regionA, regionB, distance
-  ) %>%
-  dplyr::summarise(
-    mean_sed = mean(sed, na.rm = TRUE)
-  ) %>%
-  dplyr::ungroup() %>%
-  dplyr::filter(
-    !is.na(mean_sed)
-  )
-
-save(
-  sed_spatial_distance,
-  file = "analysis/data/tmp_data/sed_simulation_regions_timeslices_spatial_distance.RData"
-)

### 2019-03-17 analysis/code/03_simulation_scripts/05_apply_mantel_test.R

@@ -1,35 +1,0 @@
-load("analysis/data/tmp_data/distance_matrix_spatial.RData")
-load("analysis/data/tmp_data/sed_simulation_regions_timeslices_matrizes.RData")
-
-mantel_test_results <- pbapply::pblapply(
-  1:length(distance_matrizes_sed), function(i, x, y, model_id) {
-      lapply(
-        1:length(x[[i]]), function(i, x, y, time, model_id) {
-          mantel_result <- vegan::mantel(
-            x[[i]], y,
-            method = "spear",
-            permutations = 999,
-            parallel = 2
-          )
-          data.frame(
-            model_id = model_id,
-            time = time[[i]],
-            statistic = mantel_result$statistic,
-            signif = mantel_result$signif
-          )
-        },
-        x = x[[i]],
-        y = y,
-        time = names(x[[i]]),
-        model_id = model_id[[i]]
-      ) %>% do.call(rbind, .)
-    },
-    x = distance_matrizes_sed,
-    y = distance_matrix_spatial,
-    model_id = names(distance_matrizes_sed)
-  ) %>% do.call(rbind, .) %>% tibble::as.tibble()
-
-save(
-  mantel_test_results,
-  file = "analysis/data/tmp_data/sed_simulation_mantel_sed_spatial.RData"
-)

### 2019-03-17 analysis/code/helper_functions/neiman_simulation.R

@@ -1,156 +1,0 @@
-#### simulation function ####
-
-#' neiman_simulation
-#'
-#' @param k Integer. Number of ideas at t = 0
-#' @param N_g Integer. Population per region
-#' @param t_start Integer. Starting time
-#' @param t_end Integer. Final timestep
-#' @param t_steps Integer. Time step length in years
-#' @param mu Double. Innovation rate
-#' @param g Integer. Number of regions
-#' @param mi Double. Degree of interregion interaction
-#' @param I Doublematrix. Interregion interaction matrix. NA means equal interaction
-#'
-neiman_simulation <- function(k, N_g, t_start, t_end, t_steps, mu, g, mi, I) {
-
-  # define variables
-  regions <- 1:g
-  population <- 1:N_g
-  ideas <- 1:k
-  timesteps <- seq(t_start + t_steps, t_end, t_steps)
-  if (is.na(I)) {
-    I <- matrix(
-      rep(1, g*g), g, g
-    )
-    diag(I) <- 0
-  }
-
-  # create starting populations
-  pop0 <- lapply(
-    regions, function(region, N, k) {
-      tibble::tibble(
-        timestep = as.integer(t_start),
-        individual = population,
-        idea = rep_len(ideas, max(population)),
-        region = region
-      )
-    },
-    k, population
-  )
-
-  # create development list
-  pop_devel <- list()
-  pop_devel[[1]] <- pop0
-
-  # determine number of ideas
-  last_idea <- max(do.call(rbind, pop_devel[[1]])$idea)
-
-  # simulation loop
-  for (p1 in 2:length(timesteps)) {
-
-    # new timestep list
-    pop_old <- pop_devel[[p1 - 1]]
-    pop_new <- pop_old
-
-    # adjust time in new timestep list
-    pop_new <- lapply(
-      pop_new, function(x, p1) {
-        x$timestep <- timesteps[p1] - t_steps
-        return(x)
-      },
-      p1
-    )
-
-    # intraregion learning
-    pop_new <- lapply(
-      pop_new, function(x) {
-        x$idea <- sample(x$idea, length(x$idea), replace = T)
-        return(x)
-      }
-    )
-
-    # interregion learning
-    if (mi != 0) {
-      pop_new <- lapply(
-        regions, function(i, pop_new, pop_old, mi, I, regions) {
[…]

### 2019-03-17 analysis/code/run_all_scripts.R

@@ -4,9 +4,8 @@ purrr::walk(
   c(
     "analysis/code/01_data_preparation_scripts",
     "analysis/code/02_cultural_distance_scripts",
-    "analysis/code/03_simulation_scripts",
-    "analysis/code/04_dataset_description_scripts",
-    "analysis/code/05_plot_scripts"
+    "analysis/code/03_dataset_description_scripts",
+    "analysis/code/04_plot_scripts"
   ),
   function(x) {
     purrr::walk(

### 2019-03-17 article/02_Materials_and_Methods.Rmd
added simple sed time series plot without simulation results -- still useful

@@
 ```
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{../analysis/figures/sed_time_series.jpeg}
+\caption{\label{simulation}Development of correlation between cultural and spatial distance.
+The results for the burial customs data (Table \ref{tab:distanceandcorrelationtable*} AB) are displayed with a solid (\textit{burial type}) and dashed (\textit{burial construction}) line.
+The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}
+\end{figure}

### 2019-03-17 analysis/code/04_plot_scripts/sed_simulation.R

@@ -1,197 +1,0 @@
-library(magrittr)
-
-load("analysis/data/tmp_data/sed_simulation_mantel_sed_spatial.RData")
-mantel_simulations <- mantel_test_results %>% dplyr::mutate(model_id = as.integer(model_id))
-load("analysis/data/tmp_data/sed_simulation_model_grid.RData")
-models_grid %<>% dplyr::mutate(I_type = ifelse(sapply(I, length) == 1, "equal distance", "spatial distance"))
-mantel_simulations %<>% dplyr::left_join(models_grid[, c("model_id", "model_group", "mi", "I_type")], by = "model_id")
-load("analysis/data/tmp_data/mantel_sed_spatial_burial_type.RData")
-mantel_burial_type <- mantel_test_results
-load("analysis/data/tmp_data/mantel_sed_spatial_burial_construction.RData")
-mantel_burial_construction <- mantel_test_results
-
-mantel_burial_type %<>%
-  dplyr::mutate(
-    context = "burial_type"
-  )
-
-mantel_burial_construction %<>%
-  dplyr::mutate(
-    context = "burial_construction"
-  )
-
-mantel_real_world <- rbind(mantel_burial_type, mantel_burial_construction)
-
-library(ggplot2)
-
-plot_mantel <- function(title, mantel_simulations, mantel_real_world) {
-  ju <- ggplot() +
-    geom_hline(
-      yintercept = 0,
-      colour = "red",
-      size = 2
-    ) +
-    geom_line(
-      data = mantel_simulations,
-      mapping = aes(
-        x = time,
-        y = statistic,
-        group = model_id
-      ),
-      position = position_nudge(x = -0.25),
-      size = 0.3,
-      alpha = 0.2
-    ) +
-    geom_point(
-      data = mantel_simulations,
-      mapping = aes(
-        x = time,
-        y = statistic
-      ),
-      position = position_nudge(x = -0.25),
-      size = 1
-    ) +
-    geom_point(
-      data = mantel_real_world,
-      mapping = aes(
-        x = time,
-        y = statistic,
-        colour = context
-      ),
-      position = position_nudge(x = -0.25),
-      size = 7
-    ) +
-    geom_line(
-      data = mantel_real_world,
-      mapping = aes(
-        x = time,
-        y = statistic,
-        colour = context,
-        group = context,
-        linetype = context
-      ),
[…]

### 2019-03-17 analysis/code/04_plot_scripts/sed_time_series.R

@@ -1,0 +1,90 @@
+library(magrittr)
+
+load("analysis/data/tmp_data/mantel_sed_spatial_burial_type.RData")
+mantel_burial_type <- mantel_test_results
+load("analysis/data/tmp_data/mantel_sed_spatial_burial_construction.RData")
+mantel_burial_construction <- mantel_test_results
+
+mantel_burial_type %<>%
+  dplyr::mutate(
+    context = "burial_type"
+  )
+
+mantel_burial_construction %<>%
+  dplyr::mutate(
+    context = "burial_construction"
+  )
+
+mantel <- rbind(mantel_burial_type, mantel_burial_construction)
+
+library(ggplot2)
+
+p <- ggplot() +
+  geom_hline(
+    yintercept = 0,
+    colour = "red",
+    size = 2
+  ) +
+  geom_point(
+    data = mantel,
+    mapping = aes(
+      x = time,
+      y = statistic,
+      colour = context
+    ),
+    size = 7
+  ) +
+  geom_line(
+    data = mantel,
+    mapping = aes(
+      x = time,
+      y = statistic,
+      colour = context,
+      group = context,
+      linetype = context
+    ),
+    size = 2
+  ) +
+  guides(
+    fill = guide_legend(override.aes = list(size = 15))
+  ) +
+  scale_colour_manual(
+    name = "Burial customs",
+    values = c(
+      "burial_type" = "#0072B2",
+      "burial_construction" = "#009E73"
+    )
+  ) +
+  scale_linetype_manual(
+    name = "Burial customs",
+    values = c(
+      "burial_type" = "solid",
+      "burial_construction" = "dashed"
+    )
+  ) +
+  theme_bw() +
+  theme(
+    legend.position = "bottom",
+    axis.text = element_text(size = 25),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title = element_text(size = 25),
+    legend.title = element_text(size = 25, face = "bold"),
+    legend.text = element_text(size = 25),
+    plot.title = element_text(size = 25, face = "bold"),
+    panel.border = element_rect(colour = "black", size = 2),
+    plot.margin = unit(c(1,1,1,2), "lines")
+  ) +
+  ylab("Correlation coefficient") +
+  xlab("Time steps in years calBC")
+
+ggsave(
+  "analysis/figures/sed_time_series.jpeg",
+  plot = p,
+  device = "jpeg",
+  scale = 1,
[…]

### 2019-03-17 analysis/code/04_plot_scripts/correlation_time_series.R
renamed plot

@@ -78,7 +78,7 @@ p <- ggplot() +
   xlab("Time steps in years calBC")
 
 ggsave(
-  "analysis/figures/sed_time_series.jpeg",
+  "analysis/figures/correlation_time_series.jpeg",
   plot = p,
   device = "jpeg",
   scale = 1,

### 2019-03-17 analysis/code/04_plot_scripts/correlation_time_series.R
no negative numbers in x-Axis labels

@@ -17,6 +17,12 @@ mantel_burial_construction %<>%
 
 mantel <- rbind(mantel_burial_type, mantel_burial_construction)
 
+# only for plot: time without sign of years
+mantel$time <- factor(
+  gsub("-(?=[0-9])", "", mantel$time, perl = TRUE),
+  levels = gsub("-(?=[0-9])", "", levels(mantel$time), perl = TRUE)
+)
+
 library(ggplot2)
 
 p <- ggplot() +

### 2019-03-17 analysis/code/04_plot_scripts/development_burial_construction.R

@@ -30,6 +30,18 @@ gl <- lapply(region_file_list, function(x) {
 })
 dummy <- tibble::tibble(region_name = unique(amount_devel$region_name), grob = gl )
 
+# only for plot: time without sign of years
+amount_devel <- amount_devel %>%
+  dplyr::ungroup() %>%
+  dplyr::mutate(
+    timestep = timestep * (-1)
+  )
+
+prop <- prop %>%
+  dplyr::mutate(
+    timestep = timestep * (-1)
+  )
+
 #### development_burial_construction_A ####
 development_burial_construction_A <- ggplot() +
   geom_area(
@@ -67,9 +79,9 @@ development_burial_construction_A <- ggplot() +
   coord_cartesian(
     ylim = c(0, 80)
   ) +
-  scale_x_continuous(
-    breaks = seq(-2200, -800, 200),
-    limits = c(-2500, -800)
+  scale_x_reverse(
+    breaks = seq(2200, 800, -200),
+    limits = c(2500, 800)
   ) +
   # guides(
   #   fill = FALSE
@@ -122,9 +134,9 @@ development_burial_construction_B <- ggplot() +
     breaks = c(0, 0.5, 1),
     labels = c("0%", "50%", "100%")
   ) +
-  scale_x_continuous(
-    breaks = seq(-2200, -800, 200),
-    limits = c(-2500, -800)
+  scale_x_reverse(
+    breaks = seq(2200, 800, -200),
+    limits = c(2500, 800)
   )
 
 development_burial_construction_B <- development_burial_construction_B +

### 2019-03-17 analysis/code/04_plot_scripts/development_burial_type.R

@@ -30,6 +30,18 @@ gl <- lapply(region_file_list, function(x) {
 })
 dummy <- tibble::tibble(region_name = unique(amount_devel$region_name), grob = gl )
 
+# only for plot: time without sign of years
+amount_devel <- amount_devel %>%
+  dplyr::ungroup() %>%
+  dplyr::mutate(
+    timestep = timestep * (-1)
+  )
+
+prop <- prop %>%
+  dplyr::mutate(
+    timestep = timestep * (-1)
+  )
+
 #### development_burial_type_A ####
 development_burial_type_A <- ggplot() +
   geom_area(
@@ -67,9 +79,9 @@ development_burial_type_A <- ggplot() +
   coord_cartesian(
     ylim = c(0, 80)
   ) +
-  scale_x_continuous(
-    breaks = seq(-2200, -800, 200),
-    limits = c(-2500, -800)
+  scale_x_reverse(
+    breaks = seq(2200, 800, -200),
+    limits = c(2500, 800)
   ) +
   # guides(
   #   fill = FALSE
@@ -123,9 +135,9 @@ development_burial_type_B <- ggplot() +
     breaks = c(0, 0.5, 1),
     labels = c("0%", "50%", "100%")
   ) +
-  scale_x_continuous(
-    breaks = seq(-2200, -800, 200),
-    limits = c(-2500, -800)
+  scale_x_reverse(
+    breaks = seq(2200, 800, -200),
+    limits = c(2500, 800)
   )
 
 development_burial_type_B <- development_burial_type_B +

### 2019-03-17 analysis/code/04_plot_scripts/sed_region_matrix.R

@@ -13,6 +13,12 @@ load("analysis/data/tmp_data/region_order.RData")
 
 regions_grid <- rbind(regions_grid_burial_construction, regions_grid_burial_type)
 
+# only for plot: time without sign of years
+regions_grid <- regions_grid %>%
+  dplyr::mutate(
+    time = time * (-1)
+  )
+
 plot_list <- list()
 combinations <- c()
 iter <- 1
@@ -22,9 +28,9 @@ for (row_region in region_order) {
     if (row_region == col_region) {
       plot_list[[iter]] <- ggplot() +
         ylim(0,2) +
-        scale_x_continuous(
-          breaks = c(-2000, -1500, -1000),
-          limits = c(-2200, -800)
+        scale_x_reverse(
+          breaks = c(2000, 1500, 1000),
+          limits = c(2200, 800)
         )
     } else {
 
@@ -75,9 +81,9 @@ for (row_region in region_order) {
       }
 
       plot_list[[iter]] <- plot_starter +
-        scale_x_continuous(
-          breaks = c(-2000, -1500, -1000),
-          limits = c(-2200, -800)
+        scale_x_reverse(
+          breaks = c(2000, 1500, 1000),
+          limits = c(2200, 800)
         ) +
         scale_y_continuous(
           limits = c(0, 2)

### 2019-03-17 analysis/code/01_data_preparation_scripts/02_prepare_c14_data.R
n() -> dplyr::n()

@@ -225,7 +225,7 @@ load("analysis/data/tmp_data/bronze16.RData")
 # take a look at the dates per feature
 # bronze16 %>%
 #   dplyr::group_by(site, feature) %>%
-#   dplyr::filter(n()>1)
+#   dplyr::filter(dplyr::n()>1)
 
 # merge information
 bronze17 <- bronze16 %>%

### 2019-03-17 analysis/code/01_data_preparation_scripts/03_construct_circle_regions.R

@@ -34,7 +34,7 @@ schnu <- sf::st_intersection(gu, region_circles)
 number_of_graves_per_circle <- schnu %>%
   dplyr::group_by(ID) %>%
   dplyr::summarise(
-    n = n()
+    n = dplyr::n()
   )
 
 regions_with_enough_graves <- number_of_graves_per_circle %>%

### 2019-03-17 analysis/code/01_data_preparation_scripts/09_calculate_burial_rite_proportions_regions_timeseries.R

@@ -43,7 +43,7 @@ proportion_per_region <- dates_probability_per_year_and_region_list %>%
         bt <- bt_basic %>%
           dplyr::group_by(age, burial_type) %>%
           dplyr::summarise(
-            count = n(), region_name = .$region_name[1]
+            count = dplyr::n(), region_name = .$region_name[1]
           ) %>%
           tidyr::spread(
             key = burial_type, value = count
@@ -77,7 +77,7 @@ proportion_per_region <- dates_probability_per_year_and_region_list %>%
         bc <- bc_basic %>%
           dplyr::group_by(age, burial_construction) %>%
           dplyr::summarise(
-            count = n(), region_name = .$region_name[1]
+            count = dplyr::n(), region_name = .$region_name[1]
           ) %>%
           tidyr::spread(
             key = burial_construction, value = count

### 2019-03-17 analysis/code/03_dataset_description_scripts/observations_during_data_preparation.R

@@ -53,20 +53,20 @@ txtstorage::store("bronze15 labnrs amount", labnrs_amount, storage_file)
 # count labnr duplicates without n/a labnrs
 labnr_doubles <- bronze15[!grepl('n/a', bronze15$labnr), ] %>%
   dplyr::group_by(labnr) %>%
-  dplyr::filter(n() > 1) %>%
+  dplyr::filter(dplyr::n() > 1) %>%
   nrow()
 txtstorage::store("bronze15 labnr doubles", labnr_doubles, storage_file)
 
 # count graves represented by multiple c14 dates
 multi_dates_one_grave <- bronze15 %>%
   dplyr::group_by(site, feature) %>%
-  dplyr::filter(n() > 1) %>%
+  dplyr::filter(dplyr::n() > 1) %>%
   nrow()
 txtstorage::store("bronze15 multi dates one grave", multi_dates_one_grave, storage_file)
 
 bronze15_burial_type_doubles <- bronze15 %>%
   dplyr::group_by(site, feature) %>%
-  dplyr::filter(n() > 1) %>%
+  dplyr::filter(dplyr::n() > 1) %>%
   dplyr::ungroup() %$%
   burial_type %>% table %>%
   unclass %>%
@@ -75,7 +75,7 @@ txtstorage::store("bronze15 burial_type doubles", bronze15_burial_type_doubles, 
 
 bronze15_burial_construction_doubles <- bronze15 %>%
   dplyr::group_by(site, feature) %>%
-  dplyr::filter(n() > 1) %>%
+  dplyr::filter(dplyr::n() > 1) %>%
   dplyr::ungroup() %$%
   burial_construction %>% table %>%
   unclass %>%
@@ -96,8 +96,8 @@ txtstorage::store("size bronze16", nrow(bronze16), storage_file)
 # count the dates per feature - get max
 max_dates_per_grave <- bronze16[grepl("[0-9]", bronze16$feature), ] %>%
   dplyr::group_by(site, feature) %>%
-  # dplyr::filter(n()>1)
-  dplyr::summarise(n = n()) %>%
+  # dplyr::filter(dplyr::n()>1)
+  dplyr::summarise(n = dplyr::n()) %>%
   # dplyr::arrange(desc(n)) %>%
   dplyr::ungroup() %$%
   max(n)
@@ -105,7 +105,7 @@ txtstorage::store("bronze16 max dates per grave", max_dates_per_grave, storage_f
 
[…]

### 2019-03-17 article/01_Introduction.Rmd
referenced the relevant work of Capuzzo et al.

@@
 The present article, in contrast, explores the former strategy.
-The first major question therefore is: *Can the main trends in the distribution of burial rites in Bronze Age Europe be detected in openly available bulk radiocarbon data?* This approach could set a precedent for a more reproducible and statistically sound research practice.
-If we could answer this first question with *yes*, then the quantitative comparison of macro regional trends could pave the way for an analysis of interaction intensity on a long-term continental scale: *Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?*
+The first major question therefore is: *Can the main trends in the distribution of burial rites in Bronze Age Europe be detected in openly available bulk radiocarbon data?*
+
+The diffusion of the cremation burial tradition in the middle and late Bronze Age with radiocarbon data was already the subject of promising research in Central and South-Western Europe with focus on Catalonia and Switzerland [@capuzzoCulturalChangesSecond2015; @capuzzoInhumacionCremacionNordeste2017; @capuzzoSpacetemporalAnalysisRadiocarbon2014].
+This approach based on the dedicated radiocarbon database EUBAR [@capuzzoEUBARDatabase14C2013] sets a precedent for a more reproducible and statistically sound research practice concerning diffusion of innovation in Bronze Age archaeology.
+
+If radiocarbon data allows to reconstruct the diffusion of burial traditions, then the quantitative comparison of macro regional trends could also pave the way for an analysis of interaction intensity on a long-term continental scale.
+The second main question: *Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?*
 

### 2019-03-17 review/answer_to_review.md
no negative numbers in x-Axis labels
referenced the relevant work of Capuzzo et al.
concerning the limitation of the data

@@
 
+I was not aware of this work and added a paragraph in the introduction to reference it. It is indeed relevant.
+
 > In the paper, it’s very high the amount of data, for which the archaeological information regarding the type of burial and funerary rite is lacking.
@@
 In particular, when the statistical analyses carried out on data aim to explore in the details the cultural distance among the different regions.
+
+This is absolutely true and I tried to emphasise the limitations of the data in the discussion.
+The challenges are not confined on the amount of data, but also concern its meaning and origin.
+The supplementary remarks are extensively long to cover these problems.
+Is it necessary to mention the limitations at other critical points of the paper?
+I'm absolutely open to do that.
 
@@
 > Regarding the images, it would be preferable not to use negative number in order to indicate years BC.
+
+Done.
 
@@
 I also don’t see much the point of testing for significance the results of each simulation run - they are effectively samples from the same generative process so I think testing them does not make much sense?
+
+Figure 6 is much simpler now without simulation results. It only contains a time series of distance correlation.

### 2019-03-24 review/answer_to_review.md
rephrased second main question

@@
 
+The decision to create discrete spatial area classes is necessary to establish subsets of dates that can be used for time series calculation.
+I'm not aware of a methodologival approach where time series could be constructed without some sort of aggregation.
+Unfortunately the size of the areas has to be very big because of the currently overall small amount of data.
+I expect this to change in the future and the analysis can be redone then with more data and much smaller areas to understand the definitely much more complex processes within the macro regions established here.
+
+So: The criticism listed here is valid and I added a paragraph in the *Data quality* section to mention it.
+I refer to this section multiple times in the article.
+Nevertheless I don't see a practical possibility
+
 > Thirdly, the author implies that the scripts used for his analysis are supplied but there are no links on the article to any repository and I did not find any electronic supplementary material.
 As the author is well aware, we need to do our best to make our work fully reproducible.
+
+This is the result of a misunderstanding in the communication with the special issue editors. I added the link to the relevant repository now.
 
@@
 > Page 1, Column 2, Lines 52-43: “Do time series ... spatial distance”. I think this question should be phrased better. Eg. What  does “meaningful” mean in this context?
+
+Rephrased to a more direct question: "How does the cultural distance between regions that results from changes in burial traditions develop, and how is this cultural distance related to spatial distance?".
 

### 2019-03-24 article/01_Introduction.Rmd
rephrased second main question

@@
 If radiocarbon data allows to reconstruct the diffusion of burial traditions, then the quantitative comparison of macro regional trends could also pave the way for an analysis of interaction intensity on a long-term continental scale.
-The second main question: *Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?*
+The second main question: *How does the cultural distance between regions that results from changes in burial traditions develop, and how is this cultural distance related to spatial distance?*
 

### 2019-03-24 article/04_Discussion.Rmd
rephrased second main question

@@
 
-The second important question posed for this paper was, *if these time series constructed from radiocarbon data* (Figure \ref{development_burial_type} and \ref{development_burial_construction}) *allow further analysis concerning cultural and spatial distance*?
-That can also be affirmed.
-While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one had a larger scope.
+The second important question posed for this paper was, *how cultural distance between regions that results from changes in burial traditions develops, and how it is related to spatial distance?* While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one had a larger scope.
 The quantitative representation of an archaeological narrative paths the way for a diverse set of methods to be applied and the analysis undertaken here is merely the beginning.

### 2019-03-24 review/answer_to_review.md
Rephrasing.

@@
 
+TODO
+
 > Thirdly, the author implies that the scripts used for his analysis are supplied but there are no links on the article to any repository and I did not find any electronic supplementary material.
@@
 
-Rephrased to a more direct question: "How does the cultural distance between regions that results from changes in burial traditions develop, and how is this cultural distance related to spatial distance?".
+Rephrased: "How does the cultural distance between regions that results from changes in burial traditions develop, and how is this cultural distance related to spatial distance?"
 
@@
 
+Rephrased: "This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries, as it assumes *Social learning* with transmission, selection and mutation of ideas [...]"
+
 > Page 2, Column 1, Line 15: “explicitly adopted the terminology of Cultural evolution”. This reads as if only the terminology is adopted? Rephrase?
+
+Rephrased: "Evolutionary archaeology, the branch of archaeology that explicitly adopted Cultural evolution theory, [...]"
 
@@
 
+Rephrased: "[...] *drift* (in case of multiple, equivalent traits a single one will dominate over time due to random effects) [...]"
+
 > Page 3, Column 1, Line 33. Please provide a link to the repository with the code.
+
+TODO
 

### 2019-03-24 article/01_Introduction.Rmd
Rephrasing.

@@
 They are therefore suitable for study using methods of *Cultural transmission theory*, a special aspect of *Cultural evolution theory* concerned with the spread of ideas in human populations [@creanzaCulturalEvolutionaryTheory2017].
-This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries, as it assumes *Social learning* to produce adaptive behaviour through transmission, selection and mutation -- a secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007; @lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
+This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries, as it assumes *Social learning* with transmission, selection and mutation of ideas -- a secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007; @lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
 
-*Evolutionary archaeology*, the branch of archaeology that explicitly adopted the terminology of Cultural evolution, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is very relevant for burial traditions: Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
+*Evolutionary archaeology*, the branch of archaeology that explicitly adopted Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is very relevant for burial traditions: Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
 Without resorting to the inappropriate simplifications of Evolutionism, their prevalence generally happens to increase over time.
@@
 @neimanStylisticVariationEvolutionary1995 takes a closer look at these processes and develops a simple framework for the simulation of non-functional idea propagation.
-He identifies the essential phenomena *drift* (random dominance of similar traits) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.
+He identifies the essential phenomena *drift* (in case of multiple, equivalent traits a single one will dominate over time due to random effects) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.
 Furthermore, he implements a measure of cultural distance to apply his theoretical insights to real-world data.

### 2019-03-24 review/answer_to_review.md
dating problems

@@
 
+Please see the reply above.
+
 > Page 3, Column 2, Line 60.
@@
 Also, the paragraph mentions “bones and teeth”, are there any risks of aquatic diet determining a reservoir effect?
+
+These dates were not excluded to avoid further reduction of the amount of data.
+It is not possible to consider these dating effects on this scale, as it would require to study all radiocarbon dates and their archaeological context individually.
+I agree that it might be necessary to do so, but this is not an endeavour that can be undertaken without long-term project funding.
+Marine reservoir effect is only one of the possible problems relevant here in addition to the Old wood effect or the human 14C-residence ("adult") effect.
+I added a paragraph abou this in the *Data quality* section.
+
+TODO
 

### 2019-03-24 review/answer_to_review.md
went through the remaining review and decided what to do today

@@
 
+This is a good idea and I will try this in the future.
+For this paper I decided not to do so, to increase the readability of the time series of absolute amounts (Figure 3&4 A).
+Each grave represents the presence of an idea during a certain period of time and the presence of this idea may very well be assumed before and after the event of the actual burial.
+
 > Page 4, Columns 1-2, Lines 58-60; 1-8.
@@
 
+I'm very interested in methods to assign a measure of confidence to the results for individual regions and times.
+Figure 3&4 are deliberately constructed to juxtapose the relative proportion with the amount of graves available.
+I understand this as an object of future research.
+I reworked Table 1 and Figure 3&4 for a better presentation of the sample size per region.
+
 > Page 4, Column 2, Lines 43-56. I think it is worth computing partial mantel tests so that the correlation between burial type and construction can take into account geography.
 
+Done and included in Table 2.
+
 > Page 5, Column 1, Line 24. Perhaps worth making this a subsection?
+
+I removed the simulation. See my reply above.
 
@@
 
+I will consider this very true questions in a future publication.
+
 > Figure 3. These are very nice plots - I would add the sample sizes for each region though.
+
+Done.
 

### 2019-03-24 article/02_Materials_and_Methods.Rmd
renamed plot
replaced deprecated function

@@
     dplyr::vars(dplyr::matches("Mean|Correlation")),
-    dplyr::funs(kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black"))) 
+    list(~kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black"))) 
   ) %>%
@@
     dplyr::vars(dplyr::matches("p-Value")),
-    dplyr::funs(kableExtra::cell_spec(., "latex", bold = . < 0.1)) 
+    list(~kableExtra::cell_spec(., "latex", bold = . < 0.1)) 
   ) %>%
@@
 \centering
-\includegraphics[width=\columnwidth]{../analysis/figures/sed_time_series.jpeg}
+\includegraphics[width=\columnwidth]{../analysis/figures/correlation_time_series.jpeg}
 \caption{\label{simulation}Development of correlation between cultural and spatial distance.

### 2019-03-24 analysis/code/01_data_preparation_scripts/11_get_burial_rite_regions_distribution_at_2200calBC.R
start distribution calculation was only necessary for the simulation

@@ -1,41 +1,0 @@
-#### burial type ####
-
-load( "analysis/data/tmp_data/development_proportions_burial_type.RData")
-
-start_proportion_burial_type <- proportion_development_burial_type %>%
-  dplyr::filter(timestep == -2200) %>%
-  tidyr::spread(idea, proportion) %>%
-  dplyr::arrange(region_name) %>%
-  magrittr::set_rownames(.$region_name) %>%
-  dplyr::select(cremation, inhumation) %>%
-  dplyr::rename(
-    idea_1 = cremation,
-    idea_2 = inhumation
-  )
-
-save(
-  start_proportion_burial_type,
-  file = "analysis/data/tmp_data/start_proportion_burial_type.RData"
-)
-
-
-
-#### burial construction ####
-
-load("analysis/data/tmp_data/development_proportions_burial_construction.RData")
-
-start_proportion_burial_construction <- proportion_development_burial_construction %>%
-  dplyr::filter(timestep == -2200) %>%
-  tidyr::spread(idea, proportion) %>%
-  dplyr::arrange(region_name) %>%
-  magrittr::set_rownames(.$region_name) %>%
-  dplyr::select(flat, mound) %>%
-  dplyr::rename(
-    idea_1 = flat,
-    idea_2 = mound
-  )
-
-save(
-  start_proportion_burial_construction,
-  file = "analysis/data/tmp_data/start_proportion_burial_construction.RData"
-)

### 2019-03-24 article/02_Materials_and_Methods.Rmd
code refactoring

@@
 
-2. Removing all dates without age information or age information outside of the limits of the Intcal13 calibration curve: 71 - 46401 BP [@reimer_intcal13_2013]. (`r r("size bronze", sf_prep)` dates)
+2.
+Removing all dates without age information or age information outside of the limits of the Intcal13 calibration curve: 71 - 46401 BP [@reimer_intcal13_2013].
+(`r r("size dates_calibrated", sf_prep)` dates)
 
@@
 Reducing date selection to all dates whose $2\sigma$ calibrated age range overlap by at least one year with the predefined window of interest: 2200 -- 800 calBC.
-(`r r("size bronze0", sf_prep)` dates)
+(`r r("size dates_time_selection", sf_prep)` dates)
 
-5. Reducing the date selection to all dates taken from graves as indicated by the *feature type* variable. (`r r("size bronze05", sf_prep)` dates)
+5. Reducing the date selection to all dates taken from graves as indicated by the *feature type* variable. (`r r("size dates_research_selection", sf_prep)` dates)
 
@@
 
-8. Reducing date selection to dates within the research area. (`r r("size bronze15", sf_prep)` dates)
+8. Reducing date selection to dates within the research area. (`r r("size dates_research_area", sf_prep)` dates)
 
-9. Removing ^14^C date doubles with tools of c14bazAAR. (`r r("size bronze16", sf_prep)` dates)
+9. Removing ^14^C date doubles with tools of c14bazAAR. (`r r("size dates_prepared", sf_prep)` dates)
 
@@
 Transforming data from the meaning *^14^C dates on graves* to *burial type at a certain time*.
-For `r r("bronze15 multi dates one grave", sf_prep)` entries in the thus far assembled subset the attribution to *site* and *feature* is not unique.
+For `r r("dates_research_area multi dates one grave", sf_prep)` entries in the thus far assembled subset the attribution to *site* and *feature* is not unique.
 That means that for many graves multiple (up to 8) ^14^C dates have been created (and entered in Radon-B) and dates therefore cannot be treated as semantically equal to graves.
@@
 The approach is compatible with the intention to create a cultural trait time series as explained below.
-(`r r("size bronze17", sf_prep)` graves)
+(`r r("size graves_prepared", sf_prep)` graves)
 
@@
     2.
-For the remaining `r r("bronze16 multi dates one grave with numbers", sf_prep)` problematic cases, groups were created by *site* and *feature*.
+For the remaining `r r("dates_prepared multi dates one grave with numbers", sf_prep)` problematic cases, groups were created by *site* and *feature*.
 The cumulated timespan covered by the $2\sigma$ age range of all these dates were taken as the relevant timespan for the respective grave.

### 2019-03-24 analysis/code/01_data_preparation_scripts/02_prepare_c14_data.R

@@ -18,24 +18,24 @@ save(radonb, file = "analysis/data/tmp_data/radonb.RData")
 
 
 
-#### calibration ####
+#### 14C calibration ####
 
 load("analysis/data/tmp_data/radonb.RData")
 
-bronze <- radonb %>%
+dates <- radonb %>%
   tibble::as.tibble() %>%
   # remove dates without age
   dplyr::filter(!is.na(c14age) & !is.na(c14std)) %>%
   # remove dates outside of theoretical calibration range
   dplyr::filter(!(c14age < 71) & !(c14age > 46401))
 
-bronze <- bronze %>%
+dates_calibrated <- dates %>%
   dplyr::mutate(
     # add list column with the age density distribution for every date
     calage_density_distribution = Bchron::BchronCalibrate(
-      ages      = bronze$c14age,
-      ageSds    = bronze$c14std,
-      calCurves = rep("intcal13", nrow(bronze)),
+      ages      = dates$c14age,
+      ageSds    = dates$c14std,
+      calCurves = rep("intcal13", nrow(dates)),
       eps       = 1e-06
     ) %>%
       # transform BchronCalibrate result to a informative tibble
@@ -61,18 +61,18 @@ bronze <- bronze %>%
 
 #### transform calBP age to calBC ####
 
-bronze$calage_density_distribution %<>% lapply(
+dates_calibrated$calage_density_distribution %<>% lapply(
   function(x) {
     x$age = -x$age + bol
     return(x)
   }
 )
 
-save(bronze, file = "analysis/data/tmp_data/bronze.RData")
+save(dates_calibrated, file = "analysis/data/tmp_data/dates_calibrated.RData")
 
-# plot to check the calibration result
+# test plot to check the calibration result
 # library(ggplot2)
-# bronze$calage_density_distribution[[3]] %>%
+# dates_calibrated$calage_density_distribution[[3]] %>%
 #   ggplot() +
 #   geom_point(aes(age, norm_dens, color = two_sigma))
 
@@ -80,16 +80,16 @@ save(bronze, file = "analysis/data/tmp_data/bronze.RData")
 
 #### filter time ####
 
-load("analysis/data/tmp_data/bronze.RData")
+load("analysis/data/tmp_data/dates_calibrated.RData")
 
-# add artifical id
-bronze <- bronze %>%
+# add artifical date id
[…]

### 2019-03-24 analysis/code/01_data_preparation_scripts/03_construct_circle_regions.R

@@ -1,99 +1,0 @@
-#### load spatial data ####
-
-load("analysis/data/tmp_data/research_area.RData")
-load("analysis/data/tmp_data/area.RData")
-
-#### define region circles ####
-
-region_circles <- tibble::tibble(
-  geometry = sf::st_make_grid(area, 400000, what = "centers", offset = c(-900000,-130000)),
-  ID = 1:length(geometry)
-  ) %>% sf::st_as_sf()
-region_circles <- sf::st_intersection(region_circles, research_area)
-region_circles %<>% sf::st_buffer(dist = 240000)
-
-load("analysis/data/tmp_data/bronze17.RData")
-bronze17 %<>% sf::st_as_sf(coords = c("lon", "lat"))
-sf::st_crs(bronze17) <- 4326
-bronze17 %<>% sf::st_transform("+proj=aea +lat_1=43 +lat_2=62 +lat_0=30 +lon_0=10 +x_0=0 +y_0=0 +ellps=intl +units=m +no_defs")
-gu <- sf::st_intersection(bronze17, research_area)
-
-# library(ggplot2)
-# ggplot() +
-#   geom_sf(data = area) +
-#   geom_sf(data = region_circles, fill = NA) +
-#   geom_sf(data = gu)
-
-schnu <- sf::st_intersection(gu, region_circles)
-
-# ggplot() +
-#   geom_sf(data = area) +
-#   geom_sf(data = region_circles, fill = NA) +
-#   geom_sf(data = schnu)
-
-number_of_graves_per_circle <- schnu %>%
-  dplyr::group_by(ID) %>%
-  dplyr::summarise(
-    n = dplyr::n()
-  )
-
-regions_with_enough_graves <- number_of_graves_per_circle %>%
-  dplyr::filter(
-    n >= 60
-  ) %$%
-  ID
-
-graves_per_region <- schnu %>% dplyr::filter(
-  ID %in% regions_with_enough_graves
-) %>% sf::st_set_geometry(NULL)
-
-save(
-  graves_per_region,
-  file = "analysis/data/tmp_data/graves_per_region.RData"
-)
-
-regions <- region_circles %>%
-  dplyr::mutate(
-    number_of_graves = number_of_graves_per_circle$n
-  ) %>%
-  dplyr::filter(ID %in% regions_with_enough_graves)
-
-
-# ggplot() +
-#   geom_sf(data = area) +
[…]

### 2019-03-24 analysis/code/01_data_preparation_scripts/03_define_regions_order_and_colors.R

@@ -1,3 +1,5 @@
+#### order of regions ####
+
 region_order <- c(
   "Southeastern Central Europe",
   "Poland",
@@ -10,6 +12,10 @@ region_order <- c(
 )
 
 save(region_order, file = "analysis/data/tmp_data/region_order.RData")
+
+
+
+#### color of regions ####
 
 region_colors <- c(
   "Southeastern Central Europe" = "#999999",

### 2019-03-24 analysis/code/01_data_preparation_scripts/04_construct_circle_regions.R

@@ -1,0 +1,64 @@
+#### load data ####
+
+load("analysis/data/tmp_data/research_area.RData")
+load("analysis/data/tmp_data/area.RData")
+load("analysis/data/tmp_data/region_order.RData")
+
+
+
+#### construct region circles ####
+
+region_circles <- tibble::tibble(
+  geometry = sf::st_make_grid(area, 400000, what = "centers", offset = c(-900000,-130000)),
+  ID = 1:length(geometry)
+  ) %>% sf::st_as_sf()
+region_circles <- sf::st_intersection(region_circles, research_area)
+region_circles %<>% sf::st_buffer(dist = 240000)
+
+
+
+#### select regions with enough graves ####
+
+load("analysis/data/tmp_data/graves_prepared.RData")
+graves_prepared %<>% sf::st_as_sf(coords = c("lon", "lat"))
+sf::st_crs(graves_prepared) <- 4326
+graves_prepared %<>% sf::st_transform("+proj=aea +lat_1=43 +lat_2=62 +lat_0=30 +lon_0=10 +x_0=0 +y_0=0 +ellps=intl +units=m +no_defs")
+
+inter_area <- sf::st_intersection(graves_prepared, research_area) %>% dplyr::select(-id)
+inter_regions <- sf::st_intersection(inter_area, region_circles) %>% dplyr::select(-id)
+
+number_of_graves_per_circle <- inter_regions %>%
+  dplyr::group_by(ID) %>%
+  dplyr::summarise(
+    n = dplyr::n()
+  )
+
+regions_with_enough_graves <- number_of_graves_per_circle %>%
+  dplyr::filter(
+    n >= 60
+  ) %$%
+  ID
+
+
+
+#### define resulting regions with name ####
+
+regions <- region_circles %>%
+  dplyr::mutate(
+    number_of_graves = number_of_graves_per_circle$n
+  ) %>%
+  dplyr::filter(ID %in% regions_with_enough_graves)
+
+regions$ID <- 1:nrow(regions)
+regions$NAME <- c(
+  "Northeastern France",
+  "Southern Germany",
+  "Southeastern Central Europe",
+  "England", "Benelux",
+  "Northern Germany",
+  "Poland",
+  "Southern Scandinavia"
+  )
+regions$NAME <- factor(regions$NAME, levels = region_order)
+
+save(regions, file = "analysis/data/tmp_data/regions.RData")

### 2019-03-24 analysis/code/01_data_preparation_scripts/04_crop_c14_data_to_regions_area.R

@@ -1,29 +1,0 @@
-#### load spatial data ####
-
-load("analysis/data/tmp_data/regions.RData")
-load("analysis/data/tmp_data/bronze16.RData")
-
-#### prepare data ####
-
-bronze16 %<>% sf::st_as_sf(coords = c("lon", "lat"))
-sf::st_crs(bronze16) <- 4326
-bronze16 %<>% sf::st_transform("+proj=aea +lat_1=43 +lat_2=62 +lat_0=30 +lon_0=10 +x_0=0 +y_0=0 +ellps=intl +units=m +no_defs")
-
-#### intersect ####
-
-schnu <- sf::st_intersection(bronze16, regions)
-
-#### make tibble ####
-
-dates_per_region <- schnu %>% sf::st_set_geometry(NULL) %>%
-  dplyr::mutate(
-    region = NAME
-  ) %>%
-  dplyr::select(
-    -id.1, -ID, -NAME
-  )
-
-save(
-  dates_per_region,
-  file = "analysis/data/tmp_data/dates_per_region.RData"
-)

### 2019-03-24 analysis/code/01_data_preparation_scripts/05_attribute_dates_and_graves_to_regions.R

@@ -1,0 +1,51 @@
+#### load data ####
+
+load("analysis/data/tmp_data/regions.RData")
+load("analysis/data/tmp_data/dates_prepared.RData")
+load("analysis/data/tmp_data/graves_prepared.RData")
+
+
+
+#### attribute dates ####
+
+dates_prepared %<>% sf::st_as_sf(coords = c("lon", "lat"))
+sf::st_crs(dates_prepared) <- 4326
+dates_prepared %<>% sf::st_transform("+proj=aea +lat_1=43 +lat_2=62 +lat_0=30 +lon_0=10 +x_0=0 +y_0=0 +ellps=intl +units=m +no_defs")
+
+inter_regions_dates <- sf::st_intersection(dates_prepared, regions)
+
+dates_per_region <- inter_regions_dates %>% sf::st_set_geometry(NULL) %>%
+  dplyr::mutate(
+    region = NAME
+  ) %>%
+  dplyr::select(
+    -id, -ID, -NAME
+  )
+
+save(
+  dates_per_region,
+  file = "analysis/data/tmp_data/dates_per_region.RData"
+)
+
+
+
+#### attribute graves ####
+
+graves_prepared %<>% sf::st_as_sf(coords = c("lon", "lat"))
+sf::st_crs(graves_prepared) <- 4326
+graves_prepared %<>% sf::st_transform("+proj=aea +lat_1=43 +lat_2=62 +lat_0=30 +lon_0=10 +x_0=0 +y_0=0 +ellps=intl +units=m +no_defs")
+
+inter_regions_graves <- sf::st_intersection(graves_prepared, regions)
+
+graves_per_region <- inter_regions_graves %>% sf::st_set_geometry(NULL) %>%
+  dplyr::mutate(
+    region = NAME
+  ) %>%
+  dplyr::select(
+    -id, -ID, -NAME
+  )
+
+save(
+  graves_per_region,
+  file = "analysis/data/tmp_data/graves_per_region.RData"
+)

### 2019-03-24 analysis/code/01_data_preparation_scripts/06_create_regions_spatial_distance_matrix.R

@@ -1,6 +1,9 @@
-#### load regions data ####
+#### load data ####
 
 load("analysis/data/tmp_data/regions.RData")
+load("analysis/data/tmp_data/region_order.RData")
+
+
 
 #### create distance information in tall format ####
 
@@ -20,6 +23,7 @@ distance_matrix_spatial_long <- region_centers %>%
   # wide matrix to tall data.frame
   tidyr::gather(key = regionB, value = distance, -regionA) %>%
   dplyr::mutate(
+    regionB = factor(regionB, levels = region_order),
     distance = as.double(distance)
   ) %>%
   # creation of distance classes
@@ -41,25 +45,20 @@ distance_matrix_spatial_long <- region_centers %>%
     )
   )
 
-# remove duplicates
+save(distance_matrix_spatial_long, file = "analysis/data/tmp_data/distance_matrix_spatial_long.RData")
+
+
+
+#### remove duplicates ####
+
 mn <- pmin(as.character(distance_matrix_spatial_long$regionA), as.character(distance_matrix_spatial_long$regionB))
 mx <- pmax(as.character(distance_matrix_spatial_long$regionA), as.character(distance_matrix_spatial_long$regionB))
 int <- as.numeric(interaction(mn, mx))
 distance_matrix_spatial_long_half <- distance_matrix_spatial_long[match(unique(int), int),]
 
-#### define factor order ####
+save(distance_matrix_spatial_long_half, file = "analysis/data/tmp_data/distance_matrix_spatial_long_half.RData")
 
-load("analysis/data/tmp_data/region_order.RData")
 
-regions_factorA <- as.factor(distance_matrix_spatial_long$regionA)
-distance_matrix_spatial_long$regionA <- factor(regions_factorA, levels = region_order)
-regions_factorB <- as.factor(distance_matrix_spatial_long$regionB)
-distance_matrix_spatial_long$regionB <- factor(regions_factorB, levels = region_order)
-
-regions_factorA <- as.factor(distance_matrix_spatial_long_half$regionA)
-distance_matrix_spatial_long_half$regionA <- factor(regions_factorA, levels = region_order)
-regions_factorB <- as.factor(distance_matrix_spatial_long_half$regionB)
-distance_matrix_spatial_long_half$regionB <- factor(regions_factorB, levels = region_order)
 
 #### transform distance information to wide format ####
 
[…]

### 2019-03-24 analysis/code/01_data_preparation_scripts/07_render_region_pictograms.R

@@ -1,11 +1,14 @@
+#### load data ####
+
 load("analysis/data/tmp_data/regions.RData")
 load("analysis/data/tmp_data/extended_area.RData")
 extended_area <- extended_area$geometry
-
-#### plot ####
-
 load("analysis/data/tmp_data/region_order.RData")
 load("analysis/data/tmp_data/region_colors.RData")
+
+
+
+#### plot loop ####
 
 path = "analysis/data/tmp_data/"
 
@@ -39,4 +42,3 @@ for (i in 1:nrow(regions)) {
   dev.off()
 
 }
-

### 2019-03-24 analysis/code/01_data_preparation_scripts/08_attribute_c14_data_to_regions.R

@@ -1,45 +1,0 @@
-#### load data ####
-
-load("analysis/data/tmp_data/regions.RData")
-load("analysis/data/tmp_data/bronze2.RData")
-load("analysis/data/tmp_data/region_order.RData")
-
-# transform to sf
-bronze_sf <- bronze2 %>%
-  sf::st_as_sf(
-    coords = c("lon", "lat"), crs = 4326
-  ) %>% sf::st_transform("+proj=aea +lat_1=43 +lat_2=62 +lat_0=30 +lon_0=10 +x_0=0 +y_0=0 +ellps=intl +units=m +no_defs")
-
-
-
-#### dates per region ####
-
-# intersect and get region id per entry
-region_index_of_date <- bronze_sf %>% sf::st_intersects(regions) %>%
-  sapply(function(z) if (length(z)==0) NA_integer_ else z[1])
-
-dates_probability_per_year_and_region_list <- bronze2 %>%
-  # add region information to bronze2
-  dplyr::mutate(
-    region_name = factor(regions$NAME[region_index_of_date], levels = region_order)
-  ) %>%
-  # remove entries without (outside of) regions
-  dplyr::filter(
-    !is.na(region_name)
-  ) %>%
-  # split datasets by region name
-  split(.$region_name)
-
-save(
-  dates_probability_per_year_and_region_list,
-  file = "analysis/data/tmp_data/dates_probability_per_year_and_region_list.RData"
-)
-
-# merge per-region data.frame list again to one dataframe
-dates_probability_per_year_and_region_df <- dates_probability_per_year_and_region_list %>%
-  dplyr::bind_rows()
-
-save(
-  dates_probability_per_year_and_region_df,
-  file = "analysis/data/tmp_data/dates_probability_per_year_and_region_df.RData"
-)

### 2019-03-24 analysis/code/01_data_preparation_scripts/08_graves_per_year_and_region.R

@@ -1,0 +1,24 @@
+#### load data ####
+
+load("analysis/data/tmp_data/graves_per_region.RData")
+
+#### unnest by year ####
+
+# unnest calage_density_distribution to have per year information:
+# a diachron perspective
+graves_per_year_and_region <- graves_per_region %>%
+  tidyr::unnest(calage_density_distribution) %>%
+  dplyr::filter(
+    two_sigma == TRUE
+  ) %>%
+  dplyr::filter(
+    age >= -2200 & age <= -800
+  ) %>%
+  dplyr::arrange(
+    desc(burial_construction)
+  )
+
+save(graves_per_year_and_region, file = "analysis/data/tmp_data/graves_per_year_and_region.RData")
+
+graves_per_year_and_region_list <- graves_per_year_and_region %>% base::split(.$region)
+save(graves_per_year_and_region_list, file = "analysis/data/tmp_data/graves_per_year_and_region_list.RData")

### 2019-03-24 analysis/code/01_data_preparation_scripts/09_calculate_burial_rite_proportions_regions_timeseries.R

@@ -1,4 +1,8 @@
-load("analysis/data/tmp_data/dates_probability_per_year_and_region_list.RData")
+#### load data ####
+
+load("analysis/data/tmp_data/graves_per_year_and_region_list.RData")
+
+
 
 #### calculate per year, per region distribution of ideas ####
 
@@ -10,7 +14,7 @@ fncols <- function(data, cname) {
 }
 
 # main loop
-proportion_per_region <- dates_probability_per_year_and_region_list %>%
+proportion_per_region <- graves_per_year_and_region_list %>%
   # apply per region data.frame
   pbapply::pblapply(function(x) {
 
@@ -34,7 +38,7 @@ proportion_per_region <- dates_probability_per_year_and_region_list %>%
 
       if (nrow(bt_basic) == 0) {
         bt <- tibble::tibble(
-          region_name = character(),
+          region = character(),
           age = integer(),
           cremation = double(),
           inhumation = double()
@@ -43,8 +47,9 @@ proportion_per_region <- dates_probability_per_year_and_region_list %>%
         bt <- bt_basic %>%
           dplyr::group_by(age, burial_type) %>%
           dplyr::summarise(
-            count = dplyr::n(), region_name = .$region_name[1]
+            count = dplyr::n(), region = .$region[1]
           ) %>%
+          dplyr::ungroup() %>%
           tidyr::spread(
             key = burial_type, value = count
           ) %>%
@@ -55,8 +60,7 @@ proportion_per_region <- dates_probability_per_year_and_region_list %>%
           ) %>%
           dplyr::mutate(
             inhumation = 1 - cremation
-          ) %>%
-          dplyr::ungroup()
+          )
       }
 
       #### burial_type: mound vs. flat ####
@@ -68,7 +72,7 @@ proportion_per_region <- dates_probability_per_year_and_region_list %>%
 
       if (nrow(bc_basic) == 0) {
         bc <- tibble::tibble(
-          region_name = character(),
+          region = character(),
           age = integer(),
           mound = double(),
           flat = double()
@@ -77,8 +81,9 @@ proportion_per_region <- dates_probability_per_year_and_region_list %>%
         bc <- bc_basic %>%
           dplyr::group_by(age, burial_construction) %>%
           dplyr::summarise(
-            count = dplyr::n(), region_name = .$region_name[1]
+            count = dplyr::n(), region = .$region[1]
           ) %>%
+          dplyr::ungroup() %>%
[…]

### 2019-03-24 analysis/code/01_data_preparation_scripts/10_calculate_burial_rite_amount_regions_timeseries.R

@@ -1,9 +1,14 @@
-#### calculate
+#### load data ####
 
-load("analysis/data/tmp_data/dates_probability_per_year_and_region_df.RData")
+load("analysis/data/tmp_data/graves_per_year_and_region.RData")
 
-amount_development_burial_type_without_zero <- dates_probability_per_year_and_region_df %>%
-  dplyr::group_by(region_name, age, burial_type) %>%
+
+
+#### calculate per year, per region appearance of graves ####
+
+# burial type
+amount_development_burial_type_without_zero <- graves_per_year_and_region %>%
+  dplyr::group_by(region, age, burial_type) %>%
   dplyr::rename(
     timestep = age,
     idea = burial_type
@@ -13,12 +18,12 @@ amount_development_burial_type_without_zero <- dates_probability_per_year_and_re
 amount_development_burial_type <- amount_development_burial_type_without_zero %>%
   dplyr::right_join(
     expand.grid(
-      region_name = unique(amount_development_burial_type_without_zero$region_name),
+      region = unique(amount_development_burial_type_without_zero$region),
       idea = unique(amount_development_burial_type_without_zero$idea),
       timestep = -2200:-800,
       stringsAsFactors = FALSE
     ),
-    by = c("region_name", "timestep", "idea")
+    by = c("region", "timestep", "idea")
   ) %>%
   dplyr::mutate(
     n = replace(n, is.na(n), 0)
@@ -29,8 +34,9 @@ save(
   file = "analysis/data/tmp_data/development_amount_burial_type.RData"
 )
 
-amount_development_burial_construction_without_zero <- dates_probability_per_year_and_region_df %>%
-  dplyr::group_by(region_name, age, burial_construction) %>%
+# burial construction
+amount_development_burial_construction_without_zero <- graves_per_year_and_region %>%
+  dplyr::group_by(region, age, burial_construction) %>%
   dplyr::rename(
     timestep = age,
     idea = burial_construction
@@ -40,12 +46,12 @@ amount_development_burial_construction_without_zero <- dates_probability_per_yea
 amount_development_burial_construction  <- amount_development_burial_construction_without_zero%>%
   dplyr::right_join(
     expand.grid(
-      region_name = unique(amount_development_burial_construction_without_zero$region_name),
+      region = unique(amount_development_burial_construction_without_zero$region),
       idea = unique(amount_development_burial_construction_without_zero$idea),
[…]

### 2019-03-24 analysis/code/02_cultural_distance_scripts/02b_burial_type_sed.R

@@ -8,7 +8,7 @@ long_prop <- prop %>%
     idea, proportion
   )
 
-regions <- prop$region_name %>% unique()
+regions <- prop$region %>% unique()
 timesteps <- prop$timestep %>% unique()
 
 regions_grid <-
@@ -19,11 +19,11 @@ regions_grid <-
   tibble::as.tibble() %>%
   dplyr::left_join(
     long_prop,
-    by = c("regionA" = "region_name", "time" = "timestep")
+    by = c("regionA" = "region", "time" = "timestep")
   ) %>%
   dplyr::left_join(
     long_prop,
-    by = c("regionB" = "region_name", "time" = "timestep"),
+    by = c("regionB" = "region", "time" = "timestep"),
     suffix = c("_regionA", "_regionB")
   )
 

### 2019-03-24 analysis/code/03_dataset_description_scripts/observations_during_data_preparation.R

@@ -2,122 +2,109 @@ storage_file <- "analysis/data/output_data/sf_prep.txt"
 
 
 
-#### bronze ####
+#### early subsets ####
+
 load("analysis/data/tmp_data/radonb.RData")
 txtstorage::store("size radonb", nrow(radonb), storage_file)
 rm(radonb)
 
+load("analysis/data/tmp_data/dates_calibrated.RData")
+txtstorage::store("size dates_calibrated", nrow(dates_calibrated), storage_file)
+rm(dates_calibrated)
 
+load("analysis/data/tmp_data/dates_time_selection.RData")
+txtstorage::store("size dates_time_selection", nrow(dates_time_selection), storage_file)
+rm(dates_time_selection)
 
-#### bronze ####
-load("analysis/data/tmp_data/bronze.RData")
-txtstorage::store("size bronze", nrow(bronze), storage_file)
-rm(bronze)
+load("analysis/data/tmp_data/dates_research_selection.RData")
+txtstorage::store("size dates_research_selection", nrow(dates_research_selection), storage_file)
+txtstorage::store("dates_research_selection variable amount", ncol(dates_research_selection), storage_file)
+rm(dates_research_selection)
 
-
-
-#### bronze0 ####
-load("analysis/data/tmp_data/bronze0.RData")
-txtstorage::store("size bronze0", nrow(bronze0), storage_file)
-rm(bronze0)
-
-
-
-#### bronze05 ####
-load("analysis/data/tmp_data/bronze05.RData")
-txtstorage::store("size bronze05", nrow(bronze05), storage_file)
-txtstorage::store("bronze05 variable amount", ncol(bronze05), storage_file)
-rm(bronze05)
+load("analysis/data/tmp_data/dates_coordinates.RData")
+txtstorage::store("size dates_coordinates", nrow(dates_coordinates), storage_file)
+txtstorage::store("dates_coordinates variable amount", ncol(dates_coordinates), storage_file)
+rm(dates_coordinates)
 
 
 
-#### bronze1 ####
-load("analysis/data/tmp_data/bronze1.RData")
-txtstorage::store("size bronze1", nrow(bronze1), storage_file)
-txtstorage::store("bronze1 variable amount", ncol(bronze1), storage_file)
-rm(bronze1)
+#### dates_research_area ####
 
-
-
-#### bronze15 ####
-
-load("analysis/data/tmp_data/bronze15.RData")
[…]

### 2019-03-24 analysis/code/03_dataset_description_scripts/observations_on_final_dataset.R

@@ -50,7 +50,20 @@ txtstorage::store(
 
 dpr$species %>% table(useNA = "always")
 
+# overview table
+# dprcrosstab <- dpr %>%
+#   dplyr::group_by(
+#     region, burial_type, burial_construction
+#   ) %>%
+#   dplyr::summarise(
+#     n = dplyr::n()
+#   ) %>%
+#   dplyr::ungroup() %>%
+#   tidyr::spread(
+#     key = burial_construction, value = n
+#   )
 dprcrosstab <- table(dpr$burial_type, dpr$burial_construction)
+
 save(
   dprcrosstab,
   file = "analysis/data/output_data/dprcrosstab.RData"

### 2019-03-24 analysis/code/04_plot_scripts/development_burial_construction.R

@@ -11,8 +11,8 @@ load("analysis/data/tmp_data/development_proportions_burial_construction.RData")
 
 #### prepare data ####
 amount_devel <- amount_development_burial_construction
-regions_factor <- as.factor(amount_devel$region_name)
-amount_devel$region_name <- factor(regions_factor, levels = region_order)
+regions_factor <- as.factor(amount_devel$region)
+amount_devel$region <- factor(regions_factor, levels = region_order)
 idea_factor <- as.factor(amount_devel$idea)
 amount_devel$idea <- factor(idea_factor, levels = rev(levels(idea_factor)))
 
@@ -20,7 +20,7 @@ prop <- proportion_development_burial_construction
 prop$idea <- as.factor(prop$idea)
 prop$idea <- factor(prop$idea , levels = rev(levels(prop$idea )))
 
-region_file_list <- unique(amount_devel$region_name) %>% gsub(" ", "_", ., fixed = TRUE)
+region_file_list <- unique(amount_devel$region) %>% gsub(" ", "_", ., fixed = TRUE)
 gl <- lapply(region_file_list, function(x) {
   img <- png::readPNG(paste0("analysis/data/tmp_data/", x, ".png"))
   g <- grid::rasterGrob(
@@ -28,7 +28,7 @@ gl <- lapply(region_file_list, function(x) {
     width = 0.14, height = 1.2
   )
 })
-dummy <- tibble::tibble(region_name = unique(amount_devel$region_name), grob = gl )
+dummy <- tibble::tibble(region = unique(amount_devel$region), grob = gl )
 
 # only for plot: time without sign of years
 amount_devel <- amount_devel %>%
@@ -51,7 +51,7 @@ development_burial_construction_A <- ggplot() +
     linetype = "blank",
     alpha = 0.8
   ) +
-  facet_wrap(~region_name, nrow = 8) +
+  facet_wrap(~region, nrow = 8) +
   xlab("Time in years calBC") +
   ylab("Amount of graves") +
   labs(fill = "Burial customs") +
@@ -106,7 +106,7 @@ development_burial_construction_B <- ggplot() +
     alpha = 0.8
   ) +
   scale_alpha_continuous(range = c(0.0, 0.7)) +
-  facet_wrap(~region_name, nrow = 8) +
+  facet_wrap(~region, nrow = 8) +
   xlab("Time in years calBC") +
   ylab("Proportion of burial customs") +
   labs(fill = "Burial customs") +

### 2019-03-24 analysis/code/04_plot_scripts/development_burial_type.R

@@ -11,8 +11,8 @@ load("analysis/data/tmp_data/development_proportions_burial_type.RData")
 
 #### prepare data ####
 amount_devel <- amount_development_burial_type
-regions_factor <- as.factor(amount_devel$region_name)
-amount_devel$region_name <- factor(regions_factor, levels = region_order)
+regions_factor <- as.factor(amount_devel$region)
+amount_devel$region <- factor(regions_factor, levels = region_order)
 idea_factor <- as.factor(amount_devel$idea)
 amount_devel$idea <- factor(idea_factor, levels = rev(levels(idea_factor)))
 
@@ -20,7 +20,7 @@ prop <- proportion_development_burial_type
 prop$idea <- as.factor(prop$idea)
 prop$idea <- factor(prop$idea , levels = rev(levels(prop$idea )))
 
-region_file_list <- unique(amount_devel$region_name) %>% gsub(" ", "_", ., fixed = TRUE)
+region_file_list <- unique(amount_devel$region) %>% gsub(" ", "_", ., fixed = TRUE)
 gl <- lapply(region_file_list, function(x) {
   img <- png::readPNG(paste0("analysis/data/tmp_data/", x, ".png"))
   g <- grid::rasterGrob(
@@ -28,7 +28,7 @@ gl <- lapply(region_file_list, function(x) {
     width = 0.14, height = 1.2
   )
 })
-dummy <- tibble::tibble(region_name = unique(amount_devel$region_name), grob = gl )
+dummy <- tibble::tibble(region = unique(amount_devel$region), grob = gl )
 
 # only for plot: time without sign of years
 amount_devel <- amount_devel %>%
@@ -51,7 +51,7 @@ development_burial_type_A <- ggplot() +
     linetype = "blank",
     alpha = 0.8
   ) +
-  facet_wrap(~region_name, nrow = 8) +
+  facet_wrap(~region, nrow = 8) +
   xlab("Time in years calBC") +
   ylab("Amount of graves") +
   labs(fill = "Burial customs") +
@@ -107,7 +107,7 @@ development_burial_type_B <- ggplot() +
     alpha = 0.8
   ) +
   scale_alpha_continuous(range = c(0.0, 0.7)) +
-  facet_wrap(~region_name, nrow = 8) +
+  facet_wrap(~region, nrow = 8) +
   xlab("Time in years calBC") +
   ylab("Proportion of burial customs") +
   labs(fill = "Burial customs") +

### 2019-03-24 analysis/code/04_plot_scripts/map.R

@@ -1,20 +1,22 @@
+#### dependencies ####
 library(ggplot2)
 library(sf)
 library(cowplot)
 library(magrittr)
 
+#### load data ####
 load("analysis/data/tmp_data/land_outline.RData")
 load("analysis/data/tmp_data/rivers.RData")
 load("analysis/data/tmp_data/lakes.RData")
 load("analysis/data/tmp_data/countries.RData")
 research_area <- sf::st_read("analysis/data/input_data/research_area_shapefile/research_area.shp")
 load("analysis/data/tmp_data/regions.RData")
-load("analysis/data/tmp_data/bronze1.RData")
+load("analysis/data/tmp_data/dates_coordinates.RData")
 load("analysis/data/tmp_data/region_order.RData")
 load("analysis/data/tmp_data/region_colors.RData")
-load("analysis/data/tmp_data/distance_matrix_spatial_long.RData")
+load("analysis/data/tmp_data/distance_matrix_spatial_long_half.RData")
 
-bronze1_sf <- bronze1 %>% sf::st_as_sf(
+dates_coordinates_sf <- dates_coordinates %>% sf::st_as_sf(
   coords = c("lon", "lat"),
   crs = 4326
 )
@@ -38,7 +40,7 @@ map_A <- ggplot() +
     fill = NA, colour = "black", size = 0.2
   ) +
   geom_sf(
-    data = bronze1_sf,
+    data = dates_coordinates_sf,
     mapping = aes(
       color = burial_type,
       shape = burial_construction,
@@ -104,7 +106,6 @@ map_A %>%
     width = 330, height = 420, units = "mm",
     limitsize = F
   )
-
 
 #### map_B ####
 
@@ -195,7 +196,7 @@ map_B <- ggplot() +
 
 #### map_C ####
 
-distance_lines <- distance_matrix_spatial_long %>%
+distance_lines <- distance_matrix_spatial_long_half %>%
   dplyr::left_join(
     region_centers,
     by = c("regionA" = "NAME")
@@ -219,11 +220,6 @@ distance_lines <- distance_matrix_spatial_long %>%
   dplyr::filter(
     regionA != regionB
   )
-
-mn <- pmin(distance_lines$regionA, distance_lines$regionB)
-mx <- pmax(distance_lines$regionA, distance_lines$regionB)
-int <- as.numeric(interaction(mn, mx))
[…]

### 2019-03-24 article/02_Materials_and_Methods.Rmd
reworked Table 1

@@
 load("../analysis/data/output_data/dprcrosstab.RData")
+colnames(dprcrosstab)[1:2] <- c("Region", "Burial type")
 dprcrosstab %>% 
-  unclass() %>% data.frame() %>% 
+  dplyr::mutate_at(
+    dplyr::vars(tidyselect::one_of("flat", "mound", "unknown")),
+    list(~kableExtra::cell_spec(., "latex", bold = . > 100)) 
+  ) %>%
   knitr::kable(
@@
     escape = F,
-    linesep = "\\addlinespace \\hline \\addlinespace",
+    #linesep = "\\addlinespace \\hline \\addlinespace",
     align = "l"
@@
   ) %>%
-  kableExtra::add_header_above(c("burial type" = 1, "burial construction" = 3))
+  kableExtra::add_header_above(c(" " = 1, " " = 1, "Burial construction" = 3)) %>%
+  kableExtra::column_spec(1, width = "2.5cm") %>%
+  kableExtra::column_spec(2, width = "1.5 cm") %>%
+  kableExtra::column_spec(3:5, width = "0.8cm") %>%
+  kableExtra::collapse_rows(1, row_group_label_position = "identity", valign = "top")
+
 ```

### 2019-03-24 analysis/code/03_dataset_description_scripts/observations_on_final_dataset.R

@@ -51,18 +51,17 @@ txtstorage::store(
 dpr$species %>% table(useNA = "always")
 
 # overview table
-# dprcrosstab <- dpr %>%
-#   dplyr::group_by(
-#     region, burial_type, burial_construction
-#   ) %>%
-#   dplyr::summarise(
-#     n = dplyr::n()
-#   ) %>%
-#   dplyr::ungroup() %>%
-#   tidyr::spread(
-#     key = burial_construction, value = n
-#   )
-dprcrosstab <- table(dpr$burial_type, dpr$burial_construction)
+dprcrosstab <- dpr %>%
+  dplyr::group_by(
+    region, burial_type, burial_construction
+  ) %>%
+  dplyr::summarise(
+    n = dplyr::n()
+  ) %>%
+  dplyr::ungroup() %>%
+  tidyr::spread(
+    key = burial_construction, value = n, fill = 0
+  )
 
 save(
   dprcrosstab,

### 2019-03-28 review/answer_to_review.md
unfortunately the changes are more time consuming then expected
further work on the text in reaction to the review: small sample size issue

@@
 The supplementary remarks are extensively long to cover these problems.
-Is it necessary to mention the limitations at other critical points of the paper?
-I'm absolutely open to do that.
+
+I now modified the Figures 3&4 as well as the Table 1 to provide a better overview over the actual amount of data available.
+I have also added several short comments/sentences in various sections of the paper to better address the small sample size issue.
 
@@
 
+TODO
+
 ## Minor points:
@@
 Marine reservoir effect is only one of the possible problems relevant here in addition to the Old wood effect or the human 14C-residence ("adult") effect.
-I added a paragraph abou this in the *Data quality* section.
+I added a paragraph about this in the *Data quality* section.
 
@@
 
+TODO
+
 > Page 4, Column 2, Lines 43-56. I think it is worth computing partial mantel tests so that the correlation between burial type and construction can take into account geography.
@@
 Done and included in Table 2.
+
+TODO
 
@@
 
-I will consider this very true questions in a future publication.
+I will consider this very important questions in a future publication.
 

### 2019-03-28 article/02_Materials_and_Methods.Rmd
further work on the text in reaction to the review: small sample size issue

@@
 The code for this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible [@marwickPackagingData2018].
-A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
+It is available in a public Repository on GitHub (\url{https://github.com/nevrome/cultrans.bronzeageburials.article2019}) and with a permanent digital object identifier (DOI) on the Open Science Framework Platform (\url{doi}) A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
 
@@
 The variable *feature type* systematically distinguishes between settlements, rock shelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, i.a..
+
 This information allows to accurately and precisely determine a moment in space and time when a grave of a particular type was created.
 In theory, the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe.
-Unfortunately, the Radon-B dataset has some quality issues mostly due to inconsistent manual data input: Incompleteness and faultiness of meta information, no well-thought-out variable classification for *sample material* and *cultural attribution* and inconsistent multilingualism even within individual entries.
+The amount of data currently available is relatively small though.
+The results presented in this paper should therefore be treated with appropriate caution.
+Several problems of representativity and data meaning are discussed in the sections below.
+
+Furthermore the Radon-B dataset has some quality issues mostly due to inconsistent manual data input: Incompleteness and faultiness of meta information, no well-thought-out variable classification for *sample material* and *cultural attribution* and inconsistent multilingualism even within individual entries.
 These issues though do not inhibit the general data analysis strategy of this paper, as the information about temporal and spatial position as well as *feature type* attribution is mostly complete.
@@
 
-From these time series another set of time series with a measure of cultural distance for each region-region relationship can be derived (Figure \ref{sed_region_matrix}).
+From these time series another set of time series with a measure of cultural distance for each region-region relationship can be derived (Figure \ref{sed_region_matrix}) -- though with a reasonable amount of caution due to the partially low number of data available to reconstruct the developments in several regions and time slots.
 Cultural distance is defined here as the squared Euclidean distance (SED) between the trait proportions of two different regions at a certain point in time [@neimanStylisticVariationEvolutionary1995].

### 2019-03-28 review/answer_to_review.md
further discussion of data quality

@@
 
-So: The criticism listed here is valid and I added a paragraph in the *Data quality* section to mention it.
-I refer to this section multiple times in the article.
-Nevertheless I don't see a practical possibility
-
-TODO
+So: The criticism listed here is valid and I added it in the discussion.
+I considered a complete revision of the time series construction methodology, but since I (1) am still convinced that it is scientifically sound and reproducible, and (2) do not know a fundamentally better approach, I suggest publishing the results as they are for the community of practice to discuss and improve.
 
@@
 
-This is the result of a misunderstanding in the communication with the special issue editors. I added the link to the relevant repository now.
-
-TODO
+I added links to the relevant repositories.
 
@@
 
-TODO
+Done.
 

### 2019-03-28 article/04_Discussion.Rmd
further discussion of data quality

@@
 The macro regions have to be very big due to a lack of data and don't allow for a more precise picture.
+The differences in the number of available dates -- sample size -- per region introduce sampling error.
+The samples are also non-independent, as large sites with multiple burials can strongly bias the proportions of a particular region.
+Furthermore the region size also causes strong risk of Simpson’s paradox (i.e.
+a trend appears in several different groups of data but disappears or reverses when these groups are combined).
 

### 2019-03-28 review/answer_to_review.md
dating problems

@@
 I'm not aware of a methodologival approach where time series could be constructed without some sort of aggregation.
-Unfortunately the size of the areas has to be very big because of the currently overall small amount of data.
-I expect this to change in the future and the analysis can be redone then with more data and much smaller areas to understand the definitely much more complex processes within the macro regions established here.
+The equal area size approach applied here for the region defition causes size of the areas to be very big due to the currently overall small amount of data.
+I expect this to change in the future and the analysis can be redone then with more samples and much smaller areas to understand the definitely much more complex processes within the macro regions established here.
 
 So: The criticism listed here is valid and I added it in the discussion.
-I considered a complete revision of the time series construction methodology, but since I (1) am still convinced that it is scientifically sound and reproducible, and (2) do not know a fundamentally better approach, I suggest publishing the results as they are for the community of practice to discuss and improve.
+I considered a complete revision of the time series construction methodology, but since I (1) am still convinced that it is scientifically sound and reproducible, and (2) do not know a fundamentally better approach that avoids these completely risks, I suggest publishing the results as they are for the community of practice to discuss and improve.
 
@@
 Marine reservoir effect is only one of the possible problems relevant here in addition to the Old wood effect or the human 14C-residence ("adult") effect.
+Not to mention cultural phenomena like secondary treatment of the body.
 I added a paragraph about this in the *Data quality* section.
-
-TODO
 

### 2019-03-28 article/06_Supplementary_Remarks.Rmd
dating problems

@@
 
-Concerning the primary variables *burial type* and *burial construction*, further special effects must be taken into account: The disturbance and robbery of graves was a widespread phenomenon in history and prehistory.
+Concerning the primary variables *burial type* and *burial construction*, further special effects must be taken into account:
+
+The disturbance and robbery of graves was a widespread phenomenon in history and prehistory.
 Burial mounds were usually attacked more severely than flat graves due to their better visibility and could, therefore, be under-represented in Radon-B.
@@
 It is difficult to estimate the impact of these effects in detail, but an over-representation of mounds seems more likely than an over-representation of flat graves.
+
 The dating of cremation graves on bone remains methodically differs from the dating of inhumations because it is based on the extraction of the mineral, not the organic, fraction in the bone [@lantingDatingCrematedBones2001].
@@
 An under-representation of cremation graves in current data collections is, therefore, more plausible.
+
+Another aspect to consider concerning the data quality is the temporal accuracy of the radiocarbon dating.
+Ideally each radiocarbon date represents the exact point in time of the creation of the respective grave.
+Unfortunately several natural and cultural effects can cause a (partially systematic) offset between the radiocarbon date and the burial ceremony, including the *Marine reservoir effect*, the *Old wood effect*, the *Human ^14^C-residence ("adult") effect* or cultural practices like a secondary treatment of the body or the deposition of burial objects that are already ancient at the time of the ceremony.
+These effects can have strong influence on the accuracy of the dating information, but can only be explored for each grave individually.
+In the present paper they therefore had to be ignored.
 

### 2019-03-28 article/06_Supplementary_Remarks.Rmd
dating accuracy: literature

@@
 Ideally each radiocarbon date represents the exact point in time of the creation of the respective grave.
-Unfortunately several natural and cultural effects can cause a (partially systematic) offset between the radiocarbon date and the burial ceremony, including the *Marine reservoir effect*, the *Old wood effect*, the *Human ^14^C-residence ("adult") effect* or cultural practices like a secondary treatment of the body or the deposition of burial objects that are already ancient at the time of the ceremony.
+Unfortunately several natural and cultural effects can cause a (partially systematic) offset between the radiocarbon date and the burial ceremony, including the *Marine reservoir effect* [@Ascough2005], the *Old wood effect* [@OLSEN201330], the *Human ^14^C-residence ("adult") effect* [@Hedges2007] or cultural practices like a secondary treatment of the body or the deposition of burial objects that are already ancient at the time of the ceremony.
 These effects can have strong influence on the accuracy of the dating information, but can only be explored for each grave individually.

### 2019-03-28 README.Rmd
update of readme with the now available identifiers

@@
   md_document:
-    variant: gfm
+    variant: markdown
 ---
@@
 
-<http://dx.doi.org/>…
+<http://dx.doi.org/10.17605/OSF.IO/B6NP2>
 
@@
 
-Not yet published.
+Not yet published: In review.
 
@@
 The Docker image includes all of the software dependencies needed to run the code in this project, including the data and code itself.
-To launch the Docker image for this project, first, [install Docker](https://docs.docker.com/installation/) on your computer and download the `.tar` file with the paper image [here](...) (not yet available).
+To launch the Docker image for this project, first, [install Docker](https://docs.docker.com/installation/) on your computer and download the `.tar` file with the paper image from the relevant repository (see DOI).
 At the Docker prompt, you can load and run the image with:
@@
 This will start a server instance of RStudio.
-Then open your web browser at localhost:8787 or run `docker-machine ip default` in the shell to find the correct IP address, and log in with rstudio/cultransbronze19.
+Then open your web browser at localhost:8787 or run `docker-machine ip default` in the shell to find the correct IP address, and log in with **rstudio/cultransbronze19**.
 Once logged in, use the Files pane (bottom right) to navigate to the script files.

### 2019-03-28 article/02_Materials_and_Methods.Rmd
added repo doi

@@
 The code for this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible [@marwickPackagingData2018].
-It is available in a public Repository on GitHub (\url{https://github.com/nevrome/cultrans.bronzeageburials.article2019}) and with a permanent digital object identifier (DOI) on the Open Science Framework Platform (\url{doi}) A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
+It is available in a public Repository on GitHub (\url{https://github.com/nevrome/cultrans.bronzeageburials.article2019}) and with a permanent digital object identifier (DOI) on the Open Science Framework Platform (\url{http://dx.doi.org/10.17605/OSF.IO/B6NP2}) A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
 

### 2019-03-28 article/02_Materials_and_Methods.Rmd
some more love for Table 1

@@
 The graves are linked to `r r("dpr period amount", sf_desc)` different *period* terms and `r r("dpr culture amount", sf_desc)` different archaeological *cultures* as they are distinguished in Radon-B.
+
 `r r("dpr material bone amount", sf_desc)` dates were measured on sample material of bones and teeth (at least `r r("dpr material cremated bones amount", sf_desc)` of them burnt), `r r("dpr material charcoal wood amount", sf_desc)` on wood and charcoal, a small part (`r r("dpr material other amount", sf_desc)`) on other materials such as nuts, resin or pitch.
 For the remaining `r r("dpr material unknown amount", sf_desc)` dates, no material information is stored.
-Regarding the variables burial type and burial construction, the conditions shown in Table \ref{tab:dprcrosstab} apply.
 
-```{r dprcrosstab, echo=FALSE, tidy=FALSE}
-load("../analysis/data/output_data/dprcrosstab.RData")
-colnames(dprcrosstab)[1:2] <- c("Region", "Burial type")
-dprcrosstab %>% 
+Regarding the variables burial type and burial construction in the grave dataset, the conditions shown in Table \ref{tab:gprcrosstab} apply.
+
+```{r gprcrosstab, echo=FALSE, tidy=FALSE}
+load("../analysis/data/output_data/gprcrosstab.RData")
+colnames(gprcrosstab)[1:2] <- c("Region", "Burial type")
+gprcrosstab %>% 
   dplyr::mutate_at(
     dplyr::vars(tidyselect::one_of("flat", "mound", "unknown")),
-    list(~kableExtra::cell_spec(., "latex", bold = . > 100)) 
+    list(
+      ~kableExtra::cell_spec(., "latex", bold = . > 100, color = ifelse(. < 10, "red", "black"))
+    ) 
   ) %>%
@@
     escape = F,
-    #linesep = "\\addlinespace \\hline \\addlinespace",
     align = "l"
@@
   kableExtra::add_header_above(c(" " = 1, " " = 1, "Burial construction" = 3)) %>%
-  kableExtra::column_spec(1, width = "2.5cm") %>%
+  kableExtra::column_spec(1, width = "2 cm") %>%
   kableExtra::column_spec(2, width = "1.5 cm") %>%

### 2019-03-28 analysis/code/03_dataset_description_scripts/observations_on_final_dataset.R

@@ -26,6 +26,23 @@ regions_graves_amounts <- graves_per_region %>%
 
 txtstorage::store("regions graves amounts", regions_graves_amounts, storage_file)
 
+gprcrosstab <- gpr %>%
+  dplyr::group_by(
+    region, burial_type, burial_construction
+  ) %>%
+  dplyr::summarise(
+    n = dplyr::n()
+  ) %>%
+  dplyr::ungroup() %>%
+  tidyr::spread(
+    key = burial_construction, value = n, fill = 0
+  )
+
+save(
+  gprcrosstab,
+  file = "analysis/data/output_data/gprcrosstab.RData"
+)
+
 
 
 #### dates analysis ####
@@ -70,20 +87,3 @@ txtstorage::store(
 )
 
 dpr$species %>% table(useNA = "always")
-
-dprcrosstab <- dpr %>%
-  dplyr::group_by(
-    region, burial_type, burial_construction
-  ) %>%
-  dplyr::summarise(
-    n = dplyr::n()
-  ) %>%
-  dplyr::ungroup() %>%
-  tidyr::spread(
-    key = burial_construction, value = n, fill = 0
-  )
-
-save(
-  dprcrosstab,
-  file = "analysis/data/output_data/dprcrosstab.RData"
-)

### 2019-03-28 article/02_Materials_and_Methods.Rmd
finished the work on the new Table 1

@@
 gprcrosstab %>% 
-  dplyr::mutate_at(
-    dplyr::vars(tidyselect::one_of("flat", "mound", "unknown")),
-    list(
-      ~kableExtra::cell_spec(., "latex", bold = . > 100, color = ifelse(. < 10, "red", "black"))
-    ) 
-  ) %>%
   knitr::kable(
-    caption = "Burial rite distribution in the active dataset.",
+    caption = "Burial rite distribution in the active grave dataset.",
     format = "latex",
@@
   kableExtra::column_spec(3:5, width = "0.8cm") %>%
-  kableExtra::collapse_rows(1, row_group_label_position = "identity", valign = "top")
+  kableExtra::collapse_rows(1, row_group_label_position = "identity", valign = "top") %>%   kableExtra::row_spec(row = 25:27, align = "center", bold = T)
 

### 2019-03-28 analysis/code/03_dataset_description_scripts/observations_on_final_dataset.R

@@ -26,7 +26,7 @@ regions_graves_amounts <- graves_per_region %>%
 
 txtstorage::store("regions graves amounts", regions_graves_amounts, storage_file)
 
-gprcrosstab <- gpr %>%
+gprcrosstab_top <- gpr %>%
   dplyr::group_by(
     region, burial_type, burial_construction
   ) %>%
@@ -37,6 +37,23 @@ gprcrosstab <- gpr %>%
   tidyr::spread(
     key = burial_construction, value = n, fill = 0
   )
+
+gprcrosstab_sum <- gpr %>%
+  dplyr::group_by(
+    burial_type, burial_construction
+  ) %>%
+  dplyr::summarise(
+    n = dplyr::n()
+  ) %>%
+  dplyr::ungroup() %>%
+  tidyr::spread(
+    key = burial_construction, value = n, fill = 0
+  ) %>%
+  dplyr::mutate(
+    region = "Total"
+  )
+
+gprcrosstab <- rbind(gprcrosstab_top, gprcrosstab_sum)
 
 save(
   gprcrosstab,

### 2019-03-28 review/answer_to_review.md
dating accuracy: literature
open todos
figure and table captions

@@
 
-This is absolutely true and I tried to emphasise the limitations of the data in the discussion.
+This is true and I tried to emphasise the limitations of the data in the discussion.
 The challenges are not confined on the amount of data, but also concern its meaning and origin.
@@
 
-I will consider this very important questions in a future publication.
+I will consider this important questions in a future publication.
 
@@
 Figure 6 is much simpler now without simulation results. It only contains a time series of distance correlation.
+
+TODO

### 2019-03-28 article/02_Materials_and_Methods.Rmd
figure and table captions

@@
   knitr::kable(
-    caption = "Burial rite distribution in the active grave dataset.",
+    caption = "Burial rite distribution in the active grave dataset. The order of the regions follows a southeast-northwest direction.",
     format = "latex",
@@
 The individual plots are separated by region, and each region is featured by a small map according to Figure \ref{map_regions} A.
-The order of the regions follows a southeast-\\northwest direction.}
+$n$ in the facet labels equals the total number of available dates with information for the variable (cf.
+Table \ref{tab:gprcrosstab}).}
 \end{figure}

### 2019-03-30 analysis/code/04_plot_scripts/correlation_time_series.R
figure height

@@ -125,6 +125,6 @@ ggsave(
   device = "jpeg",
   scale = 1,
   dpi = 300,
-  width = 300, height = 250, units = "mm",
+  width = 300, height = 280, units = "mm",
   limitsize = F
 )

### 2019-03-30 analysis/code/03_dataset_description_scripts/cultural_distance_overview_table.R
better mantel test result tables

@@ -1,66 +1,0 @@
-library(magrittr)
-
-#### burial_construction: sed - spatial distance ####
-
-load("analysis/data/tmp_data/mantel_sed_spatial_burial_construction.RData")
-load("analysis/data/tmp_data/squared_euclidian_distance_over_timeblocks_burial_construction.RData")
-burial_construction_distance <- sed_spatial_distance %>% dplyr::mutate(context = "burial_construction")
-burial_construction_mantel <- mantel_test_results
-names(burial_construction_mantel)[2:3] <- c("bc-sp: Correlation coefficient", "bc-sp: p-Value")
-
-
-
-#### burial_type: sed - spatial distance ####
-
-load("analysis/data/tmp_data/mantel_sed_spatial_burial_type.RData")
-load("analysis/data/tmp_data/squared_euclidian_distance_over_timeblocks_burial_type.RData")
-burial_type_distance <- sed_spatial_distance %>% dplyr::mutate(context = "burial_type")
-burial_type_mantel <- mantel_test_results
-names(burial_type_mantel)[2:3] <- c("bt-sp: Correlation coefficient", "bt-sp: p-Value")
-
-
-
-#### burial_construction & burial_type: sed - sed ####
-
-load("analysis/data/tmp_data/mantel_sed_spatial_burial_type_burial_construction.RData")
-load("analysis/data/tmp_data/mantel_sed_spatial_burial_type_burial_construction.RData")
-burial_type_burial_construction_mantel <- mantel_test_results
-names(burial_type_burial_construction_mantel)[2:3] <- c("bt-bc: Correlation coefficient", "bt-bc: p-Value")
-
-
-
-#### distance_and_correlation_table ####
-
-distance_and_correlation_table <- rbind(
-  burial_type_distance,
-  burial_construction_distance
-) %>%
-  tidyr::spread(context, mean_sed) %>%
-  dplyr::select(-regionA, -regionB) %>%
-  dplyr::group_by(time) %>%
-  dplyr::summarise(
-    `bc: Mean distance` = mean(burial_construction, na.rm = TRUE),
-    `bt: Mean distance` = mean(burial_type, na.rm = TRUE)
-  ) %>%
-  dplyr::left_join(
-    burial_construction_mantel,
-    by = "time"
-  )  %>%
-  dplyr::left_join(
-    burial_type_mantel,
-    by = "time"
-  ) %>%
-  dplyr::select(
-    1,3,6,7,2,4,5
[…]

### 2019-03-30 analysis/code/03_dataset_description_scripts/cultural_distance_overview_tables.R

@@ -1,0 +1,101 @@
+library(magrittr)
+
+#### burial_construction & spatial distance: sed - spatial distance ####
+
+load("analysis/data/tmp_data/mantel_sed_spatial_burial_construction.RData")
+load("analysis/data/tmp_data/squared_euclidian_distance_over_timeblocks_burial_construction.RData")
+burial_construction_distance <- sed_spatial_distance %>% dplyr::mutate(context = "burial_construction")
+burial_construction_mantel <- mantel_test_results
+names(burial_construction_mantel)[2:3] <- c("bc-sp: Correlation coefficient", "bc-sp: p-Value")
+
+
+
+#### burial_type & spatial distance: sed - spatial distance ####
+
+load("analysis/data/tmp_data/mantel_sed_spatial_burial_type.RData")
+load("analysis/data/tmp_data/squared_euclidian_distance_over_timeblocks_burial_type.RData")
+burial_type_distance <- sed_spatial_distance %>% dplyr::mutate(context = "burial_type")
+burial_type_mantel <- mantel_test_results
+names(burial_type_mantel)[2:3] <- c("bt-sp: Correlation coefficient", "bt-sp: p-Value")
+
+
+
+#### distance_and_correlation_table_A ####
+
+distance_and_correlation_table_A <- rbind(
+  burial_type_distance,
+  burial_construction_distance
+) %>%
+  tidyr::spread(context, mean_sed) %>%
+  dplyr::select(-regionA, -regionB) %>%
+  dplyr::group_by(time) %>%
+  dplyr::summarise(
+    `bc: Mean distance` = mean(burial_construction, na.rm = TRUE),
+    `bt: Mean distance` = mean(burial_type, na.rm = TRUE)
+  ) %>%
+  dplyr::left_join(
+    burial_construction_mantel,
+    by = "time"
+  )  %>%
+  dplyr::left_join(
+    burial_type_mantel,
+    by = "time"
+  ) %>%
+  dplyr::select(
+    1,3,6,7,2,4,5
+  ) %>%
+  dplyr::mutate_if(
+    is.numeric,
+    round,
+    3
+  )
+
+# hacky class manipulation: time without sign of years
+distance_and_correlation_table_A$time <- factor(
+  gsub("-(?=[0-9])", "", distance_and_correlation_table_A$time, perl = TRUE),
+  levels = gsub("-(?=[0-9])", "", levels(distance_and_correlation_table_A$time), perl = TRUE)
+)
+
[…]

### 2019-03-30 article/02_Materials_and_Methods.Rmd
embed new tables in text

@@
 The two matrices of each section were tested for correlation with the Mantel test, which uses permutations of potentially correlating matrices to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967].
-Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test (Table \ref{tab:distanceandcorrelationtable*} C).
+Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test (Table \ref{tab:dacta} C).
 
@@
 For the eight regions, four different categories of spatial distance were defined (Figure \ref{map_regions} B).
-The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the different data levels of the measurements (Table \ref{tab:distanceandcorrelationtable*} AB).
+The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the different data levels of the measurements (Table \ref{tab:dacta} AB).
 
-```{r distanceandcorrelationtable, echo=FALSE, tidy=FALSE}
-load("../analysis/data/output_data/distance_and_correlation_table.RData")
-distance_and_correlation_table %>%
+```{r dacta, echo=FALSE, tidy=FALSE}
+load("../analysis/data/output_data/distance_and_correlation_table_A.RData")
+distance_and_correlation_table_A %>%
   dplyr::mutate_at(
@@
 Listed are the overall mean squared Euclidian distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bc)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B).
-\\textbf{C} shows the Mantel test results for the two variables if compared directly.
 $p$-values below $0.1$ are printed bold.",
@@
   kableExtra::add_header_above(
-    c(" ", "burial type & spatial distance" = 3, "burial construction & spatial distance" = 3, "variable correlation" = 2)
+    c(" ", "burial type & spatial distance" = 3, "burial construction & spatial distance" = 3)
   ) %>%
   kableExtra::add_header_above(
-    c(" ", "A" = 3, "B" = 3, "C" = 2),
+    c(" ", "A" = 3, "B" = 3),
     bold = T, line = F
@@
   ) %>%
-  kableExtra::column_spec(c(2,4,5,7,9), width = "1.2cm") %>%
-  kableExtra::column_spec(c(3,6,8), width = "1.8cm") %>%
+  kableExtra::column_spec(c(1), width = "4cm") %>%
+  gsub("table", "table*", .)
+```
+
+```{r dactb, echo=FALSE, tidy=FALSE}
+load("../analysis/data/output_data/distance_and_correlation_table_B.RData")
+distance_and_correlation_table_B %>%
+  dplyr::mutate_at(
+    dplyr::vars(dplyr::matches("Correlation")),
+    list(~kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black")))
+  ) %>%
+  dplyr::mutate_at(
+    dplyr::vars(dplyr::matches("p-Value")),
+    list(~kableExtra::cell_spec(., "latex", bold = . < 0.1))
+  ) %>%
+knitr::kable(
+  "latex",
+  caption = "Like Table \\ref{tab:dacta}.
+\\textbf{C} shows the Mantel test results for the two variables if compared directly, \\textbf{D} the results of the Partial Mantel test considering spatial distance as a second independent variable.",
+  booktabs = T,
+  linesep = "\\addlinespace \\hline \\addlinespace",
+  escape = F
+) %>%
+  kableExtra::add_header_above(
+    c(" ", "variable correlation" = 2, "variable correlation with spatial correction" = 2)
+  ) %>%
+  kableExtra::add_header_above(
+    c(" ", "C" = 2, "D" = 2),
+    bold = T, line = F
+  ) %>%
+  kableExtra::kable_styling(
+    latex_options = c("hold_position"),
+    font_size = 9,
+    full_width = T
+  ) %>%
+  kableExtra::column_spec(c(1), width = "4cm") %>%
   gsub("table", "table*", .)
@@
 \caption{\label{simulation}Development of correlation between cultural and spatial distance.
-The results for the burial customs data (Table \ref{tab:distanceandcorrelationtable*} AB) are displayed with a solid (\textit{burial type}) and dashed (\textit{burial construction}) line.
+The results for the burial customs data (Table \ref{tab:dacta} AB) are displayed with a solid (\textit{burial type}) and dashed (\textit{burial construction}) line.
 The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}

### 2019-03-30 article/03_Results.Rmd
further work on the text in reaction to the review: small sample size issue
embed new tables in text

@@
 Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region-wise time series of trait occurrence.
-If we consider these results in isolation, we can construct the following narrative from them: \newline
+If we consider these results in isolation and also ignore issues of small sample size, we can construct the following narrative from them: \newline
 
@@
 
-Table \ref{tab:distanceandcorrelationtable*} C shows that the distance networks of *burial type* and *burial construction* are not equivalent.
+Table \ref{tab:dacta} C shows that the distance networks of *burial type* and *burial construction* are not equivalent.
 There is also no significant negative correlation.
@@
 
-For the relationship between the SED networks and the spatial distance network several observations result from Table \ref{tab:distanceandcorrelationtable*} AB: Cultural distance generally does not seem to correlate with spatial distance.
+For the relationship between the SED networks and the spatial distance network several observations result from Table \ref{tab:dacta} AB: Cultural distance generally does not seem to correlate with spatial distance.
 For the variable *burial type* an exception with significant correlation between the distance measures only reveals itself for the Early Bronze Age time slot 2000 to 1800 calBC.

### 2019-03-30 article/04_Discussion.Rmd
embed new tables in text

@@
 
-Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:distanceandcorrelationtable*}).
+Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta}).
 Only for small periods of time this correlation is visible in the Radon-B data.

### 2019-03-30 analysis/code/03_dataset_description_scripts/cultural_distance_overview_tables.R
forgot to rename path

@@ -64,7 +64,7 @@ save(distance_and_correlation_table_A, file = "analysis/data/output_data/distanc
 
 #### burial_construction & burial_type: sed - sed ####
 
-load("analysis/data/tmp_data/mantel_sed_spatial_burial_type_burial_construction.RData")
+load("analysis/data/tmp_data/mantel_sed_burial_type_burial_construction.RData")
 burial_type_burial_construction_mantel <- mantel_test_results
 names(burial_type_burial_construction_mantel)[2:3] <- c("bt-bc: Correlation coefficient", "bt-bc: p-Value")
 

### 2019-03-31 article/02_Materials_and_Methods.Rmd
reorganized general file structure

@@
 \centering
-\includegraphics[width=\columnwidth]{../analysis/figures/map_graves.jpeg}
+\includegraphics[width=\columnwidth]{map_graves.jpeg}
 \caption{\label{map_graves}All graves dating between 2200 and 800 calBC as extracted from the Radon-B database (\url{radon-b.ufg.uni-kiel.de}).
@@
 \centering
-\includegraphics[width=\columnwidth]{../analysis/figures/map_regions.jpeg}
+\includegraphics[width=\columnwidth]{map_regions.jpeg}
 \caption{\label{map_regions}Circular, artificial regions (\textbf{A}) and spatial distance network with classified distance values (\textbf{B}).
@@
 ```{r gprcrosstab, echo=FALSE, tidy=FALSE}
-load("../analysis/data/output_data/gprcrosstab.RData")
+load("gprcrosstab.RData")
 colnames(gprcrosstab)[1:2] <- c("Region", "Burial type")
@@
 \centering
-\includegraphics[width=\columnwidth]{../analysis/figures/development_burial_type.jpeg}
+\includegraphics[width=\columnwidth]{development_burial_type.jpeg}
 \caption{\label{development_burial_type}Regional distribution time series of the variable \textit{burial type} with the absolute amount (\textbf{A}) and proportion (\textbf{B}) of grave types.
@@
 \centering
-\includegraphics[width=\columnwidth]{../analysis/figures/development_burial_construction.jpeg}
+\includegraphics[width=\columnwidth]{development_burial_construction.jpeg}
 \caption{\label{development_burial_construction}Like Figure \ref{development_burial_type}, but for the variable \textit{burial construction}.\newline\newline\newline\newline\newline}
@@
 \centering
-\includegraphics[width=\textwidth]{../analysis/figures/sed_region_matrix.jpeg}
+\includegraphics[width=\textwidth]{sed_region_matrix.jpeg}
 \caption{\label{sed_region_matrix}Plot matrix of the region-region squared Euclidean distance time series.
@@
 ```{r dacta, echo=FALSE, tidy=FALSE}
-load("../analysis/data/output_data/distance_and_correlation_table_A.RData")
+load("distance_and_correlation_table_A.RData")
 distance_and_correlation_table_A %>%
@@
 ```{r dactb, echo=FALSE, tidy=FALSE}
-load("../analysis/data/output_data/distance_and_correlation_table_B.RData")
+load("distance_and_correlation_table_B.RData")
 distance_and_correlation_table_B %>%
@@
 \centering
-\includegraphics[width=\columnwidth]{../analysis/figures/correlation_time_series.jpeg}
+\includegraphics[width=\columnwidth]{correlation_time_series.jpeg}
 \caption{\label{simulation}Development of correlation between cultural and spatial distance.

### 2019-03-31 article/index.Rmd
reorganized general file structure

@@
   - times
-bibliography: ../references.bib
+bibliography: references.bib
 bibliographystyle: apalike
@@
 # allow easy dynamic value integration
-sf_prep <- "../analysis/data/output_data/sf_prep.txt"
-sf_desc <- "../analysis/data/output_data/sf_desc.txt"
+sf_prep <- "sf_prep.txt"
+sf_desc <- "sf_desc.txt"
 r <- txtstorage::read

### 2019-03-31 _copy_and_render.sh

@@ -1,0 +1,8 @@
+# create rendered article directory
+mkdir rendered_article
+
+# copy files to rendered_article directory
+./_copy_files.sh
+
+# do the actual rendering
+./_render_article.sh

### 2019-03-31 _copy_files.sh

@@ -1,0 +1,14 @@
+# markdown text files
+cp -r article/* rendered_article/
+
+# latex template
+cp -r latex_template/* rendered_article/
+
+# plots
+cp -r analysis/figures/* rendered_article/
+
+# data for text and tables
+cp -r analysis/data/output_data/* rendered_article/
+
+# bibliography
+cp references.bib rendered_article/

### 2019-03-31 _deploy_article.sh

@@ -4,7 +4,6 @@ git clone -b document \
   document_out
 cd document_out
 git rm -rf *
-cp -r ../analysis/figures/* ./
 cp -r ../rendered_article/* ./
 git add --all *
 git commit -m "Update the rendered article"

### 2019-03-31 _render_article.sh

@@ -1,7 +1,7 @@
 #!/usr/bin/env Rscript
 
 rmarkdown::render(
-  "article/index.Rmd",
+  "rendered_article/index.Rmd",
   output_dir = "rendered_article",
   output_file = "article.pdf"
 )

### 2019-03-31 _run_analysis.sh

@@ -1,2 +1,5 @@
+# creation of the output directories
 Rscript analysis/code/create_directories.R
+
+# run analysis
 Rscript analysis/code/run_all_scripts.R

### 2019-03-31 analysis/code/create_directories.R

@@ -1,2 +1,3 @@
 dir.create('analysis/figures')
 dir.create('analysis/data/tmp_data')
+dir.create('analysis/data/output_data')

### 2019-03-31 article/02_Materials_and_Methods.Rmd
started to incorporate the new arrangements of Figure 6 and Partial Mantel test into the text

@@
 Cultural distance is defined here as the squared Euclidean distance (SED) between the trait proportions of two different regions at a certain point in time [@neimanStylisticVariationEvolutionary1995].
-SED is a sensible metric here, because it is relatively simple, works with only a few assumptions and emphasises differences through squaring [@nakoinzFingerprintingIronAge2014].
+SED is a sensible metric in this context, because it is relatively simple, works with only a few assumptions and emphasises differences through squaring [@nakoinzFingerprintingIronAge2014].
 
@@
 For each of the $1400/200 = 7$ section and each of the two variables, a matrix of the mean SED over time was calculated.
-The two matrices of each section were tested for correlation with the Mantel test, which uses permutations of potentially correlating matrices to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967].
-Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test (Table \ref{tab:dacta} C).
+The two matrices of each section were tested for correlation with the *Mantel test*, which uses permutations of potentially correlating matrices to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967].
+Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test (Table \ref{tab:dactb} C).
 
@@
 For the eight regions, four different categories of spatial distance were defined (Figure \ref{map_regions} B).
-The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the different data levels of the measurements (Table \ref{tab:dacta} AB).
+The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904], because of the different data levels of the measurements (Table \ref{tab:dacta} AB).
+
+With a *Partial Mantel test* the spatial distance can also be considered for the direct comparison of the SED networks of *burial type* and *burial construction* (Table \ref{tab:dactb} D).
+While the Mantel test only allows a comparison among two variables, the Partial Mantel test allows one to be made among two variables while controlling for the third.
+In this case again Spearman's rank correlation coefficient has to be applied.
+
+Figure \ref{correlationtimeseries} shows the development of the correlation coefficient values obtained via Mantel statistics (Table \ref{tab:dacta} & \ref{tab:dactb}) during the succession of 200 years time slots.
 
@@
   caption = "Like Table \\ref{tab:dacta}.
-\\textbf{C} shows the Mantel test results for the two variables if compared directly, \\textbf{D} the results of the Partial Mantel test considering spatial distance as a second independent variable.",
+\\textbf{C} shows the Mantel test results for the two variables if compared directly, \\textbf{D} the results of the Partial Mantel test considering and correcting for spatial distance as a second independent variable.",
   booktabs = T,
@@
 \includegraphics[width=\columnwidth]{correlation_time_series.jpeg}
-\caption{\label{simulation}Development of correlation between cultural and spatial distance.
+\caption{\label{correlationtimeseries}Development of correlation between cultural and spatial distance.
 The results for the burial customs data (Table \ref{tab:dacta} AB) are displayed with a solid (\textit{burial type}) and dashed (\textit{burial construction}) line.

### 2019-03-31 article/04_Discussion.Rmd
started to incorporate the new arrangements of Figure 6 and Partial Mantel test into the text

@@
 
-Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta}).
+Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta}&\ref{tab:dacta}).
 Only for small periods of time this correlation is visible in the Radon-B data.

### 2019-03-31 article/02_Materials_and_Methods.Rmd
incorporating new results

@@
 For the eight regions, four different categories of spatial distance were defined (Figure \ref{map_regions} B).
-The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904], because of the different data levels of the measurements (Table \ref{tab:dacta} AB).
+The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904], because of the different data levels of the measurements (Table \ref{tab:dacta}).
 
@@
 \includegraphics[width=\columnwidth]{correlation_time_series.jpeg}
-\caption{\label{correlationtimeseries}Development of correlation between cultural and spatial distance.
-The results for the burial customs data (Table \ref{tab:dacta} AB) are displayed with a solid (\textit{burial type}) and dashed (\textit{burial construction}) line.
+\caption{\label{correlationtimeseries}Development of correlation between different measures of cultural and spatial distance.
+The results for the correlation of cultural and spatial distance (Table \ref{tab:dacta}) are displayed with a solid (\textit{burial type}) and a dashed (\textit{burial construction}) line.
+The results for the correlation of the two cultural distances (Table \ref{tab:dactb}) are displayed with a strong dotted line \textit{without} the correction for spatial distance (only here: Pearson correlation coefficient), and with a weak dotted line for the result \textit{with} the aforementioned correction.
 The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}

### 2019-03-31 article/03_Results.Rmd
started to incorporate the new arrangements of Figure 6 and Partial Mantel test into the text
incorporating new results

@@
 
-Table \ref{tab:dacta} C shows that the distance networks of *burial type* and *burial construction* are not equivalent.
+Table \ref{tab:dactb} shows that the distance networks of *burial type* and *burial construction* are not equivalent.
 There is also no significant negative correlation.
@@
 Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600 calBC the 5% significance level of the Mantle test is almost undercut, thus suggesting a statistically significant relationship.
+This moment of strong correlation is though not induced by similarity along the spatial dimension: The effect is still visible in the result of the Partial Mantel test, where the spatial distance is artificially counterbalanced.
 
-For the relationship between the SED networks and the spatial distance network several observations result from Table \ref{tab:dacta} AB: Cultural distance generally does not seem to correlate with spatial distance.
+For the relationship between the SED networks and the spatial distance network several observations result from Table \ref{tab:dacta}:
+
 For the variable *burial type* an exception with significant correlation between the distance measures only reveals itself for the Early Bronze Age time slot 2000 to 1800 calBC.
@@
 In the Middle Bronze Age the SED developments of the *burial type* variable undergo the peaks already observed.
-The correlation between cultural and spatial distance during these peaks is low and even tilts between 1400 and 1200 calBC, at the transition to the Late Bronze Age, into the negative -- although not significantly.
+The correlation between cultural and spatial distance during these peaks is low and even tilts between 1600 and 1400 calBC into the negative -- although not significantly.
 The cultural distance between several geographically distant regions is then smaller than between several directly adjacent regions.
-Looking at the SED network of *burial construction*, similar conclusions can be drawn: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated.
-As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation, though it is only significant for the time slot between 1800 and 1400 calBC.
+
+Looking at the SED network of *burial construction*, a slightly different conclusions can be drawn: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated.
+As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation, though it is only significant for the time slot between 2000 and 1400 calBC.
 There is no time period with negative mean correlation for *burial construction* cultural distance and spatial distance, but also here the correlation coefficient approaches zero in the Late Bronze Age.
+
+In general cultural distance does not seem to correlate strongly with spatial distance, especially in the Middle and Late Bronze Age.

### 2019-03-31 article/index.Rmd
fixed titles of region sections in discussion

@@
 r <- txtstorage::read
-a <- function(x) {r("regions graves amounts", sf_prep) %>% strsplit(", ") %>% unlist %>% magrittr::extract(grepl(x, .))}
+a <- function(x) {r("regions graves amounts", sf_desc) %>% strsplit(", ") %>% unlist %>% magrittr::extract(grepl(x, .))}
 ```

### 2019-03-31 article/03_Results.Rmd
small changes in the light of new results

@@
 Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600 calBC the 5% significance level of the Mantle test is almost undercut, thus suggesting a statistically significant relationship.
-This moment of strong correlation is though not induced by similarity along the spatial dimension: The effect is still visible in the result of the Partial Mantel test, where the spatial distance is artificially counterbalanced.
+This moment of strong correlation is though not induced by similarity along the spatial dimension: The effect is still visible in the result of the Partial Mantel test, where the spatial distance is artificially counterbalanced (Figure \ref{correlationtimeseries}).
 

### 2019-03-31 article/04_Discussion.Rmd
small changes in the light of new results

--- 2019-03-30 article/04_Discussion.Rmd
@@
 
-Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta}).
-Only for small periods of time this correlation is visible in the Radon-B data.
+Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta} & \ref{tab:dactB} & Figure \ref{correlationtimeseries}).
+Only for certain periods of time this correlation is visible in the Radon-B data.
 The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation.

### 2019-03-31 article/01_Introduction.Rmd
phrasing

--- 2019-03-24 article/01_Introduction.Rmd
@@
 The classic narrative of the development of these rites in Europe presented e.g.
-by @hardingEuropeanSocietiesBronze2000a (following @hauslerBestattungssittenFruhenBronzezeit1977; @hauslerGrabUndBestattungssitten1994; @hauslerTotenorientierungUndGeographischer1996) can be summarised as follows: In the early and middle Bronze Age inhumations dominated in different variations.
-Cremation burials in this time only occur in the Hungarian lowlands.
-Burial mounds are concentrated in parts of the Balkans as well as in Eastern, Western and Northern Europe, while in Central and Southern Europe flat graves -- mostly inhumations -- predominate.
+by @hardingEuropeanSocietiesBronze2000a following @hauslerBestattungssittenFruhenBronzezeit1977; @hauslerGrabUndBestattungssitten1994 and @hauslerTotenorientierungUndGeographischer1996 can be summarised as follows: In the Early and Middle Bronze Age inhumations dominated in different variations.
+Cremation burials in this time only occurred in the Hungarian lowlands.
+Burial mounds were concentrated in parts of the Balkans as well as in Eastern, Western and Northern Europe, while in Central and Southern Europe flat graves -- mostly inhumations -- predominated.
 In the Middle Bronze Age, barrows gained importance in Western and Central Europe.
 This observation is highly linked to the emergence of the *Tumulus culture* in Central Europe.
-In the late Bronze Age, the Urnfield period, cremation became the most common burial custom in the context of *Urnfield culture*.
+In the Late Bronze Age, the Urnfield period, cremation became the most common burial custom in the context of *Urnfield culture*.
 Harding's narrative is, even when taking into account its enormous simplification, partially misleading.
-Cremation, for example, was a lot more common already in the early and middle Bronze Age, though not in the context of the later Urnfield culture (e.g.
+Cremation, for example, was a lot more common already in the Early and Middle Bronze Age, though not in the context of the later Urnfield culture (e.g.
 in the Benelux area [@fokkensBronzeAgeLow2013] or Great-Britain [@robertsBritainIrelandBronze2013]).
@@
 
-The diffusion of the cremation burial tradition in the middle and late Bronze Age with radiocarbon data was already the subject of promising research in Central and South-Western Europe with focus on Catalonia and Switzerland [@capuzzoCulturalChangesSecond2015; @capuzzoInhumacionCremacionNordeste2017; @capuzzoSpacetemporalAnalysisRadiocarbon2014].
-This approach based on the dedicated radiocarbon database EUBAR [@capuzzoEUBARDatabase14C2013] sets a precedent for a more reproducible and statistically sound research practice concerning diffusion of innovation in Bronze Age archaeology.
+Similar methodology was already the subject of promising, but only regional research in Central and South-Western Europe with focus on Catalonia and Switzerland [@capuzzoCulturalChangesSecond2015; @capuzzoInhumacionCremacionNordeste2017; @capuzzoSpacetemporalAnalysisRadiocarbon2014].
+This approach based on the dedicated radiocarbon database EUBAR [@capuzzoEUBARDatabase14C2013] aimed to explore the diffusion of the cremation burial tradition in the Middle and Late Bronze Age and set a precedent for a more reproducible and statistically sound research practice in Bronze Age archaeology.
 
@@
 They are therefore suitable for study using methods of *Cultural transmission theory*, a special aspect of *Cultural evolution theory* concerned with the spread of ideas in human populations [@creanzaCulturalEvolutionaryTheory2017].
-This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries, as it assumes *Social learning* to produce adaptive behaviour through transmission, selection and mutation -- a secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007; @lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
+This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries, because it assumes a secondary heredity system in humans: *Social learning* with transmission, selection and mutation of ideas [@henrichDualinheritanceTheoryEvolution2007; @lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].
 
-*Evolutionary archaeology*, the branch of archaeology that explicitly adopted the terminology of Cultural evolution, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is very relevant for burial traditions: Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
+*Evolutionary archaeology*, the branch of archaeology that explicitly adopted Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is relevant for burial traditions.
+Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
 Without resorting to the inappropriate simplifications of Evolutionism, their prevalence generally happens to increase over time.
@@
 These expenses might have been influenced by or had their effect on the human-nature relation: Pyres, for example, require a significant amount of wood and the erection of barrows much workforce.
-Still, it can be expected that these functional disadvantages were not critical most of the time -- at least regarding the general decision which kind of burial tradition is applied.
-Most probably they influence the magnitude of invested energy for a burial, just like the level of grief [@macdonaldGriefBurialAmerican2001] or the representational intentions of the burying relatives [@henrichEvolutionCostlyDisplays2009].
+Still, it can be expected that these functional disadvantages were not critical most of the time for burial traditions -- at least regarding the general decision which ritual is performed.
+Most probably they influenced the magnitude of invested energy for a burial, just like the level of grief [@macdonaldGriefBurialAmerican2001] or the representational intentions of the burying relatives [@henrichEvolutionCostlyDisplays2009].
 
@@
 @neimanStylisticVariationEvolutionary1995 takes a closer look at these processes and develops a simple framework for the simulation of non-functional idea propagation.
-He identifies the essential phenomena *drift* (random dominance of similar traits) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.
+He identifies the essential phenomena *drift* (i.e.
+in case of multiple, equivalent traits in a population, a single one will dominate over time due to random effects) and *flow* (i.e.
+intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.
 Furthermore, he implements a measure of cultural distance to apply his theoretical insights to real-world data.

### 2019-03-31 article/02_Materials_and_Methods.Rmd
phrasing

@@
 Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the type and cultural attribution of the sampled context.
-The variable *feature type* systematically distinguishes between settlements, rock shelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, i.a..
+The variable *feature type* systematically distinguishes between settlements, rock shelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, ...
 
 This information allows to accurately and precisely determine a moment in space and time when a grave of a particular type was created.
-In theory, the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe.
+In theory, the more radiocarbon dates with this information are documented, the better our understanding and reconstruction of burial traditions in prehistoric Europe can become.
 The amount of data currently available is relatively small though.
@@
 That means that for many graves multiple (up to 8) ^14^C dates have been created (and entered in Radon-B) and dates therefore cannot be treated as semantically equal to graves.
-There is a variety of phenomena that cause this effect: In some cases, safety measurements were taken so that there are multiple dates for one single grave that all fall within the margin of errors of each other.
+There is a variety of phenomena that cause this effect: In some cases safety measurements were taken so that there are multiple dates for one single grave that all fall within the margin of errors of each other.
 In other cases, the set of dates for one grave strongly diverge because they cover the life cycle of a collective grave that was used for hundreds of years.
@@
 That means some graves are counted towards not only one but two regions.
+The creation of these macro-regions is necessary to establish discrete observation units, but they are also the source of methodological problems (see section \ref{discussion}).
 
@@
 
-The resulting dataset is structured as a table with `r r("gpr size", sf_desc)` rows each representing one grave as compiled by `r r("dpr size", sf_desc)` ^14^C dates from `r r("dpr sites amount", sf_desc)` sites.
+The resulting dataset is structured as a table with `r r("gpr size", sf_desc)` rows each representing one grave as compiled from `r r("dpr size", sf_desc)` ^14^C dates and `r r("dpr sites amount", sf_desc)` sites.
 The graves are linked to `r r("dpr period amount", sf_desc)` different *period* terms and `r r("dpr culture amount", sf_desc)` different archaeological *cultures* as they are distinguished in Radon-B.
@@
 
-Regarding the variables burial type and burial construction in the grave dataset, the conditions shown in Table \ref{tab:gprcrosstab} apply.
+Regarding the variables burial type and burial construction in the grave dataset, the distribution conditions shown in Table \ref{tab:gprcrosstab} apply.
 
@@
   knitr::kable(
-    caption = "Burial rite distribution in the active grave dataset. The order of the regions follows a southeast-northwest direction.",
+    caption = "Burial rite distribution in the active grave dataset.
+This table shows the number of graves, not the underlying dates.
+The order of the regions follows a southeast-northwest direction.",
     format = "latex",
@@
 From these time series another set of time series with a measure of cultural distance for each region-region relationship can be derived (Figure \ref{sed_region_matrix}) -- though with a reasonable amount of caution due to the partially low number of data available to reconstruct the developments in several regions and time slots.
+
 Cultural distance is defined here as the squared Euclidean distance (SED) between the trait proportions of two different regions at a certain point in time [@neimanStylisticVariationEvolutionary1995].
@@
 
-Figure \ref{correlationtimeseries} shows the development of the correlation coefficient values obtained via Mantel statistics (Table \ref{tab:dacta} & \ref{tab:dactb}) during the succession of 200 years time slots.
+Figure \ref{correlationtimeseries} shows the development of the correlation coefficient values obtained via Mantel statistics (Table \ref{tab:dacta} & \ref{tab:dactb}) during the succession of the seven relevant 200 years time slots.
 

### 2019-03-31 article/02_Materials_and_Methods.Rmd
more phrasing and minor improvements

@@
   kableExtra::add_header_above(
-    c(" ", "burial type & spatial distance" = 3, "burial construction & spatial distance" = 3)
+    c(" ", "type distance & spatial distance" = 3, "construction distance & spatial distance" = 3)
   ) %>%
@@
   caption = "Like Table \\ref{tab:dacta}.
-\\textbf{C} shows the Mantel test results for the two variables if compared directly, \\textbf{D} the results of the Partial Mantel test considering and correcting for spatial distance as a second independent variable.",
+\\textbf{C} shows the Mantel test results for the distance networks based on the two variables if compared directly, \\textbf{D} the results of the respective Partial Mantel test considering and correcting for spatial distance as a second independent variable.",
   booktabs = T,
@@
   kableExtra::add_header_above(
-    c(" ", "variable correlation" = 2, "variable correlation with spatial correction" = 2)
+    c(" ", "type & construction distance" = 2, "type & construction & spatial distance" = 2)
   ) %>%

### 2019-03-31 article/03_Results.Rmd
more phrasing and minor improvements

--- 2019-03-31 article/03_Results.Rmd
@@
 The situation was similar in Southern Scandinavia, where the proportion of cremation burials was much lower until the Late Bronze Age.
-In Poland and northern Germany, cremation became relatively suddenly and completely established with the beginning of the Late Bronze Age.
+In Poland and Southern Germany, cremation became relatively suddenly and completely established with the beginning of the Late Bronze Age.
 In Northeastern France and Northern Germany this change took place already in the Middle Bronze Age.
@@
 There is also no significant negative correlation.
-The degree of agreement between two regions in the question of inhumation and cremation is therefore not suitable for predicting the similarity of developments in these regions concerning the dominance of flat or mound graves.
-Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600 calBC the 5% significance level of the Mantle test is almost undercut, thus suggesting a statistically significant relationship.
-This moment of strong correlation is though not induced by similarity along the spatial dimension: The effect is still visible in the result of the Partial Mantel test, where the spatial distance is artificially counterbalanced.
+The degree of agreement between two regions in the question of inhumation and cremation is therefore not suitable for predicting the similarity in the dominance of flat or mound graves.
+Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600 calBC the 5% significance level of the Mantel test is almost undercut, thus suggesting a statistically significant relationship.
+This moment of strong correlation is though not induced by similarity along the spatial dimension: The effect is still visible in the result of the Partial Mantel test, where the spatial distance is artificially counterbalanced (see also Figure \ref{correlationtimeseries}).
 
@@
 In the Middle Bronze Age the SED developments of the *burial type* variable undergo the peaks already observed.
-The correlation between cultural and spatial distance during these peaks is low and even tilts between 1600 and 1400 calBC into the negative -- although not significantly.
+The correlation between cultural and spatial distance during these peaks is low and even tilts between 1600 and 1400 calBC into the negative -- but not significantly.
 The cultural distance between several geographically distant regions is then smaller than between several directly adjacent regions.
@@
 
-In general cultural distance does not seem to correlate strongly with spatial distance, especially in the Middle and Late Bronze Age.
+In general, cultural distance defined here does not seem to correlate strongly with spatial distance, especially in the Middle and Late Bronze Age.

### 2019-03-31 article/04_Discussion.Rmd
more phrasing and minor improvements

@@
 
-Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta} & \ref{tab:dactB} & Figure \ref{correlationtimeseries}).
+Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta} and \ref{tab:dactb}, Figure \ref{correlationtimeseries}).
 Only for certain periods of time this correlation is visible in the Radon-B data.

### 2019-03-31 article/04_Discussion.Rmd
and more phrasing

@@
 
-The second important question posed for this paper was, *how cultural distance between regions that results from changes in burial traditions develops, and how it is related to spatial distance?* While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one had a larger scope.
+The second important question posed for this paper was, *how cultural distance between regions develops, and how it is related to spatial distance?* While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one had a larger scope.
 The quantitative representation of an archaeological narrative paths the way for a diverse set of methods to be applied and the analysis undertaken here is merely the beginning.
@@
 Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta} and \ref{tab:dactb}, Figure \ref{correlationtimeseries}).
-Only for certain periods of time this correlation is visible in the Radon-B data.
-The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation.
+Only for certain periods of time this correlation is visible in the Radon-B data and the *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation.
 If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000 calBC) and a decline (after 2000 calBC) of cultural-spatial-distance correlation.
@@
 The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} question the credibility of the cultural distance calculation.
-The macro regions have to be very big due to a lack of data and don't allow for a more precise picture.
+The macro regions have to be very big due to a lack of data and don't represent the cultural heterogeneity of Bronze Age Europe appropriately.
 The differences in the number of available dates -- sample size -- per region introduce sampling error.

### 2019-03-31 analysis/code/04_plot_scripts/correlation_time_series.R
changed plot legend once more

@@ -14,19 +14,19 @@ mantel_burial_type_burial_construction_spatial <- mantel_test_results
 
 mantel_burial_type_spatial %<>%
   dplyr::mutate(
-    context = "burial type & spatial distance"
+    context = "A: type & spatial distance"
   )
 mantel_burial_construction_spatial %<>%
   dplyr::mutate(
-    context = "burial construction & spatial distance"
+    context = "B: construction & spatial distance"
   )
 mantel_burial_construction_burial_type %<>%
   dplyr::mutate(
-    context = "burial type & burial_construction"
+    context = "C: type & construction distance"
   )
 mantel_burial_type_burial_construction_spatial %<>%
   dplyr::mutate(
-    context = "burial type & burial_construction + spatial distance"
+    context = "D: type & construction + spatial distance"
   )
 
 mantel <- rbind(
@@ -78,28 +78,28 @@ p <- ggplot() +
   scale_colour_manual(
     name = "Distance correlations",
     values = c(
-      "burial type & spatial distance" = "#0072B2",
-      "burial construction & spatial distance" = "#009E73",
-      "burial type & burial_construction" = "black",
-      "burial type & burial_construction + spatial distance" = "black"
+      "A: type & spatial distance" = "#0072B2",
+      "B: construction & spatial distance" = "#009E73",
+      "C: type & construction distance" = "black",
+      "D: type & construction + spatial distance" = "black"
     )
   ) +
   scale_linetype_manual(
     name = "Distance correlations",
     values = c(
-      "burial type & spatial distance" = "solid",
-      "burial construction & spatial distance" = "dashed",
-      "burial type & burial_construction" = "dotted",
-      "burial type & burial_construction + spatial distance" = "dotted"
+      "A: type & spatial distance" = "solid",
+      "B: construction & spatial distance" = "dashed",
+      "C: type & construction distance" = "dotted",
+      "D: type & construction + spatial distance" = "dotted"
     )
   ) +
   scale_size_manual(
     name = "Distance correlations",
     values = c(
-      "burial type & spatial distance" = 2,
-      "burial construction & spatial distance" = 2,
-      "burial type & burial_construction" = 2,
-      "burial type & burial_construction + spatial distance" = 1
+      "A: type & spatial distance" = 2,
+      "B: construction & spatial distance" = 2,
+      "C: type & construction distance" = 2,
+      "D: type & construction + spatial distance" = 1
[…]

### 2019-03-31 article/02_Materials_and_Methods.Rmd
also the table headers

@@
   kableExtra::add_header_above(
-    c(" ", "type & construction distance" = 2, "type & construction & spatial distance" = 2)
+    c(" ", "type & construction distance" = 2, "type & construction + spatial distance" = 2)
   ) %>%

### 2019-03-31 article/06_Supplementary_Remarks.Rmd
phrasing

--- 2019-03-28 article/06_Supplementary_Remarks.Rmd
@@
 4.
-^14^C dating requires a small but sufficiently contextualised amount of organic finds.
-Due to the old wood and old water effect, short-lived sample materials such as bones of terrestrial organisms or charred cereal grains are preferred.
+^14^C dating requires a small but sufficiently contextualised amount of organic finds and short-lived sample materials such as bones of terrestrial organisms or charred cereal grains are preferred.
 The majority of ^14^C data for this paper comes from bones and teeth.
@@
 Ploughing could also result in an under-representation of the archaeological findings [@hardingEuropeanSocietiesBronze2000a].
-On the other hand, if a burial mound is still present, it is more readily visible and recognised by archaeologists today: Mounds are often clearly identifiable in aerial photographs or digital terrain models, while flat graves are much more difficult to identify for today's remote sensing methods.
+On the other hand, if a burial mound is still present, it is also more readily visible and recognised by archaeologists today: Mounds are often clearly identifiable in aerial photographs or digital terrain models, while flat graves are much more difficult to observe with modern remote sensing methods.
 It is difficult to estimate the impact of these effects in detail, but an over-representation of mounds seems more likely than an over-representation of flat graves.
@@
 
-Another aspect to consider concerning the data quality is the temporal accuracy of the radiocarbon dating.
+Another aspect to consider concerning the data quality is the temporal accuracy of radiocarbon dating.
 Ideally each radiocarbon date represents the exact point in time of the creation of the respective grave.
-Unfortunately several natural and cultural effects can cause a (partially systematic) offset between the radiocarbon date and the burial ceremony, including the *Marine reservoir effect*, the *Old wood effect*, the *Human ^14^C-residence ("adult") effect* or cultural practices like a secondary treatment of the body or the deposition of burial objects that are already ancient at the time of the ceremony.
+Unfortunately several natural and cultural effects can cause a (partially systematic) offset between the radiocarbon date and the burial ceremony, including the *Marine reservoir effect* [@Ascough2005], the *Old wood effect* [@OLSEN201330], the *Human ^14^C-residence ("adult") effect* [@Hedges2007] or cultural practices like a secondary treatment of the body or the deposition of burial objects that are already ancient at the time of the funeral.
 These effects can have strong influence on the accuracy of the dating information, but can only be explored for each grave individually.
@@
 This paper is based on ^14^C dates which were prepared for the dating of graves.
-There is a fundamental break between that great, theoretical archaeology, which wants to develop comprehensive narratives and only pay quantitative attention to the individual case, and the on-site research or emergency excavation, which has to dedicate a measure of respect and sometimes a lot of time to each grave.
-In this paper, ^14^C data are converted into a proxy of calibrated density distributions within artificial working regions, which is completely detached from the individual context.
-^14^C data on graves are used to attribute a feature to the absolute chronological context or to better describe the relative development within a site.
-The questions underlying the creation of a ^14^C date are therefore not equivalent to those imposed in this paper.
+There is a semantic divisiveness between great, theoretical archaeology, which wants to develop comprehensive narratives and only pay quantitative attention to the individual case, and the on-site research or emergency excavation, which has to dedicate a measure of respect and sometimes a lot of time to each grave.
+
+In this paper, ^14^C data are converted into a proxy of calibrated density distributions within artificial working regions.
+An endeavour which is completely detached from the individual context.
+In contrast, a single ^14^C date on a grave is usually used to attribute a feature to the absolute chronological context or to better describe the relative development within a site.
+The questions underlying the creation of a ^14^C date are therefore fundamentally different from those imposed here.
 

### 2019-03-31 review/answer_to_review.md
update of review answer
some last work on the answer to the reviewers

@@
 The challenges are not confined on the amount of data, but also concern its meaning and origin.
-The supplementary remarks are extensively long to cover these problems.
+The supplementary remarks cover these problems in detail.
 
-I now modified the Figures 3&4 as well as the Table 1 to provide a better overview over the actual amount of data available.
+I now modified the Figures 3&4 as well as the Table 1 to provide an overview over the actual amount of data available.
 I have also added several short comments/sentences in various sections of the paper to better address the small sample size issue.
@@
 
-After careful deliberation, I have decided to follow this advice.
-The simulation (1) does not add information that would be absolutely essential for the core topics, (2) reduces the overall comprehensibility due to the terminology that has to be introduced to describe it and (3) cannot be adequately presented and discussed here without leaving the initial scope of the paper.
-I will prepare another paper in the future that will present this simulation and other simulation approaches.
+After careful deliberation, I decided to follow this advice.
+The simulation (1) does not add information that would be absolutely essential for the core topics, (2) reduces the overall comprehensibility due to the terminology that has to be introduced to describe it and (3) cannot be adequately presented and discussed here without overstraining the initial scope of the paper.
+
+I removed the simulation here. A future paper will present it together with other simulation approaches.
 
@@
 The decision to create discrete spatial area classes is necessary to establish subsets of dates that can be used for time series calculation.
-I'm not aware of a methodologival approach where time series could be constructed without some sort of aggregation.
-The equal area size approach applied here for the region defition causes size of the areas to be very big due to the currently overall small amount of data.
-I expect this to change in the future and the analysis can be redone then with more samples and much smaller areas to understand the definitely much more complex processes within the macro regions established here.
+I'm not aware of a methodological approach where time series could be constructed without some sort of aggregation.
+The equal area size approach applied here for the region definition causes their size to be very big due to the currently overall small amount of data.
+I expect this to change in the future and the analysis can be redone then with more samples and much smaller areas.
 
-So: The criticism listed here is valid and I added it in the discussion.
-I considered a complete revision of the time series construction methodology, but since I (1) am still convinced that it is scientifically sound and reproducible, and (2) do not know a fundamentally better approach that avoids these completely risks, I suggest publishing the results as they are for the community of practice to discuss and improve.
+So: The criticism listed here is valid and I added it in the *Discussion* section.
+I considered a complete revision of the time series construction methodology, but since I (1) am still convinced that it is scientifically sound and reproducible, and (2) do not know a fundamentally better approach that avoids these risks completely, I suggest publishing as it is for the community of practice to discuss and improve.
 
@@
 
-I added links to the relevant repositories.
+This is a result of miscommunication with the special issue editors. I added links to the relevant repositories now.
 
@@
 
-Rephrased: "How does the cultural distance between regions that results from changes in burial traditions develop, and how is this cultural distance related to spatial distance?"
+Rephrased.
 
@@
 
-Rephrased: "This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries, as it assumes *Social learning* with transmission, selection and mutation of ideas [...]"
+Rephrased.
 
@@
 
-Rephrased: "Evolutionary archaeology, the branch of archaeology that explicitly adopted Cultural evolution theory, [...]"
+Rephrased.
 
@@
 
-Rephrased: "[...] *drift* (in case of multiple, equivalent traits a single one will dominate over time due to random effects) [...]"
+Rephrased.
 
@@
 
-TODO
-
 > Page 4, Column 2, Lines 43-56. I think it is worth computing partial mantel tests so that the correlation between burial type and construction can take into account geography.
 
-Done and included in Table 2.
-
-TODO
+A good idea! Done. I split Table 2 to present the results, added them to the new version of Figure 6 and addressed them in the text.
 
@@
 
-Figure 6 is much simpler now without simulation results. It only contains a time series of distance correlation.
-
-TODO
+Figure 6 is much simpler now without simulation results. It only contains time series of distance correlation.

### 2019-05-23 review/answer_to_review_2.md
identified TODOs in the reviews

# Associate Editor

Comments to the Author:

> Dear author, 
> I agree with the comments of the reviewer and the remaining points have to be addressed before we can accept this paper.  As some points imply adding new analyses with different tools, it may not be reasonable to  do it in the scope of this paper. But at least a stronger and clearer emphasis on the flaws and weaknesses of the statistical analysis should be done.  The alternatives approach proposed by the reviewer should be at the very least mentioned. It is not a problem if the result are, on the overall, not highly significant __as far as__ it's clear in the text. The approach and the methods are already a a valuable contribution to our special issue. I will look forward to receiving a revised manuscript from you.

TODO

# Reviewer: 1

Comments to the Author:

> The methods adopted in the paper are very innovative and have been rarely implemented to analise large 14C-dates datasets. I am strongly conviced of the quality of the manuscript, that has improved in this second version.

> Only a small remark, the publication year of the paper "Capuzzo G, Boaretto E, Barceló JA" is 2014.

TODO

# Reviewer: 2

Comments to the Author:

> I read the revised version of the manuscript and I think the authors have addressed most of the issues I raised, and acknowledged others in the text. I am still unconvinced by the creation of the artificial subsets and think that perhaps using temporal distance matrix (see for example Shennan et al 2015 on Evolution and Human Behaviour) might have been a solution, but I also see the point that such approach would not directly explore time-series. Ultimately this is an instance of the so-called MAUP (Modifiable Areal Unit Problem) in geography, and perhaps there are some more nuanced approaches out there. I think for the purpose of this paper acknowledging the problem is more than sufficient.  I also still have some issues with how sample size is effectively being ignored once proportions are computed, and I think this should be emphasised more strongly as I expect not everyone would be reading beyond the conclusion section.

TODO

> Please find below some further minor comments on the revised text. Overall, I think this is a great piece of work and I will look forward to follow-ups, including the simulation work that was excluded from this manuscript.

> Page 2, Column 1, Line 29. ‘simplify survival’ → increase?

TODO

> Page 2, Column 1, Line 42. ‘seemlingly random’  This is one of those tricky phrasing that has different meaning depending on the readership. ‘Random’ to me almost equals to unbiased transmission with drift, and stating that it is ‘seemlingly’ reads as if in reality this is not the case. This is perfectly fine, but it is slightly at odds with the definition of ‘style’ and the implicit assumption of neutrality. I would consider rephrasing, just to avoid potential misunderstandings.

TODO

> Page 2, Column 2, Line 41. the definition of drift here is more focused on its consequence. The fixation scenario presented applies only in regimes with no mutation/innovation. Perhaps specify that with no innovation a single trait will dominate by chance alone due to chance alone?

TODO

> Page 3, Column 1, Lines 2-6. Great!

> Page 3, Column 2, Line 59. ‘the region are circular’ Why this shape if it causes overlap? Why not hexagons?

TODO

> Page 5, Figure 3. Just an idea for a follow-up study. Given these are effectively adoption curves, one could fit a logistic model to assess differences in the rate of adoption and estimates on the timing of earliest uptake. It would be also very interesting to determine whether indeed there is a long-tail in southern Germany and what this means (cf. Henrich 2001 on American Anthropologist, but see also recent debates on adoption curves, e.g. https://www.biorxiv.org/content/biorxiv/early/2017/11/16/159038.full.pdf). Food for thoughts.

TODO

> Page 5, Column 1, Lines 56-. Please specify the R package used for the Mantel and Partial Mantel Tests.

TODO

> Page 5. I still think a permutation test would be more robust in this case. One could compare the observed correlation between two pairs of proportion time series against a simulation set where each iteration is the measured correlation between the proportion time-series of two artificial sets generated by randomly permuting the location label of each radiocarbon date. Pairs with a smaller number of dates will generate a wider range of correlation values and will be harder to test. One could adopt a more sophisticated solution by permuting groups of dates that occurred together in order to handle sample interdependence. Might be worth considering doing this for a follow-up study.

TODO

> Page 6, Column 2, Lines 46-49. ‘If one believes the sample, then a sudden, radical change from burial in flat graves to hill graves took place in Southern Germany at the beginning of the Middle Bronze Age.’  The sample size here is 8 dates. I appreciate the cautionary ‘if one believes the sample’ but this is a bit of a stretch to say the least, and readers that are numerically less literate might be misled. The same applies to other contexts with extremely small sample sizes and I think adding confidence intervals to those proportions would help.

TODO

> Table 3 & Figure 6. The figure nicely summarises the trend detailed on Table 3, but does not distinguish instances where the correlation were identified as signficant vs those that were not. Indeed in no cases the analysis shows a signficant correlation under the Partial Mantel test, whilst only few chronological blocks are showing signficant correlations. I suggest highlighting this by perhaps changing the the point symbology. It is also worth noting that if anything there is a bias towards type I error given: a) that we are dealing with multiple testing; and b) these analyses do not take into account sample sizes which are often very small. P-values are not the core issue here, but rather the fact that some way to at least visually convey this uncertainty is much needed.

TODO

> Page 11, Column 1, line 54. Yes, but none of the negative correlation are statistically signficant!

TODO

### 2019-05-23 article/02_Materials_and_Methods.Rmd
full stop

@@
 The code for this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible [@marwickPackagingData2018].
-It is available in a public Repository on GitHub (\url{https://github.com/nevrome/cultrans.bronzeageburials.article2019}) and with a permanent digital object identifier (DOI) on the Open Science Framework Platform (\url{http://dx.doi.org/10.17605/OSF.IO/B6NP2}) A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
+It is available in a public Repository on GitHub (\url{https://github.com/nevrome/cultrans.bronzeageburials.article2019}) and with a permanent digital object identifier (DOI) on the Open Science Framework platform (\url{http://dx.doi.org/10.17605/OSF.IO/B6NP2}).
+A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
 

### 2019-05-23 review/answer_to_review_2.md
wrong year in reference
phrasing
a hexagonal region grid?

@@
 
-TODO
+Fixed.
 
@@
 
-TODO
+Rephrased.
 
@@
 
-TODO
+This is true.
+I added a sentence to point this out.
+For this paper I will use the established regions, because changing them at this point would render the area discussions useless.
+In the future more data might be available and it should be possible to establish a lot more fine grained, hexagonal or rectangular region grid.
 

### 2019-05-23 review/answer_to_review_2.md
R package for mantel test
seemingly random: neutrality

@@
 
-TODO
+You're right.
+I thought a lot about this and finally decided to simply remove the term ‘seemlingly random’.
+In my opinion the binary distinction between style and function reaches its limit with burial traditions.
+In many cultural contexts, death and funerals are strongly reflected.
+Much more thought is given to this than to patterns on pots.
+The simulation paper will have to look at this difference more carefully, because there the neutrality is an even more important assumption.
 
@@
 
-TODO
+Done. ecodist::mantel()
 

### 2019-05-23 article/01_Introduction.Rmd
phrasing
seemingly random: neutrality

@@
 *Evolutionary archaeology*, the branch of archaeology that explicitly adopted Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is relevant for burial traditions.
-Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
+Some types of human behaviour increase the chances of survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
 Without resorting to the inappropriate simplifications of Evolutionism, their prevalence generally happens to increase over time.
@@
 Burial traditions can be described as a special form of these *stylistic traits*.
-Special because of their strong dependence on religious and spiritual beliefs and their connection with the emotionally stirring topic of death, but still subject to sometimes rapid and seemingly random fashion changes.
+Special because of their strong dependence on religious and spiritual beliefs and their connection with the emotionally stirring topic of death, but still subject to sometimes rapid fashion changes.
 

### 2019-05-23 review/answer_to_review_2.md
definition of drift
adoption model

@@
 You're right.
-I thought a lot about this and finally decided to simply remove the term ‘seemlingly random’.
+I thought a lot about this and finally decided to simply remove the term "seemlingly random".
 In my opinion the binary distinction between style and function reaches its limit with burial traditions.
@@
 
-TODO
+Rephrased. Added the important "no mutation or innovation" conditions.
 
@@
 
-TODO
+Well noted. Thank you! Incidentally for Southern Germany there is a paper by Frank Falkenstein that reconstructed the adoption rate of these burial traditions based on archaeologically dated graves:
+
+Falkenstein, F.
+(2017).
+Zum Wandel der Bestattungssitten von der Hügelgräber- zur Urnenfelderkultur in Süddeutschland.
+In Brandherm, D.
+and Nessel, B., editors, Phasenübergänge Und Umbrüche Im Bronzezeitlichen
+Europa, number 297 in UPA, pages 77–96. Bonn.
+https://www.academia.edu/36327050/Zum_Wandel_der_Bestattungssitten_von_der_H%C3%BCgelgr%C3%A4ber-_zur_Urnenfelderkultur_in_S%C3%BCddeutschland
 

### 2019-05-23 review/answer_to_review_2.md
need for future work

@@
 I think for the purpose of this paper acknowledging the problem is more than sufficient.
-I also still have some issues with how sample size is effectively being ignored once proportions are computed, and I think this should be emphasised more strongly as I expect not everyone would be reading beyond the conclusion section.
+
+Thank you very much for your review.
+You gave me some great ideas how this problem could be approached in a better way and your review will be the first thing I will go to, when I have the opportunity to continue the work on this front.
+For measuring the degree of interaction in a better way and withput regions, I recently got some inspiration from a paper by Loog et al 2017:
+
+Loog, L., Lahr, M.
+M., Kovacevic, M., Manica, A., Eriksson, A., & Thomas, M.
+G.
+(2017).
+Estimating mobility using sparse data: Application to human genetic variation.
+Proceedings of the National Academy of Sciences, 114(46), 12213–12218.
+https://doi.org/10.1073/pnas.1703642114
+
+Two major take aways: (1) Their mobility estimator could be used for a better, region independent statistic for the general degree of interaction.
+(2) I could as well use a moving window instead of 200 years time slices (Figure 6).
+
+I added some sentences in the conclusion to give an outlook on future work.
+
+TODO
+
+> I also still have some issues with how sample size is effectively being ignored once proportions are computed, and I think this should be emphasised more strongly as I expect not everyone would be reading beyond the conclusion section.
+
+Added sentence to the conclusion to make this very clear.
 
@@
 
-TODO
+P
 

### 2019-05-23 review/answer_to_review_2.md
the negative correlation is not significant

@@
 You gave me some great ideas how this problem could be approached in a better way and your review will be the first thing I will go to, when I have the opportunity to continue the work on this front.
-For measuring the degree of interaction in a better way and withput regions, I recently got some inspiration from a paper by Loog et al 2017:
+For measuring the degree of interaction in a better way and without artificial regions, I recently got some inspiration from a paper by Loog et al 2017:
 
@@
 
-P
+Well noted for the follow-up study.
 
@@
 
-TODO
+Rephrased to address this.

### 2019-05-23 article/04_Discussion.Rmd
the negative correlation is not significant

@@
 
-Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta} and \ref{tab:dactb}, Figure \ref{correlationtimeseries}).
-Only for certain periods of time this correlation is visible in the Radon-B data and the *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation.
+Against this background, the observation that there is no permanent, significant positive correlation either between the variables or with spatial distance is all the more astonishing (Table \ref{tab:dacta} and \ref{tab:dactb}, Figure \ref{correlationtimeseries}).
+Only for certain periods of time this correlation is visible in the Radon-B data.
+The *burial construction* distance generally shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation (also not statistically significant!).
 If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000 calBC) and a decline (after 2000 calBC) of cultural-spatial-distance correlation.

### 2019-05-23 analysis/code/04_plot_scripts/correlation_time_series.R
adjusted figure 6 to include statistical significance

@@ -42,7 +42,25 @@ mantel$time <- factor(
   levels = gsub("-(?=[0-9])", "", levels(mantel$time), perl = TRUE)
 )
 
+# establish significance classes and create subsets only with significant results
 
+mantel <- mantel %>%
+  dplyr::mutate(
+    signif_class = cut(
+      mantel$signif,
+      breaks = c(0, 0.05, 0.1, 1),
+      labels = c("p < 0.05", "p < 0.1", "p > 0.1 (not significant)"),
+      ordered_result = TRUE
+    )
+  )
+
+signifs_0.05 <- mantel %>% dplyr::filter(
+  signif_class == "p < 0.05"
+)
+
+signifs_0.1 <- mantel %>% dplyr::filter(
+  signif_class == "p < 0.1"
+)
 
 #### plot ####
 p <- ggplot() +
@@ -71,6 +89,42 @@ p <- ggplot() +
       linetype = context,
       size = context
     )
+  ) +
+  geom_point(
+    data = signifs_0.05,
+    mapping = aes(
+      x = time,
+      y = statistic
+    ),
+    shape = 0,
+    size = 8,
+    stroke = 2,
+    colour = "red"
+  ) +
+  annotate(
+    geom = "text",
+    x = "1800 - 1600", y = 0.48,
+    label = "p < 0.05",
+    color = "red",
+    size = 8
+  ) +
+  geom_point(
+    data = signifs_0.1,
+    mapping = aes(
+      x = time,
+      y = statistic
+    ),
+    shape = 0,
+    size = 8,
+    stroke = 2,
+    colour = "orange"
+  ) +
+  annotate(
+    geom = "text",
+    x = "1600 - 1400", y = 0.4,
+    label = "p < 0.1",
+    color = "orange",
+    size = 8
   ) +
   guides(
     fill = guide_legend(override.aes = list(size = 15))

### 2019-05-23 article/02_Materials_and_Methods.Rmd
a hexagonal region grid?
R package for mantel test
adjusted figure caption accordingly

@@
 That means some graves are counted towards not only one but two regions.
+A hexagonal grid layout could be considered for future analysis to avoid this overlap.
 The creation of these macro-regions is necessary to establish discrete observation units, but they are also the source of methodological problems (see section \ref{discussion}).
@@
 The two matrices of each section were tested for correlation with the *Mantel test*, which uses permutations of potentially correlating matrices to determine how strongly random reordering can alter an underlying correlation coefficient [@mantelDetectionDiseaseClustering1967].
+The Mantel test was calculated with the function `mantel()` in the R package *ecodist* v.2.0.1 [@ecodistrpackage2007].
 Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test (Table \ref{tab:dactb} C).
@@
 The results for the correlation of the two cultural distances (Table \ref{tab:dactb}) are displayed with a strong dotted line \textit{without} the correction for spatial distance (only here: Pearson correlation coefficient), and with a weak dotted line for the result \textit{with} the aforementioned correction.
-The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}
+The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.
+Statistically significant correlation (by p-Value) is indicated as such.}
 \end{figure}

### 2019-05-23 review/answer_to_review_2.md
answer to reviewer concerning the figure 6 problem

@@
 
+I added a visualisation of the p-values to figure 6 now.
+I understand your concerns, but I do not know how I could modify the figure with the available data to convey all the underlying uncertainty.
+A permutation test as suggested above may be the way forward here.
+I added a sentences on page 5 to explain the figure with these problems in mind.
+
 TODO

### 2019-05-23 analysis/code/01_data_preparation_scripts/02_prepare_c14_data.R
as.tibble -> as_tibble

@@ -23,7 +23,7 @@ save(radonb, file = "analysis/data/tmp_data/radonb.RData")
 load("analysis/data/tmp_data/radonb.RData")
 
 dates <- radonb %>%
-  tibble::as.tibble() %>%
+  tibble::as_tibble() %>%
   # remove dates without age
   dplyr::filter(!is.na(c14age) & !is.na(c14std)) %>%
   # remove dates outside of theoretical calibration range

### 2019-05-23 analysis/code/02_cultural_distance_scripts/02a_burial_construction_sed.R

@@ -22,7 +22,7 @@ regions_grid <-
     regionA = regions, regionB = regions, time = timesteps,
     stringsAsFactors = FALSE
   ) %>%
-  tibble::as.tibble() %>%
+  tibble::as_tibble() %>%
   dplyr::left_join(
     long_prop,
     by = c("regionA" = "region", "time" = "timestep")

### 2019-05-26 article/05_Conclusion.Rmd
some important changes to the conclusion

@@
 
-*The main trends in the distribution of burial rites in Bronze Age Europe can be detected in bulk radiocarbon data* and indeed *they allow long-term observations concerning cultural distance*.
-Radiocarbon dates can be transformed to meaningful regional time series, if they can be linked to the relevant meta information.
+*The main trends in the distribution of burial rites in Bronze Age Europe can be detected in bulk radiocarbon data* (Figure \ref{development_burial_type} B and \ref{development_burial_construction} B) and indeed *they allow long-term observations concerning cultural distance* (Figure \ref{sed_region_matrix}, Table \ref{tab:dacta} and Table \ref{tab:dactb} and Figure \ref{correlationtimeseries} B), although with the low amount of data currently available, *most findings should be taken with great caution* (Table \ref{tab:gprcrosstab}, Figure \ref{development_burial_type} A and \ref{development_burial_construction} A, Figure \ref{correlationtimeseries} A).
+In this paper sample size is effectively being ignored once proportions of burial rites are computed.
+
+Even small amounts of radiocarbon dates can be transformed to meaningful regional time series, if they can be linked to the relevant meta information.
 Although only `r r("gpr size", sf_desc)` graves -- a small fraction of maybe hundreds of thousands burials that can be estimated for the European Bronze Age [@nikulkaArchaologischeDemographieMethoden2016] -- have been available for this analysis, the resulting reconstruction of general burial tradition trends in wast areas of Central, Northern and North-western Europe comes close to the one achieved by literature research.
@@
 
-Future research is required to untangle these interactions -- of course on the basis of more precise data.
-This endeavour will not end with the simple description of social learning cascades, but with the analysis of the initial reasons for diversification.
 From the point of view of Cultural complexity -- the question, which factors accelerate or decelerate the variability of human behaviour [@henrichDemographyCulturalEvolution2004; @mesoudiWhatCumulativeCultural2018; @vaesenPopulationSizeDoes2016] -- the long term dominance fluctuations of one burial rite over the other are remarkable.
@@
 
-Beyond these intrinsic results, it is also important to ask what this paper means for the study of human behaviour in general. Two possible answers:
+Future research is required to untangle these interactions -- of course on the basis of more and more precise data.
+In this future work the methodology should be refined and extended in several ways.
+Some important ideas:
+
+1. Replacing time slices in the correlation analysis by a moving window approach to get a contious interaction proxy time series.
+
+2. Applying a model based analysis of the idea proportion adoption curve time series and their regional differences to get a better understanding of mutual influence, earliest uptake and laggards.
+
+3. Comparing the real world correlation time series with ones constructed from simulated, randomly permutating observations to test the significance of seemingly positive or negative correlation.
+
+4. Abolishing the separation into artificial regions for the cultural distance and correlation analysis in favour of an approach based on individual observations [@loogEstimatingMobilityUsing2017].
+
+Finally it is also important to ask what this paper means for the study of human behaviour in general. Two possible answers:
 

### 2019-05-26 article/05_Conclusion.Rmd
spelling

@@
 
-1. Replacing time slices in the correlation analysis by a moving window approach to get a contious interaction proxy time series.
+1. Replacing time slices in the correlation analysis by a moving window approach to get a continuous interaction proxy time series.
 
@@
 
-3. Comparing the real world correlation time series with ones constructed from simulated, randomly permutating observations to test the significance of seemingly positive or negative correlation.
+3. Comparing the real world correlation time series with ones constructed from simulated, randomly permuting observations to test the significance of seemingly positive or negative correlation.
 

### 2019-05-26 article/03_Results.Rmd
modified presentation of the time series result

@@
 
-Figure \ref{development_burial_type} and \ref{development_burial_construction} show the region-wise time series of trait occurrence.
-If we consider these results in isolation and also ignore issues of small sample size, we can construct the following narrative from them: \newline
+Figure \ref{development_burial_type} B and \ref{development_burial_construction} B show the region-wise time series of proportional trait prevalence.
+They are based on very (!) small sample quantities (Figure \ref{development_burial_type} A and \ref{development_burial_construction} A) and are therefore not statistically significant.
+Despite this obvious limitation it's possible to conduct a thought experiment and consider these results in isolation.
+A comparison with published archaeological observations based on literature study is provided in section \ref{discussion}.
+Carrying out this experiment yields the following narrative: \newline
 
-The overall picture is that during the Bronze Age in Europe there was a trend away from body burial towards cremation.
+*The overall picture is that during the Bronze Age in Europe there was a trend away from body burial towards cremation.
 Around 2200 calBC cremation burials were almost unknown in Poland, Southern Germany, Northeastern France and Northern Germany.
@@
 In Northeastern France and Northern Germany this change took place already in the Middle Bronze Age.
-In both contexts inhumation continued to play an important role.
+In both contexts inhumation continued to play an important role.*
 
-With regard to the question of grave mound construction the picture is considerably more heterogeneous.
+*With regard to the question of grave mound construction the picture is considerably more heterogeneous.
 In Austria and the Czech Republic flat graves were the dominant form of burial until the Late Bronze Age, but mounds became more relevant from the Middle Bronze Age onwards.
@@
 Burial mounds were dominant throughout the Benelux region, but in the Early and Late Bronze Ages flat graves also occurred to a significant extent.
-In England flat graves were a short phenomenon in the Early Bronze Age, which did not recur later.
-\newline
+In England flat graves were a short phenomenon in the Early Bronze Age, which did not recur later.* \newline
 

### 2019-05-26 review/answer_to_review_2.md
some important changes to the conclusion
about the changes in figure 6

--- 2019-05-23 review/answer_to_review_2.md
@@
 I will look forward to receiving a revised manuscript from you.
-
-TODO
 
@@
 
-TODO
-
 > I also still have some issues with how sample size is effectively being ignored once proportions are computed, and I think this should be emphasised more strongly as I expect not everyone would be reading beyond the conclusion section.
 
-Added sentence to the conclusion to make this very clear.
-
-TODO
+I changed the wording of the conclusion to make this very clear.
 
@@
 
-Well noted for the follow-up study.
+Well noted for the follow-up study. I also mention this approach now in the conclusion.
 
@@
 
-TODO
+I modified the small, introductory paragraph of this section to make it clear, that this is a thought experiment based on a very small amount of data. I also set the narrative in italics.
 
@@
 
-TODO
+To address this issue I modified Figure 6. I added...
+
+1. ... an additional graph Figure 6 A that shows the median (with an lower to upper quartile ribbon) of the graves per year and region.
+
+2. ... an indication of the mean of this medians for each 200 years time window in Figure 6 B. The square markers show on how much data the correlation observations are based.
+
+3. ... an indication of the p-values in Figure 6 B to show which correlations are significant.
+
+I also modified the description of this figure in the Materials and Methods as well as in the Results and the Discussion sections accordingly.
+I believe these changes are sufficient to visually convey the underlying uncertainty.
 

### 2019-05-26 analysis/code/04_plot_scripts/correlation_time_series.R
added labels to correlation plot

@@ -384,11 +384,12 @@ correlation_time_series_plot <- ggplot() +
 combined_plot <- plot_grid(
   data_amount_plot,
   correlation_time_series_plot,
+  labels = c("A", "B"),
   nrow = 2,
   rel_heights = c(1, 3),
   align = "v",
   label_size = 35,
-  vjust = 1.0,
+  vjust = 1.5,
   hjust = 0
 )
 

### 2019-05-26 article/02_Materials_and_Methods.Rmd
modified figure caption

@@
 \includegraphics[width=\columnwidth]{correlation_time_series.jpeg}
-\caption{\label{correlationtimeseries}Development of correlation between different measures of cultural and spatial distance.
+\caption{\label{correlationtimeseries}(\textbf{A}) Development of the median numbers of graves per region and year for the variables \textit{burial type} and \textit{burial construction}.
+The ribbon indicates the range between the lower and the upper quartile.
+(\textbf{B}) Development of the correlation between different measures of cultural and spatial distance.
 The results for the correlation of cultural and spatial distance (Table \ref{tab:dacta}) are displayed with a solid (\textit{burial type}) and a dashed (\textit{burial construction}) line.
@@
 The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.
-Statistically significant correlation (by p-Value) is indicated as such.}
+Statistically significant correlation (by p-Value) is indicated as such.
+The intensity of the small rectangle around each observation visualizes the mean of the medians in (A) for each 200 years time window and the respective variables.}
 \end{figure}

### 2019-05-26 article/05_Conclusion.Rmd
phrasing

@@
 
-4. Abolishing the separation into artificial regions for the cultural distance and correlation analysis in favour of an approach based on individual observations [@loogEstimatingMobilityUsing2017].
+4.
+Abolishing the separation into artificial regions for the cultural distance and correlation analysis in favour of an approach based on individual observations [e.g.
+@loogEstimatingMobilityUsing2017].
 

### 2019-05-26 article/02_Materials_and_Methods.Rmd
reflecting the changes in Figure 6 in the text

@@
 
-Figure \ref{correlationtimeseries} shows the development of the correlation coefficient values obtained via Mantel statistics (Table \ref{tab:dacta} & \ref{tab:dactb}) during the succession of the seven relevant 200 years time slots.
+Figure \ref{correlationtimeseries} B shows the development of the correlation coefficient values obtained via Mantel statistics (Table \ref{tab:dacta} & \ref{tab:dactb}) during the succession of the seven relevant 200 years time slots.
+Because of the highly suggestive nature of this figure, it is complemented by Figure \ref{correlationtimeseries} A, which shows the (very low!) median number of graves available for each region and year to calculate these correlations.
 

### 2019-05-26 article/03_Results.Rmd
reflecting the changes in Figure 6 in the text

@@
 
-Table \ref{tab:dactb} shows that the distance networks of *burial type* and *burial construction* are not equivalent.
+Table \ref{tab:dactb} and Figure \ref{correlationtimeseries} show that the distance networks of *burial type* and *burial construction* are not equivalent.
 There is also no significant negative correlation.
@@
 Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600 calBC the 5% significance level of the Mantel test is almost undercut, thus suggesting a statistically significant relationship.
-This moment of strong correlation is though not induced by similarity along the spatial dimension: The effect is still visible in the result of the Partial Mantel test, where the spatial distance is artificially counterbalanced (see also Figure \ref{correlationtimeseries}).
+This moment of strong correlation is though not induced by similarity along the spatial dimension alone: The effect is still partially visible in the result of the Partial Mantel test, where the spatial distance is artificially counterbalanced.
 
-For the relationship between the SED networks and the spatial distance network several observations result from Table \ref{tab:dacta}:
+For the relationship between the SED networks and the spatial distance network several observations result from Table \ref{tab:dacta} and Figure \ref{correlationtimeseries}:
 
@@
 
-Looking at the SED network of *burial construction*, a slightly different conclusions can be drawn: The distance values here are larger overall, and the SED differences within the individual spatial distance classes are clearly more accentuated.
-As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation, though it is only significant for the time slot between 2000 and 1400 calBC.
+Looking at the SED network of *burial construction*, a slightly different conclusions can be drawn: The distance values here are larger, and the SED differences within the individual spatial distance classes are clearly more accentuated.
+As well as in *burial type* the Early Bronze Age is by far the phase with the most substantial spatial correlation.
+But it is only significant for the time slot between 2000 and 1400 calBC.
 There is no time period with negative mean correlation for *burial construction* cultural distance and spatial distance, but also here the correlation coefficient approaches zero in the Late Bronze Age.
+These latter results in particular, however, are based on unsatisfactory few observations.
 
-In general, cultural distance defined here does not seem to correlate strongly with spatial distance, especially in the Middle and Late Bronze Age.
+In general, cultural distance as defined here does not seem to correlate strongly with spatial distance, especially in the Middle and Late Bronze Age.
+However, this hypothesis is not conclusive at the moment as it is derived from a very small sample size.

### 2019-05-26 article/04_Discussion.Rmd
reflecting the changes in Figure 6 in the text

@@
 The quantitative representation of an archaeological narrative paths the way for a diverse set of methods to be applied and the analysis undertaken here is merely the beginning.
+Unfortunately, the small sample size makes it very difficult to obtain meaningful, significant results in this context.
 
@@
 
-Against this background, the observation that there is no permanent, significant positive correlation either between the variables or with spatial distance is all the more astonishing (Table \ref{tab:dacta} and \ref{tab:dactb}, Figure \ref{correlationtimeseries}).
+Against this background, the observation that there does not seem to be permanent, significant positive correlation neither in between the variables nor with spatial distance is all the more astonishing (Table \ref{tab:dacta} and \ref{tab:dactb}, Figure \ref{correlationtimeseries} B).
 Only for certain periods of time this correlation is visible in the Radon-B data.
-The *burial construction* distance generally shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation (also not statistically significant!).
+The *burial construction* distance generally shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation (not statistically significant!).
 If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000 calBC) and a decline (after 2000 calBC) of cultural-spatial-distance correlation.
@@
 
-Keeping this in mind, the surprising result is still valid and has to be explained.
+Keeping this in mind, the surprising result still can be discussed.
 One interpretation might be, that the Early Bronze Age was a time of rising cultural exchange and religious homogenization, while in later periods different domains secluded themselves more strongly.
@@
 The latter may have been of subordinate relevance in the Tumulus culture context.
-With the rise of the Urnfield culture this phenomenon could have been reversed as *burial type* became an expression of something that could have been one of the first formalized religions in Europe: The spread of the Urnfield Culture was an extraordinary development and possibly accompanied by a structured, religious dogma that could have built up around the symbolic content of cult wagons, sun barks and waterfowl [@bilicSwanChariotSolar2016; @kimmig_seevolkerbewegung_1964; @molloy_european_2018].
+With the rise of the Urnfield culture this phenomenon could have been reversed as *burial type* became an expression of something that could have been one of the first formalized religions in Europe: The spread of the Urnfield Culture was an extraordinary development and possibly accompanied by a structured, religious dogma built up around the symbolic content of cult wagons, sun barks and waterfowl [@bilicSwanChariotSolar2016; @kimmig_seevolkerbewegung_1964; @molloy_european_2018].
 Of course these explanations could, if correct, only be considered on a big scale.

### 2019-05-26 article/05_Conclusion.Rmd
reflecting the changes in Figure 6 in the text

@@
 They yield the surprising observation that cultural distance and spatial distance do not significantly correlate.
-Although this interpretation must be treated with caution, it seems as if the spread of burial traditions are not simple *Diffusion of Innovation* events as introduced by @rogersDiffusionInnovations1962, but the result of more complex interactions between competing ideas.
+Although this result must be treated with caution because of the very small sample size, it seems as if the spread of burial traditions are not simple *Diffusion of Innovation* events as introduced by @rogersDiffusionInnovations1962, but the result of more complex interactions between competing ideas.
 

### 2019-05-26 article/05_Conclusion.Rmd
wording

@@
 
-*The main trends in the distribution of burial rites in Bronze Age Europe can be detected in bulk radiocarbon data* (Figure \ref{development_burial_type} B and \ref{development_burial_construction} B) and indeed *they allow long-term observations concerning cultural distance* (Figure \ref{sed_region_matrix}, Table \ref{tab:dacta} and Table \ref{tab:dactb} and Figure \ref{correlationtimeseries} B), although with the low amount of data currently available, *most findings should be taken with great caution* (Table \ref{tab:gprcrosstab}, Figure \ref{development_burial_type} A and \ref{development_burial_construction} A, Figure \ref{correlationtimeseries} A).
+*The main trends in the distribution of burial rites in Bronze Age Europe can be detected in bulk radiocarbon data* (Figure \ref{development_burial_type} B and \ref{development_burial_construction} B) and indeed *they allow long-term observations concerning cultural distance* (Figure \ref{sed_region_matrix}, Table \ref{tab:dacta} and \ref{tab:dactb} and Figure \ref{correlationtimeseries} B), although with the low amount of data currently available, *most findings should be taken with great caution* (Table \ref{tab:gprcrosstab}, Figure \ref{development_burial_type} A, \ref{development_burial_construction} A and \ref{correlationtimeseries} A).
 In this paper sample size is effectively being ignored once proportions of burial rites are computed.

### 2019-05-26 review/answer_to_review_2.md
final work on the answer to the reviewers

@@
-# Associate Editor
-
-Comments to the Author:
-
-> Dear author, 
-> I agree with the comments of the reviewer and the remaining points have to be addressed before we can accept this paper.
-As some points imply adding new analyses with different tools, it may not be reasonable to  do it in the scope of this paper.
-But at least a stronger and clearer emphasis on the flaws and weaknesses of the statistical analysis should be done.
-The alternatives approach proposed by the reviewer should be at the very least mentioned.
-It is not a problem if the result are, on the overall, not highly significant __as far as__ it's clear in the text.
-The approach and the methods are already a a valuable contribution to our special issue.
-I will look forward to receiving a revised manuscript from you.
-
 # Reviewer: 1
@@
 I am strongly conviced of the quality of the manuscript, that has improved in this second version.
+
+Great!
 
@@
 
-I changed the wording of the conclusion to make this very clear.
+I made some major changes to the conclusion to make this very clear.
 
@@
 For this paper I will use the established regions, because changing them at this point would render the area discussions useless.
-In the future more data might be available and it should be possible to establish a lot more fine grained, hexagonal or rectangular region grid.
+In the future more data might be available and it should be possible to establish an a lot more fine grained, hexagonal or rectangular region grid.
 
@@
 
-1. ... an additional graph Figure 6 A that shows the median (with an lower to upper quartile ribbon) of the graves per year and region.
+1. ... an additional graph Figure 6 A that shows the medians (with an lower to upper quartile range ribbon) of the graves per year and region.
 
-2. ... an indication of the mean of this medians for each 200 years time window in Figure 6 B. The square markers show on how much data the correlation observations are based.
+2. ... an indication of the mean of this medians for each 200 years time window in Figure 6 B. The square markers show, on how much data the correlation observations are based.
 
-3. ... an indication of the p-values in Figure 6 B to show which correlations are significant.
+3. ... an indication of the p-Values in Figure 6 B to show which correlations are significant.
 

### 2019-06-29 article/01_Introduction.Rmd
definition of drift
proofreading

@@
 Some types of human behaviour increase the chances of survival, like specific hunting techniques, superior tools or sowing more fruitful crops.
-Without resorting to the inappropriate simplifications of Evolutionism, their prevalence generally happens to increase over time.
-On the other hand, some behaviours are not related to survival and reproductive capacity, like e.g.
-ceramic decorations, traditional costume fashion or jewellery design.
+Without resorting to the inappropriate simplifications of *Evolutionism*, their prevalence generally happens to increase over time.
+On the other hand, some behaviours are not related to survival and reproductive capacity, such as ceramic decorations, traditional costume fashion or jewellery design.
 Despite or because their seemingly subordinate importance, these characteristics often pass through a breathtakingly fast development.
@@
 He identifies the essential phenomena *drift* (i.e.
-in case of multiple, equivalent traits in a population, a single one will dominate over time due to random effects) and *flow* (i.e.
+in case of multiple, equivalent traits in a population with no mutation or innovation, a single trait will dominate over time due to chance alone) and *flow* (i.e.
 intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology.

### 2019-06-29 article/02_Materials_and_Methods.Rmd
proofreading

@@
 The code for this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible [@marwickPackagingData2018].
-It is available in a public Repository on GitHub (\url{https://github.com/nevrome/cultrans.bronzeageburials.article2019}) and with a permanent digital object identifier (DOI) on the Open Science Framework platform (\url{http://dx.doi.org/10.17605/OSF.IO/B6NP2}).
+It is available in a public repository on GitHub (\url{https://github.com/nevrome/cultrans.bronzeageburials.article2019}) and with a permanent digital object identifier (DOI) on the Open Science Framework platform (\url{http://dx.doi.org/10.17605/OSF.IO/B6NP2}).
 A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].
@@
 
-Regarding the variables burial type and burial construction in the grave dataset, the distribution conditions shown in Table \ref{tab:gprcrosstab} apply.
+Regarding the variables *burial type* and *burial construction* in the grave dataset, the distribution conditions shown in Table \ref{tab:gprcrosstab} apply.
 

### 2019-06-29 article/04_Discussion.Rmd
proofreading

@@
 
-One major question asked in this paper was, *if the main trends in the distribution of burial rites in Bronze Age Europe can be detected in openly available bulk radiocarbon data?*.
+One major question asked in this paper was, if *the main trends in the distribution of burial rites in Bronze Age Europe can be detected in openly available bulk radiocarbon data?*.
 This is the case.
@@
 `r a("Southern Scandinavia")`:
-The region *Southern Scandinavia* includes all of Denmark with Jutland, Funen, Zeeland and all smaller islands as well as large parts of East of Southern Sweden, and coastal regions in Schleswig-Holstein and Mecklenburg-Western Pomerania.
+The region *Southern Scandinavia* includes all of Denmark with Jutland, Funen, Zeeland and all smaller islands as well as large parts of Southern Sweden, and coastal regions in Schleswig-Holstein and Mecklenburg-Western Pomerania.
 The ^14^C data are very well distributed across Jutland, Funen and the Swedish Skåne region.

### 2019-06-29 article/04_Discussion.Rmd
proofreading

@@
 
-The second important question posed for this paper was, *how cultural distance between regions develops, and how it is related to spatial distance?* While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one had a larger scope.
+The second important question posed for this paper was, *how cultural distance between regions develops, and how it is related to spatial distance?* While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one has a larger scope.
 The quantitative representation of an archaeological narrative paths the way for a diverse set of methods to be applied and the analysis undertaken here is merely the beginning.
@@
 2.
-The data used for this analysis was collected for another purpose and its application in this context requires a problematic shift in data meaning as explained in section \ref{data-meaning-discussion}.
+The data used for this analysis was collected for another purpose and their application in this context requires a problematic shift in data meaning as explained in section \ref{data-meaning-discussion}.
 

### 2019-06-29 article/05_Conclusion.Rmd
proofreading

@@
 
-From the point of view of Cultural complexity -- the question, which factors accelerate or decelerate the variability of human behaviour [@henrichDemographyCulturalEvolution2004; @mesoudiWhatCumulativeCultural2018; @vaesenPopulationSizeDoes2016] -- the long term dominance fluctuations of one burial rite over the other are remarkable.
+From the point of view of *Cultural complexity* -- the question, which factors accelerate or decelerate the variability of human behaviour [@henrichDemographyCulturalEvolution2004; @mesoudiWhatCumulativeCultural2018; @vaesenPopulationSizeDoes2016] -- the long term dominance fluctuations of one burial rite over the other are remarkable.
 Its standard explanation models (among others population density and environmental risk) can hardly be applied to stylistic variables.

### 2019-06-29 article/06_Supplementary_Remarks.Rmd
proofreading

@@
 The quantity and nature of the known graves cannot be representative.
-In this case it can be expected that further, deviant burial rituals were practised, which left no archaeologically visible remains.
+In this case it can be expected that other burial rituals were practised, which left no archaeologically visible remains.
 These rituals are invisible in the ^14^C data from Radon-B.
@@
 The respective contexts can be multifarious: similar rituals in different cultural contexts can have different meanings.
-For this paper, it is assumed -- with reference to *Cultural Evolution Theory* -- that despite this restriction, actions such as burning the dead or burying in burial mounds can be passed on across cultural and ethnic group boundaries through imitation and teaching.
+For this paper, it is assumed -- with reference to *Cultural Evolution Theory* -- that despite this restriction, behaviour such as burning the dead or burying in burial mounds can be passed on across cultural and ethnic group boundaries through imitation and teaching.
 To put it provocatively: cremation burial in the Hungarian Early Bronze Age is a manifestation of the same idea as cremation in the Northern Late Bronze Age.

### 2019-07-16 article/index.Rmd
marked rendered article as preprint

@@
 
+This is a preprint: the version of the paper that was accepted for publication. Please see the following reference for the final, published version:
+
+Schmid, C.
+(2019).
+Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and Northwestern Europe using radiocarbon data.
+Adaptive Behavior.
+https://doi.org/10.1177/1059712319860842
+
+Please note that the numerical results of the Mantel statistics are not exactly the same here compared to the final, published version. The Mantel test is a nondeterministic algorithm.
+
 ```{r child = '01_Introduction.Rmd'}
